### 2020-09-08 twccc.tex
twccc talk.

\documentclass[xcolor=dvipsnames, 8pt]{beamer} %
%\setbeamertemplate{navigation symbols}{}

\usetheme{SantaBarbara}

\definecolor{black}{HTML}{0A0A0A}
\definecolor{green}{HTML}{0420eb} 
\definecolor{darkgreen}{HTML}{008000}
\definecolor{gold}{HTML}{FFD000}
\setbeamercolor{normal text}{fg=black,bg=white}
\setbeamercolor{alerted text}{fg=red}
\setbeamercolor{example text}{fg=black}
\setbeamercolor{palette primary}{fg=black, bg=gray!20}
\setbeamercolor{palette secondary}{fg=black, bg=gray!20}
\setbeamercolor{palette tertiary}{fg=white, bg=green!80}
\setbeamercolor{block title}{fg=black,bg=gold!40}
\setbeamercolor{frametitle}{fg=white, bg=green!80}
\setbeamercolor{title}{fg=white, bg=green!80}

\newcommand{\svec}{\operatorname{svec}}

\usepackage[utf8]{inputenc}

\usepackage{tikz, tikzsettings}
\usepackage{verbatim}
\usepackage{amssymb}
\usepackage{amsmath}
%\usepackage{amsfonts}
\usepackage{algorithmic}
\graphicspath{{./}{./figures/}{./figures/presentation/}}
%\usepackage[version=4]{mhchem}
\usepackage{subcaption}
\usepackage[authoryear,round]{natbib}
%\usepackage{fancyvrb}
\usepackage{color}
\usepackage{colortbl}
\usepackage{xcolor}
\usepackage{physics}
\usepackage{pgfplots}
\usepackage{ragged2e}
%\pgfplotsset{compat=newest} 
%\pgfplotsset{plot coordinates/math parser=false}
%\usepackage{environ}
%\usetikzlibrary{decorations.markings}
%\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{shapes,calc,spy, calc, backgrounds,arrows, fit, decorations.pathmorphing, decorations.pathreplacing, matrix}
%\usepackage[absolute,overlay]{textpos}
\usepackage{caption}
\usepackage{graphicx}
%\usepackage[controls=true,poster=first]{animate}

\AtBeginSection[]
{\frame<beamer>{\frametitle{Outline}   
	\tableofcontents[currentsection, currentsection]}
	\addtocounter{framenumber}{-1}}
	%
%{
%	\frame<beamer>{ 
%		\frametitle{Outline}   
%		\tableofcontents[currentsection,currentsubsection]}

%}

\makeatother
\setbeamertemplate{footline}
{
	\leavevmode%
	\hbox{%
		\begin{beamercolorbox}[wd=.3\paperwidth,ht=2.25ex,dp=1ex,center]{author in head/foot}%
			\usebeamerfont{author in head/foot}\insertshortauthor
		\end{beamercolorbox}%
		\begin{beamercolorbox}[wd=.6\paperwidth,ht=2.25ex,dp=1ex,center]{title in head/foot}%
			\usebeamerfont{title in head/foot}\insertshorttitle
		\end{beamercolorbox}%
		\begin{beamercolorbox}[wd=.1\paperwidth,ht=2.25ex,dp=1ex,center]{date in head/foot}%
			\insertframenumber{} / \inserttotalframenumber\hspace*{1ex}
	\end{beamercolorbox}}%
	\vskip0pt%
}

\title[Data completion of initial gray box models for process control and optimization]{Data completion of initial gray box models for process control and optimization}
\date{September 15, 2020}
\author[Kumar and Rawlings]{\large Pratyush Kumar and James B. Rawlings}
\institute[UCSB]{
	\begin{minipage}{4in}
		\vspace{-10pt}
		\centering
		\raisebox{-0.1\height}{\includegraphics[width=0.25\textwidth]{UCSB_seal}}
		%\hspace*{.2in}
		%\raisebox{-0.5\height}{\includegraphics[width=0.25\textwidth]{jci_logo}}
	\end{minipage}
	\vspace{10pt}
	\newline
	{\large Department of Chemical Engineering}
	\vspace{10pt}
	\newline
	{\large TWCCC Meeting (Virtual)}}
	%\vspace{10pt}
	%\newline
	%{\large Austin, TX}}

\begin{document}

\frame{\titlepage}

% MPC intro frame
\begin{frame}{Idea.}
Hi.
\end{frame}

%\begin{frame}[allowframebreaks]{References}
%\bibliographystyle{abbrvnat}
%\bibliography{articles,proceedings,books,unpub}
%\end{frame}

\end{document}

### 2020-09-11 twccc.tex
talk structure.

@@
 
-\title[Data completion of initial gray box models for process control and optimization]{Data completion of initial gray box models for process control and optimization}
+\title[Data-based completion of initial grey-box models for process control]{Data-based completion of initial grey-box models for process control}
 \date{September 15, 2020}
@@
 % MPC intro frame
-\begin{frame}{Idea.}
+\begin{frame}{Grey-box models and undermodeling}
+Hi.
+\end{frame}
+
+\begin{frame}{Completing grey-box models from operational data}
+	Hi.
+\end{frame}
+
+\begin{frame}{Simulation example}
+	Hi.
+\end{frame}
+
+\begin{frame}{Simulation example -- validation data}
 	Hi.

### 2020-09-14 twccc.tex
ok, change the flash talk.

@@
 \newcommand{\svec}{\operatorname{svec}}
+\usepackage[version=4]{mhchem}
 
@@
 \begin{frame}{Grey-box models and undermodeling}
-Hi.
+\begin{block}{Process models}
+\begin{itemize}
+	\item Most first principle based grey-box modeling of chemical processes
+	are incomplete. For e.g, due to unknown reactions, incorrect 
+	quasi steady state assumption, etc.
+	\item Routine operational data can be used in conjuction 
+	with first principle knowledge to improve the prediction 
+	quality of the initial grey-box models.
+\end{itemize}
+\end{block}
+
+\begin{block}{Motivating example}
+	\begin{itemize}
+		\item Consider the reaction system
+		\begin{center}
+			\ce{A ->[k1] B},\hspace{0.2in} \ce{B ->[k2] C}, \hspace{0.2in} \ce{2B ->[k3] D}
+		\end{center}
+		\item A is the primary reactant, B is an intermediate, 
+		C and D are products.
+		\item Depending on the ratios of rate constants and measured 
+		concentrations, the reaction model may not be 
+		completely characterized with only grey-box models.
+	\end{itemize}
+\end{block}
 \end{frame}
@@
 \begin{frame}{Completing grey-box models from operational data}
-	Hi.
+	\begin{block}{Augment grey-box models with 
+				  black-box parameterized models}
+		\begin{align*}
+			\begin{bmatrix}
+			  x_G \\
+			  x_B
+			\end{bmatrix}^+ = \begin{bmatrix}
+			  f_G(x_G, \theta_G) + g_B(x_G, x_B, \theta_B)\\
+			  f_B(x_G, x_B, \theta_B) \\
+			\end{bmatrix}, \quad
+			y = h_G(x_G) + h_B(x_G, x_B)
+		  \end{align*}
+		  \begin{itemize}
+			\item $f_G(\cdot)$ and $h_G(\cdot)$ is the initial grey-box model. 
+			$g_B(\cdot)$, $f_B(\cdot)$, and $h_B(\cdot)$ is the parameterized black-box model (for e.g., a neural network).
+			\item $x_G$ and $x_B$ are the grey-box and black-box model states
+			respectively.
+		\end{itemize}
+	\end{block}
+	\begin{block}{System identification framework}
+		\begin{itemize}
+			\item Minimize the following 
+			multi-step-ahead prediction-error to compute the 
+			parameters $\theta_B$
+			\begin{align*}
+		\min_{x_G(0), \ x_B(0), \ \theta_B} \sum_{k=0}^{N_s-1} (\hat{y}(k) - y(k))^2
+			\end{align*}
+			\item $x_G(0)$, $x_B(0)$ are the initial states at the 
+			start of the forecasting horizon and $\hat{y}$ is the model forecast.
+		\end{itemize}
+	\end{block}
 \end{frame}
 
-\begin{frame}{Simulation example}
-	Hi.
+\begin{frame}{Reaction system example}
+	\begin{itemize}
+		\item Ethene reacts with chlorine to form Dichloroethane and Dichlorobutane.
+	\end{itemize}
+	\begin{align*}
+		\ce{3C2H4 + 2Cl2 -> C2H4Cl2 + C4H8Cl2}
+	\end{align*} 
+	\begin{itemize}
+		\item The reaction mechanism is -
+	\end{itemize}
+	\small{\begin{align*}
+		\ce{Cl2 + M &<=>[k1][k_{-1}] 2Cl* + M}\\
+		\ce{Cl* + C2H4 &->[k2] C2H4Cl*}\\
+		\ce{C2H4Cl* + Cl2 &->[k3] C2H4Cl2 + Cl*}\\
+		\ce{C2H4Cl* + C2H4Cl* &->[k4] C4H8Cl2}\\
+		\ce{Cl* +  C2H4Cl* &->[k5] C2H4Cl2}
+	\end{align*}}
+\normalsize{\begin{itemize}
+	\item The rate constants are such that a quasi steady state assumption (QSSA) on the species $\ce{Cl*}$ and $\ce{C2H4Cl*}$ is incorrect, and neural networks are used to improve the predictions of a  
+	grey-box model based on the QSSA assumption.
+\end{itemize}}
 \end{frame}
 
-\begin{frame}{Simulation example -- validation data}
-	Hi.
+\begin{frame}{Data requirements and model validation}
+	Show two plots. \\
+	On the left, show a performance metric
+	that shows the prediction of the overall model improves with data
+	relative to the grey-box model prediction. \\
+	On the right, show the model predictions vs actual states. \\
+	At the bottom, add two bullets discussing future work. \\
 \end{frame}

### 2020-09-14 twccc.tex
talk.

--- 2020-09-08 twccc.tex
@@
 
-\title[Data completion of initial gray box models for process control and optimization]{Data completion of initial gray box models for process control and optimization}
+\title[Nonlinear system identification with neural networks for 
+process control]{Nonlinear system identification with neural networks for 
+process control}
 \date{September 15, 2020}
@@
 % MPC intro frame
-\begin{frame}{Idea.}
-Hi.
+\begin{frame}{System identification for model predictive control (MPC)}
+
+MPC requires a dynamic plant model ($x^+ = f(x, u)$) for use in online optimization.
+
+	\begin{block}{Standard industrial practice}
+		\begin{itemize}
+			\item Estimate a linear model from data and use integrating disturbance models to maintain zero offset 
+			in controlled measurements.
+			\item For highly nonlinear plants, the closed-loop controller performance can deteriorate due to significant model mismatch. 
+		\end{itemize}
+	\end{block}
+
+	\begin{block}{Neural networks and system identification\footnote[frame]{\cite{hussain:1999, draeger:engell:ranke:1995, prasad:bequette:2003, eaton:rawlings:ungar:1994, bhat:minderman:mcavoy:wang:1990}}}
+		\begin{itemize}
+			\item \textbf{Black-box identification:} Estimate the entire predictive model ($x^+ =  f(x, u)$) from actuators to
+			sensors using the data collected from open-loop identification experiments.
+			\item \textbf{Hybrid model identification:} Model the dynamics from actuators to sensors using first-principle knowledge, and use
+			neural networks to estimate the mismatch or any unmeasured disturbance dynamics.  
+		\end{itemize}
+	\end{block}
 \end{frame}
 
-%\begin{frame}[allowframebreaks]{References}
-%\bibliographystyle{abbrvnat}
-%\bibliography{articles,proceedings,books,unpub}
-%\end{frame}
+\begin{frame}{Nonlinear HVAC example}
+	\begin{align*}
+	\dot{x} = Ax + Bg(u) + B_pp
+	\end{align*}
+	Nonlinearity between the valve (actuator) to the cooling duty. $x \in \{ T_{A1}, T_{C1}, T_{A2}, T_{C2} \}$, $u \in \{v_1, v_2\}$,
+	and $p$ is the ambient temperature $T_a$ (disturbance).
+	\pause
+	\begin{figure}[!h]
+		\centering
+		\includegraphics[page = 5,width=0.6\textwidth]{plotshvacnn.pdf}
+	\end{figure}
+	\alert{Objective: Test the performance of linear MPC in regions of high plant-model mismatch.}
+	\end{frame}
+
+\begin{frame}{Linear MPC vs NN based MPC}
+	\begin{figure}[!h]
+		\centering
+		\includegraphics[page = 2, width=0.48\textwidth]{plotshvacnn.pdf} \hfill
+		\includegraphics[page = 3, width=0.48\textwidth]{plotshvacnn.pdf}
+	\end{figure}
+
+	\begin{itemize}
+		\item NN trained on noisy open-loop system identification data (5000 samples).
+		\item Change in ambient temperature at 10 hours, and setpoint changes in air-zone 
+		temperatures at 50 and 100 hours. 
+		\item The NN model provides superior closed-loop performance than the linear model.
+	\end{itemize}
+\end{frame}
+
+	\begin{frame}{References}
+\bibliographystyle{abbrvnat}
+\bibliography{articles,proceedings,books,unpub}
+\end{frame}
 

### 2020-09-14 twccc.tex
add files for the talk.

@@
 			\item \textbf{Black-box identification:} Estimate the entire predictive model ($x^+ =  f(x, u)$) from actuators to
-			sensors using the data collected from open-loop identification experiments.
+			sensors using data collected from open-loop identification experiments.
 			\item \textbf{Hybrid model identification:} Model the dynamics from actuators to sensors using first-principle knowledge, and use
-			neural networks to estimate the mismatch or any unmeasured disturbance dynamics.  
+			neural networks (NNs) to estimate the mismatch or any unmeasured disturbance dynamics.  
 		\end{itemize}
@@
 	and $p$ is the ambient temperature $T_a$ (disturbance).
-	\pause
 	\begin{figure}[!h]
@@
 	\end{figure}
-	\alert{Objective: Test the performance of linear MPC in regions of high plant-model mismatch.}
+	\vspace{-0.1in}
+	\begin{itemize}
+		\item Test the performance of linear MPC in regions of high plant-model mismatch.
+	\end{itemize}
 \end{frame}
@@
 	\end{figure}
-
 	\begin{itemize}

### 2020-09-14 plotshvacnn.py

@@ -1,0 +1,233 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+# [depends] hvacnntest.mat
+# [makes] figure
+"""
+Created on Wed Oct 10 18:21:30 2018
+@author: pkumar
+"""
+import custompath
+custompath.add()
+
+# Importing the required packages
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+from plottools.matio import loadmat
+from matplotlib.backends.backend_pdf import PdfPages
+from matplotlib.ticker import FormatStrFormatter
+import plottools
+
+########### Setting for plots ############
+ms=0.9
+lw=0.5
+yax=-0.25
+
+
+# Figure Size
+figx=10
+figy=8
+
+#White Spacing
+ws=1.2
+hs=0.8
+
+# Grid Spacing
+gsy=4
+gsx=4
+
+# fonts of plots
+plt.rcParams.update({'font.size': 16})
+
+# Labels for X and Y axis
+ylabel=['$T_{A1}$ (K)','$T_{C1}$ (K)','$T_{A2}$ (K)','$T_{C2}$ (K)']
+ulabel=['$v_{1}$','$v_{2}$']
+xlabel = 'Time (hr)'
+
+# Do not rotate ylabel unless required
+ylabelRotation = False
+
+# Color of plots
+ucolor = 'g'
+ycolor = 'b-'
+xspcolor = 'r'
+
+# Colors in the ID data color
+plantcolor = 'r'
+lcolor = 'k'
+nncolor = 'b'
+
+# Colors in the gain response plots
+nonlincolor = 'r'
+lincolor = 'b'
+sscolor = 'k'
+sslstyle = '--'
+
+# legends for the gain response plots
+linlabel = 'Linear model'
+plantlabel = 'Plant'
+
+
+# Limits on plots which use valve positions
+vlim = [0, 0.7]
+###########################################
+
+def iddata(x, xlin, xnn, u, time):
+    
+
+    gs=gridspec.GridSpec(gsy,gsx)
+    gs.update(wspace=ws, hspace=hs)
+    fig1=plt.figure(figsize=(figx,figy))
+        
+    # Plot the measurements
+    Nx = x.shape[0]
+    for i in range(0, Nx):
+        
+        ax=fig1.add_subplot(gs[i,gsx-2:gsx]) 
[…]

### 2020-09-15 twccc.tex
ok, the talk is almost done.

@@
 
-	\begin{block}{Neural networks and system identification\footnote[frame]{\cite{hussain:1999, draeger:engell:ranke:1995, prasad:bequette:2003, eaton:rawlings:ungar:1994, bhat:minderman:mcavoy:wang:1990}}}
+	\begin{block}{Neural networks and system identification\footnote[frame]{\cite{hussain:1999, draeger:engell:ranke:1995, lovelett:avalos:kevrekidis:2019, tuor:drgona:vrabie:2020}}}
 		\begin{itemize}
@@
 
-\begin{frame}{Nonlinear HVAC example}
+\begin{frame}{Nonlinear HVAC model}
+
+	\centering
+	\colorlet{CoolFlow}{NavyBlue!60!Cyan}
+	\colorlet{OtherFlow}{black}
+	\colorlet{HeatFlow}{Red}
+	\colorlet{BuildingColor}{Brown}
+	\colorlet{HVACColor}{Cerulean}
+	\colorlet{SunColor}{yellow}
+
+	\tikzset{>=stealth'}
+	\resizebox{.6\textwidth}{!}{
+	\begin{tikzpicture}
+	[
+	scale=1,
+	flow/.style={->, thick, line cap=rect},
+	flowlabel/.style={black, align=center, font=\scriptsize\itshape},
+	masszonestyle/.style={draw,rectangle,minimum width=2.5cm,minimum height=3cm,color=black,ultra thick,fill=BuildingColor,fill opacity=0.6,text=black,text opacity=1},
+	airzonestyle/.style={draw,rectangle,minimum width=1.5cm,minimum height=2cm,color=black,ultra thick,fill=BuildingColor!5,fill opacity=0.8,text=black,text opacity=1},
+	hvacstyle/.style={draw,rectangle,minimum width=2cm,minimum height=1.5cm,color=black,ultra thick,fill=HVACColor,fill opacity=0.65,text=black,text opacity=1},
+	]
+	% Grid for drawing.
+	% \draw (-5,-3) to[grid with coordinates] (5,3);
+
+	% Draw the sun.
+	\draw (-4,2) node[line width=1pt,draw=black,fill=SunColor,circle,minimum size=1cm] (TheSun) {};
+	\foreach \angle in {0,45,...,315}
+	{\draw [rotate around={\angle:(TheSun.center)}] ($(TheSun.center) + (0.7,0)$) node[shape border rotate=\angle-90,line width=1pt,draw=black,fill=SunColor,regular polygon,regular polygon sides=3,inner sep=0.05cm] {};}
+
+	% Building.
+	\node (building_rect) at (-1.25,-1.5) [masszonestyle] {};
+	\node (building_rect) at (1.25,-1.5) [masszonestyle] {};
+	\node (building_rect) at (-1.25,-1.5) [airzonestyle,label={[label distance=-1cm]90:Zone 1},label={[label distance =-1.4cm]90:$T_{A1}$}] {};
+	\node (building_rect) at (1.25,-1.5) [airzonestyle,label={[label distance=-1cm]90:Zone 2},label={[label distance =-1.4cm]90:$T_{A2}$}] {};
+	\node (hvac_rect) at (0,0.75) [hvacstyle] {HVAC};
+
+	% Arrows.
+	\draw[flow, HeatFlow] (-3.5, 1) -- (-2.75, 0) node[midway,below left,flowlabel] {Radiation};
+	\draw[flow, HeatFlow] (-4.5, -1.5) -- (-2.75, -1.5) node[midway,below,flowlabel] {Convection} node[midway,above,flowlabel] {Ambient};
+	\draw[flow, HeatFlow] (-1.8, -2.4) -- ++(0, 0.8) node[align=center,midway,right,flowlabel,xshift=-0.05cm, font=\tiny\itshape] {Heat\\Generation};
+	\draw[flow, HeatFlow] (1.8, -2.4) -- ++(0, 0.8) node[align=center,midway,right,flowlabel,xshift=-0.05cm] {};
+	\draw[flow, HeatFlow, <->] (-0.5, -2.8) -- ++(1, 0) node[align=center,above,right,flowlabel,yshift=0.05cm, font=\tiny\itshape] {Heat Transfer \\ Between Zones};
+	\draw[flow, HeatFlow, <->] (-0.8, -1.6) -- ++(0.6, 0) node[align=center,above,right,flowlabel,yshift=0.05cm] {};
+	\draw[flow, HeatFlow, <->] (0.8, -1.6) -- ++(-0.6, 0) node[align=center,above,right,flowlabel,yshift=0.05cm] {};
+	\draw[flow, CoolFlow] (-0.75, -0.8) node[right,flowlabel,yshift=0.5cm] {} -- ++(0, 1.2);
+	\draw[flow, CoolFlow] (0.75, -0.8) node[left,flowlabel,yshift=0.5cm] {} -- ++(0, 1.2);
+	\draw[flow, OtherFlow, dashed] (-0.5, 2) node[above,flowlabel] {Power} -- ++(0, -1);
+	\draw[flow, HeatFlow] (hvac_rect.east) -- ++(1, 0) -- ++(0, 1) node[align=center,midway,right,flowlabel] {Rejection \\ to Ambient};
+
+	% Labels.
+	\node at (0,0) [above, flowlabel] {Cooling};
+	\node at (1.6,-0.5) [below, flowlabel] {Air};
+	\node at (1.3,-0.5) [above, flowlabel] {Mass};
+	\node at (-3.5,-1.2) [above, flowlabel] {$T_a$};
+	\node at (-2,-0.5) [above, flowlabel] {$T_{C1}$};
+	\node at (2,-0.5) [above, flowlabel] {$T_{C2}$};
+	\node at (-0.4,-0.55) [above, flowlabel] {$\dot Q_{C1}$};
+	\node at (0.4,-0.55) [above, flowlabel] {$\dot Q_{C2}$};
+	\end{tikzpicture}}
 	\begin{align*}
-	\dot{x} = Ax + Bg(u) + B_pp
+		C_i\dfrac{dT_{i}}{dt} &= -H_{i}(T_i-T_a)-\sum_{j \neq i}\beta_{ij}(T_i-T_j)-\dot Q_{ci}(v_i)+\dot Q_{oi}
 		\end{align*} 
-	Nonlinearity between the valve (actuator) to the cooling duty. $x \in \{ T_{A1}, T_{C1}, T_{A2}, T_{C2} \}$, $u \in \{v_1, v_2\}$,
-	and $p$ is the ambient temperature $T_a$ (disturbance).
+		Nonlinearity between the valve ($v_i$) to the cooling duty 
+		($\dot Q_{ci}$). $x \in \{ T_{A1}, T_{C1}, T_{A2}, T_{C2} \}$, $u \in \{v_1, v_2\}$, and $p$ is the ambient temperature $T_a$ (disturbance).	
+\end{frame}
+
+\begin{frame}{Nonlinearity in the plant-model}
 	\begin{figure}[!h]
@@
 
-\begin{frame}{Linear MPC vs NN based MPC}
+\begin{frame}{Linear MPC vs NN-based MPC}
 	\begin{figure}[!h]
@@
 	\begin{itemize}
-		\item NN trained on noisy open-loop system identification data (5000 samples).
-		\item Change in ambient temperature at 10 hours, and setpoint changes in air-zone 
-		temperatures at 50 and 100 hours. 
+		\item NN trained on noisy open-loop identification data (5000 samples).
 		\item The NN model provides superior closed-loop performance than the linear model.
+		\item \textbf{Future work --} \\
+		i) Develop identification frameworks to build
+		hybrid grey-box + black-box models. \\
+		ii) Use neural networks to model unmeasured disturbance dynamics (e.g., heat disturbances in HVAC applications).
 	\end{itemize}

### 2020-09-16 hvacnntrain.py
add tex.

@@ -1,37 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-# [depends] hvacnndata.mat
-# [makes] mat
-"""
-@author: pkumar
-
-Code for a 3 state 2 input nonLinear MPC example.
-"""
-from plottools.matio import loadmat, savemat
-import custompath
-custompath.add()
-from tfnet import tfnet
-
-data=loadmat('hvacnndata.mat',scalararray=False,asdict=True)
-X = data['xnn']
-Y = data['ynn']
-del data
-
-tfeg=tfnet(data={'X':X,'Y':Y},          
-           options={'layer_dims':[6, 12, 4],
-                    'frac_train':0.9,
-                    'learning_rate':1e-2,
-                    'num_epochs':20000,
-                    'normalize_data':True,
-                    'optimization_algorithm':'adam',
-                    'activation':'tanh',
-                    'minibatch_size':X.shape[1],
-                    'print_cost':True,
-                    'np_seed':10,
-                    'tf_seed':1023
-                    }           
-           )
-
-data=tfeg.model()
-savemat('hvacnntrain.mat', data, appendmat=False, oned_as="column",
-            list_as_cell=True)

### 2020-09-16 plotshvacnn.py

@@ -1,233 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-# [depends] hvacnntest.mat
-# [makes] figure
-"""
-Created on Wed Oct 10 18:21:30 2018
-@author: pkumar
-"""
-import custompath
-custompath.add()
-
-# Importing the required packages
-import matplotlib.pyplot as plt
-import matplotlib.gridspec as gridspec
-from plottools.matio import loadmat
-from matplotlib.backends.backend_pdf import PdfPages
-from matplotlib.ticker import FormatStrFormatter
-import plottools
-
-########### Setting for plots ############
-ms=0.9
-lw=0.5
-yax=-0.25
-
-
-# Figure Size
-figx=10
-figy=8
-
-#White Spacing
-ws=1.2
-hs=0.8
-
-# Grid Spacing
-gsy=4
-gsx=4
-
-# fonts of plots
-plt.rcParams.update({'font.size': 16})
-
-# Labels for X and Y axis
-ylabel=['$T_{A1}$ (K)','$T_{C1}$ (K)','$T_{A2}$ (K)','$T_{C2}$ (K)']
-ulabel=['$v_{1}$','$v_{2}$']
-xlabel = 'Time (hr)'
-
-# Do not rotate ylabel unless required
-ylabelRotation = False
-
-# Color of plots
-ucolor = 'g'
-ycolor = 'b-'
-xspcolor = 'r'
-
-# Colors in the ID data color
-plantcolor = 'r'
-lcolor = 'k'
-nncolor = 'b'
-
-# Colors in the gain response plots
-nonlincolor = 'r'
-lincolor = 'b'
-sscolor = 'k'
-sslstyle = '--'
-
-# legends for the gain response plots
-linlabel = 'Linear model'
-plantlabel = 'Plant'
-
-
-# Limits on plots which use valve positions
-vlim = [0, 0.7]
-###########################################
-
-def iddata(x, xlin, xnn, u, time):
-    
-
-    gs=gridspec.GridSpec(gsy,gsx)
-    gs.update(wspace=ws, hspace=hs)
-    fig1=plt.figure(figsize=(figx,figy))
-        
-    # Plot the measurements
-    Nx = x.shape[0]
-    for i in range(0, Nx):
-        
-        ax=fig1.add_subplot(gs[i,gsx-2:gsx]) 
[…]

### 2020-09-16 systemID_NNs.tex
add files.

\documentclass{article}

%% The graphicx and color packages should give you all the color,
%% figure inclusion, and text/figure resizing capabilities you need.
\usepackage{graphicx}
\usepackage[dvipsnames]{xcolor}
\usepackage{tikz,tikzsettings,bm}
\usepackage{almostfull}

\definecolor{pblue}{HTML}{1F77B4}
\definecolor{pgold}{HTML}{FF7f0E}
\definecolor{pgreen}{HTML}{2CA02C}
\definecolor{pred}{HTML}{D62728}
\definecolor{ppurple}{HTML}{9467BD}
\definecolor{ppink}{HTML}{E377C2}
\definecolor{pyellow}{HTML}{BCBD22}
\definecolor{pcyan}{HTML}{17BECF}

\usepackage{authblk}

%% You should not need more than this for fancy math.
\usepackage{amsmath}   % Extra math commands and environments from the AMS
\usepackage{amssymb}   % Special symbols from the AMS
\usepackage{amsthm}    % Enhanced theorem and proof environments from the AMS
\usepackage{latexsym}  % A few extra LaTeX symbols

\usepackage{lucidbry}
\input{stanacce}
\renewcommand{\baselinestretch}{1.0666}

%% The URL package is handy for typesetting URLs.  It does not define
%% an \email command because so many document styles already do that.
%% So we define one here that uses a typewriter font.
\usepackage{url}
%\DeclareRobustCommand{\email}{\begingroup \urlstyle{tt}\Url}
\providecommand{\email}{}
\renewcommand{\email}[1]{\texttt{#1}}

%% This provides various customized verbatim commands and
%% environments.  You probably don't need it.
\usepackage{fancyvrb}
\DefineShortVerb{\|}
\VerbatimFootnotes
\DefineVerbatimEnvironment{code}{Verbatim}{%
  frame=single,
  framesep=1em,
  xleftmargin=1em,
  xrightmargin=1em,
  samepage=true,
  fontsize=\footnotesize}
\usepackage{upquote}

%% Your document may require a different bibliography style.  I've
%% come to prefer Author (YYYY) styles because it makes it easier for
%% a person who knows something about the literature to understand
%% what is being cited without having to skip to the list of
%% references.
\usepackage[authoryear,round,longnamesfirst]{natbib}

%% You won't normally need this definition in your documents, but it
%% is here so we can typeset the BibTeX logo correctly.
\makeatletter
\@ifundefined{BibTeX}
   {\def\BibTeX{{\rmfamily B\kern-.05em%
    \textsc{i\kern-.025em b}\kern-.08em%
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}}{}

\def\blfootnote{\xdef\@thefnmark{}\@footnotetext}
\makeatother

%% Allow more of the pages to be occupied by graphs.  These parameters
%% are described in section c.9.1 of the LaTeXbook.  Your document may
%% not benefit from these parameters (it could make things worse) so
%% you should only change these from the defaults if you need to.
\setcounter{topnumber}{2}              %% 2
\setcounter{bottomnumber}{1}           %% 1
\setcounter{totalnumber}{3}            %% 3
\renewcommand{\topfraction}{0.9}       %% 0.7
\renewcommand{\bottomfraction}{0.9}    %% 0.3
\renewcommand{\textfraction}{0.1}      %% 0.2
\renewcommand{\floatpagefraction}{.7}  %% 0.5
\newlength{\graphwidth}
\setlength{\graphwidth}{0.8\columnwidth}

\usepackage{subfiles}

\graphicspath{{./}{./figures/}{./figures/paper/}}

\newcommand{\Is}{\bm{\mathrm{I}}} % Resources.
\newcommand{\Js}{\bm{\mathrm{J}}} % Units.
\newcommand{\Ks}{\bm{\mathrm{K}}} % Resources.
\newcommand{\Ms}{\bm{\mathrm{M}}} % Transition catalog.
\newcommand{\Ns}{\bm{\mathrm{N}}} % Transition counter.

\title{System ID and control with NNs}
\author{Problem formulation}
\date{\today}

\begin{document}

\maketitle

\section{Process modeling}
Consider a nonlinear plant evolving as
\begin{align*}
  x^+ &= f(x, u) \\
  y &= h(x)
\end{align*}
in which $x$ is the plant state (vector), $u$ is the manipulated control input,
and $y$ is the measurement. Assume that we don't have any 
disturbances/measurement and process noise 
for now, and we'll add those after we 
understand the issues in this nominal case.

\subsection{Completion of the u-y model}
In this class of problems, we assume that grey-box modeling is performed 
for the plant, and we wish to augment the grey-box model with neural networks
for improvement in the accuracy of the model. From first-principles modeling, 
we have the following (in continuous time, as grey-box modeling will 
almost always be in continuous time).
\begin{align*}
  \dot{x}_g &= f_g(x_g, u) \\
  y &= h_g(x_g)
\end{align*}

From Taken's embedding theorem, we know that the original plant state
can be reconstructed using a history of measurements and control inputs.
Denote this history as 
\begin{align*}
  \mathbf{y}_{k:k-\tau} &= [y(k)', y(k-1)', ..., y(k-\tau)']' \\
  \mathbf{u}_{k-1:k-\tau} &= [u(k-1)', u(k-2)', ..., u(k-\tau)']'
\end{align*}

So from the embedding theorem for observable nonlinear plants, 
there exists a $\phi(\cdot)$ and number of delay 
time steps $\tau$, such that 
$x(k) = \phi(\mathbf{y}_{k:k-\tau}, \mathbf{u}_{k-1:k-\tau})$.

Now depending on the application, an approximation of $\phi(\cdot)$
with a neural network (say $\phi_{NN}(\cdot)$) 
can be used to augment the grey-box model 
in several ways. The following is proposed in the recent 
paper by Kevrekidis.
\begin{align*}
  x_g^+ &= f_g(x_g, u, 
  \theta(\phi_{NN}(\mathbf{y}_{k:k-\tau}, \mathbf{u}_{k-1:k-\tau})))\\
  y &= h_g(x_g)
\end{align*}

in which $\theta(\cdot)$ are a set of grey-box parameters which 
depend on the plant states. For example, rate constants which 
depend on temperature, etc. An RK4 discretization is performed
to get this model in discrete time.

A different type of problems could be say, there are unknown
side reactions in the plant we don't about, which are affecting the 
physical grey-box 
states. In these problems, we can augment the grey-box 
model with a NN as follows
\begin{align*}
x_g^+ &= f_g(x_g, u) + f_{NN}(\mathbf{y}_{k:k-\tau}, \mathbf{u}_{k-1:k-\tau})\\
y &= h_g(x_g) 
\end{align*}
in which, we are anticipating that $f_{NN}$,
a neural network can learn the difference between the 
true grey-box state dynamics and what is modeled with 
the function ($f_g(\cdot)$). We assume that the measurements
are directly some subset of the physical grey-box states.

\textbf{Need to discuss the following issues:}
\begin{itemize}
  \item What are the example problems/applications 
  where both kind of modeling 
  strategies make sense, and where should we start.
  Reproduce results from the Kevrekidis paper with measurement 
  noise?
  \item The model training step will be the optimization of 
  the following multi-step-ahead prediction-error
  \begin{equation*}
    \underset{W_i, b_i}{\textnormal{min}} \sum_{k=0}^{t-1} (\hat{y}(k) - y(k))^2
  \end{equation*}
  in which ($W_i, b_i$) are the weights of the NN. This optimization 
  is for one ID experiment, and requires an initial grey-box 
  state $x_g(0)$ to make the forecast. If we have data 
  for multiple ID experiments, we'll need initial states for 
  all the ID experiments. In theory, we can add the initial 
  states as decision variables in the optimization, but 
  unfortunately tensorflow 
  does not has a mechanism to do this.
\end{itemize}

\subsection{Adding feedforward disturbance forecasts}
In this class of problems, we assume that we have a pretty 
good u-y model from first-principles knowledge, and we'll use 
neural networks to make feedforward disturbance forecasts
for use in online MPC. HVAC is a pretty good application.

\section{Optimization in MPC with neural networks as model}
This is probably a late fall quarter or a winter break project. \\
a) Try Casadi/MPC tools. 
If tanh is the activation function (optimization is smooth),
this should be fine. \\
b) Try a global optimization solver. \\
c) Train another NN with solutions from a global optimization solver.

\end{document}

### 2020-09-22 systemID_NNs.tex
add things.

@@
 
-\title{System ID and control with NNs}
+\title{Data-based completion of grey-box models with neural networks}
 \author{Problem formulation}
@@
 \section{Process modeling}
-Consider a nonlinear plant evolving as
+Consider a nonlinear plant evolving in continuous time as
 \begin{align*}
-  x^+ &= f(x, u) \\
-  y &= h(x)
+  \dot x &= f_p(x, u) \\
+  y &= C_px
 \end{align*}
@@
 for now, and we'll add those after we 
-understand the issues in this nominal case.
+understand the issues in the nominal case.
 
@@
   \dot{x}_g &= f_g(x_g, u) \\
-  y &= h_g(x_g)
+  y &= C_gx_g
 \end{align*}
 
-From Taken's embedding theorem, we know that the original plant state
-can be reconstructed using a history of measurements and control inputs.
+in which $x_g$ is some set of physical plant states chosen to be 
+modeled. Denoted these physical states as $x_g=Cx$.
+For observable nonlinear systems, 
+we can construct the original plant state
+using a history of measurements and control inputs.
 Denote this history as 
 \begin{align*}
-  \mathbf{y}_{k:k-\tau} &= [y(k)', y(k-1)', ..., y(k-\tau)']' \\
-  \mathbf{u}_{k-1:k-\tau} &= [u(k-1)', u(k-2)', ..., u(k-\tau)']'
+  \mathbf{y}_{k:k-N} &= [y(k)', y(k-\Delta)', ..., y(k-N\Delta)']' \\
+  \mathbf{u}_{k-1:k-N} &= [u(k-1)', u(k-2)', ..., u(k-N\Delta)']'
 \end{align*}
 
-So from the embedding theorem for observable nonlinear plants, 
-there exists a $\phi(\cdot)$ and number of delay 
-time steps $\tau$, such that 
-$x(k) = \phi(\mathbf{y}_{k:k-\tau}, \mathbf{u}_{k-1:k-\tau})$.
+There exists a $\phi(\cdot)$ and a number of delay 
+time steps $N$, such that 
+$x(k) = \phi(\mathbf{y}_{k:k-N}, \mathbf{u}_{k-1:k-N})$.
 
-Now depending on the application, an approximation of $\phi(\cdot)$
-with a neural network (say $\phi_{NN}(\cdot)$) 
-can be used to augment the grey-box model 
-in several ways. The following is proposed in the recent 
-paper by Kevrekidis.
+We augment the grey-box model in continuous time 
+as follows -
 \begin{align*}
-  x_g^+ &= f_g(x_g, u, 
-  \theta(\phi_{NN}(\mathbf{y}_{k:k-\tau}, \mathbf{u}_{k-1:k-\tau})))\\
-  y &= h_g(x_g)
+\dot x_g &= f_g(x_g, u) + 
+f_{NN}(\mathbf{y}_{k:k-N}, \mathbf{u}_{k:k-N})\\
+y &= C_gx_g 
+\end{align*}
+in which $f_{NN}(\cdot)$,
+is a standard feedforward neural network. 
+We assume that the measurements
+are directly some subset of the physical grey-box states.
+
+It is expected that the function $f_{NN}(\cdot)$ approximates
+the difference between the true grey-box state dynamics
+with what is modeled. 
+\begin{align*}
+  f_{NN}(\mathbf{y}_{k:k-N}, \mathbf{u}_{k:k-N}) \approx 
+      f_p(x, u) - f_g(x_g, u) = 
+      Cf_p(\phi(\mathbf{y}_{k:k-N}, \mathbf{u}_{k-1:k-N}), u) - f_g(x_g, u)
 \end{align*}
 
-in which $\theta(\cdot)$ are a set of grey-box parameters which 
-depend on the plant states. For example, rate constants which 
-depend on temperature, etc. An RK4 discretization is performed
-to get this model in discrete time.
+\subsection{Conversion to discrete time and training}
 
-A different type of problems could be say, there are unknown
-side reactions in the plant we don't about, which are affecting the 
-physical grey-box 
-states. In these problems, we can augment the grey-box 
-model with a NN as follows
-\begin{align*}
-x_g^+ &= f_g(x_g, u) + f_{NN}(\mathbf{y}_{k:k-\tau}, \mathbf{u}_{k-1:k-\tau})\\
-y &= h_g(x_g) 
-\end{align*}
-in which, we are anticipating that $f_{NN}$,
-a neural network can learn the difference between the 
-true grey-box state dynamics and what is modeled with 
-the function ($f_g(\cdot)$). We assume that the measurements
-are directly some subset of the physical grey-box states.
+\subsection{Simple example}
 
-\textbf{Need to discuss the following issues:}
-\begin{itemize}
-  \item What are the example problems/applications 
-  where both kind of modeling 
-  strategies make sense, and where should we start.
-  Reproduce results from the Kevrekidis paper with measurement 
-  noise?
-  \item The model training step will be the optimization of 
-  the following multi-step-ahead prediction-error
-  \begin{equation*}
-    \underset{W_i, b_i}{\textnormal{min}} \sum_{k=0}^{t-1} (\hat{y}(k) - y(k))^2
-  \end{equation*}
-  in which ($W_i, b_i$) are the weights of the NN. This optimization 
-  is for one ID experiment, and requires an initial grey-box 
-  state $x_g(0)$ to make the forecast. If we have data 
-  for multiple ID experiments, we'll need initial states for 
-  all the ID experiments. In theory, we can add the initial 
-  states as decision variables in the optimization, but 
-  unfortunately tensorflow 
-  does not has a mechanism to do this.
-\end{itemize}
+%\textbf{Need to discuss the following issues:}
+%\begin{itemize}
+%  \item What are the example problems/applications 
+%  where both kind of modeling 
+%  strategies make sense, and where should we start.
+%  Reproduce results from the Kevrekidis paper with measurement 
+%  noise?
+%  \item The model training step will be the optimization of 
+%  the following multi-step-ahead prediction-error
+%  \begin{equation*}
+%    \underset{W_i, b_i}{\textnormal{min}} \sum_{k=0}^{t-1} (\hat{y}(k) - y(k))^2
+%  \end{equation*}
+%  in which ($W_i, b_i$) are the weights of the NN. This optimization 
+%  is for one ID experiment, and requires an initial grey-box 
+%  state $x_g(0)$ to make the forecast. If we have data 
+%  for multiple ID experiments, we'll need initial states for 
+%  all the ID experiments. In theory, we can add the initial 
+%  states as decision variables in the optimization, but 
+%  unfortunately tensorflow 
+%  does not has a mechanism to do this.
+%\end{itemize}
 
-\subsection{Adding feedforward disturbance forecasts}
-In this class of problems, we assume that we have a pretty 
-good u-y model from first-principles knowledge, and we'll use 
-neural networks to make feedforward disturbance forecasts
-for use in online MPC. HVAC is a pretty good application.
+%\subsection{Adding feedforward disturbance forecasts}
+%In this class of problems, we assume that we have a pretty 
+%good u-y model from first-principles knowledge, and we'll use 
+%neural networks to make feedforward disturbance forecasts
+%for use in online MPC. HVAC is a pretty good application.
 
-\section{Optimization in MPC with neural networks as model}
-This is probably a late fall quarter or a winter break project. \\
-a) Try Casadi/MPC tools. 
-If tanh is the activation function (optimization is smooth),
-this should be fine. \\
-b) Try a global optimization solver. \\
-c) Train another NN with solutions from a global optimization solver.
+%\section{Optimization in MPC with neural networks as model}
+%This is probably a late fall quarter or a winter break project. \\
+%a) Try Casadi/MPC tools. 
+%If tanh is the activation function (optimization is smooth),
+%this should be fine. \\
+%b) Try a global optimization solver. \\
+%c) Train another NN with solutions from a global optimization solver.
 

### 2020-09-22 tworeac_parameters.py

@@ -17,6 +17,7 @@ def _tworeac_plant_ode(x, u, p, parameters):
     # Extract the parameters.
     k1 = parameters['k1']
     k2 = parameters['k2']
+    k3 = parameters['k3']
 
     # Extract the plant states into meaningful names.
     (Ca, Cb, Cc) = x[0:3]
@@ -25,8 +26,8 @@ def _tworeac_plant_ode(x, u, p, parameters):
 
     # Write the ODEs.
     dCabydt = (Ca0-Ca)/tau - k1*Ca
-    dCbbydt = k1*Ca - 2*k2*Cb - Cb/tau
-    dCcbydt = k2*Cb - Cc/tau
+    dCbbydt = k1*Ca - 2*k2*Cb + k3*Cc - Cb/tau
+    dCcbydt = k2*Cb - k3*Cc - Cc/tau
 
     # Return the derivative.
     return np.array([dCabydt, dCbbydt, dCcbydt])
@@ -35,7 +36,7 @@ def _tworeac_greybox_ode(x, u, p, parameters):
     """ Simple ODE describing the grey-box plant. """
     # Extract the parameters.
     k1 = parameters['k1']
-    k2 = parameters['k2']
+
     # Extract the plant states into meaningful names.
     (Ca, Cb) = x[0:2]
     Ca0 = u[0:1]
@@ -60,6 +61,7 @@ def _get_tworeac_parameters():
     parameters = {}
     parameters['k1'] = 1 # m^3/min.
     parameters['k2'] = 0.05 # m^3/min.
+    parameters['k3'] = 0.01 # m^3/min.
 
     #parameters['tau'] = 2. # m^3 
     #parameters['k3'] = 0.05 # m^3/min.
@@ -189,10 +191,27 @@ def _get_greybox_validation_predictions(*, parameters, training_data):
                 Ca0=np.asarray(model.u).squeeze())
     return data
 
+def _check_cont_time_obsv(*, parameters):
+    """ Check the observability of the continuous time linear system. """
+    # Measurement matrix for the plant.
+    C = np.array([[1., 0., 0.], 
+                  [0., 1., 0.]])
+    tau = parameters['ps'][0]
+    (k1, k2, k3) = (parameters['k1'], parameters['k2'], parameters['k3'])
+    A = np.array([[-k1-(1/tau), 0., 0.], 
+                  [k1, -2*k2 - (1/tau), k3], 
[…]

### 2020-09-23 systemID_NNs.tex
ok, working on the system id.

@@
   \dot x &= f_p(x, u) \\
-  y &= C_px
+  y &= h_p(x)
 \end{align*}
@@
 \subsection{Completion of the u-y model}
-In this class of problems, we assume that grey-box modeling is performed 
+We assume that grey-box modeling is performed 
 for the plant, and we wish to augment the grey-box model with neural networks
 for improvement in the accuracy of the model. From first-principles modeling, 
-we have the following (in continuous time, as grey-box modeling will 
-almost always be in continuous time).
+we have the following grey-box model in continuous time
 \begin{align*}
   \dot{x}_g &= f_g(x_g, u) \\
-  y &= C_gx_g
+  y &= h_g(x_g)
 \end{align*}
 
-in which $x_g$ is some set of physical plant states chosen to be 
-modeled. Denoted these physical states as $x_g=Cx$.
+in which $x_g$ are the grey-box states, which are some subset 
+of the plant states (denote $x_g = Gx$) chosen to 
+be modeled. The dynamics of these grey-box states 
+potentially have some missing terms 
+in their ODEs.
 For observable nonlinear systems, 
@@
 
-There exists a $\phi(\cdot)$ and a number of delay 
-time steps $N$, such that 
+in which $\Delta$ is the sample 
+time between two consecutive
+prediction instants of the desired discrete time 
+model to be used in MPC. There exists a $\phi(\cdot)$ 
+and a number of delay 
+time steps $N$, such that the plant state
+at a given time can be written as
 $x(k) = \phi(\mathbf{y}_{k:k-N}, \mathbf{u}_{k-1:k-N})$.
 
-We augment the grey-box model in continuous time 
-as follows -
+We augment the grey-box model using 
+neural networks as follows -
 \begin{align*}
@@
 f_{NN}(\mathbf{y}_{k:k-N}, \mathbf{u}_{k:k-N})\\
-y &= C_gx_g 
+y &= h_g(x_g) + h_{NN}(\mathbf{y}_{k:k-N}, \mathbf{u}_{k-1:k-N})
 \end{align*}
-in which $f_{NN}(\cdot)$,
-is a standard feedforward neural network. 
-We assume that the measurements
-are directly some subset of the physical grey-box states.
+in which $f_{NN}(\cdot)$ and $h_{NN}(\cdot)$
+are standard feedforward neural networks. We add the current 
+control $u(k)$ as an input to the network $f_{NN}(\cdot)$ because 
+there may be a mismatch in the actuator to the grey-box dynamics
+as well.
 
-It is expected that the function $f_{NN}(\cdot)$ approximates
-the difference between the true grey-box state dynamics
-with what is modeled. 
+It is expected that the functions $f_{NN}(\cdot)$ 
+and $h_{NN}(\cdot)$ approximate
+the following differences between the true grey-box state dynamics
+and true measurement function with what is modeled.
 \begin{align*}
-  f_{NN}(\mathbf{y}_{k:k-N}, \mathbf{u}_{k:k-N}) \approx 
-      f_p(x, u) - f_g(x_g, u) = 
-      Cf_p(\phi(\mathbf{y}_{k:k-N}, \mathbf{u}_{k-1:k-N}), u) - f_g(x_g, u)
+  f_{NN}(\mathbf{y}_{k:k-N}, \mathbf{u}_{k:k-N}) &\approx 
+      Gf_p(\phi(\mathbf{y}_{k:k-N}, \mathbf{u}_{k-1:k-N}), u) - f_g(x_g, u) = 
+      Gf_p(x, u) - f_g(x_g, u) \\
+      h_{NN}(\mathbf{y}_{k:k-N}, \mathbf{u}_{k-1:k-N}) &\approx 
+      h_p(\phi(\mathbf{y}_{k:k-N}, \mathbf{u}_{k-1:k-N})) - h_g(x_g) = 
+      h_p(x) - h_g(x_g) 
 \end{align*}
@@
 \subsection{Conversion to discrete time and training}
+We note that the network $f_{NN}(\cdot)$ is an addition 
+to the grey-box dynamics in continuous time, whereas 
+$h_{NN}(\cdot)$ is an addition directly to the 
+measurement function. The Runge-Kutta-4 (RK4) method 
+is used for the discretization as follows
+
+\begin{align*}
+  k_1 &= f_g(x_g(k), u(k)) + f_{NN}(\mathbf{y}_{k:k-N}, \mathbf{u}_{k:k-N}) \\
+  k_2 &= f_g(x_g(k) + (\Delta/2)k_1, u(k)) + 
+         f_{NN}(\mathbf{y}_{k+0.5:k-N+0.5}, 
+                \mathbf{u}_{k+0.5:k-N+0.5})\\
+  k_3 &= f_g(x_g(k) + (\Delta/2)k_2, u(k)) + 
+                f_{NN}(\mathbf{y}_{k+0.5:k-N+0.5}, 
+                      \mathbf{y}_{k+0.5:k-N+0.5})\\
+  k_4 &= f_g(x_g(k) + \Delta k_3, u(k)) + 
+                       f_{NN}(\mathbf{y}_{k+1:k-N+1}, 
+                              \mathbf{u}_{k+1:k-N+1})\\
+  x_g(k+\Delta) &= x_g(k) + (\Delta/6)(k_1 + 2k_2 + 2k_3 + k_4) \\
+  \hat{y}(k+\Delta) &= h_g(x_g(k+\Delta)) + h_{NN}(\mathbf{y}_{k+\Delta:k-N+\Delta}, 
+                                     \mathbf{u}_{k-1+\Delta:k-N+\Delta})
+\end{align*}
+in which the sequences $\mathbf{y}_{k+0.5:k-N+0.5}$ 
+and $\mathbf{u}_{k+0.5:k-N+0.5}$ are defined as 
+\begin{align*}
+  \mathbf{y}_{k+0.5:k-N+0.5} &= 
+    [y(k + \Delta/2)', y(k-\Delta/2)', ..., y(k-N\Delta + \Delta/2)']' \\
+  \mathbf{u}_{k+0.5:k-N+0.5} &= 
+  [u(k + \Delta/2)', u(k-\Delta/2)', ..., u(k-N\Delta + \Delta/2)']' 
+\end{align*}
+We assume a zero order hold on the control inputs between two consecutive 
+MPC execution instants, such that $u(k + \Delta/2) = u(k)$. To get the 
+measurements between two sample instants ($y(k + \Delta/2)$) for 
+the RK4 predictions, a linear interpolation can be used 
+or we can assume that additional 
+measurements from the plant are available 
+between two MPC execution instants as well.
+
+The following optimization problem is solved to compute 
+the weights in the neural networks 
+\begin{align*}
+ \underset{W_{fi}, b_{fi}, W_{hi}, b_{hi}}{\textnormal{min}} \sum_{i=0}^{N_s-1} 
+ \dfrac{1}{2}(\hat{y}(i) - y(i))^2
+\end{align*}
+in which ($W_{fi}, b_{fi}, W_{hi}, b_{hi}$) are the weights in the 
+two neural networks ($f_{NN}(\cdot)$ and $h_{NN}(\cdot)$), 
+$\hat{y}$ are predictions of the augmented model, and $y$
+are the measurements from the plant.
+Note that this optimization requires an estimate 
+of the initial grey-box state. We assume that 
+before the open-loop system identification experiment 
+performed for data collection, 
+the plant is being operated at some 
+steady-state such that input to output
+responses of the plant and grey-box model are aligned with 
+the use of an integrating disturbance model. We can use 
+the grey-box state at that steady state as the initial condition 
+for the above multi-step-ahead prediction-error
+optimization.
 
 \subsection{Simple example}
+
+Consider the following plant, model,
+and the expected function to be approximated 
+with a neural network.
+
+\textbf{Plant --}
+\begin{align*}
+  \dfrac{dC_A}{dt} &= \dfrac{C_{A0} - C_A}{\tau} - k_1C_A\\
+  \dfrac{dC_B}{dt} &= k_1C_A - k_2C_B + k_3C_C- \dfrac{C_B}{\tau}\\
+  \dfrac{dC_C}{dt} &= k_2C_B - k_3C_C - \dfrac{C_C}{\tau}
+\end{align*}
+These ODEs represent the dynamics of the reaction system
+$A \rightarrow B \rightleftharpoons C$ in a CSTR with constant 
+height and temperature. The control input is $u = C_{A0}$, 
+and the measurements are $y = [C_A, C_B]'$. The plant model 
+is observable with these choices of the measurements and dynamics.
+
+\textbf{Model --}
+\begin{align*}
+  \dfrac{dC_A}{dt} &= \dfrac{C_{A0} - C_A}{\tau} - k_1C_A\\
+  \dfrac{dC_B}{dt} &= k_1C_A - \dfrac{C_B}{\tau}
+\end{align*}
+i.e, we assume that we don't know about the second reaction while 
+performing grey-box modeling.
 

### 2020-09-24 systemID_NNs.tex
ok.

@@
 \title{Data-based completion of grey-box models with neural networks}
-\author{Problem formulation}
+\author{Pratyush Kumar}
 \date{\today}
@@
 height and temperature. The control input is $u = C_{A0}$,
+the states are $x = [C_A, C_B, C_C]'$, 
 and the measurements are $y = [C_A, C_B]'$. The plant model 
@@
 i.e, we assume that we don't know about the second reaction while 
-performing grey-box modeling.
+performing grey-box modeling. Assume the following 
+parameter choices $k_1 = 1 \ \textnormal{m}^3/\textnormal{min}, 
+\ k_2 = 10^{-2} \ \textnormal{m}^3/\textnormal{min}
+\ k_3 = 10^{-3} \ \textnormal{m}^3/\textnormal{min}$, 
+and $\tau = 20 \ \textnormal{min}$.
+
+Since the measurements are a simple linear function 
+of the grey-box and plant states, we use only one network to 
+augment the grey-box dynamics. The expected function 
+to be approximated with a neural network is
+\begin{align}
+  f_{NN}(\mathbf{y}_{k:k-N}, \mathbf{u}_{k-1:k-N}) \approx (e^{A\Delta})^N
+  \mathcal{O}(C, e^{A\Delta})^{\dagger}
+\end{align}
+
+\begin{figure}
+  \centering
+  \includegraphics[page=2, width=\textwidth]{tworeac_plots.pdf}
+\end{figure}
 

### 2020-10-15 systemID_NNs.tex
will work slightly to modify the plant cost curve.

--- 2020-09-23 systemID_NNs.tex
@@
 \begin{align*}
-  \mathbf{y}_{k:k-N} &= [y(k)', y(k-\Delta)', ..., y(k-N\Delta)']' \\
-  \mathbf{u}_{k-1:k-N} &= [u(k-1)', u(k-2)', ..., u(k-N\Delta)']'
+  \mathbf{y}_{k+1-N_p:k} &= [y(k)', y(k-\Delta)', ..., y(k-(N_p-1)\Delta)']' \\
+  \mathbf{u}_{k+1-N_p:k-1} &= [u(k-\Delta)', u(k-2\Delta)', 
+                               ..., u(k-(N_p-1)\Delta)']'
 \end{align*}
 
-in which $\Delta$ is the sample 
-time between two consecutive
-prediction instants of the desired discrete time 
-model to be used in MPC. There exists a $\phi(\cdot)$ 
+in which $N_p$ is the number of 
+past measurements and
+$\Delta$ is the sample time between two consecutive
+measurements. There exists a $\phi(\cdot)$ 
 and a number of delay 
-time steps $N$, such that the plant state
+time steps $N_p$, such that the plant state
 at a given time can be written as
-$x(k) = \phi(\mathbf{y}_{k:k-N}, \mathbf{u}_{k-1:k-N})$.
+$x(k) = \phi(\mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k+1-N_p:k-1})$.
+For linear plants, this function $\phi(\cdot)$
+is known analytically in terms of the plant dynamic
+matrices.
 
@@
 \dot x_g &= f_g(x_g, u) + 
-f_{NN}(\mathbf{y}_{k:k-N}, \mathbf{u}_{k:k-N})\\
-y &= h_g(x_g) + h_{NN}(\mathbf{y}_{k:k-N}, \mathbf{u}_{k-1:k-N})
+f_{NN}(\mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k+1-N_p:k})\\
+y &= h_g(x_g) + h_{NN}(\mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k+1-N_p:k-1})
 \end{align*}
@@
 \begin{align*}
-  f_{NN}(\mathbf{y}_{k:k-N}, \mathbf{u}_{k:k-N}) &\approx 
-      Gf_p(\phi(\mathbf{y}_{k:k-N}, \mathbf{u}_{k-1:k-N}), u) - f_g(x_g, u) = 
-      Gf_p(x, u) - f_g(x_g, u) \\
-      h_{NN}(\mathbf{y}_{k:k-N}, \mathbf{u}_{k-1:k-N}) &\approx 
-      h_p(\phi(\mathbf{y}_{k:k-N}, \mathbf{u}_{k-1:k-N})) - h_g(x_g) = 
-      h_p(x) - h_g(x_g) 
+  f_{NN}(\mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k+1-N_p:k}) &\approx 
+      Gf_p(x, u) - f_g(Gx, u) \\
+      h_{NN}(\mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k+1-N_p:k-1}) &\approx
+      h_p(x) - h_g(Gx) 
 \end{align*}  
+in which $x(k) = \phi(\mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k+1-N_p:k-1})$
+is an unknown function of past measurements and inputs.
 
@@
 \begin{align*}
-  k_1 &= f_g(x_g(k), u(k)) + f_{NN}(\mathbf{y}_{k:k-N}, \mathbf{u}_{k:k-N}) \\
+  k_1 &= f_g(x_g(k), u(k)) + 
+  f_{NN}(\mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k+1-N_p:k}) \\
   k_2 &= f_g(x_g(k) + (\Delta/2)k_1, u(k)) + 
-         f_{NN}(\mathbf{y}_{k+0.5:k-N+0.5}, 
-                \mathbf{u}_{k+0.5:k-N+0.5})\\
+  f_{NN}(\mathbf{y}_{k+1.5-N_p:k+0.5}, \mathbf{u}_{k+1.5-N_p:k+0.5})\\
   k_3 &= f_g(x_g(k) + (\Delta/2)k_2, u(k)) + 
-                f_{NN}(\mathbf{y}_{k+0.5:k-N+0.5}, 
-                      \mathbf{y}_{k+0.5:k-N+0.5})\\
+  f_{NN}(\mathbf{y}_{k+1.5-N_p:k+0.5}, \mathbf{u}_{k+1.5-N_p:k+0.5})\\
   k_4 &= f_g(x_g(k) + \Delta k_3, u(k)) + 
-                       f_{NN}(\mathbf{y}_{k+1:k-N+1}, 
-                              \mathbf{u}_{k+1:k-N+1})\\
+  f_{NN}(\mathbf{y}_{k+2-N_p:k+1}, \mathbf{u}_{k+2-N_p:k+1})\\
   x_g(k+\Delta) &= x_g(k) + (\Delta/6)(k_1 + 2k_2 + 2k_3 + k_4) \\
-  \hat{y}(k+\Delta) &= h_g(x_g(k+\Delta)) + h_{NN}(\mathbf{y}_{k+\Delta:k-N+\Delta}, 
-                                     \mathbf{u}_{k-1+\Delta:k-N+\Delta})
+  \hat{y}(k+\Delta) &= h_g(x_g(k+\Delta)) + 
+  h_{NN}(\mathbf{y}_{k+2-N_p:k+1}, \mathbf{u}_{k+2-N_p:k+1})
 \end{align*}
-in which the sequences $\mathbf{y}_{k+0.5:k-N+0.5}$ 
-and $\mathbf{u}_{k+0.5:k-N+0.5}$ are defined as 
+in which the sequences $\mathbf{y}_{k+1.5-N_p:k+0.5}$ 
+and $\mathbf{u}_{k+1.5-N_p:k+0.5}$ are defined as 
 \begin{align*}
-  \mathbf{y}_{k+0.5:k-N+0.5} &= 
-    [y(k + \Delta/2)', y(k-\Delta/2)', ..., y(k-N\Delta + \Delta/2)']' \\
-  \mathbf{u}_{k+0.5:k-N+0.5} &= 
-  [u(k + \Delta/2)', u(k-\Delta/2)', ..., u(k-N\Delta + \Delta/2)']' 
+  \mathbf{y}_{k+1.5-N_p:k+0.5} &= 
+    [y(k + \Delta/2)', y(k-\Delta/2)', ..., y(k-(N_p-1)\Delta + \Delta/2)']' \\
+    \mathbf{u}_{k+1.5-N_p:k+0.5} &= 
+  [u(k + \Delta/2)', u(k-\Delta/2)', ..., u(k-(N_p-1)\Delta + \Delta/2)']' 
 \end{align*}
@@
 MPC execution instants, such that $u(k + \Delta/2) = u(k)$. To get the 
-measurements between two sample instants ($y(k + \Delta/2)$) for 
-the RK4 predictions, a linear interpolation can be used 
-or we can assume that additional 
-measurements from the plant are available 
-between two MPC execution instants as well.
+measurements between two sample instants (for e.g., $y(k + \Delta/2)$) for 
+the RK4 predictions, linear interpolation is used.
 
@@
 \begin{align*}
- \underset{W_{fi}, b_{fi}, W_{hi}, b_{hi}}{\textnormal{min}} \sum_{i=0}^{N_s-1} 
- \dfrac{1}{2}(\hat{y}(i) - y(i))^2
+ \underset{W_{fi}, b_{fi}, W_{hi}, b_{hi}}{\textnormal{min}} \sum_{k=0}^{N_s-1} 
+ \dfrac{1}{2}(\hat{y}(k) - y(k))^2
 \end{align*}
@@
 two neural networks ($f_{NN}(\cdot)$ and $h_{NN}(\cdot)$), 
-$\hat{y}$ are predictions of the augmented model, and $y$
+$\hat{y}$ are the predictions of the augmented
+grey-box + neural network model, and $y$
 are the measurements from the plant.
@@
 
-\subsection{Simple example}
+\subsection{Linear plant}
 
@@
 height and temperature. The control input is $u = C_{A0}$,
-and the measurements are $y = [C_A, C_B]'$. The plant model 
-is observable with these choices of the measurements and dynamics.
+the states are $x = [C_A, C_B, C_C]'$, 
+and the measurements are $y = [C_A, C_B]'$. 
+Assume the following 
+parameters for this model, $k_1 = 1 \ \textnormal{m}^3/\textnormal{min}, 
+\ k_2 = 0.3 \ \textnormal{m}^3/\textnormal{min},
+\ k_3 = 0.2 \ \textnormal{m}^3/\textnormal{min}$, 
+and $\tau = 5 \ \textnormal{min}$.
+The plant model 
+is linear ($dx/dt = Ax + Bu$) and 
+observable with the choices of 
+the parameters and measurements.
 
@@
 performing grey-box modeling.
+
+The current state $x(k)$ can be constructed using the history of 
+past measurements and inputs as 
+\begin{equation*}
+  x(k) = A^{N_p-1}_d
+  \mathcal{O}^{\dagger}(\mathbf{y}_{k+1-N_p:k} - 
+                        \mathcal{B}\mathbf{u}_{k+1-N_p:k-1}) + \\ 
+      \mathcal{B}_{nN_p-n+1:nN_p}\mathbf{u}_{k+1-N_p:k-1}
+\end{equation*}
+
+in which $A_d$ is the plant state transition matrix in discrete-time, 
+$\mathcal{O}$ is the observability matrix, 
+and $\mathcal{B}$ is a combination of the 
+discrete time matrices ($A_d$ and $B_d$) such that it 
+denotes the response of a vector of inputs to states, 
+and $\mathcal{B}_{nN_p-n+1:nN_p}$ denotes the last $n$
+rows of $\mathcal{B}$.
+
+Since the measurements are a simple linear function 
+of the grey-box and plant states, we use only one 
+neural network to 
+augment the grey-box dynamics. 
+The expected function 
+to be approximated with the network is
+\begin{align*}
+  f_{NN}(\mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k+1-N_p:k-1}) \approx 
+  \begin{bmatrix}
+  0 & 0 & 0 \\  
+  0 & -k_2 & k_3 
+  \end{bmatrix}x(k)
+\end{align*}
+
+In this example, the network just approximates a linear function.
+The figure below shows validation performances of the grey-box 
+and augmented hybrid models. The structure 
+of the network is restricted to just a matrix and
+the augmented model was trained on 120 samples.
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[page=2, width=\textwidth]{tworeac_plots_lin.pdf}
+\end{figure}
+
+\subsection{Nonlinear plant}
+The plant is made nonlinear by changing the side reaction 
+to a third order reaction ($A \rightarrow B $ and  $3B \rightleftharpoons C$),
+and is now simulated with the following ODEs
+\begin{align*}
+  \dfrac{dC_A}{dt} &= \dfrac{C_{A0} - C_A}{\tau} - k_1C_A\\
+  \dfrac{dC_B}{dt} &= k_1C_A - 3k_2C^3_B + 3k_3C_C- \dfrac{C_B}{\tau}\\
+  \dfrac{dC_C}{dt} &= k_2C^3_B - k_3C_C - \dfrac{C_C}{\tau}
+\end{align*}
+
+The grey-box model is kept the same. The parameters 
+in the model are modified to 
+$k_1 = 1 \ \textnormal{m}^3/\textnormal{min}, 
+\ k_2 = 0.05 \ \textnormal{m}^3/\textnormal{min},
+\ k_3 = 0.1 \ \textnormal{m}^3/\textnormal{min}$, 
+and $\tau = 5 \ \textnormal{min}$.
+An augmented model with a neural
+network architecture [8, 4, 2] is trained on 180 samples. The 
+tanh activation is used in the network. The figure below shows 
+validation performances of the grey-box and augmented 
+hybrid models.
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[page=2, width=\textwidth]{tworeac_plots_nonlin.pdf}
+\end{figure}
 

### 2020-10-15 lib/hybridid.py

@@ -38,6 +38,22 @@ class PickleTool:
         with open(filename, "wb") as stream:
             pickle.dump(data_object, stream)
 
+def c2d(A, B, sample_time):
+    """ Custom c2d function for linear systems."""
+    
+    # First construct the incumbent matrix
+    # to take the exponential.
+    (Nx, Nu) = B.shape
+    M1 = np.concatenate((A, B), axis=1)
+    M2 = np.zeros((Nu, Nx+Nu))
+    M = np.concatenate((M1, M2), axis=0)
+    Mexp = scipy.linalg.expm(M*sample_time)
+
+    # Return the extracted matrices.
+    Ad = Mexp[:Nx, :Nx]
+    Bd = Mexp[:Nx, -Nu:]
+    return (Ad, Bd)
+    
 class NonlinearPlantSimulator:
     """Custom class for simulating non-linear plants."""
     def __init__(self, *, fxup, hx, Rv, Nx, Nu, Np, Ny, 

### 2020-10-15 tworeac_ssopt_nonlin.py

@@ -1,62 +1,96 @@
 # [depends] %LIB%/hybridid.py tworeac_parameters_nonlin.pickle
 # [makes] pickle
-""" Script to train the hybrid model for the 
-    three reaction system. 
+""" Script to use the trained hybrid model for 
+    steady-state optimization.
     Pratyush Kumar, pratyushkumar@ucsb.edu """
 
 import sys
 sys.path.append('lib/')
-import tensorflow as tf
-import time
 import numpy as np
-from hybridid import (PickleTool, get_tworeac_train_val_data, 
-                      SimData)
-from HybridModelLayers import TwoReacModel
+import casadi
+import matplotlib.pyplot as plt
+import mpctools as mpc
+from matplotlib.backends.backend_pdf import PdfPages
+from hybridid import (PickleTool, PAPER_FIGSIZE)
+from tworeac_parameters_nonlin import (_tworeac_plant_ode, 
+                                       _tworeac_greybox_ode)
 
-# Set the tensorflow graph-level seed.
-tf.random.set_seed(1)
+def _plant_ys(*, us, parameters):
+    """ Return the plant xs. 
+        Solve: f(x, u) = 0 (continuous time)
+    """
+    # Construct and return the plant.
+    ps = parameters['ps']
+    Nx, Ny = parameters['Nx'], parameters['Ny']
+    plant = lambda x: _tworeac_plant_ode(x, us, ps, parameters)
+    xs = casadi.SX.sym('xs', Nx)
+    plant = mpc.getCasadiFunc(plant, [Nx], ['xs'])
+    rootfinder = casadi.rootfinder('fxu', 'newton', plant)
+    xguess = 0.5*np.ones((Nx, 1))
+    xs = np.asarray(rootfinder(xguess))
+    # Return.
+    return xs[0:Ny]
 
-def create_tworeac_model(*, Np, fnn_dims, tworeac_parameters):
-    """ Create/compile the two reaction model for training. """
-    tworeac_model = TwoReacModel(Np=Np,
-                                 fnn_dims=fnn_dims,
-                                 tworeac_parameters=tworeac_parameters)
-    # Compile the nn controller.
-    tworeac_model.compile(optimizer='adam', 
-                          loss='mean_squared_error')
+def _greybox_ys(*, us, parameters):
+    """ Return the plant xs. 
+        Solve: Ax + Bu = 0 (continuous time SS) 
+    """
[…]

### 2020-10-16 tworeac_ssopt_nonlin.py
make some writing edits in a little bit.

@@ -1,11 +1,12 @@
 # [depends] %LIB%/hybridid.py tworeac_parameters_nonlin.pickle
-# [makes] pickle
+# [depends] tworeac_train_nonlin.pickle
 """ Script to use the trained hybrid model for 
     steady-state optimization.
     Pratyush Kumar, pratyushkumar@ucsb.edu """
 
 import sys
 sys.path.append('lib/')
+sys.path.append('../')
 import numpy as np
 import casadi
 import matplotlib.pyplot as plt
@@ -146,8 +147,9 @@ def solve_plant_hybrid_nlps(*, Np, fnn_weights, parameters):
      hybrid_us) = (hybrid_nlp_soln['f'].full().squeeze(axis=-1)[0], 
                     hybrid_nlp_soln['g'].full().squeeze()[-1])
 
-    print("Plant cost: " + str(plant_cost) + "Plant u: " + str(plant_us))
-    print("Hybrid cost: " + str(hybrid_cost) + "Plant u: " + str(hybrid_us))
+    print("Plant cost: " + str(plant_cost) + ", Plant u: " + str(plant_us))
+    print("Hybrid cost: " + str(hybrid_cost) + ", Plant u: " + str(hybrid_us))
+    return 
 
 def main():
     """ Main function to be executed. """
@@ -165,7 +167,9 @@ def main():
       cost_hybrid, us) = compute_cost_curves(Np=Np, fnn_weights=fnn_weights,
                                     parameters=tworeac_parameters['parameters'])
     figures = plot_profit_curve(us=us, 
-                                costs=[cost_plant, cost_greybox, cost_hybrid])
+                                costs=[cost_plant, cost_greybox, cost_hybrid],
+                                ylabel_xcoordinate=-0.18,
+                                left_label_frac=0.18)
     solve_plant_hybrid_nlps(Np=Np, fnn_weights=fnn_weights,
                             parameters=tworeac_parameters['parameters'])
     # Save data.

### 2020-10-23 systemID_NNs.tex
make some writing edits in a little bit.
getting closer.

@@
 In this example, the network just approximates a linear function.
-The figure below shows validation performances of the grey-box 
+The Figure \ref{fig:validation_linear} 
+below shows validation performances of the grey-box 
 and augmented hybrid models. The structure 
 of the network is restricted to just a matrix and
-the augmented model was trained on 120 samples.
+the augmented model was trained on 240 samples (4 hours).
 
@@
   \includegraphics[page=2, width=\textwidth]{tworeac_plots_lin.pdf}
+  \caption{Model validation for the linear plant}
+  \label{fig:validation_linear}
+\end{figure}
+
+We notice that the augmented model accurately predicts the plant 
+behavior. The steady cost profile (Cost = $c_1C_{A0} - c_2C_B$)
+is shown in Figure \ref{fig:cost_linear}.
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[page=1, width=\textwidth]{tworeac_ssopt_lin.pdf}
+  \caption{Steady state cost curve for 
+           the linear plant, grey-box, and hybrid model.}
+  \label{fig:cost_linear}
 \end{figure}
@@
 $k_1 = 1 \ \textnormal{m}^3/\textnormal{min}, 
-\ k_2 = 0.05 \ \textnormal{m}^3/\textnormal{min},
-\ k_3 = 0.1 \ \textnormal{m}^3/\textnormal{min}$, 
+\ k_2 = 0.01 \ \textnormal{m}^3/\textnormal{min},
+\ k_3 = 0.05 \ \textnormal{m}^3/\textnormal{min}$, 
 and $\tau = 5 \ \textnormal{min}$.
 An augmented model with a neural
-network architecture [8, 4, 2] is trained on 180 samples. The 
-tanh activation is used in the network. The figure below shows 
+network architecture [8, 4, 2] is trained on 360 samples (6 hours). 
+The tanh activation is used in the network. The figure below shows 
 validation performances of the grey-box and augmented 
@@
   \includegraphics[page=2, width=\textwidth]{tworeac_plots_nonlin.pdf}
+  \caption{Model validation for the non linear plant}
+  \label{fig:validation_nonlinear}
+\end{figure}
+
+We notice that the augmented model accurately predicts the plant 
+behavior. The steady cost profile (Cost = $c_1C_{A0} - c_2C_B$)
+is shown in Figure \ref{fig:cost_nonlinear}. Upon optimization 
+using casadi, the optimum cost and control for the plant are: $-19.62 \$$ 
+and 1.13 $\textnormal{mol}/\textnormal{m}^3$, and for the hybrid 
+grey-box and NN model are: $-19.91 \$$ 
+and 1.11 $\textnormal{mol}/\textnormal{m}^3$.
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[page=1, width=\textwidth]{tworeac_ssopt_nonlin.pdf}
+  \caption{Steady state cost curve for the nonlinear plant, grey-box, and hybrid model.}
+  \label{fig:cost_nonlinear}
 \end{figure}

### 2020-10-23 lib/hybridid.py

@@ -154,7 +154,8 @@ def get_tworeac_train_val_data(*, Np, parameters, data_list):
                     outputs=outputs[-1])
     return (train_data, trainval_data, val_data)
 
-def plot_profit_curve(*, us, costs, figure_size=PAPER_FIGSIZE, 
+def plot_profit_curve(*, us, costs, colors, legends, 
+                         figure_size=PAPER_FIGSIZE,
                          ylabel_xcoordinate=-0.12, 
                          left_label_frac=0.15):
     """ Plot the profit curves. """
@@ -162,10 +163,8 @@ def plot_profit_curve(*, us, costs, figure_size=PAPER_FIGSIZE,
                                         sharex=True, 
                                         figsize=figure_size, 
                                     gridspec_kw=dict(left=left_label_frac))
-    xlabel = r'$C_{A0} \ (\textnormal{mol/m}^3)$'
+    xlabel = r'$C_{Af} \ (\textnormal{mol/m}^3)$'
     ylabel = r'Cost ($\$ $)'
-    colors = ['b', 'g', 'm']
-    legends = ['Plant', 'Grey-box', 'Hybrid']
     for (cost, color) in zip(costs, colors):
         # Plot the corresponding data.
         axes.plot(us, cost, color)

### 2020-10-23 tworeac_ssopt_nonlin.py

@@ -47,22 +47,24 @@ def _greybox_ys(*, us, parameters):
     # Return.
     return xs[0:Ng]
 
+def _blackbox_func(xs, us, Np, fnn_weights):
+    """ Blackbox fNN. """
+    xs = xs[:, np.newaxis]
+    nn_input = np.concatenate((np.tile(xs, (Np+1, 1)), 
+                                np.tile(us, (Np+1, 1))), axis=0)
+    for i in range(0, len(fnn_weights)-1, 2):
+        (W, b) = fnn_weights[i:i+2]
+        nn_input = np.tanh(W.T @ nn_input + b[:, np.newaxis])
+    nn_output = fnn_weights[-1].T @ nn_input
+    return -xs[:, 0] + nn_output[:, 0]
+
 def _blackbox_ys(*, us, Np, fnn_weights, parameters):
     """ Return the plant xs. 
         Solve: -xs + fnn(xs, 2*xs, 2*us, us) = 0 (continuous time SS) 
     """
-    def _fnn(xs, us, Np, fnn_weights):
-        xs = xs[:, np.newaxis]
-        nn_input = np.concatenate((np.tile(xs, (Np+1, 1)), 
-                                   np.tile(us, (Np+1, 1))), axis=0)
-        for i in range(0, len(fnn_weights)-1, 2):
-            (W, b) = fnn_weights[i:i+2]
-            nn_input = np.tanh(W.T @ nn_input + b[:, np.newaxis])
-        nn_output = fnn_weights[-1].T @ nn_input
-        return nn_output[:, 0]
     # Construct and return the plant.
     Ng = parameters['Ng']
-    blackbox = lambda xs: -xs + _fnn(xs, us, Np, fnn_weights)
+    blackbox = lambda xs: _blackbox_func(xs, us, Np, fnn_weights)
     blackbox = mpc.getCasadiFunc(blackbox, [Ng], ['xs'])
     rootfinder = casadi.rootfinder('fxu', 'newton', blackbox)
     xguess = 0.5*np.ones((Ng, 1))
@@ -70,10 +72,8 @@ def _blackbox_ys(*, us, Np, fnn_weights, parameters):
     # Return.
     return xGsys[-Ng:]
 
-def _residual_ys(*, us, Np, fnn_weights, parameters):
-    """ Return the plant xs. 
-        Solve: Ax + Bu + fnn() = 0 (continuous time SS) 
-    """
[…]

### 2020-10-23 tworeac_train_nonlin.py

@@ -69,16 +69,17 @@ def main():
                                    tworeac_parameters['training_data'])
 
     # Number of samples.
-    num_samples = [num_sample for num_sample in range(24*60, 24*60+1, 12*60)]
-    
+    #num_samples = [num_sample for num_sample in range(24*60, 24*60+1, 12*60)]
+    num_samples = [60, 3*60]
+
     # Create lists.
     Nps = [2, 3, 2]
-    fnn_dims = [[9, 8, 2], [9, 16, 2], [8, 4, 2]]
+    fnn_dims = [[9, 16, 16, 2], [9, 16, 16, 2], [8, 4, 2]]
     model_types = ['black-box', 'residual', 'hybrid']
     trained_weights = []
     val_metrics = []
     val_predictions = []
-    
+
     # Filenames.
     ckpt_path = 'tworeac_train_nonlin.ckpt'
     stdout_filename = 'tworeac_train_nonlin.txt'
@@ -103,11 +104,14 @@ def main():
             tworeac_model = create_tworeac_model(Np=Np, fnn_dims=fnn_dim,
                                                  tworeac_parameters=parameters, 
                                                  model_type=model_type)
+            train_samples = dict(x0=train_data['x0'],
+                            inputs=train_data['inputs'][:, :num_sample, :],
+                            outputs=train_data['outputs'][:, :num_sample, :]) 
             (tworeac_model, 
              val_prediction, 
              val_metric) = train_model(tworeac_model, 
-                                            train_data, trainval_data, val_data,
-                                            stdout_filename, ckpt_path)
+                                        train_samples, trainval_data, val_data,
+                                        stdout_filename, ckpt_path)
             fnn_weights = tworeac_model.get_weights()
 
             # Save info.

### 2020-11-23 systemID_NNs.tex
add scripts.

@@
 
+\newcommand{\useq}{\mathbf{u}}
+\newcommand{\xseq}{\mathbf{x}}
+\newcommand{\bbR}{\mathbb{R}}
+\newcommand{\bbW}{\mathbb{W}}
+\newcommand{\bbU}{\mathbb{U}}
+\newcommand{\bbI}{\mathbb{I}}
+\newcommand{\bbX}{\mathbb{X}}
+\newcommand{\bbP}{\mathbb{P}}
+
 \title{Data-based completion of grey-box models with neural networks}
@@
 \begin{align*}
-  \dot x &= f_p(x, u) \\
-  y &= h_p(x)
+  \dot x_p &= f_p(x_p, u, w) \\
+  y &= h_p(x_p) + v
 \end{align*}
-in which $x$ is the plant state (vector), $u$ is the manipulated control input,
-and $y$ is the measurement. Assume that we don't have any 
-disturbances/measurement and process noise 
-for now, and we'll add those after we 
-understand the issues in the nominal case.
+in which $x_p \in \bbR^{n_p}$ is the plant state, 
+$u \in \bbU$ is the manipulated control input,
+and $y \in \bbR^p$ is the measurement.
 
@@
 
-in which $x_g$ are the grey-box states, which are some subset 
-of the plant states (denote $x_g = Gx$) chosen to 
+in which $x_g \in \bbR^{n_g}$ are the grey-box states, 
+which are some subset 
+of the plant states (denote $x_g = Gx_p$) chosen to 
 be modeled. The dynamics of these grey-box states 
@@
 \begin{align*}
-  \mathbf{y}_{k+1-N_p:k} &= [y(k)', y(k-\Delta)', ..., y(k-(N_p-1)\Delta)']' \\
-  \mathbf{u}_{k+1-N_p:k-1} &= [u(k-\Delta)', u(k-2\Delta)', 
-                               ..., u(k-(N_p-1)\Delta)']'
+  \mathbf{y}_{k-N_p:k-1} &= [y(k-\Delta)', ..., y(k-N_p\Delta)']' \\
+  \mathbf{u}_{k-N_p:k-1} &= [u(k-\Delta)', 
+                               ..., u(k-N_p\Delta)']'
 \end{align*}
@@
 at a given time can be written as
-$x(k) = \phi(\mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k+1-N_p:k-1})$.
+$x_p(k) = \phi(\mathbf{y}_{k-N_p:k}, \mathbf{u}_{k-N_p:k-1})$.
 For linear plants, this function $\phi(\cdot)$
@@
 \dot x_g &= f_g(x_g, u) + 
-f_{NN}(\mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k+1-N_p:k})\\
-y &= h_g(x_g) + h_{NN}(\mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k+1-N_p:k-1})
+f_N(h_g(x_g), z, u)\\
+y &= h_g(x_g) 
 \end{align*}
-in which $f_{NN}(\cdot)$ and $h_{NN}(\cdot)$
-are standard feedforward neural networks. We add the current 
-control $u(k)$ as an input to the network $f_{NN}(\cdot)$ because 
-there may be a mismatch in the actuator to the grey-box dynamics
-as well.
+in which $f_N(\cdot)$ is a feedforward neural network
+and $z(k) = [\mathbf{y}'_{k-N_p:k-1}, \mathbf{u}'_{k-N_p:k-1}]'$.
 
-It is expected that the functions $f_{NN}(\cdot)$ 
-and $h_{NN}(\cdot)$ approximate
-the following differences between the true grey-box state dynamics
-and true measurement function with what is modeled.
-\begin{align*}
-  f_{NN}(\mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k+1-N_p:k}) &\approx 
-      Gf_p(x, u) - f_g(Gx, u) \\
-      h_{NN}(\mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k+1-N_p:k-1}) &\approx
-      h_p(x) - h_g(Gx) 
-\end{align*}  
-in which $x(k) = \phi(\mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k+1-N_p:k-1})$
-is an unknown function of past measurements and inputs.
+Alternative modeling choices are:
+\begin{itemize}
+	\item Black-box: $y^+ = f_N(y, z, u)$, (purely black-box)
+	\item Hybrid 1: $\dot x_g = f_N(x_g, z, u), \ y = h_g(x_g)$, 
+  (states have meaning)
+  \item Hybrid 2: $\dot x_g = f_g(x_g, u), \ y = h_g(x_g) + h_N(z)$, 
+	(model residual)
+	\item Hybrid 3: $\dot x_g = f_g(x_g, u, p(h_g(x_g), z)), \ y = h_g(x_g)$, (use NNs to learn some parameters)
+\end{itemize}
+
+%It is expected that the functions $f_{NN}(\cdot)$ 
+%and $h_{NN}(\cdot)$ approximate
+%the following differences between the true grey-box state dynamics
+%and true measurement function with what is modeled.
+%\begin{align*}
+%  f_{NN}(\mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k+1-N_p:k}) &\approx 
+%      Gf_p(x, u) - f_g(Gx, u) \\
+%      h_{NN}(\mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k+1-N_p:k-1}) &\approx
+%      h_p(x) - h_g(Gx) 
+%\end{align*}  
+%in which $x(k) = \phi(\mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k+1-N_p:k-1})$
+%is an unknown function of past measurements and inputs.
 
 \subsection{Conversion to discrete time and training}
-We note that the network $f_{NN}(\cdot)$ is an addition 
-to the grey-box dynamics in continuous time, whereas 
-$h_{NN}(\cdot)$ is an addition directly to the 
-measurement function. The Runge-Kutta-4 (RK4) method 
+We note that the network $f_N(\cdot)$ is an addition 
+to the grey-box dynamics in continuous time. 
+The Runge-Kutta-4 (RK4) method 
 is used for the discretization as follows
@@
 \begin{align*}
-  k_1 &= f_g(x_g(k), u(k)) + 
-  f_{NN}(\mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k+1-N_p:k}) \\
-  k_2 &= f_g(x_g(k) + (\Delta/2)k_1, u(k)) + 
-  f_{NN}(\mathbf{y}_{k+1.5-N_p:k+0.5}, \mathbf{u}_{k+1.5-N_p:k+0.5})\\
-  k_3 &= f_g(x_g(k) + (\Delta/2)k_2, u(k)) + 
-  f_{NN}(\mathbf{y}_{k+1.5-N_p:k+0.5}, \mathbf{u}_{k+1.5-N_p:k+0.5})\\
-  k_4 &= f_g(x_g(k) + \Delta k_3, u(k)) + 
-  f_{NN}(\mathbf{y}_{k+2-N_p:k+1}, \mathbf{u}_{k+2-N_p:k+1})\\
+  k_1 &= f_g(x_g, u) + 
+  f_N(h_g(x_g), \mathbf{y}_{k-N_p:k-1}, \mathbf{u}_{k+1-N_p:k}) \\
+  x_g & := x_g + (\Delta/2)k_1 \\
+  k_2 &= f_g(x_g, u) + 
+  f_N(h_g(x_g), \mathbf{y}_{k+0.5-N_p:k-0.5}, \mathbf{u}_{k+1-N_p:k})\\
+  x_g & := x_g + (\Delta/2)k_2 \\
+  k_3 &= f_g(x_g, u) + 
+  f_N(h_g(x_g), \mathbf{y}_{k+0.5-N_p:k-0.5}, \mathbf{u}_{k+1-N_p:k})\\
+  x_g & := x_g + \Delta k_3 \\
+  k_4 &= f_g(x_g, u) + 
+  f_N(h_g(x_g), \mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k+1-N_p:k})\\
   x_g(k+\Delta) &= x_g(k) + (\Delta/6)(k_1 + 2k_2 + 2k_3 + k_4) \\
-  \hat{y}(k+\Delta) &= h_g(x_g(k+\Delta)) + 
-  h_{NN}(\mathbf{y}_{k+2-N_p:k+1}, \mathbf{u}_{k+2-N_p:k+1})
+  \hat{y}(k+\Delta) &= h_g(x_g(k+\Delta)) 
 \end{align*}
-in which the sequences $\mathbf{y}_{k+1.5-N_p:k+0.5}$ 
-and $\mathbf{u}_{k+1.5-N_p:k+0.5}$ are defined as 
-\begin{align*}
-  \mathbf{y}_{k+1.5-N_p:k+0.5} &= 
-    [y(k + \Delta/2)', y(k-\Delta/2)', ..., y(k-(N_p-1)\Delta + \Delta/2)']' \\
-    \mathbf{u}_{k+1.5-N_p:k+0.5} &= 
-  [u(k + \Delta/2)', u(k-\Delta/2)', ..., u(k-(N_p-1)\Delta + \Delta/2)']' 
-\end{align*}
-We assume a zero order hold on the control inputs between two consecutive 
-MPC execution instants, such that $u(k + \Delta/2) = u(k)$. To get the 
-measurements between two sample instants (for e.g., $y(k + \Delta/2)$) for 
-the RK4 predictions, linear interpolation is used.
+in which the sequence $\mathbf{y}_{k+0.5-N_p:k-0.5}$ is obtained by 
+linear interpolation.
 
 The following optimization problem is solved to compute 
-the weights in the neural networks 
+the weights in the neural network
 \begin{align*}
- \underset{W_{fi}, b_{fi}, W_{hi}, b_{hi}}{\textnormal{min}} \sum_{k=0}^{N_s-1} 
+\underset{W_i, b_i}{\textnormal{min}} \sum_{k=0}^{N_s-1} 
 \dfrac{1}{2}(\hat{y}(k) - y(k))^2
  \end{align*}
-in which ($W_{fi}, b_{fi}, W_{hi}, b_{hi}$) are the weights in the 
-two neural networks ($f_{NN}(\cdot)$ and $h_{NN}(\cdot)$), 
+in which $W_i$, $b_i$ are the weights in each hidden layer of the NN, 
+$N_s$ is the number of training samples, 
 $\hat{y}$ are the predictions of the augmented
@@
 \begin{align*}
-  \dfrac{dC_A}{dt} &= \dfrac{C_{A0} - C_A}{\tau} - k_1C_A\\
+  \dfrac{dC_A}{dt} &= \dfrac{C_{Af} - C_A}{\tau} - k_1C_A\\
   \dfrac{dC_B}{dt} &= k_1C_A - k_2C_B + k_3C_C- \dfrac{C_B}{\tau}\\
@@
 $A \rightarrow B \rightleftharpoons C$ in a CSTR with constant 
-height and temperature. The control input is $u = C_{A0}$,
+height and temperature. The control input is $u = C_{Af}$,
 the states are $x = [C_A, C_B, C_C]'$, 
@@
 \begin{align*}
-  \dfrac{dC_A}{dt} &= \dfrac{C_{A0} - C_A}{\tau} - k_1C_A\\
+  \dfrac{dC_A}{dt} &= \dfrac{C_{Af} - C_A}{\tau} - k_1C_A\\
   \dfrac{dC_B}{dt} &= k_1C_A - \dfrac{C_B}{\tau}
@@
 
-The current state $x(k)$ can be constructed using the history of 
+The current plant state $x_p(k)$ can be constructed using the history of 
 past measurements and inputs as 
 \begin{equation*}
-  x(k) = A^{N_p-1}_d
+  x_p(k) = A^{N_p-1}_d
   \mathcal{O}^{\dagger}(\mathbf{y}_{k+1-N_p:k} - 
@@
 \begin{align*}
-  f_{NN}(\mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k+1-N_p:k-1}) \approx 
+  f_N(\mathbf{y}_{k-N_p:k}, \mathbf{u}_{k+1-N_p:k-1}) \approx 
   \begin{bmatrix}
@@
   0 & -k_2 & k_3 
-  \end{bmatrix}x(k)
+  \end{bmatrix}x_p(k)
 \end{align*}
@@
 
+THe Figures \ref{fig:cost_nonlinear_5hours} and 
+\ref{fig:cost_nonlinear_5hours}
+
 \begin{figure}[!h]
@@
 We notice that the augmented model accurately predicts the plant 
-behavior. The steady cost profile (Cost = $c_1C_{A0} - c_2C_B$)
+behavior. The steady cost profile (Cost = $c_1C_{Af} - c_2C_B$)
 is shown in Figure \ref{fig:cost_linear}.
@@
   \centering
-  \includegraphics[page=1, width=\textwidth]{tworeac_ssopt_lin.pdf}
+  \includegraphics[page=3, width=\textwidth]{tworeac_plots_lin.pdf}
   \caption{Steady state cost curve for 
@@
 \begin{align*}
-  \dfrac{dC_A}{dt} &= \dfrac{C_{A0} - C_A}{\tau} - k_1C_A\\
+  \dfrac{dC_A}{dt} &= \dfrac{C_{Af} - C_A}{\tau} - k_1C_A\\
   \dfrac{dC_B}{dt} &= k_1C_A - 3k_2C^3_B + 3k_3C_C- \dfrac{C_B}{\tau}\\
@@
 and $\tau = 5 \ \textnormal{min}$.
-An augmented model with a neural
-network architecture [8, 4, 2] is trained on 360 samples (6 hours). 
-The tanh activation is used in the network. The figure below shows 
-validation performances of the grey-box and augmented 
-hybrid models.
+
+\newpage
+Three models are compared: black-box, residual, and hybrid model with additive 
+NN to the grey-box dynamics. The chosen model architectures 
+are [9, 16, 16, 2], [9, 16, 16, 2], and [8, 16, 2] respectively.
+The Figure \ref{fig:validation_nonlinear} shows model validation on open loop 
+data after training with 10 hours of data.
+
+The Figures \ref{fig:cost_nonlinear_5hours} and \ref{fig:cost_nonlinear_10hours}
+show the steady state cost curve after training with 5 and 10 hours 
+of data respectively. We notice that the hybrid model is more data efficient 
+compared to the alternatives.
 
@@
 
-We notice that the augmented model accurately predicts the plant 
-behavior. The steady cost profile (Cost = $c_1C_{A0} - c_2C_B$)
-is shown in Figure \ref{fig:cost_nonlinear}. Upon optimization 
-using casadi, the optimum cost and control for the plant are: $-19.62 \$$ 
-and 1.13 $\textnormal{mol}/\textnormal{m}^3$, and for the hybrid 
-grey-box and NN model are: $-19.91 \$$ 
-and 1.11 $\textnormal{mol}/\textnormal{m}^3$.
+%We notice that the augmented model accurately predicts the plant 
+%behavior. The steady cost profile (Cost = $c_1C_{A0} - c_2C_B$)
+%is shown in Figure \ref{fig:cost_nonlinear}. Upon optimization 
+%using casadi, the optimum cost and control for the plant are: $-19.62 \$$ 
+%and 1.13 $\textnormal{mol}/\textnormal{m}^3$, and for the hybrid 
+%grey-box and NN model are: $-19.91 \$$ 
+%and 1.11 $\textnormal{mol}/\textnormal{m}^3$.
 
@@
   \centering
-  \includegraphics[page=1, width=\textwidth]{tworeac_ssopt_nonlin.pdf}
-  \caption{Steady state cost curve for the nonlinear plant, grey-box, and hybrid model.}
-  \label{fig:cost_nonlinear}
+  \includegraphics[page=3, width=\textwidth]{tworeac_plots_nonlin.pdf}
+  \caption{Steady state cost curve for the nonlinear plant, grey-box, 
+  black-box, residual, and hybrid model after training with 5 hours of data.}
+  \label{fig:cost_nonlinear_5hours}
+\end{figure}
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[page=4, width=\textwidth]{tworeac_plots_nonlin.pdf}
+  \caption{Steady state cost curve for the nonlinear plant, grey-box, 
+  black-box, residual, and hybrid model after training with 10 hours of data.}
+  \label{fig:cost_nonlinear_10hours}
+\end{figure}
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[page=1, width=\textwidth]{tworeac_distmodels_nonlin.pdf}
+  \caption{Steady state grey-box states and disturbance estimates based on 
+  output and input disturbance models}
+  \label{fig:cost_nonlinear_10hours}
 \end{figure}

### 2020-11-23 lib/HybridModelLayers.py

@@ -73,7 +73,7 @@ class TwoReacHybridCell(tf.keras.layers.AbstractRNNCell):
         Delta = self.tworeac_parameters['sample_time']
         
         # Get k1.
-        k1 = self._fg(xG, u) + self._fnn(xg, z)
+        k1 = self._fg(xG, u) + self._fnn(xG, z)
         
         # Interpolate for k2 and k3.
         ypseq_interp = self.interp_layer(tf.concat((ypseq, xG), axis=-1))

### 2020-11-23 lib/hybridid.py

@@ -3,7 +3,6 @@
 import sys
 import numpy as np
 import mpctools as mpc
-import pandas as pd
 import scipy.linalg
 import cvxopt as cvx
 import matplotlib.pyplot as plt
@@ -126,32 +125,36 @@ def get_tworeac_train_val_data(*, Np, parameters, data_list):
     """ Get the data for training in appropriate format. """
     tsteps_steady = parameters['tsteps_steady']
     (Ny, Nu) = (parameters['Ny'], parameters['Nu'])
-    (inputs, x0, outputs) = ([], [], [])
+    (inputs, xGz0, outputs) = ([], [], [])
     for data in data_list:
-        Nsim = len(data.t)
-        u_traj = data.u[tsteps_steady:][np.newaxis, :, np.newaxis]
-        x0_traj = data.y[tsteps_steady, :][np.newaxis, :]
-        y_traj = data.y[tsteps_steady:, :][np.newaxis, ...]
-        (ypseq_traj, upseq_traj) = ([], [])
-        for t in range(tsteps_steady, Nsim):
-            ypseq = data.y[t-Np:t, :].reshape(Np*Ny, )
-            upseq = data.u[t-Np:t]
-            ypseq_traj.append(ypseq)
-            upseq_traj.append(upseq)
-        ypseq_traj = np.asarray(ypseq_traj)[np.newaxis, ...]
-        upseq_traj = np.asarray(upseq_traj)[np.newaxis, ...]
-        # Collect the trajectories in list. 
-        inputs.append(np.concatenate((u_traj, ypseq_traj, upseq_traj), axis=-1))
-        x0.append(x0_traj)
+        t = tsteps_steady
+        
+        # Get input trajectory.
+        u_traj = data.u[t:][np.newaxis, :, np.newaxis]
+        
+        # Get initial state.
+        x0 = data.y[t, :][np.newaxis, :]
+        yp0seq = data.y[t-Np:t, :].reshape(Np*Ny, )[np.newaxis, :]
+        up0seq = data.u[t-Np:t][np.newaxis, :]
+        xGz0_traj = np.concatenate((x0, yp0seq, up0seq), axis=-1)
+
+        # Get output trajectory.       
+        y_traj = data.y[t:, :][np.newaxis, ...]
+        
+        # Collect the trajectories in list.
+        inputs.append(u_traj)
+        xGz0.append(xGz0_traj)
         outputs.append(y_traj)
+    
[…]

### 2020-11-23 tworeac_parameters_lin.py

@@ -88,7 +88,7 @@ def _get_tworeac_parameters():
     parameters['tsteps_steady'] = 60
 
     # Measurement noise.
-    parameters['Rv'] = np.diag([1e-3, 1e-2])
+    parameters['Rv'] = 0*np.diag([1e-3, 1e-2])
 
     # Return the parameters dict.
     return parameters
@@ -157,7 +157,7 @@ def _gen_train_val_data(*, parameters,
     for _ in range(num_traj):
         plant = _get_tworeac_model(parameters=parameters, plant=True)
         us_init = np.tile(np.random.uniform(ulb, uub), (tsteps_steady, 1))
-        u = sample_prbs_like(num_change=8, num_steps=Nsim, 
+        u = sample_prbs_like(num_change=4, num_steps=Nsim, 
                              lb=ulb, ub=uub,
                              mean_change=30, sigma_change=2, seed=seed+1)
         u = np.concatenate((us_init, u), axis=0)
@@ -251,7 +251,7 @@ def main():
     delta_matrix = _check_obsv_compute_delta(parameters=parameters)
     # Generate training data.
     training_data = _gen_train_val_data(parameters=parameters, 
-                                        num_traj=3, Nsim=240, seed=10)
+                                        num_traj=3, Nsim=120, seed=10)
     greybox_val_data = _get_greybox_val_preds(parameters=
                                             parameters, 
                                             training_data=training_data)

### 2020-11-23 tworeac_parameters_nonlin.py

@@ -173,13 +173,13 @@ def _gen_train_val_data(*, parameters, num_traj,
         elif traj == num_traj-2:
             "Get input for validation simulation."
             Nsim = Nsim_trainval
-            u = sample_prbs_like(num_change=12, num_steps=Nsim_trainval, 
+            u = sample_prbs_like(num_change=6, num_steps=Nsim_trainval, 
                                  lb=ulb, ub=uub,
-                                 mean_change=10, sigma_change=2, seed=seed+2)
+                                 mean_change=20, sigma_change=2, seed=seed+2)
         else:
             "Get input for training simulation."
             Nsim = Nsim_train
-            u = sample_prbs_like(num_change=96, num_steps=Nsim_train, 
+            u = sample_prbs_like(num_change=48, num_steps=Nsim_train, 
                                  lb=ulb, ub=uub,
                                  mean_change=30, sigma_change=2, seed=seed+3)
 
@@ -221,7 +221,7 @@ def main():
     
     # Generate training data.
     training_data = _gen_train_val_data(parameters=parameters, 
-                                        num_traj=3, Nsim_train=2880,
+                                        num_traj=3, Nsim_train=1440,
                                         Nsim_trainval=120, Nsim_val=720, 
                                         seed=100)
     greybox_val_data = _get_greybox_val_preds(parameters=

### 2020-11-23 tworeac_plots_lin.py

@@ -88,7 +88,7 @@ def plot_val_model_predictions(*, plantsim_data,
     axes.set_xlim([np.min(time), np.max(time)])
     figure.legend(handles = legend_handles,
                   labels = ('Plant', 'Grey-box', 'Hybrid'), 
-                  loc = (0.2, 0.9), ncol=3)
+                  loc = (0.18, 0.9), ncol=3)
     # Return the figure object.
     return [figure]
 

### 2020-11-23 tworeac_plots_nonlin.py

@@ -58,7 +58,7 @@ def plot_val_model_predictions(*, plantsim_data,
     ylabels = [r'$C_A \ (\textnormal{mol/m}^3)$', 
                r'$C_B \ (\textnormal{mol/m}^3)$',
                r'$C_{Af} \ (\textnormal{mol/m}^3)$']
-    model_legend_colors = ['green', 'dimgray', 'orange', 'tomato']
+    model_legend_colors = ['green', 'tomato']
     legend_handles = []
     plant_data_list = [plantsim_data.y[tsteps_steady:, 0], 
                        plantsim_data.y[tsteps_steady:, 1], 
@@ -86,9 +86,9 @@ def plot_val_model_predictions(*, plantsim_data,
     legend_handles.insert(0, plant_legend_handle[0])
     axes.set_xlabel('Time (hr)')
     axes.set_xlim([np.min(time[start:end]), np.max(time[start:end])])
-    labels = ('Plant', 'Grey-box', 'Black-box', 'Residual', 'Hybrid')
+    labels = ('Plant', 'Grey-box', 'Hybrid')
     figure.legend(handles = legend_handles, labels = labels, 
-                  loc = (0.16, 0.9), ncol=3)
+                  loc = (0.13, 0.9), ncol=3)
     # Return the figure object.
     return [figure]
 
@@ -102,11 +102,11 @@ def plot_sub_gaps(*, num_samples, sub_gaps, colors, legends,
                                   figsize=figure_size, 
                                   gridspec_kw=dict(left=left_label_frac))
     ylabel = r'$\% \ $ Suboptimality Gap'
-    xlabel = 'Hours training samples'
+    xlabel = 'Hours of training samples'
     num_samples = num_samples/60
     for (sub_gap, color) in zip(sub_gaps, colors):
         # Plot the corresponding data.
-        axes.semilogy(num_samples, sub_gap, color)
+        axes.plot(num_samples, sub_gap, color)
     axes.legend(legends)
     axes.set_xlabel(xlabel)
     axes.set_ylabel(ylabel)
@@ -117,14 +117,14 @@ def plot_sub_gaps(*, num_samples, sub_gaps, colors, legends,
 
 def plot_val_metrics(*, num_samples, val_metrics, colors, legends, 
                      figure_size=PAPER_FIGSIZE,
-                     ylabel_xcoordinate=-0.13, 
+                     ylabel_xcoordinate=-0.11, 
                      left_label_frac=0.15):
     """ Plot validation metric on open loop data. """
[…]

### 2020-11-23 tworeac_ssopt_nonlin.py

@@ -276,6 +276,20 @@ def compute_suboptimality_gaps(*, Nps, model_types,
         sub_gaps.append(model_sub_gaps)
     return sub_gaps
 
+def compute_cost_mses(costss):
+    """ Get the absolute difference between the 
+        plant cost and model costs. """
+    num_models = len(costss[0][2:])
+    cost_mses = []
+    for i in range(num_models):
+        model_cost_mse = []
+        for costs in costss:
+            plant = costs[0]
+            Delta = np.abs(plant - costs[i+2]).squeeze()
+            model_cost_mse.append(Delta)
+        cost_mses.append(model_cost_mse)
+    return cost_mses
+
 def main():
     """ Main function to be executed. """
     # Load data.
@@ -286,21 +300,31 @@ def main():
     tworeac_train = PickleTool.load(filename='tworeac_train_nonlin.pickle',
                                     type='read')
     (Nps, model_types, 
-     trained_weights) = (tworeac_train['Nps'], 
-                         tworeac_train['model_types'],
-                         tworeac_train['trained_weights'])
-    
+     trained_weights,
+     num_samples) = (tworeac_train['Nps'], 
+                     tworeac_train['model_types'],
+                     tworeac_train['trained_weights'],
+                     tworeac_train['num_samples'])
+
     # Create cost curves.
-    fnn_weights = []
-    for model_trained_weights in trained_weights:
-        fnn_weights.append(model_trained_weights[-1])
-    (costs, us) = compute_cost_curves(Nps=Nps, model_types=model_types,
-                                      fnn_weights=fnn_weights,
-                                      parameters=parameters)
-    sub_gaps = compute_suboptimality_gaps(Nps=Nps, model_types=model_types,
-                                          trained_weights=trained_weights, 
+    costss = []
+    num_samples = list(num_samples)
+    for (i, num_sample) in enumerate(num_samples):
+        fnn_weights = []
+        for model_trained_weights in trained_weights:
+            fnn_weights.append(model_trained_weights[i])
+        (costs, us) = compute_cost_curves(Nps=Nps, model_types=model_types,
+                                          fnn_weights=fnn_weights,
                                           parameters=parameters)
-    PickleTool.save(data_object=dict(us=us, costs=costs, sub_gaps=sub_gaps),
+        costss.append(costs)
+    
+    cost_mses = compute_cost_mses(costss=costss)
+    sub_gaps = compute_suboptimality_gaps(Nps=Nps, model_types=model_types,
[…]

### 2020-11-23 tworeac_train_lin.py

@@ -43,7 +43,7 @@ def train_model(model, train_data, trainval_data, val_data,
     # Call the fit method to train.
     model.fit(x=[train_data['inputs'], train_data['x0']], 
               y=train_data['outputs'], 
-            epochs=5000, batch_size=1,
+            epochs=1000, batch_size=1,
             validation_data = ([trainval_data['inputs'], trainval_data['x0']], 
                                 trainval_data['outputs']),
             callbacks = [checkpoint_callback])

### 2020-11-23 tworeac_train_nonlin.py

@@ -14,8 +14,8 @@ from hybridid import (PickleTool, get_tworeac_train_val_data,
                       SimData)
 from HybridModelLayers import TwoReacModel
 
-# Set the tensorflow graph-level seed.
-tf.random.set_seed(1)
+# Set the tensorflow global and graph-level seed.
+tf.random.set_seed(123)
 
 def create_tworeac_model(*, Np, fnn_dims, tworeac_parameters, model_type):
     """ Create/compile the two reaction model for training. """
@@ -41,20 +41,20 @@ def train_model(model, train_data, trainval_data, val_data,
                                                     save_weights_only=True,
                                                     verbose=1)
     # Call the fit method to train.
-    model.fit(x=[train_data['inputs'], train_data['x0']], 
+    model.fit(x=[train_data['inputs'], train_data['xGz0']], 
               y=train_data['outputs'], 
-            epochs=3000, batch_size=1,
-            validation_data = ([trainval_data['inputs'], trainval_data['x0']], 
-                                trainval_data['outputs']),
+              epochs=2000, batch_size=1,
+        validation_data = ([trainval_data['inputs'], trainval_data['xGz0']], 
+                            trainval_data['outputs']),
             callbacks = [checkpoint_callback])
-    
+
     # Get predictions on validation data.
     model.load_weights(ckpt_path)
-    model_predictions = model.predict(x=[val_data['inputs'], val_data['x0']])
+    model_predictions = model.predict(x=[val_data['inputs'], val_data['xGz0']])
     val_predictions = SimData(t=None, x=None, u=None,
                               y=model_predictions.squeeze())
     # Get prediction error on the validation data.
-    val_metric = model.evaluate(x=[val_data['inputs'], val_data['x0']], 
+    val_metric = model.evaluate(x=[val_data['inputs'], val_data['xGz0']], 
                                y=val_data['outputs'])
     # Return the NN controller.
     return (model, val_predictions, val_metric)
@@ -69,12 +69,12 @@ def main():
                                    tworeac_parameters['training_data'])
 
     # Number of samples.
-    num_samples = [num_sample for num_sample in range(48*60, 48*60+1, 60)]
+    num_samples = [hour*60 for hour in [6]]
 
     # Create lists.
-    Nps = [2, 3, 2]
-    fnn_dims = [[9, 16, 16, 2], [9, 16, 16, 2], [8, 16, 2]]
-    model_types = ['black-box']
+    Nps = [2, 2]
[…]

### 2020-11-23 systemID_NNs.tex
edit.

@@
 \documentclass{article}
-
+\usepackage[total={6in, 8in}]{geometry}
 %% The graphicx and color packages should give you all the color,
@@
 
-\title{Data-based completion of grey-box models with neural networks}
+\title{Hybrid nonlinear system identification using neural networks}
 \author{Pratyush Kumar}
@@
 \begin{align*}
-  \dot x_p &= f_p(x_p, u, w) \\
+  \dot{x}_p &= f_p(x_p, u, w) \\
   y &= h_p(x_p) + v
@@
 $u \in \bbU$ is the manipulated control input,
-and $y \in \bbR^p$ is the measurement.
+$y \in \bbR^p$ is the measurement, $w$ is the process noise, 
+and $v$ is the measurement noise.
 
@@
 \begin{align*}
-  \mathbf{y}_{k-N_p:k-1} &= [y(k-\Delta)', ..., y(k-N_p\Delta)']' \\
-  \mathbf{u}_{k-N_p:k-1} &= [u(k-\Delta)', 
-                               ..., u(k-N_p\Delta)']'
+  \mathbf{y}_{k-N_p:k-1} &= [y(k-N_p\Delta)', ... \ , y(k-\Delta)']' \\
+  \mathbf{u}_{k-N_p:k-1} &= [u(k-N_p\Delta)', 
+                               ... \ , u(k-\Delta)']'
 \end{align*}
@@
 neural networks as follows -
+\begin{align} \label{eq:hybrid_cont_time}
+\dot x_g &= f_g(x_g, u) + 
+f_N(x_g, z, u) \notag \\
+y &= h_g(x_g) 
+\end{align}
+in which $f_N(\cdot)$ is a feedforward neural network
+and $z(k) = [\mathbf{y}'_{k-N_p:k-1}, \mathbf{u}'_{k-N_p:k-1}]'$.
+This model is converted to discrete-time using the 
+Runge-Kutta 4 method, and the vector of past controls
+and measurements ($z$) is also viewed as a state. The discrete-time 
+model is represented as 
 \begin{align*}
-\dot x_g &= f_g(x_g, u) + 
-f_N(h_g(x_g), z, u)\\
+  x_g^+ &= f_h(x_g, z, u) \\
+  z^+ &= f_z(z, x_g, u)\\
   y &= h_g(x_g) 
 \end{align*}
-in which $f_N(\cdot)$ is a feedforward neural network
-and $z(k) = [\mathbf{y}'_{k-N_p:k-1}, \mathbf{u}'_{k-N_p:k-1}]'$.
+in which, $f_h(\cdot)$ is a obtained by discretization
+of \eqref{eq:hybrid_cont_time} using the Runge-Kutta 4 method, 
+and $f_z$ is known analytically.
 
-Alternative modeling choices are:
+A few alternative modeling choices are:
 \begin{itemize}
 	\item Black-box: $y^+ = f_N(y, z, u)$, (purely black-box)
-	\item Hybrid 1: $\dot x_g = f_N(x_g, z, u), \ y = h_g(x_g)$, 
+  \item Hybrid 1: $x_g^+ = f_N(x_g, z, u), 
+                   \ z^+ = f_z(z, x_g, u), \ y = h_g(x_g)$, 
   (states have meaning)
-  \item Hybrid 2: $\dot x_g = f_g(x_g, u), \ y = h_g(x_g) + h_N(z)$, 
+  \item Hybrid 2: $x_g^+ = f_g(x_g, u), 
+                   \ z^+ = f_z(z, x_g, u), \ y = h_g(x_g) + h_N(z)$, 
 	(model residual)
-	\item Hybrid 3: $\dot x_g = f_g(x_g, u, p(h_g(x_g), z)), \ y = h_g(x_g)$, (use NNs to learn some parameters)
+  \item Hybrid 3: $\dot x_g = f_g(x_g, u, p(h_g(x_g), z)),
+  \ z^+ = f_z(z, x_g, u), \ y = h_g(x_g)$, (use NNs to learn some parameters)
 \end{itemize}
@@
 We note that the network $f_N(\cdot)$ is an addition 
-to the grey-box dynamics in continuous time. 
+to the grey-box dynamics in continuous time in \eqref{eq:hybrid_cont_time}. 
 The Runge-Kutta-4 (RK4) method 
@@
   k_1 &= f_g(x_g, u) + 
-  f_N(h_g(x_g), \mathbf{y}_{k-N_p:k-1}, \mathbf{u}_{k+1-N_p:k}) \\
+  f_N(x_g, \mathbf{y}_{k-N_p:k-1}, \mathbf{u}_{k-N_p:k-1}, u) \\
   x_g & := x_g + (\Delta/2)k_1 \\
   k_2 &= f_g(x_g, u) + 
-  f_N(h_g(x_g), \mathbf{y}_{k+0.5-N_p:k-0.5}, \mathbf{u}_{k+1-N_p:k})\\
+  f_N(x_g, \mathbf{y}_{k+0.5-N_p:k-0.5}, \mathbf{u}_{k-N_p:k-1}, u)\\
   x_g & := x_g + (\Delta/2)k_2 \\
   k_3 &= f_g(x_g, u) + 
-  f_N(h_g(x_g), \mathbf{y}_{k+0.5-N_p:k-0.5}, \mathbf{u}_{k+1-N_p:k})\\
+  f_N(x_g, \mathbf{y}_{k+0.5-N_p:k-0.5}, \mathbf{u}_{k-N_p:k-1}, u)\\
   x_g & := x_g + \Delta k_3 \\
   k_4 &= f_g(x_g, u) + 
-  f_N(h_g(x_g), \mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k+1-N_p:k})\\
+  f_N(x_g, \mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k-N_p:k-1}, u)\\
   x_g(k+\Delta) &= x_g(k) + (\Delta/6)(k_1 + 2k_2 + 2k_3 + k_4) \\
@@
 
-\subsection{Linear plant}
+%\subsection{Linear plant}
 
-Consider the following plant, model,
-and the expected function to be approximated 
-with a neural network.
+%Consider the following plant, model,
+%and the expected function to be approximated 
+%with a neural network.
 
-\textbf{Plant --}
-\begin{align*}
-  \dfrac{dC_A}{dt} &= \dfrac{C_{Af} - C_A}{\tau} - k_1C_A\\
-  \dfrac{dC_B}{dt} &= k_1C_A - k_2C_B + k_3C_C- \dfrac{C_B}{\tau}\\
-  \dfrac{dC_C}{dt} &= k_2C_B - k_3C_C - \dfrac{C_C}{\tau}
-\end{align*}
-These ODEs represent the dynamics of the reaction system
-$A \rightarrow B \rightleftharpoons C$ in a CSTR with constant 
-height and temperature. The control input is $u = C_{Af}$,
-the states are $x = [C_A, C_B, C_C]'$, 
-and the measurements are $y = [C_A, C_B]'$. 
-Assume the following 
-parameters for this model, $k_1 = 1 \ \textnormal{m}^3/\textnormal{min}, 
-\ k_2 = 0.3 \ \textnormal{m}^3/\textnormal{min},
-\ k_3 = 0.2 \ \textnormal{m}^3/\textnormal{min}$, 
-and $\tau = 5 \ \textnormal{min}$.
-The plant model 
-is linear ($dx/dt = Ax + Bu$) and 
-observable with the choices of 
-the parameters and measurements.
+%\textbf{Plant --}
+%\begin{align*}
+%  \dfrac{dC_A}{dt} &= \dfrac{C_{Af} - C_A}{\tau} - k_1C_A\\
+%  \dfrac{dC_B}{dt} &= k_1C_A - k_2C_B + k_3C_C- \dfrac{C_B}{\tau}\\
+%  \dfrac{dC_C}{dt} &= k_2C_B - k_3C_C - \dfrac{C_C}{\tau}
+%\end{align*}
+%These ODEs represent the dynamics of the reaction system
+%$A \rightarrow B \rightleftharpoons C$ in a CSTR with constant 
+%height and temperature. The control input is $u = C_{Af}$,
+%the states are $x = [C_A, C_B, C_C]'$, 
+%and the measurements are $y = [C_A, C_B]'$. 
+%Assume the following 
+%parameters for this model, $k_1 = 1 \ \textnormal{m}^3/\textnormal{min}, 
+%\ k_2 = 0.3 \ \textnormal{m}^3/\textnormal{min},
+%\ k_3 = 0.2 \ \textnormal{m}^3/\textnormal{min}$, 
+%and $\tau = 5 \ \textnormal{min}$.
+%The plant model 
+%is linear ($dx/dt = Ax + Bu$) and 
+%observable with the choices of 
+%the parameters and measurements.
 
-\textbf{Model --}
-\begin{align*}
-  \dfrac{dC_A}{dt} &= \dfrac{C_{Af} - C_A}{\tau} - k_1C_A\\
-  \dfrac{dC_B}{dt} &= k_1C_A - \dfrac{C_B}{\tau}
-\end{align*}
-i.e, we assume that we don't know about the second reaction while 
-performing grey-box modeling.
+%\textbf{Model --}
+%\begin{align*}
+%  \dfrac{dC_A}{dt} &= \dfrac{C_{Af} - C_A}{\tau} - k_1C_A\\
+%  \dfrac{dC_B}{dt} &= k_1C_A - \dfrac{C_B}{\tau}
+%\end{align*}
+%i.e, we assume that we don't know about the second reaction while 
+%performing grey-box modeling.
 
-The current plant state $x_p(k)$ can be constructed using the history of 
-past measurements and inputs as 
-\begin{equation*}
-  x_p(k) = A^{N_p-1}_d
-  \mathcal{O}^{\dagger}(\mathbf{y}_{k+1-N_p:k} - 
-                        \mathcal{B}\mathbf{u}_{k+1-N_p:k-1}) + \\ 
-      \mathcal{B}_{nN_p-n+1:nN_p}\mathbf{u}_{k+1-N_p:k-1}
-\end{equation*}
+%The current plant state $x_p(k)$ can be constructed using the history of 
+%past measurements and inputs as 
+%\begin{equation*}
+%  x_p(k) = A^{N_p-1}_d
+%  \mathcal{O}^{\dagger}(\mathbf{y}_{k+1-N_p:k} - 
+%                        \mathcal{B}\mathbf{u}_{k+1-N_p:k-1}) + \\ 
+%      \mathcal{B}_{nN_p-n+1:nN_p}\mathbf{u}_{k+1-N_p:k-1}
+%\end{equation*}
 
-in which $A_d$ is the plant state transition matrix in discrete-time, 
-$\mathcal{O}$ is the observability matrix, 
-and $\mathcal{B}$ is a combination of the 
-discrete time matrices ($A_d$ and $B_d$) such that it 
-denotes the response of a vector of inputs to states, 
-and $\mathcal{B}_{nN_p-n+1:nN_p}$ denotes the last $n$
-rows of $\mathcal{B}$.
+%in which $A_d$ is the plant state transition matrix in discrete-time, 
+%$\mathcal{O}$ is the observability matrix, 
+%and $\mathcal{B}$ is a combination of the 
+%discrete time matrices ($A_d$ and $B_d$) such that it 
+%denotes the response of a vector of inputs to states, 
+%and $\mathcal{B}_{nN_p-n+1:nN_p}$ denotes the last $n$
+%rows of $\mathcal{B}$.
 
-Since the measurements are a simple linear function 
-of the grey-box and plant states, we use only one 
-neural network to 
-augment the grey-box dynamics. 
-The expected function 
-to be approximated with the network is
-\begin{align*}
-  f_N(\mathbf{y}_{k-N_p:k}, \mathbf{u}_{k+1-N_p:k-1}) \approx 
-  \begin{bmatrix}
-  0 & 0 & 0 \\  
-  0 & -k_2 & k_3 
-  \end{bmatrix}x_p(k)
-\end{align*}
+%Since the measurements are a simple linear function 
+%of the grey-box and plant states, we use only one 
+%neural network to 
+%augment the grey-box dynamics. 
+%The expected function 
+%to be approximated with the network is
+%\begin{align*}
+%  f_N(\mathbf{y}_{k-N_p:k}, \mathbf{u}_{k+1-N_p:k-1}) \approx 
+%  \begin{bmatrix}
+%  0 & 0 & 0 \\  
+%  0 & -k_2 & k_3 
+%  \end{bmatrix}x_p(k)
+%\end{align*}
 
-In this example, the network just approximates a linear function.
-The Figure \ref{fig:validation_linear} 
-below shows validation performances of the grey-box 
-and augmented hybrid models. The structure 
-of the network is restricted to just a matrix and
-the augmented model was trained on 240 samples (4 hours).
+%In this example, the network just approximates a linear function.
+%The Figure \ref{fig:validation_linear} 
+%below shows validation performances of the grey-box 
+%and augmented hybrid models. The structure 
+%of the network is restricted to just a matrix and
+%the augmented model was trained on 240 samples (4 hours).
 
-THe Figures \ref{fig:cost_nonlinear_5hours} and 
-\ref{fig:cost_nonlinear_5hours}
+%THe Figures \ref{fig:cost_nonlinear_5hours} and 
+%\ref{fig:cost_nonlinear_5hours}
 
-\begin{figure}[!h]
-  \centering
-  \includegraphics[page=2, width=\textwidth]{tworeac_plots_lin.pdf}
-  \caption{Model validation for the linear plant}
-  \label{fig:validation_linear}
-\end{figure}
+%\begin{figure}[!h]
+%  \centering
+%  \includegraphics[page=2, width=\textwidth]{tworeac_plots_lin.pdf}
+%  \caption{Model validation for the linear plant}
+%  \label{fig:validation_linear}
+%\end{figure}
 
-We notice that the augmented model accurately predicts the plant 
-behavior. The steady cost profile (Cost = $c_1C_{Af} - c_2C_B$)
-is shown in Figure \ref{fig:cost_linear}.
+%We notice that the augmented model accurately predicts the plant 
+%behavior. The steady cost profile (Cost = $c_1C_{Af} - c_2C_B$)
+%is shown in Figure \ref{fig:cost_linear}.
 
-\begin{figure}[!h]
-  \centering
-  \includegraphics[page=3, width=\textwidth]{tworeac_plots_lin.pdf}
-  \caption{Steady state cost curve for 
-           the linear plant, grey-box, and hybrid model.}
-  \label{fig:cost_linear}
-\end{figure}
+%\begin{figure}[!h]
+%  \centering
+%  \includegraphics[page=3, width=\textwidth]{tworeac_plots_lin.pdf}
+%  \caption{Steady state cost curve for 
+%           the linear plant, grey-box, and hybrid model.}
+%  \label{fig:cost_linear}
+%\end{figure}
 
-\subsection{Nonlinear plant}
-The plant is made nonlinear by changing the side reaction 
-to a third order reaction ($A \rightarrow B $ and  $3B \rightleftharpoons C$),
-and is now simulated with the following ODEs
+\subsection{Nonlinear reactor example}
+We consider a nonlinear reactor with 
+two reactions ($A \rightarrow B $ and  $3B \rightleftharpoons C$),
+simulated with the following ODEs
 \begin{align*}

### 2020-11-24 systemID_NNs.tex
ok.

@@
 \documentclass{article}
-\usepackage[total={6in, 8in}]{geometry}
-%% The graphicx and color packages should give you all the color,
-%% figure inclusion, and text/figure resizing capabilities you need.
+\usepackage[total={4.5in, 8in}]{geometry}
+%% The graphicx and color packages should give you all the color, % figure
+%inclusion, and text/figure resizing capabilities you need.
 \usepackage{graphicx}
@@
 \usepackage{latexsym}  % A few extra LaTeX symbols
-
+\usepackage{cleveref}
 \usepackage{lucidbry}
@@
 
-%% The URL package is handy for typesetting URLs.  It does not define
-%% an \email command because so many document styles already do that.
-%% So we define one here that uses a typewriter font.
+%% The URL package is handy for typesetting URLs.  It does not define % an
+%\email command because so many document styles already do that. % So we define
+%one here that uses a typewriter font.
 \usepackage{url}
@@
 
-%% This provides various customized verbatim commands and
-%% environments.  You probably don't need it.
+%% This provides various customized verbatim commands and % environments.  You
+%probably don't need it.
 \usepackage{fancyvrb}
-\DefineShortVerb{\|}
-\VerbatimFootnotes
+\DefineShortVerb{\|} \VerbatimFootnotes
 \DefineVerbatimEnvironment{code}{Verbatim}{%
-  frame=single,
-  framesep=1em,
-  xleftmargin=1em,
-  xrightmargin=1em,
-  samepage=true,
+  frame=single, framesep=1em, xleftmargin=1em, xrightmargin=1em, samepage=true,
   fontsize=\footnotesize}
@@
 
-%% Your document may require a different bibliography style.  I've
-%% come to prefer Author (YYYY) styles because it makes it easier for
-%% a person who knows something about the literature to understand
-%% what is being cited without having to skip to the list of
-%% references.
+%% Your document may require a different bibliography style.  I've % come to
+%prefer Author (YYYY) styles because it makes it easier for % a person who knows
+%something about the literature to understand % what is being cited without
+%having to skip to the list of % references.
 \usepackage[authoryear,round,longnamesfirst]{natbib}
 
-%% You won't normally need this definition in your documents, but it
-%% is here so we can typeset the BibTeX logo correctly.
+%% You won't normally need this definition in your documents, but it % is here
+%so we can typeset the BibTeX logo correctly.
 \makeatletter
-\@ifundefined{BibTeX}
-   {\def\BibTeX{{\rmfamily B\kern-.05em%
+\@ifundefined{BibTeX} {\def\BibTeX{{\rmfamily B\kern-.05em%
     \textsc{i\kern-.025em b}\kern-.08em%
@@
 
-%% Allow more of the pages to be occupied by graphs.  These parameters
-%% are described in section c.9.1 of the LaTeXbook.  Your document may
-%% not benefit from these parameters (it could make things worse) so
-%% you should only change these from the defaults if you need to.
+%% Allow more of the pages to be occupied by graphs.  These parameters % are
+%described in section c.9.1 of the LaTeXbook.  Your document may % not benefit
+%from these parameters (it could make things worse) so % you should only change
+%these from the defaults if you need to.
 \setcounter{topnumber}{2}              %% 2
@@
 
-\newcommand{\useq}{\mathbf{u}}
-\newcommand{\xseq}{\mathbf{x}}
-\newcommand{\bbR}{\mathbb{R}}
-\newcommand{\bbW}{\mathbb{W}}
-\newcommand{\bbU}{\mathbb{U}}
-\newcommand{\bbI}{\mathbb{I}}
-\newcommand{\bbX}{\mathbb{X}}
-\newcommand{\bbP}{\mathbb{P}}
+\newcommand{\useq}{\mathbf{u}} \newcommand{\xseq}{\mathbf{x}}
+\newcommand{\bbR}{\mathbb{R}} \newcommand{\bbW}{\mathbb{W}}
+\newcommand{\bbU}{\mathbb{U}} \newcommand{\bbI}{\mathbb{I}}
+\newcommand{\bbX}{\mathbb{X}} \newcommand{\bbP}{\mathbb{P}}
 
@@
 \end{align*}
-in which $x_p \in \bbR^{n_p}$ is the plant state, 
-$u \in \bbU$ is the manipulated control input,
-$y \in \bbR^p$ is the measurement, $w$ is the process noise, 
-and $v$ is the measurement noise.
+in which $x_p \in \bbR^{n_p}$ is the plant state, $u \in \bbU$ is the
+manipulated control input, $y \in \bbR^p$ is the measurement, $w$ is the process
+noise, and $v$ is the measurement noise.
 
 \subsection{Completion of the u-y model}
-We assume that grey-box modeling is performed 
-for the plant, and we wish to augment the grey-box model with neural networks
-for improvement in the accuracy of the model. From first-principles modeling, 
-we have the following grey-box model in continuous time
+We assume that grey-box modeling is performed for the plant, and we wish to
+augment the grey-box model with neural networks for improvement in the accuracy
+of the model. From first-principles modeling, we have the following grey-box
+model in continuous time
 \begin{align*}
@@
 
-in which $x_g \in \bbR^{n_g}$ are the grey-box states, 
-which are some subset 
-of the plant states (denote $x_g = Gx_p$) chosen to 
-be modeled. The dynamics of these grey-box states 
-potentially have some missing terms 
-in their ODEs.
-For observable nonlinear systems, 
-we can construct the original plant state
-using a history of measurements and control inputs.
-Denote this history as 
+in which $x_g \in \bbR^{n_g}$ are the grey-box states, which are some subset of
+the plant states (denote $x_g = Gx_p$) chosen to be modeled. The dynamics of
+these grey-box states potentially have some missing terms in their ODEs. For
+observable nonlinear systems, we can construct the original plant state using a
+history of measurements and control inputs. Denote this history as 
 \begin{align*}
@@
 
-in which $N_p$ is the number of 
-past measurements and
-$\Delta$ is the sample time between two consecutive
-measurements. There exists a $\phi(\cdot)$ 
-and a number of delay 
-time steps $N_p$, such that the plant state
-at a given time can be written as
-$x_p(k) = \phi(\mathbf{y}_{k-N_p:k}, \mathbf{u}_{k-N_p:k-1})$.
-For linear plants, this function $\phi(\cdot)$
-is known analytically in terms of the plant dynamic
-matrices.
+in which $N_p$ is the number of past measurements and $\Delta$ is the sample
+time between two consecutive measurements. There exists a $\phi(\cdot)$ and a
+number of delay time steps $N_p$, such that the plant state at a given time can
+be written as $x_p(k) = \phi(\mathbf{y}_{k-N_p:k}, \mathbf{u}_{k-N_p:k-1})$. For
+linear plants, this function $\phi(\cdot)$ is known analytically in terms of the
+plant dynamic matrices.
 
-We augment the grey-box model using
-neural networks as follows -
+We augment the grey-box model using neural networks as follows -
 \begin{align} \label{eq:hybrid_cont_time}
@@
 \end{align}
-in which $f_N(\cdot)$ is a feedforward neural network
-and $z(k) = [\mathbf{y}'_{k-N_p:k-1}, \mathbf{u}'_{k-N_p:k-1}]'$.
-This model is converted to discrete-time using the 
-Runge-Kutta 4 method, and the vector of past controls
-and measurements ($z$) is also viewed as a state. The discrete-time 
-model is represented as 
-\begin{align*}
+in which $f_N(\cdot)$ is a feedforward neural network and $z(k) =
+[\mathbf{y}'_{k-N_p:k-1}, \mathbf{u}'_{k-N_p:k-1}]'$. This model is converted to
+discrete-time using the Runge-Kutta 4 method, and the vector of past controls
+and measurements ($z$) is also viewed as a state. The discrete-time model is
+represented as 
+\begin{equation} \label{eq:hybrid_discrete_time}
+  \begin{split}
     x_g^+ &= f_h(x_g, z, u)  \\
-  z^+ &= f_z(z, x_g, u)\\
+    z^+ &= f_z(z, x_g, u) \\
     y &= h_g(x_g)       
-\end{align*}
-in which, $f_h(\cdot)$ is a obtained by discretization
-of \eqref{eq:hybrid_cont_time} using the Runge-Kutta 4 method, 
-and $f_z$ is known analytically.
+  \end{split}
+\end{equation}
+in which, $f_h(\cdot)$ is a obtained by discretization of
+\eqref{eq:hybrid_cont_time} using the Runge-Kutta 4 method, and $f_z$ is known
+analytically.
 
@@
 	\item Black-box: $y^+ = f_N(y, z, u)$, (purely black-box)
-  \item Hybrid 1: $x_g^+ = f_N(x_g, z, u), 
-                   \ z^+ = f_z(z, x_g, u), \ y = h_g(x_g)$, 
-  (states have meaning)
-  \item Hybrid 2: $x_g^+ = f_g(x_g, u), 
-                   \ z^+ = f_z(z, x_g, u), \ y = h_g(x_g) + h_N(z)$, 
-	(model residual)
-  \item Hybrid 3: $\dot x_g = f_g(x_g, u, p(h_g(x_g), z)),
-  \ z^+ = f_z(z, x_g, u), \ y = h_g(x_g)$, (use NNs to learn some parameters)
+  \item Hybrid 1: $x_g^+ = f_N(x_g, z, u), \ z^+ = f_z(z, x_g, u), \ y =
+                   h_g(x_g)$, (states have meaning)
+  \item Hybrid 2: $x_g^+ = f_g(x_g, u), \ z^+ = f_z(z, x_g, u), \ y = h_g(x_g) +
+                   h_N(z)$, (model residual)
+  \item Hybrid 3: $\dot x_g = f_g(x_g, u, p(h_g(x_g), z)), \ z^+ = f_z(z, x_g,
+  u), \ y = h_g(x_g)$, (use NNs to learn some parameters)
 \end{itemize}
 
-%It is expected that the functions $f_{NN}(\cdot)$ 
-%and $h_{NN}(\cdot)$ approximate
-%the following differences between the true grey-box state dynamics
-%and true measurement function with what is modeled.
-%\begin{align*}
-%  f_{NN}(\mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k+1-N_p:k}) &\approx 
-%      Gf_p(x, u) - f_g(Gx, u) \\
-%      h_{NN}(\mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k+1-N_p:k-1}) &\approx
-%      h_p(x) - h_g(Gx) 
-%\end{align*}  
-%in which $x(k) = \phi(\mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k+1-N_p:k-1})$
-%is an unknown function of past measurements and inputs.
-
 \subsection{Conversion to discrete time and training}
-We note that the network $f_N(\cdot)$ is an addition 
-to the grey-box dynamics in continuous time in \eqref{eq:hybrid_cont_time}. 
-The Runge-Kutta-4 (RK4) method 
+We note that the network $f_N(\cdot)$ is an addition to the grey-box dynamics in
+continuous time in \eqref{eq:hybrid_cont_time}. The Runge-Kutta-4 (RK4) method
 is used for the discretization as follows
@@
 \end{align*}
-in which the sequence $\mathbf{y}_{k+0.5-N_p:k-0.5}$ is obtained by 
-linear interpolation.
+in which the sequence $\mathbf{y}_{k+0.5-N_p:k-0.5}$ is obtained by linear
+interpolation.
 
-The following optimization problem is solved to compute 
-the weights in the neural network
+The following optimization problem is solved to compute the weights in the
+neural network
 \begin{align*}
@@
  \end{align*}
-in which $W_i$, $b_i$ are the weights in each hidden layer of the NN, 
-$N_s$ is the number of training samples, 
-$\hat{y}$ are the predictions of the augmented
-grey-box + neural network model, and $y$
-are the measurements from the plant.
-Note that this optimization requires an estimate 
-of the initial grey-box state. We assume that 
-before the open-loop system identification experiment 
-performed for data collection, 
-the plant is being operated at some 
-steady-state such that input to output
-responses of the plant and grey-box model are aligned with 
-the use of an integrating disturbance model. We can use 
-the grey-box state at that steady state as the initial condition 
-for the above multi-step-ahead prediction-error
-optimization.
-
-%\subsection{Linear plant}
-
-%Consider the following plant, model,
-%and the expected function to be approximated 
-%with a neural network.
-
-%\textbf{Plant --}
-%\begin{align*}
-%  \dfrac{dC_A}{dt} &= \dfrac{C_{Af} - C_A}{\tau} - k_1C_A\\
-%  \dfrac{dC_B}{dt} &= k_1C_A - k_2C_B + k_3C_C- \dfrac{C_B}{\tau}\\
-%  \dfrac{dC_C}{dt} &= k_2C_B - k_3C_C - \dfrac{C_C}{\tau}
-%\end{align*}
-%These ODEs represent the dynamics of the reaction system
-%$A \rightarrow B \rightleftharpoons C$ in a CSTR with constant 
-%height and temperature. The control input is $u = C_{Af}$,
-%the states are $x = [C_A, C_B, C_C]'$, 
-%and the measurements are $y = [C_A, C_B]'$. 
-%Assume the following 
-%parameters for this model, $k_1 = 1 \ \textnormal{m}^3/\textnormal{min}, 
-%\ k_2 = 0.3 \ \textnormal{m}^3/\textnormal{min},
-%\ k_3 = 0.2 \ \textnormal{m}^3/\textnormal{min}$, 
-%and $\tau = 5 \ \textnormal{min}$.
-%The plant model 
-%is linear ($dx/dt = Ax + Bu$) and 
-%observable with the choices of 
-%the parameters and measurements.
-
-%\textbf{Model --}
-%\begin{align*}
-%  \dfrac{dC_A}{dt} &= \dfrac{C_{Af} - C_A}{\tau} - k_1C_A\\
-%  \dfrac{dC_B}{dt} &= k_1C_A - \dfrac{C_B}{\tau}
-%\end{align*}
-%i.e, we assume that we don't know about the second reaction while 
-%performing grey-box modeling.
-
-%The current plant state $x_p(k)$ can be constructed using the history of 
-%past measurements and inputs as 
-%\begin{equation*}
-%  x_p(k) = A^{N_p-1}_d
-%  \mathcal{O}^{\dagger}(\mathbf{y}_{k+1-N_p:k} - 
-%                        \mathcal{B}\mathbf{u}_{k+1-N_p:k-1}) + \\ 
-%      \mathcal{B}_{nN_p-n+1:nN_p}\mathbf{u}_{k+1-N_p:k-1}
-%\end{equation*}
-
-%in which $A_d$ is the plant state transition matrix in discrete-time, 
-%$\mathcal{O}$ is the observability matrix, 
-%and $\mathcal{B}$ is a combination of the 
-%discrete time matrices ($A_d$ and $B_d$) such that it 
-%denotes the response of a vector of inputs to states, 
-%and $\mathcal{B}_{nN_p-n+1:nN_p}$ denotes the last $n$
-%rows of $\mathcal{B}$.
-
-%Since the measurements are a simple linear function 
-%of the grey-box and plant states, we use only one 
-%neural network to 
-%augment the grey-box dynamics. 
-%The expected function 
-%to be approximated with the network is
-%\begin{align*}
-%  f_N(\mathbf{y}_{k-N_p:k}, \mathbf{u}_{k+1-N_p:k-1}) \approx 
-%  \begin{bmatrix}
-%  0 & 0 & 0 \\  
-%  0 & -k_2 & k_3 
-%  \end{bmatrix}x_p(k)
-%\end{align*}
-
-%In this example, the network just approximates a linear function.
-%The Figure \ref{fig:validation_linear} 
-%below shows validation performances of the grey-box 
-%and augmented hybrid models. The structure 
-%of the network is restricted to just a matrix and
-%the augmented model was trained on 240 samples (4 hours).
-
-%THe Figures \ref{fig:cost_nonlinear_5hours} and 
-%\ref{fig:cost_nonlinear_5hours}
-
-%\begin{figure}[!h]
-%  \centering
-%  \includegraphics[page=2, width=\textwidth]{tworeac_plots_lin.pdf}
-%  \caption{Model validation for the linear plant}
-%  \label{fig:validation_linear}
-%\end{figure}
-
-%We notice that the augmented model accurately predicts the plant 
-%behavior. The steady cost profile (Cost = $c_1C_{Af} - c_2C_B$)
-%is shown in Figure \ref{fig:cost_linear}.
-
-%\begin{figure}[!h]
-%  \centering
-%  \includegraphics[page=3, width=\textwidth]{tworeac_plots_lin.pdf}
-%  \caption{Steady state cost curve for 
-%           the linear plant, grey-box, and hybrid model.}
-%  \label{fig:cost_linear}
-%\end{figure}
+in which $W_i$, $b_i$ are the weights in each hidden layer of the NN, $N_s$ is
+the number of training samples, $\hat{y}$ are the predictions of the augmented
+grey-box + neural network model, and $y$ are the measurements from the plant.
+Note that this optimization requires an estimate of the initial grey-box state.
+We assume that before the open-loop system identification experiment performed
+for data collection, the plant is being operated at some steady-state such that
+input to output responses of the plant and grey-box model are aligned with the
+use of an integrating disturbance model. We can use the grey-box state at that
+steady state as the initial condition for the above multi-step-ahead
+prediction-error optimization.
 
 \subsection{Nonlinear reactor example}
-We consider a nonlinear reactor with 
-two reactions ($A \rightarrow B $ and  $3B \rightleftharpoons C$),
-simulated with the following ODEs
+We consider a nonlinear reactor with two reactions ($A \rightarrow B $ and  $3B
+\rightleftharpoons C$), simulated using the following ordinary differential
+equations (ODEs)
 \begin{align*}
@@
 
-The grey-box model is kept the same. The parameters 
-in the model are modified to 
-$k_1 = 1 \ \textnormal{m}^3/\textnormal{min}, 
-\ k_2 = 0.01 \ \textnormal{m}^3/\textnormal{min},
-\ k_3 = 0.05 \ \textnormal{m}^3/\textnormal{min}$, 
-and $\tau = 5 \ \textnormal{min}$.
+The grey-box model is chosen as follows with an assumption that knowledge of the
+second side reaction ($3B \rightleftharpoons C $) is not available
+\begin{align*}
+  \dfrac{dC_A}{dt} &= \dfrac{C_{Af} - C_A}{\tau} - k_1C_A\\
+  \dfrac{dC_B}{dt} &= k_1C_A - \dfrac{C_B}{\tau}
+\end{align*}
+The control input is $u = C_{Af}$, the states in the plant model are $x = [C_A,
+C_B, C_C]'$, and the measurements are $y = [C_A, C_B]'$. We assume the following
+parameters values, $k_1 = 1 \ \textnormal{m}^3/\textnormal{min}, \ k_2 = 0.01 \
+\textnormal{m}^3/\textnormal{min}, \ k_3 = 0.05 \
+\textnormal{m}^3/\textnormal{min}$, and $\tau = 5 \ \textnormal{min}$.
 
-\newpage
-Three models are compared: black-box, residual, and hybrid model with additive 
-NN to the grey-box dynamics. The chosen model architectures 
-are [9, 16, 16, 2], [9, 16, 16, 2], and [8, 16, 2] respectively.
-The Figure \ref{fig:validation_nonlinear} shows model validation on open loop 
-data after training with 10 hours of data.
-
-The Figures \ref{fig:cost_nonlinear_5hours} and \ref{fig:cost_nonlinear_10hours}
-show the steady state cost curve after training with 5 and 10 hours 
-of data respectively. We notice that the hybrid model is more data efficient 
-compared to the alternatives.
+We examine the performance of the hybrid modeling framework
+\eqref{eq:hybrid_discrete_time} in simulation studies. First, we generate
+training data by simulating the plant model using a pseudo-random-binary-signal
+(PRBS) manipulated input profile. Second, we train a black-box neural network
+and the hybrid model \eqref{eq:hybrid_discrete_time} with network architectures
+of [9, 16, 2] and [8, 16, 2]. Third, we examine trained model predictions and a
+mean squared error (MSE) performance metric on validation data generated using
+the plant model. Finally, we analyze the steady state cost curves obtained using
+the trained models and their differences with the steady state cost 
+curve of the plant model. These
+simulations and analysis shed light on the data requirements of the hybrid and black-box
+modeling frameworks, and enable an understanding of their performance if used
+for real-time economic optimization in chemical plants.
 
@@
 
-%We notice that the augmented model accurately predicts the plant 
-%behavior. The steady cost profile (Cost = $c_1C_{A0} - c_2C_B$)
-%is shown in Figure \ref{fig:cost_nonlinear}. Upon optimization 
-%using casadi, the optimum cost and control for the plant are: $-19.62 \$$ 
-%and 1.13 $\textnormal{mol}/\textnormal{m}^3$, and for the hybrid 
-%grey-box and NN model are: $-19.91 \$$ 
-%and 1.11 $\textnormal{mol}/\textnormal{m}^3$.
-
 \begin{figure}[!h]
@@
   \includegraphics[page=3, width=\textwidth]{tworeac_plots_nonlin.pdf}
-  \caption{Steady state cost curve for the nonlinear plant, grey-box, 
-  black-box, residual, and hybrid model after training with 5 hours of data.}
+  \caption{Steady state cost curve for the nonlinear plant, grey-box, black-box,
+  residual, and hybrid model after training with 5 hours of data.}
   \label{fig:cost_nonlinear_5hours}
@@
   \includegraphics[page=4, width=\textwidth]{tworeac_plots_nonlin.pdf}
-  \caption{Steady state cost curve for the nonlinear plant, grey-box, 
-  black-box, residual, and hybrid model after training with 10 hours of data.}
+  \caption{Steady state cost curve for the nonlinear plant, grey-box, black-box,
+  residual, and hybrid model after training with 10 hours of data.}
   \label{fig:cost_nonlinear_10hours}
@@
 
-\begin{figure}[!h]
-  \centering
-  \includegraphics[page=1, width=\textwidth]{tworeac_distmodels_nonlin.pdf}
-  \caption{Steady state grey-box states and disturbance estimates based on 
-  output and input disturbance models}
-  \label{fig:cost_nonlinear_10hours}
-\end{figure}
-
-%\textbf{Need to discuss the following issues:}
-%\begin{itemize}
-%  \item What are the example problems/applications 
-%  where both kind of modeling 
-%  strategies make sense, and where should we start.
-%  Reproduce results from the Kevrekidis paper with measurement 
-%  noise?
-%  \item The model training step will be the optimization of 
-%  the following multi-step-ahead prediction-error
-%  \begin{equation*}
-%    \underset{W_i, b_i}{\textnormal{min}} \sum_{k=0}^{t-1} (\hat{y}(k) - y(k))^2
-%  \end{equation*}
-%  in which ($W_i, b_i$) are the weights of the NN. This optimization 
-%  is for one ID experiment, and requires an initial grey-box 
-%  state $x_g(0)$ to make the forecast. If we have data 
-%  for multiple ID experiments, we'll need initial states for 
-%  all the ID experiments. In theory, we can add the initial 
-%  states as decision variables in the optimization, but 
-%  unfortunately tensorflow 
-%  does not has a mechanism to do this.
-%\end{itemize}
-
-%\subsection{Adding feedforward disturbance forecasts}
-%In this class of problems, we assume that we have a pretty 
-%good u-y model from first-principles knowledge, and we'll use 
-%neural networks to make feedforward disturbance forecasts
-%for use in online MPC. HVAC is a pretty good application.
-
-%\section{Optimization in MPC with neural networks as model}
-%This is probably a late fall quarter or a winter break project. \\
-%a) Try Casadi/MPC tools. 
-%If tanh is the activation function (optimization is smooth),
-%this should be fine. \\
-%b) Try a global optimization solver. \\
-%c) Train another NN with solutions from a global optimization solver.
+%\begin{figure}[!h] \centering \includegraphics[page=1,
+%  width=\textwidth]{tworeac_distmodels_nonlin.pdf} \caption{Steady state
+%  grey-box states and disturbance estimates based on output and input
+%  disturbance models} \label{fig:cost_nonlinear_10hours} \end{figure}
 

### 2020-11-24 systemID_NNs.tex
making progress in writing.

@@
 \underset{W_i, b_i}{\textnormal{min}} \sum_{k=0}^{N_s-1} 
-\dfrac{1}{2}(\hat{y}(k) - y(k))^2
+\dfrac{1}{N_s}(\hat{y}(k) - y(k))^2
  \end{align*}
@@
 the number of training samples, $\hat{y}$ are the predictions of the augmented
-grey-box + neural network model, and $y$ are the measurements from the plant.
-Note that this optimization requires an estimate of the initial grey-box state.
-We assume that before the open-loop system identification experiment performed
-for data collection, the plant is being operated at some steady-state such that
-input to output responses of the plant and grey-box model are aligned with the
-use of an integrating disturbance model. We can use the grey-box state at that
-steady state as the initial condition for the above multi-step-ahead
-prediction-error optimization.
+grey-box + neural network model, $N_s$ is the number of training samples, and
+$y$ are the measurements from the plant. Note that this optimization requires an
+estimate of the initial grey-box state. We assume that before the open-loop
+system identification experiment performed for data collection, the plant is
+being operated at some steady-state such that input to output responses of the
+plant and grey-box model are aligned with the use of an integrating disturbance
+model. We can use the grey-box state at that steady state as the initial
+condition for the above multi-step-ahead prediction-error optimization.
 
@@
 and the hybrid model \eqref{eq:hybrid_discrete_time} with network architectures
-of [9, 16, 2] and [8, 16, 2]. Third, we examine trained model predictions and a
-mean squared error (MSE) performance metric on validation data generated using
-the plant model. Finally, we analyze the steady state cost curves obtained using
-the trained models and their differences with the steady state cost 
-curve of the plant model. These
-simulations and analysis shed light on the data requirements of the hybrid and black-box
-modeling frameworks, and enable an understanding of their performance if used
-for real-time economic optimization in chemical plants.
+of [9, 16, 2] and [8, 16, 2]. Third, we examine the trained model predictions
+and a mean squared error (MSE) performance metric on validation data generated
+using the plant model. Finally, we analyze the steady-state cost curves obtained
+using the trained models and their differences with the steady-state cost curve
+of the plant model. These simulations and analyses shed light on the data
+requirements of the hybrid and black-box modeling frameworks, and enable an
+understanding of their performance if used for real-time economic optimization
+in chemical plants.
+
+The Figure \ref{fig:validation_pred} shows predictions of the trained black-box
+and hybrid models after training with 8 hours (480 samples) of data. We observe
+that both the models can match the plant data accurately. The black-box model
+shows slightly worse predictions in a few places. However, both the trained
+models are adequate for real-time optimization in a feedback control system.
 
@@
   \includegraphics[page=2, width=\textwidth]{tworeac_plots_nonlin.pdf}
-  \caption{Model validation for the non linear plant}
-  \label{fig:validation_nonlinear}
+  \caption{Model validation for the nonlinear reactor model. Both the black-box
+           and hybrid model can predict the plant responses accurately. The
+           model predictions shown in this plot correspond to training with 8
+           hours of data.}
+  \label{fig:validation_pred}
 \end{figure}
+
+We examine the prediction accuracy of the models as a function of the number of
+training samples using the mean squared error (MSE) metric computed as follows
+on the validation data
+\begin{align*}
+  \textnormal{MSE} =  \sum_{k=0}^{N_s-1} 
+  \dfrac{1}{N_s}(\hat{y}(k) - y(k))^2
+\end{align*}
+in which $\hat{y}$ are the model predictions and $y$ denotes the plant
+measurements. The Figure \ref{fig:validation_mse} shows the variation of this
+MSE metric as a function of the number of training data obtained with both the
+black-box and hybrid models. We observe that the hybrid model provides better
+performance than the black-box model in terms of overall prediction accuracy and
+data requirements.
 
@@
   \centering
-  \includegraphics[page=3, width=\textwidth]{tworeac_plots_nonlin.pdf}
-  \caption{Steady state cost curve for the nonlinear plant, grey-box, black-box,
+  \includegraphics[page=9, width=\textwidth]{tworeac_plots_nonlin.pdf}
+  \caption{Mean squared error (MSE) performance metric obtained with the
+           black-box and hybrid model with increase in the number of training
+           samples. We notice that the hybrid model slightly outperforms the
+           black-box model in data requirements.}
+  \label{fig:validation_mse}
+\end{figure}
+
+Next, we study the performance of the trained models in a steady-state economic
+optimization problem, and compare the cost curves obtained with the trained
+models with the cost curve of the plant model. The motivation is to understand
+the performance of these models if used in real-time economic optimization
+layers common in chemical process industries. The optimization problem solved is
+
+\begin{align*}
+  \underset{u_s}{\textnormal{min}} & \quad \ell(y_s, u_s) \\
+  x_s &= f(x_s, u_s)\\ 
+  y_s &= h(x_s)
+\end{align*}
+in which $x_s$, $y_s$, $u_s$ are the state, output, and control respectively at
+steady-states. For the nonlinear reactor example, the cost is chosen as
+$\ell(y_s, u_s) = c_1C_{Af} - c_2C_B$, which is a measure of the raw material
+cost and product profit. The Figure  
+\ref{fig:cost_curves}
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[page=6, width=\textwidth]{tworeac_plots_nonlin.pdf}
+  \caption{Steady-state cost curve for the nonlinear plant, grey-box, black-box,
   residual, and hybrid model after training with 5 hours of data.}
-  \label{fig:cost_nonlinear_5hours}
+  \label{fig:cost_curves}
 \end{figure}
@@
   \includegraphics[page=4, width=\textwidth]{tworeac_plots_nonlin.pdf}
-  \caption{Steady state cost curve for the nonlinear plant, grey-box, black-box,
+  \caption{Steady-state cost curve for the nonlinear plant, grey-box, black-box,
   residual, and hybrid model after training with 10 hours of data.}
@@
 
+\section{Conclusion}
+Hey.
 %\begin{figure}[!h] \centering \includegraphics[page=1,

### 2020-11-24 systemID_NNs.tex
ok, add things.

@@
 steady-states. For the nonlinear reactor example, the cost is chosen as
-$\ell(y_s, u_s) = c_1C_{Af} - c_2C_B$, which is a measure of the raw material
-cost and product profit. The Figure  
-\ref{fig:cost_curves}
+$\ell(y_s, u_s) = c_1C_{Af} - c_2C_B$, which measures the raw material
+cost and product profit. The Figure \ref{fig:cost_curves} shows the 
+cost curves of the plant, grey-box, black-box, and hybrid model 
+obtained with 8 hours of training data.
 
@@
   \caption{Steady-state cost curve for the nonlinear plant, grey-box, black-box,
-  residual, and hybrid model after training with 5 hours of data.}
+  and hybrid model after training with 8 hours of data.}
   \label{fig:cost_curves}

### 2020-11-24 systemID_NNs.tex
sysID.

@@
 steady-states. For the nonlinear reactor example, the cost is chosen as
-$\ell(y_s, u_s) = c_1C_{Af} - c_2C_B$, which measures the raw material
-cost and product profit. The Figure \ref{fig:cost_curves} shows the 
-cost curves of the plant, grey-box, black-box, and hybrid model 
-obtained with 8 hours of training data.
+$\ell(y_s, u_s) = c_1C_{Af} - c_2C_B$, which measures the raw material cost and
+product profit. The Figure \ref{fig:cost_curves} shows the cost curves of the
+plant, grey-box, black-box, and hybrid model obtained with 8 hours of training
+data. We first notice from the cost curve of the grey-box model that an
+optimization with the grey-box model results in the optimizer recommending a
+full supply of the raw material $A$. Since some of the product $B$ reacts to
+form an undesired side product $C$ in the plant, this recommendation incurres a
+cost as observed from the plant cost curve. An intuitive approach to correct
+this type of plant-model mismatch is to augment the grey-box model with
+integrating disturbance models and estimate the plant model mismatch from data
+at one steady state. This strategy however, cannot recover the plant cost curve
+in the entire actuator constraint set as the integrating disturbance models are
+often linear. We notice that the cost curve of the hybrid model is almost the
+same as the cost curve of the plant model in the entire actuator constraint set.
+The cost curve of the black-box model is also shown and we notice that it is not
+as closely aligned with the plant as compared to the hybrid model.
 
@@
   \centering
-  \includegraphics[page=4, width=\textwidth]{tworeac_plots_nonlin.pdf}
+  \includegraphics[page=7, width=\textwidth]{tworeac_plots_nonlin.pdf}
   \caption{Steady-state cost curve for the nonlinear plant, grey-box, black-box,
   residual, and hybrid model after training with 10 hours of data.}
-  \label{fig:cost_nonlinear_10hours}
+  \label{fig:cost_error_bb}
 \end{figure}
 
-\section{Conclusion}
-Hey.
-%\begin{figure}[!h] \centering \includegraphics[page=1,
-%  width=\textwidth]{tworeac_distmodels_nonlin.pdf} \caption{Steady state
-%  grey-box states and disturbance estimates based on output and input
-%  disturbance models} \label{fig:cost_nonlinear_10hours} \end{figure}
+Finally, we analyze the variation in the cost curves of the black-box and hybrid
+models with increasing number of training samples. We analyze the following cost
+error metric in the entire actuator constraint set
+\begin{align*}
+  e(u_s) = \vert \ell(y_p(u_s), u_s) - \ell(y_m(u_s), u_s) \vert
+\end{align*}
+in which $y_p(\cdot)$ and $y_m(\cdot)$ denote the plant and model steady state
+output respectively. The Figures \ref{fig:cost_error_bb} and
+\ref{fig:cost_error_hybrid} show the variation in this cost error metric in the
+actuator constraint set with increasing number of training samples for the
+black-box and hybrid model respectively. We observe that with an increase in the
+number of training samples, the error metric for the hybrid model approaches
+near zero values for a wide range of actuator values in the constraint set. The
+error metric of the black-box model remains relatively larger than that of the
+hybrid model in the constraint set with increase in training data. The
+implication of this error metric going to zero for all actuator values in the
+constraint set is that the optimum computed for the steady-state economic
+optimization problem with the trained model will be the same as the true plant
+optimum.
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[page=8, width=\textwidth]{tworeac_plots_nonlin.pdf}
+  \caption{Steady-state cost curve for the nonlinear plant, grey-box, black-box,
+  residual, and hybrid model after training with 10 hours of data.}
+  \label{fig:cost_error_hybrid}
+\end{figure}
 

### 2020-11-24 systemID_NNs.tex
edits.:

@@
 in which $N_p$ is the number of past measurements and $\Delta$ is the sample
-time between two consecutive measurements. There exists a $\phi(\cdot)$ and a
-number of delay time steps $N_p$, such that the plant state at a given time can
-be written as $x_p(k) = \phi(\mathbf{y}_{k-N_p:k}, \mathbf{u}_{k-N_p:k-1})$. For
-linear plants, this function $\phi(\cdot)$ is known analytically in terms of the
-plant dynamic matrices.
+time between two consecutive measurements. There exists a function $\phi(\cdot)$
+and a number of delay time steps $N_p$, such that the plant state at a given
+time can be written as $x_p(k) = \phi(\mathbf{y}_{k-N_p:k},
+\mathbf{u}_{k-N_p:k-1})$. For linear plants, this function $\phi(\cdot)$ is
+known analytically in terms of the plant dynamic matrices.
 
@@
 \begin{align*}
+  x_{gk} &= x_g(k) \\
   k_1 &= f_g(x_g, u) + 
@@
   f_N(x_g, \mathbf{y}_{k+1-N_p:k}, \mathbf{u}_{k-N_p:k-1}, u)\\
-  x_g(k+\Delta) &= x_g(k) + (\Delta/6)(k_1 + 2k_2 + 2k_3 + k_4) \\
+  x_g(k+\Delta) &= x_{gk} + (\Delta/6)(k_1 + 2k_2 + 2k_3 + k_4) \\
   \hat{y}(k+\Delta) &= h_g(x_g(k+\Delta)) 
@@
 \end{align*}
-in which $W_i$, $b_i$ are the weights in each hidden layer of the NN, $N_s$ is
-the number of training samples, $\hat{y}$ are the predictions of the augmented
-grey-box + neural network model, $N_s$ is the number of training samples, and
-$y$ are the measurements from the plant. Note that this optimization requires an
-estimate of the initial grey-box state. We assume that before the open-loop
-system identification experiment performed for data collection, the plant is
-being operated at some steady-state such that input to output responses of the
-plant and grey-box model are aligned with the use of an integrating disturbance
-model. We can use the grey-box state at that steady state as the initial
-condition for the above multi-step-ahead prediction-error optimization.
+in which $W_i$, $b_i$ are the weights in each hidden layer of the network, $N_s$
+is the number of training samples, $\hat{y}$ are the predictions of the
+hybrid grey-box + neural network model, $N_s$ is the number of training
+samples, and $y$ are the measurements from the plant. Note that this
+optimization requires an estimate of the initial grey-box state. We assume that
+before the open-loop system identification experiment is performed for data
+collection, the plant is being operated at some steady-state such that input to
+output responses of the plant and grey-box model can be aligned with the use of an
+integrating disturbance model. We can use the grey-box state at that steady
+state as the initial condition for the above multi-step-ahead prediction-error
+optimization.
 
 \subsection{Nonlinear reactor example}
-We consider a nonlinear reactor with two reactions ($A \rightarrow B $ and  $3B
+We now demonstrate the application of the above hybrid modeling framework
+via a simulation study on a simple chemical process example. We consider a nonlinear reactor with two reactions ($A \rightarrow B $ and  $3B
 \rightleftharpoons C$), simulated using the following ordinary differential
@@
 C_B, C_C]'$, and the measurements are $y = [C_A, C_B]'$. We assume the following
-parameters values, $k_1 = 1 \ \textnormal{m}^3/\textnormal{min}, \ k_2 = 0.01 \
+parameter values, $k_1 = 1 \ \textnormal{m}^3/\textnormal{min}, \ k_2 = 0.01 \
 \textnormal{m}^3/\textnormal{min}, \ k_3 = 0.05 \
@@
 training data by simulating the plant model using a pseudo-random-binary-signal
-(PRBS) manipulated input profile. Second, we train a black-box neural network
+(PRBS) of the manipulated control input. Second, we train a black-box neural network
 and the hybrid model \eqref{eq:hybrid_discrete_time} with network architectures
@@
 understanding of their performance if used for real-time economic optimization
-in chemical plants.
+purposes in chemical plants.
 
@@
 shows slightly worse predictions in a few places. However, both the trained
-models are adequate for real-time optimization in a feedback control system.
+models are adequate for real-time optimization purpose in a feedback control system.
 
@@
            samples. We notice that the hybrid model slightly outperforms the
-           black-box model in data requirements.}
+           black-box model in data requirements and 
+           overall prediction accuracy.}
   \label{fig:validation_mse}
@@
   x_s &= f(x_s, u_s)\\ 
-  y_s &= h(x_s)
+  y_s &= h(x_s) \\
+  \underline{u} \leq &u_s \leq \overline{u}
 \end{align*}
@@
 full supply of the raw material $A$. Since some of the product $B$ reacts to
-form an undesired side product $C$ in the plant, this recommendation incurres a
+form an undesired side product $C$ in the plant, this recommendation incurs a
 cost as observed from the plant cost curve. An intuitive approach to correct
@@
   \includegraphics[page=7, width=\textwidth]{tworeac_plots_nonlin.pdf}
-  \caption{Steady-state cost curve for the nonlinear plant, grey-box, black-box,
-  residual, and hybrid model after training with 10 hours of data.}
+  \caption{Variation of the cost error metric ($e(C_{Af})$) in the actuator
+           constraint set with increase in the training data for the 
+           black-box neural network. The metric improves in the 
+           constraint set with an increase in training data.}
   \label{fig:cost_error_bb}
@@
   \includegraphics[page=8, width=\textwidth]{tworeac_plots_nonlin.pdf}
-  \caption{Steady-state cost curve for the nonlinear plant, grey-box, black-box,
-  residual, and hybrid model after training with 10 hours of data.}
+  \caption{Variation of the cost error metric ($e(C_{Af})$) in the actuator
+  constraint set with increase in the training data for the 
+  hybrid grey-box and neural network model. The metric improves in the 
+  constraint set with an increase in training data.}
   \label{fig:cost_error_hybrid}

### 2020-11-24 lib/hybridid.py

@@ -176,9 +176,9 @@ def plot_profit_curve(*, us, costs, colors, legends,
     axes.set_ylabel(ylabel, rotation=False)
     axes.get_yaxis().set_label_coords(ylabel_xcoordinate, 0.5) 
     axes.set_xlim([np.min(us), np.max(us)])
-    figlabel = r'$\ell(y, u), \ \textnormal{subject to} \ f(x, u)=0, y=h(x)$'
-    figure.suptitle(figlabel,
-                    x=0.55, y=0.94)
+    #figlabel = r'$\ell(y, u), \ \textnormal{subject to} \ f(x, u)=0, y=h(x)$'
+    #figure.suptitle(figlabel,
+    #                x=0.55, y=0.94)
     # Return the figure object.
     return [figure]
 

### 2020-11-24 tworeac_plots_nonlin.py

@@ -58,7 +58,7 @@ def plot_val_model_predictions(*, plantsim_data,
     ylabels = [r'$C_A \ (\textnormal{mol/m}^3)$', 
                r'$C_B \ (\textnormal{mol/m}^3)$',
                r'$C_{Af} \ (\textnormal{mol/m}^3)$']
-    model_legend_colors = ['green', 'tomato']
+    model_legend_colors = ['green', 'dimgray', 'tomato']
     legend_handles = []
     plant_data_list = [plantsim_data.y[tsteps_steady:, 0], 
                        plantsim_data.y[tsteps_steady:, 1], 
@@ -86,9 +86,9 @@ def plot_val_model_predictions(*, plantsim_data,
     legend_handles.insert(0, plant_legend_handle[0])
     axes.set_xlabel('Time (hr)')
     axes.set_xlim([np.min(time[start:end]), np.max(time[start:end])])
-    labels = ('Plant', 'Grey-box', 'Hybrid')
+    labels = ('Plant', 'Grey-box', 'Black-box', 'Hybrid')
     figure.legend(handles = legend_handles, labels = labels, 
-                  loc = (0.13, 0.9), ncol=3)
+                  loc = (0.07, 0.9), ncol=4)
     # Return the figure object.
     return [figure]
 
@@ -188,7 +188,7 @@ def main():
                                   plot_range=(0, 6*60))
 
     # Plot predictions on validation data.
-    val_predictions.pop(0)
+    #val_predictions.pop(0)
     modelsim_datum = [greybox_validation_data] + val_predictions
     figures += plot_val_model_predictions(plantsim_data=training_data[-1],
                                     modelsim_datum=modelsim_datum,

### 2020-12-17 systemID_NNs.tex
commit the files.

@@
 
-\subsection{Conversion to discrete time and training}
+\section{Conversion to discrete time and training}
 We note that the network $f_N(\cdot)$ is an addition to the grey-box dynamics in
@@
 
-\subsection{Nonlinear reactor example}
+\section{Nonlinear reactor example}
 We now demonstrate the application of the above hybrid modeling framework
@@
 
+\section{Large, CSTR and flash in series example}
+\begin{figure}[!h]
+  \centering
+  \includegraphics[width=\textwidth, height=0.4\textheight]{cstr_flash.pdf}
+  \caption{Schematic of the CSTR and flash plant. The CSTR facilitates two 
+  reactions, the first reaction ($A \rightarrow B$) forms the primary product $B$, and the second reaction ($3B \rightarrow C$) 
+  consumes the desired product $B$ and forms the undesired product $C$.
+  The non-adiabatic flash separates the reactant $A$ and 
+  products $B$ and $C$. The major component in the vapor phase 
+  is $A$, and it is recycled back after cooling to the CSTR.}
+  \label{fig:cstr_flash}
+\end{figure}
+\newpage
+\textbf{Plant -- }
+\begin{align*}
+  \dfrac{dH_r}{dt} &= \dfrac{F + D -F_r}{A_r}\\
+  \dfrac{dC_{Ar}}{dt} &= \dfrac{F(C_{Af} -C_{Ar}) + 
+                         D(C_{Ad} -C_{Ar})}{A_rH_r} - r_1 \\
+  \dfrac{dC_{Br}}{dt} &= \dfrac{-FC_{Br} + 
+                          D(C_{Bd} -C_{Br})}{A_rH_r} + r_1 -3r_2\\
+  \dfrac{dC_{Cr}}{dt} &= \dfrac{-FC_{Cr} + 
+  D(C_{Cd} -C_{Cr})}{A_rH_r} + r_2\\
+  \dfrac{dT_r}{dt} &= \dfrac{F(T_f - T_r) + D(T_d -T_r)}{A_rH_r} + 
+                      \dfrac{r_1\Delta H_1 + r_2\Delta H_2}{\rho C_p} + 
+                      \dfrac{Q_r}{\rho A_r C_p H_r}\\
+  \dfrac{dH_b}{dt} &= \dfrac{F_r - F_b - D}{A_b} \\
+  \dfrac{dC_{Ab}}{dt} &= \dfrac{F_r(C_{Ar} -C_{Ab}) + 
+                          D(C_{Ab} -C_{Ad})}{A_bH_b} \\
+  \dfrac{dC_{Bb}}{dt} &= \dfrac{F_r(C_{Br} -C_{Bb}) + 
+                          D(C_{Bb} -C_{Bd})}{A_bH_b} \\
+  \dfrac{dC_{Cb}}{dt} &= \dfrac{F_r(C_{Cr} -C_{Cb}) + 
+                          D(C_{Cb} -C_{Cd})}{A_bH_b} \\
+  \dfrac{dT_b}{dt} &= \dfrac{F_r(T_r - T_b)}{A_bH_b} +
+                      \dfrac{Q_b}{\rho A_b C_p H_b}\\
+\end{align*}
+
+The sample time for plant measurements is $1$ min, the 
+plant states are ($H_r, C_{Ar}, C_{Br}, C_{Cr}, T_r, 
+H_b, C_{Ab}, C_{Bb}, C_{Cb}, T_b$), and the measurements are just 
+the heights and temperatures of the CSTR and flash. \\
+
+\textbf{Grey-box model -- }
+\begin{align*}
+  \dfrac{dH_r}{dt} &= \dfrac{F + D -F_r}{A_r}\\
+  \dfrac{dC_{Ar}}{dt} &= \dfrac{F(C_{Af} -C_{Ar}) + 
+                         D(C_{Ad} -C_{Ar})}{A_rH_r} - r_1 \\
+  \dfrac{dC_{Br}}{dt} &= \dfrac{-FC_{Br} + 
+                          D(C_{Bd} -C_{Br})}{A_rH_r} + r_1\\
+  \dfrac{dT_r}{dt} &= \dfrac{F(T_f - T_r) + D(T_d -T_r)}{A_rH_r} + 
+                      \dfrac{r_1\Delta H_1}{\rho C_p} + 
+                      \dfrac{Q_r}{\rho A_r C_p H_r}\\
+  \dfrac{dH_b}{dt} &= \dfrac{F_r - F_b - D}{A_b} \\
+  \dfrac{dC_{Ab}}{dt} &= \dfrac{F_r(C_{Ar} -C_{Ab}) + 
+                          D(C_{Ab} -C_{Ad})}{A_bH_b} \\
+  \dfrac{dC_{Bb}}{dt} &= \dfrac{F_r(C_{Br} -C_{Bb}) + 
+                          D(C_{Bb} -C_{Bd})}{A_bH_b} \\
+  \dfrac{dT_b}{dt} &= \dfrac{F_r(T_r - T_b)}{A_bH_b} +
+                      \dfrac{Q_b}{\rho A_b C_p H_b}\\
+\end{align*}
+
+The grey-box model assumes no knowledge of the second 
+reaction ($3B \rightarrow C$), which produces the undesired 
+product $C$. Note that since only heights and temperatures are 
+measured, all the states in the plant model are 
+not observable and only asymptotically detectable.
+
+The economic objective is -- 
+\begin{align*}
+  \underset{u_s}{\textnormal{min}} \quad 
+  \ell(u_s, x_s) \quad := & \quad c_1F + c_2Q_r + 
+                                      c_3Q_b - c_4C_{Bb} \\
+  x_s &= f(x_s, u_s)\\ 
+  y_s &= h(x_s) \\
+  \underline{u} \leq &u_s \leq \overline{u}
+\end{align*}
+
 \end{document}

### 2020-12-17 cstr_flash_plots.py

@@ -34,34 +34,18 @@ ulabels = [r'$F \ (\textnormal{m}^3/\textnormal{min})$',
            r'$D \ (\textnormal{m}^3/\textnormal{min})$',
            r'$Q_b \ (\textnormal{kJ/min})$']
 
-def get_plotting_arrays(data, plot_range):
-    """ Get data and return for plotting. """
-    start, end = plot_range
-    u = data.u[start:end, :]
-    x = data.x[start:end, :]
-    y = data.y[start:end, :]
-    t = data.t[start:end]/60 # Convert to hours.
-    return (t, x, y, u)
-
-def plot_inputs(t, udatum, figure_size, ylabel_xcoordinate,
-                legend_names, legend_colors):
+def plot_inputs(t, u, figure_size, ylabel_xcoordinate):
     """ Plot the training input data."""
     nrow = len(ulabels)
     (figure, axes) = plt.subplots(nrows=nrow, ncols=1,
                                   sharex=True, figsize=figure_size,
                                   gridspec_kw=dict(wspace=0.4))
-    legend_handles = []
-    for (u, color) in zip(udatum, legend_colors):
-        for row in range(nrow):
-            handle = axes[row].step(t, u[:, row], color)
-            axes[row].set_ylabel(ulabels[row])
-            axes[row].get_yaxis().set_label_coords(ylabel_xcoordinate, 0.5)
-        axes[row].set_xlabel('Time (hr)')
-        axes[row].set_xlim([np.min(t), np.max(t)])
-        legend_handles += [handle]
-    figure.legend(handles = legend_handles,
-                  labels = legend_names,
-                  loc = (0.2, 0.9), ncol=len(legend_names))
+    for row in range(nrow):
+        handle = axes[row].step(t, u[:, row], 'b')
+        axes[row].set_ylabel(ulabels[row])
+        axes[row].get_yaxis().set_label_coords(ylabel_xcoordinate, 0.5)
+    axes[row].set_xlabel('Time (hr)')
+    axes[row].set_xlim([np.min(t), np.max(t)])
     return [figure]
 
 def plot_outputs(t, ydatum, figure_size, ylabel_xcoordinate,
@@ -79,51 +63,67 @@ def plot_outputs(t, ydatum, figure_size, ylabel_xcoordinate,
             axes[row].get_yaxis().set_label_coords(ylabel_xcoordinate, 0.5)
         axes[row].set_xlabel('Time (hr)')
         axes[row].set_xlim([np.min(t), np.max(t)])
[…]

### 2020-12-17 lib/hybridid.py

@@ -13,7 +13,7 @@ import plottools
 
 FIGURE_SIZE_A4 = (9, 10)
 PRESENTATION_FIGSIZE = (6, 6)
-PAPER_FIGSIZE = (4, 4)
+PAPER_FIGSIZE = (5, 6)
 
 SimData = collections.namedtuple('SimData', 
                                 ['t', 'x', 'u', 'y'])

### 2020-12-28 cstr_flash_train.py
try to do with noise now.

@@ -44,7 +44,7 @@ def train_model(model, x0key, xuyscales, train_data, trainval_data, val_data,
     # Call the fit method to train.
     model.fit(x=[train_data['inputs'], train_data[x0key]],
               y=train_data['outputs'], 
-              epochs=1000, batch_size=8,
+              epochs=1000, batch_size=2,
         validation_data = ([trainval_data['inputs'], trainval_data[x0key]], 
                             trainval_data['outputs']),
         callbacks = [checkpoint_callback])
@@ -74,9 +74,9 @@ def main():
     num_samples = [hour*60 for hour in [4]]
 
     # Create lists.
-    Nps = [9]
+    Nps = [5]
     #fnn_dims = [[100, 128, 6], [102, 128, 8], [102, 128, 8]]
-    fnn_dims = [[102, 64, 64, 8]]
+    fnn_dims = [[102, 32, 32, 8]]
     model_types = ['hybrid']
     trained_weights = []
     val_metrics = []
@@ -113,9 +113,9 @@ def main():
                 x0key = 'yz0'
             else:
                 x0key = 'xGz0'
-            train_samples=dict(inputs=train_data['inputs'][:128,:num_sample, :],
-                            outputs=train_data['outputs'][:128,:num_sample, :])
-            train_samples[x0key] = train_data[x0key][:128, :]
+            train_samples=dict(inputs=train_data['inputs'][:8,:num_sample, :],
+                            outputs=train_data['outputs'][:8,:num_sample, :])
+            train_samples[x0key] = train_data[x0key][:8, :]
             (cstr_flash_model,
              val_prediction,
              val_metric) = train_model(cstr_flash_model, x0key, xuyscales,

### 2021-01-01 cstr_flash_parameters.py
ok, working on the empc script.

@@ -13,129 +13,9 @@ import scipy.linalg
 from hybridid import (PickleTool, NonlinearPlantSimulator,
                       c2d, sample_prbs_like, SimData,
                       NonlinearMHEEstimator)
-
-def _plant_ode(x, u, p, parameters):
-    """ ODEs describing the 10-D system. """
-
-    # Extract the parameters.
-    alphaA = parameters['alphaA']
-    alphaB = parameters['alphaB']
-    alphaC = parameters['alphaC']
-    pho = parameters['pho']
-    Cp = parameters['Cp']
-    Ar = parameters['Ar']
-    Ab = parameters['Ab']
-    kr = parameters['kr']
-    kb = parameters['kb']
-    delH1 = parameters['delH1']
-    delH2 = parameters['delH2']
-    EbyR = parameters['EbyR']
-    k1star = parameters['k1star']
-    k2star = parameters['k2star']
-    Td = parameters['Td']
-
-    # Extract the plant states into meaningful names.
-    (Hr, CAr, CBr, CCr, Tr) = x[0:5]
-    (Hb, CAb, CBb, CCb, Tb) = x[5:10]
-    (F, Qr, D, Qb) = u[0:4]
-    (CAf, Tf) = p[0:2]
-
-    # The flash vapor phase mass fractions.
-    den = alphaA*CAb + alphaB*CBb + alphaC*CCb
-    CAd = alphaA*CAb/den
-    CBd = alphaB*CBb/den
-    CCd = alphaB*CCb/den
-
-    # The outlet mass flow rates.
-    Fr = kr*np.sqrt(Hr)
-    Fb = kb*np.sqrt(Hb)
-
-    # The rate constants.
-    k1 = k1star*np.exp(-EbyR/Tr)
-    k2 = k2star*np.exp(-EbyR/Tr)
-
-    # The rate of reactions.
-    r1 = k1*CAr
-    r2 = k2*(CBr**3)
-
-    # Write the CSTR odes.
-    dHrbydt = (F + D - Fr)/Ar
-    dCArbydt = (F*(CAf - CAr) + D*(CAd - CAr))/(Ar*Hr) - r1
-    dCBrbydt = (-F*CBr + D*(CBd - CBr))/(Ar*Hr) + r1 - 3*r2
[…]

### 2021-01-01 lib/hybridid.py

@@ -625,6 +625,129 @@ def get_cstr_flash_train_val_data(*, Np, parameters,
     # Return.
     return (train_data, trainval_data, val_data, xuyscales)
 
+def _cstr_flash_plant_ode(x, u, p, parameters):
+    """ ODEs describing the 10-D system. """
+
+    # Extract the parameters.
+    alphaA = parameters['alphaA']
+    alphaB = parameters['alphaB']
+    alphaC = parameters['alphaC']
+    pho = parameters['pho']
+    Cp = parameters['Cp']
+    Ar = parameters['Ar']
+    Ab = parameters['Ab']
+    kr = parameters['kr']
+    kb = parameters['kb']
+    delH1 = parameters['delH1']
+    delH2 = parameters['delH2']
+    EbyR = parameters['EbyR']
+    k1star = parameters['k1star']
+    k2star = parameters['k2star']
+    Td = parameters['Td']
+
+    # Extract the plant states into meaningful names.
+    (Hr, CAr, CBr, CCr, Tr) = x[0:5]
+    (Hb, CAb, CBb, CCb, Tb) = x[5:10]
+    (F, Qr, D, Qb) = u[0:4]
+    (CAf, Tf) = p[0:2]
+
+    # The flash vapor phase mass fractions.
+    den = alphaA*CAb + alphaB*CBb + alphaC*CCb
+    CAd = alphaA*CAb/den
+    CBd = alphaB*CBb/den
+    CCd = alphaB*CCb/den
+
+    # The outlet mass flow rates.
+    Fr = kr*np.sqrt(Hr)
+    Fb = kb*np.sqrt(Hb)
+
+    # The rate constants.
+    k1 = k1star*np.exp(-EbyR/Tr)
+    k2 = k2star*np.exp(-EbyR/Tr)
+
+    # The rate of reactions.
+    r1 = k1*CAr
+    r2 = k2*(CBr**3)
+
+    # Write the CSTR odes.
+    dHrbydt = (F + D - Fr)/Ar
+    dCArbydt = (F*(CAf - CAr) + D*(CAd - CAr))/(Ar*Hr) - r1
+    dCBrbydt = (-F*CBr + D*(CBd - CBr))/(Ar*Hr) + r1 - 3*r2
[…]

### 2021-01-09 cstr_flash_empc.py
need to change the NEMPC classes to take funcs in discrete time.

@@ -79,7 +79,7 @@ def get_controller(model_ode, model_pars, model_type,
     # Horizon lengths.
     Nmpc = 60
     Nmhe = 30
-
+    breakpoint()
     # Return the NN controller.
     return NonlinearEMPCController(fxu=fxu, hx=hx,
                                    lxup=lxup, Bd=Bd, Cd=Cd,
@@ -90,7 +90,7 @@ def get_controller(model_ode, model_pars, model_type,
                                    ulb=ulb, uub=uub, Nmpc=Nmpc,
                                    Qwx=Qwx, Qwd=Qwd, Rv=Rv, Nmhe=Nmhe)
 
-def get_hybrid_pars(*, greybox_pars, Npast, fnn_weights):
+def get_hybrid_pars(*, greybox_pars, Npast, fnn_weights, xuyscales):
     """ Get the hybrid model parameters. """
 
     hybrid_pars = copy.deepcopy(greybox_pars)
@@ -110,6 +110,11 @@ def get_hybrid_pars(*, greybox_pars, Npast, fnn_weights):
     hybrid_pars['Npast'] = Npast
     hybrid_pars['fnn_weights'] = fnn_weights 
 
+    # Scaling.
+    hybrid_pars['xscale'] = xuyscales['xscale']
+    hybrid_pars['uscale'] = xuyscales['uscale']
+    hybrid_pars['yscale'] = xuyscales['yscale']
+
     # Return.
     return hybrid_pars
 
@@ -117,21 +122,28 @@ def _hybrid_ode(xGz, u, p, parameters):
     """ The augmented continuous time model. """
 
     Ng = parameters['Ng']
+    Npast = parameters['Npast']
     fnn_weights = parameters['fnn_weights']
+    xscale = parameters['xscale']
+    uscale = parameters['uscale']
+    yscale = parameters['yscale']
+    xGzscale = np.concatenate((xscale,
+                               np.tile(yscale, (Npast, )), 
+                               np.tile(uscale, (Npast, ))))
 
     # Compute the NN output.
-    nn_output = np.concatenate((xGz, u))
+    nn_output = np.concatenate((xGz/xGzscale, u/uscale))[:, np.newaxis]
     for i in range(0, len(fnn_weights)-2, 2):
         (W, b) = fnn_weights[i:i+2]
         nn_output = np.tanh(W.T @ nn_output + b[:, np.newaxis])
     (Wf, bf) = fnn_weights[-2:]
-    nn_output = Wf.T @ nn_output + bf[:, np.newaxis]
[…]

### 2021-01-09 lib/HybridModelLayers.py

@@ -115,7 +115,7 @@ class CstrFlashHybridCell(tf.keras.layers.AbstractRNNCell):
         (self.Ng, self.Ny, self.Nu) = (cstr_flash_parameters['Ng'],
                                        cstr_flash_parameters['Ny'],
                                        cstr_flash_parameters['Nu'])
-
+    
     @property
     def state_size(self):
         return self.Ng + self.Np*(self.Ny + self.Nu)
@@ -139,7 +139,7 @@ class CstrFlashHybridCell(tf.keras.layers.AbstractRNNCell):
         kr = self.parameters['kr']
         kb = self.parameters['kb']
         delH1 = self.parameters['delH1']
-        EbyR = self.parameters['EbyR']
+        E1byR = self.parameters['E1byR']
         k1star = self.parameters['k1star']
         Td = self.parameters['Td']
         ps = self.parameters['ps']
@@ -169,7 +169,7 @@ class CstrFlashHybridCell(tf.keras.layers.AbstractRNNCell):
         Fb = kb*tf.math.sqrt(Hb)
 
         # Rate constant and reaction rate.
-        k1 = k1star*tf.math.exp(-EbyR/Tr)
+        k1 = k1star*tf.math.exp(-E1byR/Tr)
         r1 = k1*CAr
 
         # Write the CSTR odes.

### 2021-01-09 lib/hybridid.py

@@ -218,6 +218,10 @@ class NonlinearEMPCRegulator:
                                   guess=guess)
         self.regulator.solve()
         self.regulator.saveguess()
+        useq = np.asarray(casadi.horzcat(*self.regulator.var['u'])).T
+        xseq = np.asarray(casadi.horzcat(*self.regulator.var['x'])).T
+        breakpoint()
+        self._append_data(x0, useq, xseq)
 
     def solve(self, x0, empc_pars):
         """Setup and the solve the dense QP, output is
@@ -351,10 +355,9 @@ class NonlinearEMPCController:
         Bd is in continuous time.
     """
     def __init__(self, *, fxu, hx, lxup, Bd, Cd, sample_time,
-                     Nx, Nu, Ny, Nd, Np,
+                     Nx, Nu, Ny, Nd,
                      xs, us, ds, ys,
-                     empc_pars,
-                     ulb, uub, Nmpc,
+                     empc_pars, ulb, uub, Nmpc,
                      Qwx, Qwd, Rv, Nmhe):
         
         # Model and stage cost.
@@ -370,7 +373,7 @@ class NonlinearEMPCController:
         self.Nu = Nu
         self.Ny = Ny
         self.Nd = Nd
-        self.Np = Np
+        self.Np = empc_pars.shape[1]
 
         # Steady states of the system.
         self.xs = xs
@@ -481,30 +484,33 @@ def online_simulation(plant, controller, *, plant_lxup, Nsim=None,
     sys.stdout = open(stdout_filename, 'w')
     measurement = plant.y[0] # Get the latest plant measurement.
     disturbances = disturbances[..., np.newaxis]
-    stage_costs = []
+    avg_stage_costs = [0.]
     for (simt, disturbance) in zip(range(Nsim), disturbances):
         print("Simulation Step:" + f"{simt}")
         control_input = controller.control_law(simt, measurement)
         print("Computation time:" + str(controller.computation_times[-1]))
         stage_cost = plant_lxup(plant.x[-1], control_input,
                                 controller.empc_pars[simt:simt+1, :].T)[0]
-        stage_costs += [stage_cost]
+        avg_stage_costs += [(avg_stage_costs[-1]*simt + stage_cost)/(simt+1)]
         measurement = plant.step(control_input, disturbance)
     # Create a sim data and stage cost array.
     cl_data = SimData(t=np.asarray(plant.t[0:-1]).squeeze(),
[…]

### 2021-01-19 cstr_flash_plots.py
need to add a few things in the talk.

@@ -162,8 +162,8 @@ def plot_cost_pars(t, cost_pars,
                                   gridspec_kw=dict(left=0.18))
     xlabel = 'Time (hr)'
     ylabels = ['Energy Price ($\$$/kW)',
-                'Raw Material Cost ($\$$/mol-A)',
-                'Product Price ($\$$/mol-B)']
+                'Raw Mat ($\$$/mol-A)',
+                'Prod Price ($\$$/mol-B)']
     for (axes, pari, ylabel) in zip(axes_list, range(num_pars), ylabels):
         # Plot the corresponding data.
         cost_pars[:, 0] = 60*cost_pars[:, 0]

### 2021-01-19 tworeac_plots_nonlin.py

@@ -82,7 +82,7 @@ def plot_cost_pars(t, cost_pars,
                                   figsize=figure_size,
                                   gridspec_kw=dict(left=0.18))
     xlabel = 'Time (hr)'
-    ylabels = [r'$c_1$', r'$c_2$']
+    ylabels = [r'$c_a$', r'$c_b$']
     for (axes, pari, ylabel) in zip(axes_list, range(num_pars), ylabels):
         # Plot the corresponding data.
         axes.plot(t, cost_pars[:len(t), pari])

### 2021-01-20 lib/HybridModelLayers.py
slightly postive outcome.

@@ -415,10 +415,13 @@ class TwoReacModel(tf.keras.Model):
         # Dense layers for the NN.
         fnn_layers = []
         for dim in fnn_dims[1:-1]:
-            fnn_layers.append(tf.keras.layers.Dense(dim, activation='tanh'))
+            fnn_layers.append(tf.keras.layers.Dense(dim, activation='sigmoid',
+                              kernel_regularizer=tf.keras.regularizers.l2(1e-3),
+                              bias_regularizer=tf.keras.regularizers.l2(1e-3)))
         fnn_layers.append(tf.keras.layers.Dense(fnn_dims[-1],
                                                 kernel_initializer='zeros',
-                                                use_bias=False))
+                            kernel_regularizer=tf.keras.regularizers.l2(1e-3),
+                            use_bias=False))
 
         # Build model depending on option.
         if model_type == 'black-box':

### 2021-01-20 lib/linNonlinMPC.py

@@ -140,7 +140,10 @@ class NonlinearEMPCRegulator:
         # Some parameters for the regulator.
         empc_pars = self.init_empc_pars
         guess = self.init_guess
-        x0 = guess['x']
+        if len(guess['x'].shape) == 2:
+            x0 = guess['x'][0, :]
+        else:
+            x0 = guess['x']
         lb = dict(u=self.ulb)
         ub = dict(u=self.uub)
         self.regulator = mpc.nmpc(f=self.fxu, l=lxup, N=N, funcargs=funcargs,
@@ -287,7 +290,8 @@ class NonlinearEMPCController:
                      Nx, Nu, Ny, Nd,
                      xs, us, ds, ys,
                      empc_pars, ulb, uub, Nmpc,
-                     Qwx, Qwd, Rv, Nmhe):
+                     Qwx, Qwd, Rv, Nmhe,
+                     regulator_guess):
         
         # Model and stage cost.
         self.fxu = fxu
@@ -315,6 +319,7 @@ class NonlinearEMPCController:
         self.ulb = ulb
         self.uub = uub
         self.Nmpc = Nmpc
+        self.regulator_guess = regulator_guess
         self._setup_regulator()
 
         # MHE Parameters.
@@ -376,8 +381,17 @@ class NonlinearEMPCController:
         build the regulator. """
         fxud = mpc.getCasadiFunc(self._aug_ss_model(),
                                 [self.Nx + self.Nd, self.Nu], ["x", "u"])
-        init_guess = dict(x=np.concatenate((self.xs, self.ds), axis=0), 
-                          u=self.us)
+        if self.regulator_guess is None:
+            init_guess = dict(x=np.concatenate((self.xs, self.ds), axis=0), 
+                              u=self.us)
+        else:
+            xseq = [np.concatenate((self.xs, self.ds))]
+            useq = self.regulator_guess[1]
+            for t in range(self.Nmpc):
+                x = np.asarray(fxud(xseq[-1], useq[t, :]))[:, 0]
+                xseq.append(x)
+            xseq = np.asarray(xseq)
+            init_guess = dict(x=xseq, u=useq)
         init_empc_pars = self.empc_pars[0:self.Nmpc, :]
         self.regulator  = NonlinearEMPCRegulator(fxu=fxud,
                                      lxup = self.lxup,

### 2021-01-20 tworeac_train_nonlin.py

@@ -82,9 +82,9 @@ def main():
     num_samples = [hours*60 for hours in [6]]
 
     # Create lists.
-    Nps = [2, 2]
-    fnn_dims = [[10, 16, 2], [8, 16, 2]]
-    model_types = ['black-box', 'hybrid']
+    Nps = [2]
+    fnn_dims = [[8, 32, 2]]
+    model_types = ['hybrid']
     trained_weights = []
     val_metrics = []
     val_predictions = []

### 2021-02-23 systemID_NNs.tex
working on the tensorflow script.
ok, working.
try to do with noise now.
ok, working on the empc script.
need to change the NEMPC classes to take funcs in discrete time.
add scripts.

@@
 \begin{itemize}
-	\item Black-box: $y^+ = f_N(y, z, u)$, (purely black-box)
+  \item Black-box: $y^+ = f_N(y, z, u), \ z^+ = f_z(z, y, u)$
+        (purely black-box).
   \item Hybrid 1: $x_g^+ = f_N(x_g, z, u), \ z^+ = f_z(z, x_g, u), \ y =
@@
 
-The Figure \ref{fig:validation_pred} shows predictions of the trained black-box
+The Figure shows predictions of the trained black-box
 and hybrid models after training with 8 hours (480 samples) of data. We observe
@@
 models are adequate for real-time optimization purpose in a feedback control system.
-
-\begin{figure}[!h]
-  \centering
-  \includegraphics[page=2, width=\textwidth]{tworeac_plots_nonlin.pdf}
-  \caption{Model validation for the nonlinear reactor model. Both the black-box
-           and hybrid model can predict the plant responses accurately. The
-           model predictions shown in this plot correspond to training with 8
-           hours of data.}
-  \label{fig:validation_pred}
-\end{figure}
 
@@
 in which $\hat{y}$ are the model predictions and $y$ denotes the plant
-measurements. The Figure \ref{fig:validation_mse} shows the variation of this
+measurements. The Figure  shows the variation of this
 MSE metric as a function of the number of training data obtained with both the
@@
 data requirements.
-
-\begin{figure}[!h]
-  \centering
-  \includegraphics[page=9, width=\textwidth]{tworeac_plots_nonlin.pdf}
-  \caption{Mean squared error (MSE) performance metric obtained with the
-           black-box and hybrid model with increase in the number of training
-           samples. We notice that the hybrid model slightly outperforms the
-           black-box model in data requirements and 
-           overall prediction accuracy.}
-  \label{fig:validation_mse}
-\end{figure}
 
@@
 $\ell(y_s, u_s) = c_1C_{Af} - c_2C_B$, which measures the raw material cost and
-product profit. The Figure \ref{fig:cost_curves} shows the cost curves of the
+product profit. The Figure shows the cost curves of the
 plant, grey-box, black-box, and hybrid model obtained with 8 hours of training
@@
 
-\begin{figure}[!h]
-  \centering
-  \includegraphics[page=6, width=\textwidth]{tworeac_plots_nonlin.pdf}
-  \caption{Steady-state cost curve for the nonlinear plant, grey-box, black-box,
-  and hybrid model after training with 8 hours of data.}
-  \label{fig:cost_curves}
-\end{figure}
-
-\begin{figure}[!h]
-  \centering
-  \includegraphics[page=7, width=\textwidth]{tworeac_plots_nonlin.pdf}
-  \caption{Variation of the cost error metric ($e(C_{Af})$) in the actuator
-           constraint set with increase in the training data for the 
-           black-box neural network. The metric improves in the 
-           constraint set with an increase in training data.}
-  \label{fig:cost_error_bb}
-\end{figure}
-
 Finally, we analyze the variation in the cost curves of the black-box and hybrid
@@
 in which $y_p(\cdot)$ and $y_m(\cdot)$ denote the plant and model steady state
-output respectively. The Figures \ref{fig:cost_error_bb} and
-\ref{fig:cost_error_hybrid} show the variation in this cost error metric in the
+output respectively. The Figures and show the variation in this cost error metric in the
 actuator constraint set with increasing number of training samples for the
@@
 optimum.
-
-\begin{figure}[!h]
-  \centering
-  \includegraphics[page=8, width=\textwidth]{tworeac_plots_nonlin.pdf}
-  \caption{Variation of the cost error metric ($e(C_{Af})$) in the actuator
-  constraint set with increase in the training data for the 
-  hybrid grey-box and neural network model. The metric improves in the 
-  constraint set with an increase in training data.}
-  \label{fig:cost_error_hybrid}
-\end{figure}
 
@@
 \end{figure}
-\newpage
+%\newpage
 \textbf{Plant -- }
@@
                       \dfrac{Q_r}{\rho A_r C_p H_r}\\
+\end{align*}
+
+\begin{align*}
   \dfrac{dH_b}{dt} &= \dfrac{F_r - F_b - D}{A_b} \\
@@
   \underset{u_s}{\textnormal{min}} \quad 
-  \ell(u_s, x_s) \quad := & \quad c_1F + c_2Q_r + 
-                                      c_3Q_b - c_4C_{Bb} \\
+  \ell(u_s, x_s) \quad := & \quad c_1FC_{Af} + c_2Q_r + 
+                                      c_2Q_b + c_2D\rho C_p(T_d-T_b) - 
+                                      c_3F_bC_{Bb} \\
   x_s &= f(x_s, u_s)\\ 
@@
 
+in which $c_1$ is the raw material cost of unit $\$$/mol-A,
+$c_2$ is the energy price of unit $\$$/kW, 
+and $c_3$ is the product selling price of unit $\$$/mol-B.
 \end{document}

### 2021-02-23 cstr_flash_funcs.py

@@ -8,63 +8,138 @@
 import sys
 sys.path.append('lib/')
 import mpctools as mpc
+import copy
 import numpy as np
 import scipy.linalg
-from hybridid import (PickleTool, c2d, sample_prbs_like, SimData)
+from hybridid import (PickleTool, c2d, sample_prbs_like, SimData,
+                      get_scaling, interpolate_yseq, _resample_fast)
 from linNonlinMPC import NonlinearPlantSimulator, NonlinearMHEEstimator
-from hybridid import _cstr_flash_plant_ode as _plant_ode
-from hybridid import _cstr_flash_greybox_ode as _greybox_ode
-from hybridid import _cstr_flash_measurement as _measurement
 
-def _get_greybox_parameters(*, plant_pars):
-    """ Get the parameter values for the
-        CSTRs with flash example. """
+def plant_ode(x, u, p, parameters):
+    """ ODEs describing the 10-D system. """
 
-    # Parameters.
-    parameters = {}
-    parameters['alphaA'] = 6.
-    parameters['alphaB'] = 1.
-    parameters['pho'] = 6. # Kg/m^3
-    parameters['Cp'] = 6. # KJ/(Kg-K)
-    parameters['Ar'] = 2. # m^2
-    parameters['Ab'] = 2. # m^2
-    parameters['kr'] = 2. # m^2
-    parameters['kb'] = 1.5 # m^2
-    parameters['delH1'] = 150. # kJ/mol
-    parameters['E1byR'] = 200 # K
-    parameters['k1star'] = 0.3 # 1/min
-    parameters['Td'] = 300 # K
+    # Extract the parameters.
+    alphaA = parameters['alphaA']
+    alphaB = parameters['alphaB']
+    alphaC = parameters['alphaC']
+    pho = parameters['pho']
+    Cp = parameters['Cp']
+    Ar = parameters['Ar']
+    Ab = parameters['Ab']
+    kr = parameters['kr']
+    kb = parameters['kb']
+    delH1 = parameters['delH1']
+    delH2 = parameters['delH2']
+    E1byR = parameters['E1byR']
+    E2byR = parameters['E2byR']
+    k1star = parameters['k1star']
+    k2star = parameters['k2star']
+    Td = parameters['Td']
 
-    # Store the dimensions.
-    parameters['Ng'] = 8
-    parameters['Nu'] = plant_pars['Nu']
[…]

### 2021-02-23 cstr_flash_parameters.py

@@ -12,162 +12,13 @@ import numpy as np
 import scipy.linalg
 from hybridid import (PickleTool, c2d, sample_prbs_like, SimData)
 from linNonlinMPC import NonlinearPlantSimulator, NonlinearMHEEstimator
-from hybridid import _cstr_flash_plant_ode as _plant_ode
-from hybridid import _cstr_flash_greybox_ode as _greybox_ode
-from hybridid import _cstr_flash_measurement as _measurement
+from cstr_flash_funcs import plant_ode, greybox_ode, measurement, get_model
+from cstr_flash_funcs import get_plant_parameters, get_greybox_parameters
+from cstr_flash_funcs import get_rectified_xs
 
-def _get_greybox_parameters(*, plant_pars):
-    """ Get the parameter values for the
-        CSTRs with flash example. """
-
-    # Parameters.
-    parameters = {}
-    parameters['alphaA'] = 6.
-    parameters['alphaB'] = 1.
-    parameters['pho'] = 6. # Kg/m^3
-    parameters['Cp'] = 6. # KJ/(Kg-K)
-    parameters['Ar'] = 2. # m^2
-    parameters['Ab'] = 2. # m^2
-    parameters['kr'] = 2. # m^2
-    parameters['kb'] = 1.5 # m^2
-    parameters['delH1'] = 150. # kJ/mol
-    parameters['E1byR'] = 200 # K
-    parameters['k1star'] = 0.3 # 1/min
-    parameters['Td'] = 300 # K
-
-    # Store the dimensions.
-    parameters['Ng'] = 8
-    parameters['Nu'] = plant_pars['Nu']
-    parameters['Ny'] = plant_pars['Ny']
-    parameters['Np'] = plant_pars['Np']
-
-    # Sample time.
-    parameters['Delta'] = 1. # min
-
-    # Get the steady states.
-    gb_indices = [0, 1, 2, 4, 5, 6, 7, 9]
-    parameters['xs'] = plant_pars['xs'][gb_indices]
-    parameters['us'] = plant_pars['us']
-    parameters['ps'] = np.array([5., 320.])
-
-    # The C matrix for the grey-box model.
-    parameters['tsteps_steady'] = plant_pars['tsteps_steady']
-    parameters['yindices'] = [0, 1, 2, 3, 4, 5, 6, 7]
-
-    # Get the constraints.
-    parameters['ulb'] = plant_pars['ulb']
-    parameters['uub'] = plant_pars['uub']
-
-    # Noise for MHE.
[…]

### 2021-02-23 cstr_flash_plots.py

@@ -187,6 +187,8 @@ def main():
                                        type='read')
     cstr_flash_empc = PickleTool.load(filename="cstr_flash_empc.pickle",
                                        type='read')
+    cstr_flash_rto = PickleTool.load(filename="cstr_flash_rto.pickle",
+                                       type='read')
 
     # Collect data to plot open-loop predictions.
     val_predictions = cstr_flash_train['val_predictions']
@@ -203,8 +205,8 @@ def main():
     ulist += ulist_train
     ylist += ylist_train
     xlist += xlist_train
-    legend_names = ['Plant', 'Grey-Box', 'Black-box', 'Hybrid']
-    legend_colors = ['b', 'g', 'dimgrey', 'm']
+    legend_names = ['Plant', 'Grey-Box','Hybrid']
+    legend_colors = ['b', 'g', 'm']
     figures = []
     figures += plot_data(t=t, udatum=ulist, ydatum=ylist,
                               xdatum=xlist, data_type='open_loop',
@@ -220,19 +222,19 @@ def main():
     #                            legends=['Black-box', 'Hybrid'])
 
     # Plot the open-loop solutions.
-    legend_names = ['Plant', 'Grey-box', 'Hybrid']
-    legend_colors = ['b', 'g', 'm']
-    openloop_sols = cstr_flash_empc['openloop_sols']
-    udatum = [openloop_sols[0][0], openloop_sols[1][0], openloop_sols[2][0]]
-    xdatum = [openloop_sols[0][1], openloop_sols[1][1], openloop_sols[2][1]]
-    figures += plot_openloop_sols(t=t, udatum=udatum, xdatum=xdatum,
-                              legend_names=legend_names,
-                              legend_colors=legend_colors)
+    #legend_names = ['Plant', 'Grey-box', 'Hybrid']
+    #legend_colors = ['b', 'g', 'm']
+    #openloop_sols = cstr_flash_empc['openloop_sols']
+    #udatum = [openloop_sols[0][0], openloop_sols[1][0], openloop_sols[2][0]]
+    #xdatum = [openloop_sols[0][1], openloop_sols[1][1], openloop_sols[2][1]]
+    #figures += plot_openloop_sols(t=t, udatum=udatum, xdatum=xdatum,
+    #                          legend_names=legend_names,
+    #                          legend_colors=legend_colors)
     
     # Plot the closed-loop simulation.
-    legend_names = ['Plant', 'Grey-box', 'Hybrid']
-    legend_colors = ['b', 'g', 'm']
[…]

### 2021-02-23 cstr_flash_train.py

@@ -10,9 +10,9 @@ sys.path.append('lib/')
 import tensorflow as tf
 import time
 import numpy as np
-from hybridid import (PickleTool, get_cstr_flash_train_val_data, 
-                      SimData)
+from hybridid import (PickleTool, SimData)
 from HybridModelLayers import CstrFlashModel
+from cstr_flash_funcs import get_train_val_data
 
 # Set the tensorflow global and graph-level seed.
 tf.random.set_seed(2)
@@ -87,7 +87,7 @@ def train_model(model, x0key, xuyscales, train_data, trainval_data, val_data,
     # Call the fit method to train.
     model.fit(x = [train_data['inputs'], train_data[x0key]],
               y = train_outputs,
-              epochs=6000, batch_size=12,
+              epochs=2000, batch_size=2,
         validation_data = ([trainval_data['inputs'], trainval_data[x0key]], 
                            trainval_outputs),
         callbacks = [checkpoint_callback])
@@ -126,7 +126,7 @@ def main():
 
     # Create lists.
     Nps = [5]
-    fnn_dims = [[102, 32, 32, 8]]
+    fnn_dims = [[102, 32, 8]]
     model_types = ['hybrid']
     trained_weights = []
     val_metrics = []
@@ -145,7 +145,7 @@ def main():
         # Get the training data.
         (train_data,
          trainval_data,
-         val_data, xuyscales) = get_cstr_flash_train_val_data(Np=Np,
+         val_data, xuyscales) = get_train_val_data(Np=Np,
                                 parameters=greybox_pars,
                                 greybox_processed_data=greybox_processed_data)
 

### 2021-02-23 lib/HybridModelLayers.py

@@ -459,12 +459,109 @@ class CstrFlashModel(tf.keras.Model):
         # Dense layers for the NN.
         fnn_layers = []
         for dim in fnn_dims[1:-1]:
-            fnn_layers.append(tf.keras.layers.Dense(dim, activation='sigmoid',
-                              kernel_regularizer=tf.keras.regularizers.l2(1e-3),
-                              bias_regularizer=tf.keras.regularizers.l2(1e-3)))
+            fnn_layers.append(tf.keras.layers.Dense(dim, activation='tanh'))
         fnn_layers.append(tf.keras.layers.Dense(fnn_dims[-1], 
-                          kernel_initializer='zeros',
-                          kernel_regularizer=tf.keras.regularizers.l2(1e-3)))
+                          kernel_initializer='zeros'))
+
+        # Build model depending on option.
+        if model_type == 'black-box':
+            cstr_flash_cell = BlackBoxCell(Np, Ny, Nu, fnn_layers)
+        if model_type == 'hybrid':
+            interp_layer = InterpolationLayer(p=Ny, Np=Np)
+            cstr_flash_cell = CstrFlashHybridCell(Np, interp_layer, fnn_layers,
+                                            xuyscales, cstr_flash_parameters)
+        #if model_type == 'grey-black':
+        #    cstr_flash_cell = CstrFlashGreyBlackCell(Np, fnn_layers,
+        #                                             cstr_flash_parameters)
+
+        # Construct the RNN layer and the computation graph.
+        cstr_flash_layer = tf.keras.layers.RNN(cstr_flash_cell,
+                                               return_sequences=True)
+        layer_output = cstr_flash_layer(inputs=layer_input,
+                                        initial_state=[initial_state])
+        if model_type == 'black-box':
+            outputs = [layer_output]
+        else:
+            y, xG = tf.split(layer_output, [Ny, Ng], axis=-1)
+            outputs = [y, xG]
+        # Construct model.
+        super().__init__(inputs=[layer_input, initial_state],
+                         outputs=outputs)
+
+class KoopmanCell(tf.keras.layers.AbstractRNNCell):
+    """
+    RNN Cell.
+    z = [y_{k-N_p:k-1}', u_{k-N_p:k-1}'];
+    x_{kp} = NN(y, z)
+    x_{kp}^+  = A_x_{kp} + Bu
+    [y';z']^+ = Hx_{kp}^+
+    y = [I, 0][y';z']
[…]

### 2021-02-23 lib/hybridid.py

@@ -121,47 +121,6 @@ def get_plotting_array_list(*, simdata_list, plot_range):
     # Return lists.
     return (t, ulist, ylist, xlist)
 
-def _get_energy_price(*, num_days, sample_time):
-    """ Get a two day heat disturbance profile. """
-    energy_price = np.zeros((24, 1))
-    energy_price[0:8, :] = np.ones((8, 1))
-    energy_price[8:16, :] = 70*np.ones((8, 1))
-    energy_price[16:24, :] = np.ones((8, 1))
-    energy_price = 1e-2*np.tile(energy_price, (num_days, 1))
-    return _resample_fast(x=energy_price,
-                          xDelta=60,
-                          newDelta=sample_time,
-                          resample_type='zoh')
-
-def get_cstr_flash_empc_pars(*, num_days, sample_time, plant_pars):
-    """ Get the parameters for Empc and RTO simulations. """
-
-    # Get the cost parameters.
-    energy_price = _get_energy_price(num_days=num_days, sample_time=sample_time)
-    raw_mat_price = _resample_fast(x = np.array([[1000.], [1000.], 
-                                                 [1000.], [1000.], 
-                                                 [1000.], [1000.], 
-                                                 [1000.], [1000.]]), 
-                                   xDelta=6*60,
-                                   newDelta=sample_time,
-                                   resample_type='zoh')
-    product_price = _resample_fast(x = np.array([[6000.], [7000.], 
-                                                 [5000.], [4000.], 
-                                                 [4000.], [4000.], 
-                                                 [4000.], [4000.]]),
-                                   xDelta=6*60,
-                                   newDelta=sample_time,
-                                   resample_type='zoh')
-    cost_pars = np.concatenate((energy_price,
[…]

### 2021-02-23 tworeac_nonlin_funcs.py

@@ -18,7 +18,7 @@ def plant_ode(x, u, p, parameters):
     k1 = parameters['k1']
     k2 = parameters['k2']
     k3 = parameters['k3']
-
+    
     # Extract the plant states into meaningful names.
     (Ca, Cb, Cc) = x[0:3]
     Ca0 = u[0:1]

### 2021-02-23 tworeac_plots_nonlin.py

@@ -147,25 +147,22 @@ def main():
                                          "tworeac_parameters_nonlin.pickle",
                                          type='read')
     (parameters, training_data,
-     greybox_val_data) = (tworeac_parameters['parameters'], 
+     greybox_val_data) = (tworeac_parameters['parameters'],
                           tworeac_parameters['training_data'],
                           tworeac_parameters['greybox_val_data'])
     
     # Load data after training.
     tworeac_train = PickleTool.load(filename=
-                                    "tworeac_train_nonlin.pickle", 
+                                    "tworeac_train_nonlin.pickle",
                                     type='read')
     val_predictions = tworeac_train['val_predictions']
 
     # Load data after economic MPC/RTO simulation.
-    tworeac_empc = PickleTool.load(filename=
-                                    "tworeac_empc_nonlin.pickle", 
-                                    type='read')
-    tworeac_rto = PickleTool.load(filename=
-                                    "tworeac_rto_nonlin.pickle", 
-                                    type='read')
-    cl_data_list = tworeac_empc['cl_data_list']
-    cost_pars = tworeac_empc['cost_pars']
+    #tworeac_empc = PickleTool.load(filename=
+    #                                "tworeac_empc_nonlin.pickle", 
+    #                                type='read')
+    #cl_data_list = tworeac_empc['cl_data_list']
+    #cost_pars = tworeac_empc['cost_pars']
 
     # Create a figures list.
     figures = []
@@ -173,7 +170,7 @@ def main():
     # Plot validation data.
     legend_names = ['Plant', 'Grey-box', 'Hybrid']
     legend_colors = ['b', 'g', 'm']
-    valdata_list = [training_data[-1], greybox_val_data] 
+    valdata_list = [training_data[-1], greybox_val_data]
     valdata_list += val_predictions
     t, ulist, ylist, xlist = get_plotting_array_list(simdata_list=
                                                      valdata_list[:2],
@@ -210,34 +207,34 @@ def main():
     # Plot closed-loop simulation data.
     #legend_names = ['Plant', 'Grey-box', 'Hybrid']
     #legend_colors = ['b', 'g', 'm']
-    t, ulist, ylist, xlist = get_plotting_array_list(simdata_list=
-                                                     cl_data_list,
-                                                     plot_range=(0, 8*60))
-    figures += plot_xudata(t=t, xlist=xlist, ulist=ulist,
-                           legend_names=legend_names,
-                           legend_colors=legend_colors)
[…]

### 2021-02-23 tworeac_train_nonlin.py

@@ -1,4 +1,5 @@
 # [depends] %LIB%/hybridid.py %LIB%/HybridModelLayers.py
+# [depends] %LIB%/../tworeac_nonlin_funcs.py
 # [depends] tworeac_parameters_nonlin.pickle
 # [makes] pickle
 """ Script to train the hybrid model for the 
@@ -45,7 +46,7 @@ def train_model(model, train_data, trainval_data, val_data,
     # Call the fit method to train.
     model.fit(x=[train_data['inputs'], train_data['xGz0']], 
               y=train_data['outputs'], 
-              epochs=8000, batch_size=1,
+              epochs=2000, batch_size=1,
         validation_data = ([trainval_data['inputs'], trainval_data['xGz0']], 
                             trainval_data['outputs']),
             callbacks = [checkpoint_callback])
@@ -143,11 +144,11 @@ def main():
                                  trained_weights=trained_weights,
                                  val_predictions=val_predictions,
                                  val_metrics=val_metrics,
-                                 num_samples=num_samples, 
+                                 num_samples=num_samples,
                                  xuscales=xuscales)
     
     # Save data.
-    PickleTool.save(data_object=tworeac_training_data, 
+    PickleTool.save(data_object=tworeac_training_data,
                     filename='tworeac_train_nonlin.pickle')
 
 main()

### 2021-03-05 cstr_flash_blackbox_train.py
will work on writing for now.

@@ -45,7 +45,7 @@ def train_model(model, train_data, trainval_data, stdout_filename, ckpt_path):
     # Call the fit method to train.
     model.fit(x = [train_data['inputs'], train_data['yz0']],
               y = train_data['outputs'],
-              epochs=3000, batch_size=12,
+              epochs=3000, batch_size=24,
         validation_data = ([trainval_data['inputs'], trainval_data['yz0']], 
                            trainval_data['outputs']),
         callbacks = [checkpoint_callback])

### 2021-03-05 cstr_flash_empc.py

@@ -17,39 +17,39 @@ from linNonlinMPC import (NonlinearPlantSimulator, NonlinearEMPCController,
                          online_simulation)
 from cstr_flash_funcs import plant_ode, greybox_ode, measurement
 from cstr_flash_funcs import get_economic_opt_pars, get_model
-from hybridid import get_koopman_pars_check_func
+from hybridid import get_koopman_pars_check_func, koopman_func
 
-def get_controller(model_func, model_pars, model_type, 
+def get_controller(model_func, model_pars, model_type,
                    cost_pars, mhe_noise_tuning, regulator_guess):
     """ Construct the controller object comprised of 
         EMPC regulator and MHE estimator. """
 
+    # Get the sizes.
+    if model_type == 'grey-box':
+        Nx = model_pars['Ng']
+    else:
+        Nx = model_pars['Nx']
+    Nu, Ny = model_pars['Nu'], model_pars['Ny']
+    Nd = Ny
+
     # Get models.
-    if model_type == 'hybrid':
+    if model_type == 'koopman':
         fxu = lambda x, u: model_func(x, u, model_pars)
+        ymean, ystd = model_pars['xuyscales']['yscale']
+        hx = lambda x: x[:Ny]*ystd + ymean
     else:
         ps = model_pars['ps']
         Delta = model_pars['Delta']
         fxu = lambda x, u: model_func(x, u, ps, model_pars)
         fxu = c2dNonlin(fxu, Delta)
-
-    # Measurement function.
-    hx = lambda x: measurement(x, model_pars)
+        hx = lambda x: measurement(x, model_pars)
 
     # Get the stage cost.
-    if model_type == 'grey-box':
-        Nx = model_pars['Ng']
-    else:
-        Nx = model_pars['Nx']
     lyup = lambda y, u, p: stage_cost(y, u, p, model_pars)
     
-    # Get the sizes/sample time.
-    (Nu, Ny) = (model_pars['Nu'], model_pars['Ny'])
-    Nd = Ny
-
     # Get the disturbance models.
     Bd = np.zeros((Nx, Nd))
-    if model_type == 'plant':
+    if model_type == 'plant' or model_type == 'koopman':
         Bd[0, 0] = 1.
         Bd[2, 1] = 1.
         Bd[4, 2] = 1.
[…]

### 2021-03-05 cstr_flash_plots.py

@@ -49,12 +49,12 @@ def plot_inputs(t, udatum, figure_size, ylabel_xcoordinate,
     legend_handles = []
     for (u, color) in zip(udatum, legend_colors):
         for row in range(nrow):
-            handle = axes[row].step(t, u[:, row], color)
+            handle = axes[row].step(t, u[:, row], color, where='post')
             axes[row].set_ylabel(ulabels[row])
             axes[row].get_yaxis().set_label_coords(ylabel_xcoordinate, 0.5)
         axes[row].set_xlabel('Time (hr)')
         axes[row].set_xlim([np.min(t), np.max(t)])
-        if data_type == 'closed_loop' and row == 3:
+        if data_type == 'closed_loop' and row in [1, 3]:
             axes[row].set_ylim([100., 500.])
         legend_handles += handle
     if data_type == 'closed_loop':
@@ -73,7 +73,7 @@ def plot_outputs(t, ydatum, figure_size, ylabel_xcoordinate,
     legend_handles = []
     for (y, color) in zip(ydatum, legend_colors):
         for row in range(nrow):
-            handle = axes[row].step(t, y[:, row], color)
+            handle = axes[row].step(t, y[:, row], color, where='post')
             axes[row].set_ylabel(ylabels[row])
             axes[row].get_yaxis().set_label_coords(ylabel_xcoordinate, 0.5)
         axes[row].set_xlabel('Time (hr)')
@@ -130,10 +130,10 @@ def get_openloop_xtrajs(xdatum):
         if i==0:
             x_trajs.append(x_traj[:-1, :])
         else:
-            x_traj = x_traj[:-1, :8]
-            x_traj = np.insert(x_traj, [3, 7], 
-                               np.nan*np.ones((x_traj.shape[0], 2)), axis=1)
-            x_trajs.append(x_traj)
+            #x_traj = x_traj[:-1, :8]
+            x_traj = np.insert(x_traj, [1, 2, 4, 5], 
+                               np.nan*np.ones((x_traj.shape[0], 4)), axis=1)
+            x_trajs.append(x_traj[:-1, :])
     # Return.
     return x_trajs
 
@@ -187,30 +187,34 @@ def main():
     cstr_flash_kooptrain = PickleTool.load(filename=
                                             "cstr_flash_kooptrain.pickle",
                                             type='read')
-    #cstr_flash_empc = PickleTool.load(filename="cstr_flash_empc.pickle",
[…]

### 2021-03-05 lib/HybridModelLayers.py

@@ -496,12 +496,12 @@ class KoopmanCell(tf.keras.layers.AbstractRNNCell):
     x_{kp}^+  = Ax_{kp} + Bu
     #[y';z']^+ = Hx_{kp}^+
     [y;z] = H*xkp
-    y = [I, 0]*[y;z]
+    #y = [I, 0]*[y;z]
     """
-    def __init__(self, Nxkp, Ny, A, B, H, **kwargs):
+    def __init__(self, Nxkp, Ny, Nz, A, B, **kwargs):
         super(KoopmanCell, self).__init__(**kwargs)
-        self.Nxkp, self.Ny = Nxkp, Ny
-        self.A, self.B, self.H = A, B, H
+        self.Nxkp, self.Ny, self.Nz = Nxkp, Ny, Nz
+        self.A, self.B = A, B
 
     @property
     def state_size(self):
@@ -509,13 +509,13 @@ class KoopmanCell(tf.keras.layers.AbstractRNNCell):
     
     @property
     def output_size(self):
-        return self.Ny
+        return self.Ny + self.Nz
 
     def call(self, inputs, states):
         """ Call function of the hybrid RNN cell.
             Dimension of states: (None, Nxkp)
             Dimension of input: (None, Nu)
-            Dimension of output: (None, Ny)
+            Dimension of output: (None, Nz)
         """
         # Extract important variables.
         [xkp] = states
@@ -525,26 +525,25 @@ class KoopmanCell(tf.keras.layers.AbstractRNNCell):
         xkplus = self.A(xkp) + self.B(u)
         
         # Get the output at the current timestep.
-        yz = self.H(xkp)
-        y = yz
+        [yz, _] = tf.split(xkp, [self.output_size, self.Nxkp-self.output_size], 
+                           axis=-1)
 
         # Return output and states at the next time-step.
-        return (y, xkplus)
+        return (yz, xkplus)
 
 class KoopmanModel(tf.keras.Model):
     """ Custom model for the Deep Koopman operator model. """
-    def __init__(self, Np, Ny, Nu, fnn_dims):
+    def __init__(self, Np, Ny, Nu, fnn_dims, kernel_reg=1e-4):
         
         # Get a few sizes.
         Nz = Np*(Ny+Nu)
-        Nxkp = fnn_dims[-1]
+        Nxkp = Ny + Nz + fnn_dims[-1]
 
         # Create inputs to the layers.
[…]

### 2021-03-05 lib/hybridid.py

@@ -281,15 +281,15 @@ def get_koopman_pars_check_func(*, parameters, training_data, train):
 
     # Get weights.
     trained_weights = train['trained_weights'][-1]
-    fnn_weights = trained_weights[:-3]
-    A = trained_weights[-3].T
-    B = trained_weights[-2].T
-    H = trained_weights[-1].T
+    fnn_weights = trained_weights[:-2]
+    A = trained_weights[-2].T
+    B = trained_weights[-1].T
+    #H = trained_weights[-1].T
 
     # Get sizes.
     Np = train['Np']
     Ny, Nu = parameters['Ny'], parameters['Nu']
-    Nx = train['fnn_dims'][-1]
+    Nx = Ny + Np*(Ny + Nu) + train['fnn_dims'][-1]
 
     # Get scaling.
     xuyscales = train['xuyscales']
@@ -306,16 +306,15 @@ def get_koopman_pars_check_func(*, parameters, training_data, train):
     ys = parameters['xs'][yindices]
     yzs = np.concatenate((np.tile(ys, (Np+1, )), 
                           np.tile(us, (Np, ))))[:, np.newaxis]
-    breakpoint()
     yzs = (yzs - yzmean)/yzstd
-    xkps = fnn_koopman(yzs, fnn_weights)[:, 0]
+    xkps = np.concatenate((yzs, fnn_koopman(yzs, fnn_weights)))[:, 0]
 
     # Constraints for MPC.
     ulb, uub = parameters['ulb'], parameters['uub']
 
     # Make the parameter dictionary.
     koopman_pars  = dict(Np=Np, Nx=Nx, Nu=Nu, Ny=Ny,
-                         A=A, B=B, H=H, xs=xkps, us=us,
+                         A=A, B=B, xs=xkps, us=us,
                          ulb=ulb, uub=uub, xuyscales=xuyscales)
 
     # Get the control input profile for the simulation.
@@ -331,11 +330,11 @@ def get_koopman_pars_check_func(*, parameters, training_data, train):
     y0 = y[ts, :, np.newaxis]
     yz0 = np.concatenate((y0, yp0seq, up0seq))
     yz0 = (yz0 - yzmean)/yzstd
-    xkp0 = fnn_koopman(yz0, fnn_weights)[:, 0]
+    xkp0 = np.concatenate((yz0, fnn_koopman(yz0, fnn_weights)))[:, 0]
 
     # Get the functions.
[…]

### 2021-03-05 tworeac_empc_nonlin.py

@@ -39,9 +39,8 @@ def get_controller(model_func, model_pars, model_type,
     # Get state space and input models.
     if model_type == 'koopman':
         fxu = lambda x, u: model_func(x, u, model_pars)
-        H = model_pars['H']
         ymean, ystd = model_pars['xuyscales']['yscale']
-        hx = lambda x: (H @ x[:Nx])[:Ny]*ystd + ymean
+        hx = lambda x: x[:Ny]*ystd + ymean
     else:
         ps = model_pars['ps']
         Delta = model_pars['Delta']
@@ -79,7 +78,7 @@ def get_controller(model_func, model_pars, model_type,
     Qwx, Qwd, Rv = mhe_noise_tuning
 
     # Horizon lengths.
-    Nmpc = 120
+    Nmpc = 60
     Nmhe = 30
 
     # Return the NN controller.
@@ -99,17 +98,18 @@ def stage_cost(y, u, p):
     ca, cb = p[0:2]
     CA, CB = y[0:2]
     # Compute and return cost.
-    return ca*CAf - cb*CB
+    return (CB - 1.2)**2
+    # return ca*CAf - cb*CB
 
 def get_mhe_noise_tuning(model_type, model_par):
     # Get MHE tuning.
     if model_type == 'plant' or model_type =='koopman':
         Qwx = 1e-6*np.eye(model_par['Nx'])
-        Qwd = 1e-6*np.eye(model_par['Ny'])
+        Qwd = 1e-2*np.eye(model_par['Ny'])
         Rv = 1e-3*np.eye(model_par['Ny'])
     if model_type == 'grey-box':
         Qwx = 1e-6*np.eye(model_par['Ng'])
-        Qwd = np.eye(model_par['Ny'])
+        Qwd = 1e-2*np.eye(model_par['Ny'])
         Rv = 1e-3*np.eye(model_par['Ny'])
     return (Qwx, Qwd, Rv)
 

### 2021-03-05 tworeac_kooptrain_nonlin.py

@@ -45,7 +45,7 @@ def train_model(model, train_data, trainval_data,
     # Call the fit method to train.
     model.fit(x=[train_data['inputs'], train_data['yz0']], 
               y=[train_data['yz'], train_data['outputs']], 
-              epochs=6000, batch_size=11,
+              epochs=5000, batch_size=11,
         validation_data = ([trainval_data['inputs'], trainval_data['yz0']], 
                            [trainval_data['yz'], trainval_data['outputs']]),
             callbacks = [checkpoint_callback])

### 2021-03-05 tworeac_plots_nonlin.py

@@ -45,7 +45,7 @@ def plot_xudata(*, t, xlist, ulist,
         legend_handles += handle
     figure.legend(handles = legend_handles,
                   labels = legend_names,
-                  loc = (0.15, 0.9), ncol=len(legend_names))
+                  loc = (0.08, 0.9), ncol=len(legend_names))
     # Return figure.
     return [figure]
 
@@ -163,15 +163,24 @@ def main():
                                     type='read')
     koopval_predictions = tworeac_kooptrain['val_predictions']
 
+    # Load data after Koopman training.
+    tworeac_encdeckooptrain = PickleTool.load(filename=
+                                    "tworeac_encdeckooptrain_nonlin.pickle",
+                                    type='read')
+    encdeckoopval_predictions = tworeac_encdeckooptrain['val_predictions']
+
+
     # Create a figures list.
     figures = []
 
     # Plot validation data.
-    legend_names = ['Plant', 'Grey-box', 'Black-box', 'Koopman']
-    legend_colors = ['b', 'g', 'dimgrey', 'm']
+    legend_names = ['Plant', 'Grey-box', 'Black-box', 
+                    'Koopman', 'Koopman-ENC-DEC']
+    legend_colors = ['b', 'g', 'dimgrey', 'm', 'tomato']
     valdata_list = [training_data[-1], greybox_val_data]
     valdata_list += blackbox_predictions
     valdata_list += koopval_predictions
+    valdata_list += encdeckoopval_predictions
     t, ulist, ylist, xlist = get_plotting_array_list(simdata_list=
                                                      valdata_list[:2],
                                                      plot_range=(10, 24*60+10))

### 2021-03-10 sysID.tex
now focus on the Koopman model for the smaller example.

\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{authblk}
\usepackage{amssymb}
%\usepackage{lucidbry}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{subcaption}
\usepackage{physics}
\usepackage[dvipsnames]{xcolor}
\usepackage{tikz,tikzsettings}
\usepackage{verbatim}
\usetikzlibrary{arrows,shapes}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{verbatim}
\usepackage{float}
\usepackage[round]{natbib}
\usepackage[english]{babel}
\bibliographystyle{plainnat}
%\setcitestyle{numeric,open={(},close={)}} \usepackage{color}
%\usepackage{colortbl}
\usetikzlibrary{decorations.markings}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{shapes,calc,backgrounds,arrows,fit,decorations.pathmorphing,matrix}
\usepackage[margin=0.79in]{geometry}

\graphicspath{{./}{./figures/}{./figures/paper/}}

\newcommand{\Is}{\bm{\mathrm{I}}} % Resources.
\newcommand{\Js}{\bm{\mathrm{J}}} % Units.
\newcommand{\Ks}{\bm{\mathrm{K}}} % Resources.
\newcommand{\Ms}{\bm{\mathrm{M}}} % Transition catalog.
\newcommand{\Ns}{\bm{\mathrm{N}}} % Transition counter.

\newcommand{\useq}{\mathbf{u}} \newcommand{\xseq}{\mathbf{x}}
\newcommand{\bbR}{\mathbb{R}} \newcommand{\bbW}{\mathbb{W}}
\newcommand{\bbU}{\mathbb{U}} \newcommand{\bbI}{\mathbb{I}}
\newcommand{\bbX}{\mathbb{X}} \newcommand{\bbP}{\mathbb{P}}

\title{Nonlinear system identification using machine learning tools}
\author{Pratyush Kumar}
\date{\today}

\begin{document}

\maketitle

\section{Process modeling}
We consider the problem of system identification of the following nonlinear plant.
\begin{align*}
  x_p^+ &= f_p(x_p, u, w) \\
  y &= h_p(x_p) + v
\end{align*}
in which $x_p \in \bbR^{n_p}$ is the plant state, $u \in \bbU \subset \bbR^m$ is the
manipulated control input, $y \in \bbR^p$ is the measurement, $w$ is the process
noise, and $v$ is the measurement noise. The objective of developing the model is to use it subsequently in model predictive control (MPC). Therefore, we consider several types of grey-box, hybrid, and black-box models and analyze their predictive capabilities as well as their usefulness for online optimization in MPC.

\section{Model architectures for system identification}
We describe several process modeling frameworks which determine parameters in their proposed model architecture by minimizing the multi-step ahead prediction error of the model on training data. A nonlinear optimization problem is solved for this identification step. The estimated parameters define the model, which is then used subsequently to perform model validation and for use in MPC.

\subsection{Grey-box model}
We first consider the grey-box modeling strategy, which is the current industrial state-of-the-art system identification technique. A parameterized model is defined based on first-principle knowledge about the plant, and parameters in the model are estimated from training data by solving the prediction-error minimization problem. The grey-box model is typically defined in continuous time and is of the following form:

\begin{align*}
  \dot{x}_g &= f_g(x_g, u, \theta_g) \\
  y &= h_g(x_g, \theta_g)
\end{align*}

The following nonlinear program (NLP) is solved to identify the parameters $\theta_g$:

\begin{align} \label{eq:gb_training}
  \underset{x_g^i(0), \ \theta_g}{\textnormal{min}} \sum_{i=1}^{N_{tr}} \sum_{k=1}^{N_t} 
  \dfrac{1}{N_tN_{tr}} &\norm{\tilde{y}^i(k) - y^i(k)}^2_{R_v^{-1}}  \\
  \textnormal{subject to} \quad \quad x_g^{i+} &= f_{gd}(x_g^i, u^i, \theta_g) \nonumber \\
   \tilde{y}^i &= h_g(x_g^i, \theta_g) \nonumber
\end{align}

in which $\theta_g$ and $x_g^i(0)$ are the decision variables in the optimization problem, which are parameters in the grey-box model and initial state used for forecasting the predicted measurements. We assume that multiple trajectories can be used in the training data, and the superscript $i$ is used to denote the trajectory number. We use $N_{tr}$ to denote the total number of trajectories in training data and $N_t$ to denote the number of samples in each trajectory. The function $f_{g}(\cdot)$ is discretized at the measurement sample time to obtain the function $f_{gd}(\cdot)$ that is used to make forecasts in the above optimization.

\subsection{Hybrid model}

Often in industrial applications, first-principle knowledge about the plant is incomplete. Hence, grey-box system identification and subsequent use of the grey-box model in real-time optimization can lead to suboptimal economic performance. We propose to augment the grey-box model using a neural network (NN) such that the overall hybrid model can capture any unknown process dynamics in the plant, therefore leading to improved economic performance. We define this hybrid model as follows:

\begin{align*}
  x_g^+ &= f_{gd}(x_g, u, \theta_g) + f_N(z, u, \theta_N) \\
  z^+ &= f_z(x_g, z, u) \\
  y &= h_g(x_g, \theta_g)
\end{align*}
in which $f_N(\cdot)$ is a NN and $\theta_N$ denotes the parameters in the network. We use $z$ to denote the vector of past measurements and control inputs. This vector is considered as an additional state in the overall dynamic model and is defined as follows:
\begin{align} \label{eq:pastyu}
  \mathbf{y}_{k-N_p:k-1} &= [y(k-N_p)', ... \ , y(k-1)']' \nonumber \\
  \mathbf{u}_{k-N_p:k-1} &= [u(k-N_p)', ... \ , u(k-1)']' \nonumber \\
  z(k) &= [\mathbf{y}_{k-N_p:k-1}', \mathbf{u}_{k-N_p:k-1}']'
\end{align}
in which $N_p$ is the number of past measurements and control inputs used. The function $f_z(\cdot)$ describes the dynamics of the state $z$ and its structure is defined as follows: 
\begin{align*}
  f_z(x_g, z, u) = \begin{bmatrix}
    y(k-N_p+1) \\
    y(k-N_p+2) \\
    \vdots \\
    h_g(x_g, \theta_g) \\
    u(k-N_p+1) \\ 
    u(k-N_p+2) \\
    \vdots \\
    u(k)
  \end{bmatrix}
\end{align*}
We assume that the parameters in the grey-box model are first estimated by solving the NLP \eqref{eq:gb_training}, and the following similar NLP is solved to determine parameters ($\theta_N$) in the NN.

\begin{align} \label{eq:hybrid_training}
  \underset{\theta_N}{\textnormal{min}} \sum_{i=1}^{N_{tr}} \sum_{k=1}^{N_t} 
  \dfrac{1}{N_tN_{tr}} &\norm{\tilde{y}^i(k) - y^i(k)}^2  \\
  \textnormal{subject to} \quad \quad x_g^{i+} &= f_{gd}(x_g^i, u^i, \theta_g) + f_z(z^i, u^i, \theta_N) \nonumber \\
  z^{i+} &= f_z(x_g^i, z^i, u^i) \nonumber \\
   \tilde{y}^i &= h_g(x_g^i, \theta_g) \nonumber
\end{align}
The stochastic gradient descent algorithm Adam is used to solve the above optimization using the software tensorflow. Note that we do not estimate the initial state for each trajectory, and fix those states heuristically to make forecasts during the optimization process. The control inputs and measurements are scaled as $u := (u - u_{\textnormal{MEAN}})/u_{\textnormal{STD}}$ and 
$y := (y - y_{\textnormal{MEAN}})/y_{\textnormal{STD}}$, in which $u_{\textnormal{MEAN}}$ and $y_{\textnormal{MEAN}}$ are the respective means computed on training data. Similarly, $u_{\textnormal{STD}}$ and $y_{\textnormal{STD}}$ are the respective standard deviations. Due to this scaling, we do not consider the penalty $R_v^{-1}$ in the training objective.

\subsection{Black-box models}
Next, we consider black-box system identification strategies which do not impose any a-prior knowledge about the plant and develop the dynamic model purely from data. We consider two types of black-box models in this section: NNs and dynamic models based on the Koopman operator theory. The latter type of model is considered based on the motivation that it imposes a linear structure on the dynamics in a high-dimensional space, therefore leading to a linear or quadratic program to be solved online in MPC.

\subsubsection{Neural network}
The black-box NN uses the vector of past measurements and control inputs $z$ \eqref{eq:pastyu} as the state in the dynamic model and is structured as follows
\begin{align*}
  z^+ &= f_z(z, u) \\ 
  y &= h_N(z, \theta_N)
\end{align*}
in which $h_N(\cdot)$ is a NN and $\theta_N$ are the parameters in the network. The function $f_z(\cdot)$ denotes the dynamics of the state $z$ and is defined as

\begin{align*}
  f_z(z, u) = \begin{bmatrix}
    y(k-N_p+1) \\
    y(k-N_p+2) \\
    \vdots \\
    h_N(z, \theta_N) \\
    u(k-N_p+1) \\ 
    u(k-N_p+2) \\
    \vdots \\
    u(k)
  \end{bmatrix}
\end{align*}

\subsubsection{Koopman operator model}

\subsubsection{Training}
The training problem is to find the parameters in the neural network
$f_N$ and the linear model matrices $A$ and $B$ by minimizing the following mean squared error:

\begin{align*}
  \underset{A, B, W_i, b_i}{\textnormal{min}} \sum_{k=0}^{N_t-1} 
  \dfrac{1}{N_t}(\hat{x}(k) - x(k))^2
\end{align*}

\section{Two reaction system plant model}

\section{Model validation}

\section{Steady state optimization}

\section{Dynamic economic optimization}

\section{Future work}

\end{document}

### 2021-03-10 cstr_flash_bbtrain.py

@@ -35,8 +35,8 @@ def main():
     # Create some parameters.
     xinsert_indices = [1, 2, 4, 5]
     tthrow = 120
-    Np = 6
-    hN_dims = [Np*(Ny+Nu), 256, 256, 6]
+    Np = 3
+    hN_dims = [Np*(Ny+Nu), 32, 6]
 
     # Create lists to store data.
     trained_weights = []
@@ -65,7 +65,7 @@ def main():
                              outputs=train_data['outputs'])
 
         # Train.
-        train_bbmodel(model=model, epochs=12000, batch_size=96, 
+        train_bbmodel(model=model, epochs=30000, batch_size=12, 
                       train_data=train_samples, trainval_data=trainval_data, 
                       stdout_filename=stdout_filename, ckpt_path=ckpt_path)
 

### 2021-03-10 cstr_flash_kooptrain.py

@@ -41,7 +41,7 @@ def train_model(model, train_data, trainval_data, stdout_filename, ckpt_path):
     # Call the fit method to train.
     model.fit(x = [train_data['inputs'], train_data['yz0']],
               y = [train_data['yz'], train_data['outputs']],
-              epochs=10000, batch_size=12,
+              epochs=10000, batch_size=54,
         validation_data = ([trainval_data['inputs'], trainval_data['yz0']], 
                            [trainval_data['yz'], trainval_data['outputs']]),
         callbacks = [checkpoint_callback])
@@ -88,8 +88,8 @@ def main():
     num_samples = [batch*Nsim_train for batch in num_batches]
 
     # Create lists.
-    Np = 2
-    fnn_dims = [102, 64, 16]
+    Np = 6
+    fnn_dims = [102, 128, 32]
     trained_weights = []
     val_metrics = []
     val_predictions = []

### 2021-03-10 cstr_flash_openloop.py

@@ -33,10 +33,10 @@ def get_openloop_sol(fxu, hx, model_pars, xuguess):
     # Some sizes. 
     Np = 3
     Nx, Nu = model_pars['Nx'], model_pars['Nu']
-    Nmpc = 120
+    Nmpc = 60
 
     # Initial parameters. 
-    init_empc_pars = np.tile(np.array([[1, 1000, 10000]]), (Nmpc, 1))
+    init_empc_pars = np.tile(np.array([[1, 2000, 12000]]), (Nmpc, 1))
 
     # Get upper and lower bounds.
     ulb = model_pars['ulb']
@@ -70,12 +70,19 @@ def main():
                                          'cstr_flash_parameters.pickle',
                                          type='read')
     plant_pars = cstr_flash_parameters['plant_pars']
-    #tworeac_bbtrain = PickleTool.load(filename='tworeac_bbtrain.pickle',
-    #                                 type='read')
+    cstr_flash_bbtrain = PickleTool.load(filename='cstr_flash_bbtrain.pickle',
+                                     type='read')
 
     # Get the black-box model parameters and function handles.
-    #bb_pars, blackb_fxu, blackb_hx = get_bbpars_fxu_hx(train=tworeac_bbtrain, 
-    #                                                   parameters=parameters) 
+    bb_pars, blackb_fxu, blackb_hx = get_bbpars_fxu_hx(train=
+                                                       cstr_flash_bbtrain, 
+                                                       parameters=plant_pars)
+    # Add some more parameters to bb_pars.
+    bb_pars['ps'] = plant_pars['ps']
+    bb_pars['Td'] = plant_pars['Td']
+    bb_pars['pho'] = plant_pars['pho']
+    bb_pars['Cp'] = plant_pars['Cp']
+    bb_pars['kb'] = plant_pars['kb']
 
     # Get the plant function handle.
     Delta = plant_pars['Delta']
@@ -91,11 +98,11 @@ def main():
     #gb_pars['Nx'] = len(parameters['gb_indices'])
 
     # Lists to loop over for the three problems.  
-    model_types = ['plant']#, 'grey-box', 'black-box']
-    fxu_list = [plant_fxu]#, gb_fxu, blackb_fxu]
-    hx_list = [plant_hx]#, plant_hx, blackb_hx]
-    par_list = [plant_pars]#, gb_pars, bb_pars]
-    Nps = [None]#, None, bb_pars['Np']]
[…]

### 2021-03-10 cstr_flash_plots.py

@@ -15,34 +15,6 @@ from hybridid import PickleTool
 from plotting_funcs import (PAPER_FIGSIZE, get_plotting_array_list, 
                             CstrFlashPlots)
 
-def get_openloop_xtrajs(xdatum):
-    """ Clean up open-loop data for plotting. """
-    x_trajs = []
-    for (i, x_traj) in enumerate(xdatum):
-        if i==0:
-            x_trajs.append(x_traj[:-1, :])
-        else:
-            #x_traj = x_traj[:-1, :8]
-            x_traj = np.insert(x_traj, [1, 2, 4, 5], 
-                               np.nan*np.ones((x_traj.shape[0], 4)), axis=1)
-            x_trajs.append(x_traj[:-1, :])
-    # Return.
-    return x_trajs
-
-def plot_openloop_sols(*, t, udatum, xdatum,
-                          legend_names, legend_colors,
-                          figure_size=PAPER_FIGSIZE):
-    """ Plot the open-loop EMPC solutions. """
-    xdatum = get_openloop_xtrajs(xdatum)
-    figures = []
-    t = t[:udatum[0].shape[0]]
-    figures += plot_inputs(t, udatum, figure_size, -0.1,
-                           'closed_loop', legend_names, legend_colors)
-    figures += plot_states(t, xdatum, figure_size, -0.25, 
-                           legend_names, legend_colors)
-    # Return the figure object.
-    return figures
-
 def plot_cost_pars(t, cost_pars,
                    figure_size=PAPER_FIGSIZE,
                    ylabel_xcoordinate=-0.15):
@@ -73,9 +45,8 @@ def main():
     cstr_flash_parameters = PickleTool.load(filename=
                                             "cstr_flash_parameters.pickle",
                                             type='read')
-    # cstr_flash_blackbox_train = PickleTool.load(filename=
-    #                                         "cstr_flash_blackbox_train.pickle",
-    #                                         type='read')
+    cstr_flash_bbtrain = PickleTool.load(filename="cstr_flash_bbtrain.pickle",
+                                         type='read')
     # cstr_flash_kooptrain = PickleTool.load(filename=
     #                                         "cstr_flash_kooptrain.pickle",
     #                                         type='read')
@@ -88,26 +59,26 @@ def main():
     #                                   type='read')
 
     # Collect data to plot open-loop predictions.
-    # blackbox_val_predictions = cstr_flash_blackbox_train['val_predictions']
[…]

### 2021-03-10 lib/HybridModelLayers.py

@@ -388,13 +388,15 @@ class BlackBoxModel(tf.keras.Model):
         useq = tf.keras.Input(name='u', shape=(None, Nu))
         z0 = tf.keras.Input(name='z0', shape=(Np*(Ny+Nu), ))
 
-        def simplesig(x):
-            return 1./(1. + tf.math.exp(-0.4*x))
+        def scaledtanh(x, a=0.1):
+            num = tf.math.exp(a*x) - tf.math.exp(-a*x)
+            den = tf.math.exp(a*x) + tf.math.exp(-a*x)
+            return num/den
 
         # Dense layers for the NN.
         hN_layers = []
         for dim in hN_dims[1:-1]:
-            hN_layers += [tf.keras.layers.Dense(dim, activation=simplesig)]
+            hN_layers += [tf.keras.layers.Dense(dim, activation=scaledtanh)]
         hN_layers += [tf.keras.layers.Dense(hN_dims[-1])]
 
         # Build model.

### 2021-03-10 lib/economicopt.py

@@ -49,12 +49,18 @@ def fnn(nn_input, nn_weights):
         Assume that the input is a vector with shape size 
         1, and return output with shape size 1.
         """
+    def scaledtanh(x, a=3e-2):
+        num = np.exp(a*x) - np.exp(-a*x)
+        den = np.exp(a*x) + np.exp(-a*x)
+        return num/den
+
     nn_output = nn_input[:, np.newaxis]
     for i in range(0, len(nn_weights)-2, 2):
         (W, b) = nn_weights[i:i+2]
         nn_output = W.T @ nn_output + b[:, np.newaxis]
         #nn_output = np.tanh(nn_output)
-        nn_output = 1./(1. + np.exp(-0.4*nn_output))
+        #nn_output = 1./(1. + np.exp(-0.4*nn_output))
+        nn_output = scaledtanh(nn_output)
     (Wf, bf) = nn_weights[-2:]
     nn_output = (Wf.T @ nn_output + bf[:, np.newaxis])[:, 0]
     # Return.

### 2021-03-10 lib/hybridid.py

@@ -99,15 +99,15 @@ def get_plotting_array_list(*, simdata_list, plot_range):
     # Return lists.
     return (t, ulist, ylist, xlist)
 
-# def _resample_fast(*, x, xDelta, newDelta, resample_type):
-#     """ Resample with either first of zero-order hold. """
-#     Delta_ratio = int(xDelta/newDelta)
-#     if resample_type == 'zoh':
-#         return np.repeat(x, Delta_ratio, axis=0)
-#     else:
-#         x = np.concatenate((x, x[-1, np.newaxis, :]), axis=0)
-#         return np.concatenate([np.linspace(x[t, :], x[t+1, :], Delta_ratio)
-#                                for t in range(x.shape[0]-1)], axis=0)
+def resample_fast(*, x, xDelta, newDelta, resample_type):
+    """ Resample with either first of zero-order hold. """
+    Delta_ratio = int(xDelta/newDelta)
+    if resample_type == 'zoh':
+        return np.repeat(x, Delta_ratio, axis=0)
+    else:
+        x = np.concatenate((x, x[-1, np.newaxis, :]), axis=0)
+        return np.concatenate([np.linspace(x[t, :], x[t+1, :], Delta_ratio)
+                               for t in range(x.shape[0]-1)], axis=0)
 
 # def interpolate_yseq(yseq, Npast, Ny):
 #     """ y is of dimension: (None, (Npast+1)*p)

### 2021-03-10 lib/plotting_funcs.py

@@ -10,6 +10,7 @@ import collections
 import pickle
 import plottools
 import time
+import itertools
 
 FIGURE_SIZE_A4 = (9, 10)
 PRESENTATION_FIGSIZE = (6, 6)
@@ -95,6 +96,117 @@ class TwoReacPlots:
         axes.get_yaxis().set_label_coords(ylabel_xcoordinate, 0.5) 
         axes.set_xlim([np.min(us), np.max(us)])
         return [figure]
+
+class CstrFlashPlots:
+
+    ylabels = [r'$H_r \ (\textnormal{m})$',
+               r'$C_{Br} \ (\textnormal{mol/m}^3)$', 
+               r'$T_r \ (K)$',
+               r'$H_b \ (\textnormal{m})$',
+               r'$C_{Bb} \ (\textnormal{mol/m}^3)$',
+               r'$T_b \ (K)$']
+
+    xlabels = [r'$H_r \ (\textnormal{m})$',
+               r'$C_{Ar} \ (\textnormal{mol/m}^3)$', 
+               r'$C_{Br} \ (\textnormal{mol/m}^3)$',
+               r'$C_{Cr} \ (\textnormal{mol/m}^3)$',
+               r'$T_r \ (K)$',
+               r'$H_b \ (\textnormal{m})$',
+               r'$C_{Ab} \ (\textnormal{mol/m}^3)$',
+               r'$C_{Bb} \ (\textnormal{mol/m}^3)$',
+               r'$C_{Cb} \ (\textnormal{mol/m}^3)$',
+               r'$T_b \ (K)$']
+
+    ulabels = [r'$F \ (\textnormal{m}^3/\textnormal{min})$',
+               r'$D \ (\textnormal{m}^3/\textnormal{min})$']
+
+    def plot_inputs(t, ulist, figure_size, ylabel_xcoordinate, 
+                    data_type, legend_names, legend_colors, 
+                    title_loc):
+        """ Plot the training input data. """
+        nrow = len(CstrFlashPlots.ulabels)
+        (figure, axes) = plt.subplots(nrows=nrow, ncols=1,
+                                      sharex=True, figsize=figure_size,
+                                      gridspec_kw=dict(wspace=0.5))
+        legend_handles = []
+        for (u, color) in zip(ulist, legend_colors):
+            for row in range(nrow):
+                handle = axes[row].step(t, u[:, row], color, where='post')
+                axes[row].set_ylabel(CstrFlashPlots.ulabels[row])
[…]

### 2021-03-10 tworeac_bbtrain.py

@@ -35,7 +35,7 @@ def main():
     xinsert_indices = [2]
     tthrow = 10
     Np = 2
-    hN_dims = [Np*(Ny+Nu), 32, 2]
+    hN_dims = [Np*(Ny+Nu), 16, 2]
 
     # Create lists to store data.
     trained_weights = []
@@ -64,7 +64,7 @@ def main():
                              outputs=train_data['outputs'])
 
         # Train.
-        train_bbmodel(model=model, epochs=10000, batch_size=3, 
+        train_bbmodel(model=model, epochs=10, batch_size=3, 
                       train_data=train_samples, trainval_data=trainval_data, 
                       stdout_filename=stdout_filename, ckpt_path=ckpt_path)
 

### 2021-03-10 tworeac_empc.py

@@ -91,16 +91,6 @@ def get_controller(model_func, model_pars, model_type,
                                    Qwx=Qwx, Qwd=Qwd, Rv=Rv, Nmhe=Nmhe,
                                    guess=regulator_guess), hx
 
-def stage_cost(y, u, p):
-    """ Custom stage cost for the tworeac system. """
-    # Get inputs, parameters, and states.
-    CAf = u[0:1]
-    ca, cb = p[0:2]
-    CA, CB = y[0:2]
-    # Compute and return cost.
-    return (CB - 1.2)**2
-    # return ca*CAf - cb*CB
-
 def get_mhe_noise_tuning(model_type, model_par):
     # Get MHE tuning.
     if model_type == 'plant' or model_type =='koopman':

### 2021-03-13 sysID.tex
writing updates for the smaller example.

@@
 \section{Process modeling}
-We consider the problem of system identification of the following nonlinear plant.
+We consider the problem of system identification of the following nonlinear plant:
 \begin{align*}
@@
 
-in which $\theta_g$ and $x_g^i(0)$ are the decision variables in the optimization problem, which are parameters in the grey-box model and initial state used for forecasting the predicted measurements.
+in which $\theta_g$ and $x_g^i(0)$ are the decision variables in the optimization problem, which are parameters in the grey-box model and initial state used for forecasting the predicted measurements respectively.
 We assume that multiple trajectories can be used in the training data, and the superscript $i$ is used to denote the trajectory number.
@@
 Hence, grey-box system identification and subsequent use of the grey-box model in real-time optimization can lead to suboptimal economic performance.
-We propose to augment the grey-box model using a neural network (NN) such that the overall hybrid model can capture any unknown process dynamics in the plant, therefore leading to improved economic performance.
+We propose to augment the grey-box model using a neural network (NN) such that the overall hybrid model can capture any unknown process dynamics in the plant, therefore leading to improved economic performance when used subsequently in real-time optimization.
 We define this hybrid model as follows:
@@
 The stochastic gradient descent algorithm Adam is used to solve the above optimization using the software tensorflow.
-Note that we do not estimate the initial state for each trajectory, and fix those states heuristically to make forecasts during the optimization process.
+Note that we do not estimate the initial grey-box state for each trajectory in the optimization problem, and fix those states heuristically to make forecasts during the optimization process.
 The control inputs and measurements are scaled as $u := (u - u_{\textnormal{MEAN}})/u_{\textnormal{STD}}$ and
@@
 The black-box NN uses the vector of past measurements and control inputs $z$ \eqref{eq:pastyu} as the state in the dynamic model and is structured as follows
-\begin{align*}
+\begin{align} \label{eq:bbnn_model}
   z^+ &= f_z(z, u) \\ 
-  y &= h_N(z, \theta_N)
-\end{align*}
+  y &= h_N(z, \theta_N) \nonumber
+\end{align}
 in which $h_N(\cdot)$ is a NN and $\theta_N$ are the parameters in the network. The function $f_z(\cdot)$ denotes the dynamics of the state $z$ and is defined as
@@
 
+The Koopman model represents the dynamics as a linear system in a high-dimensional state-space that is a nonlinear transformation of the original state-space.
+This transformation is chosen such that the input-output behavior is linear.
+The overall dynamic model is represented as:
+
+\begin{align} \label{eq:koop_model}
+  x_{kp} &= [y', z', f_N(y, z, \theta_N)']' \\
+  x_{kp}^+ &= Ax_{kp} + Bu \nonumber \\
+  y &= [I, \ 0] x_{kp} \nonumber
+\end{align}
+in which $x_{kp} \in \bbR^N$ is the state in a high-dimensional state-space, $f_N(\cdot)$ is a NN used for the nonlinear transformation, and $\theta_N$ denotes the parameters in the network.
+
 \subsubsection{Training}
-The training problem is to find the parameters in the neural network
-$f_N$ and the linear model matrices $A$ and $B$ by minimizing the following mean squared error:
+The training optimization problem for both the black-box NN and the Koopman operator model is to minimize the mean-squared-error (MSE) similar to the nonlinear program \eqref{eq:hybrid_training}.
+The model equality constraints are replaced with the equations \eqref{eq:bbnn_model} and \eqref{eq:koop_model} for the black-box and Koopman operator model respectively.
+The measurements and control inputs are scaled using the mean and standard deviation computed from the training data.
+The initial state to make model predictions in the optimization process is set based on a window of past measurements and control inputs in the training data.
 
+\section{Two reaction system example}
+
+We next study the performance of the modeling frameworks on a simple two reaction system example in three steps.
+First, we examine the predictive capability of the developed models on open-loop validation data.
+Second, we analyze the steady-state optimum with an economic stage cost.
+Finally, we analyze the optimal control input profiles on a dynamic economic optimization problem.
+These three steps shed light on the predictive abilities of the models and their usefulness for real-time optimization in the operation of chemical processes.
+
+The following plant model with two reactions $A \rightarrow B$ and $3B \rightleftharpoons C$ is considered:
 \begin{align*}
-  \underset{A, B, W_i, b_i}{\textnormal{min}} \sum_{k=0}^{N_t-1} 
-  \dfrac{1}{N_t}(\hat{x}(k) - x(k))^2
+  \dfrac{dC_A}{dt} &= \dfrac{C_{Af} - C_A}{\tau} - k_1C_A\\
+  \dfrac{dC_B}{dt} &= k_1C_A - 3k_2C^3_B + 3k_3C_C- \dfrac{C_B}{\tau}\\
+  \dfrac{dC_C}{dt} &= k_2C^3_B - k_3C_C - \dfrac{C_C}{\tau}
 \end{align*}
 
-\section{Two reaction system plant model}
+in which, $C_A$, $C_B$, and $C_C$ are the concentrations of the three species $A$, $B$, and $C$. The parameters $k_1$, $k_2$, and $k_3$ are the rate constants and the $\tau$ is the residence time.
 
-\section{Model validation}
+The grey-box model is built as follows based on an assumption that the knowledge of the second side reaction ($3B \rightleftharpoons C $) is not available:
+\begin{align*}
+  \dfrac{dC_A}{dt} &= \dfrac{C_{Af} - C_A}{\tau} - k_1C_A\\
+  \dfrac{dC_B}{dt} &= k_1C_A - \dfrac{C_B}{\tau}
+\end{align*}
 
-\section{Steady state optimization}
+The rate constants and residence time in both the models are set to 
+$k_1 = 1 \ \textnormal{m}^3/\textnormal{min}$, $k_2 = 0.01 \ \textnormal{m}^3/\textnormal{min}$, $k_3 = 0.05 \ \textnormal{m}^3/\textnormal{min}$, and $\tau = 5 \ \textnormal{min}$.
+The sample time for the measurements is $1 \ \textnormal{min}$, and we assume that only the concentrations of the species $A$ and $B$ are measured in real-time.
 
-\section{Dynamic economic optimization}
+\subsection{Model validation}
+In the simulations for this example, we fix the grey-box model parameters heuristically and only train the two black-box models: NN and Koopman operator model.
+We generate 1 day (1440 samples) of data by simulating the plant model with a Pseudo-random-binary-signal (PRBS) control input profile.
+The generated data is split into 12 hours of training data, 6 hours of buffer data, and 6 hours of validation data.
+The MSE metric is monitored on the buffer data during the training process, and model weights are updated only when MSE decreases on the buffer dataset.
+After the training step, the black-box NN and Koopman operator models are used to make predictions on the validation data shown in Figure \ref{fig:tworeac_validation}.
+We observe good quality measurement predictions with both types of dynamic black-box models.
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[page=1, height=0.5\textheight, 
+                   width=0.7\textwidth]{tworeac_plots.pdf}
+  \caption{Model validation on open-loop data with the black-box NN and Koopman operator model.}
+  \label{fig:tworeac_validation}
+\end{figure}
+
+\subsection{Steady state optimization}
+The two trained black-box NN and Koopman operator model are used in a steady-state optimization problem to determine the economically optimal
+steady state of the system.
+In a typical chemical plant, this type of optimization is solved periodically every few hours based on changes in raw material costs and product selling prices.
+The obtained steady state is sent to lower-level control layers such as model predictive control (MPC) or the proportional-integral-derivative (PID) controller.
+The optimization problem solved is:
+\begin{align} \label{eq:ss_opt}
+  \underset{u_s}{\textnormal{min}} & \quad \ell(y_s, u_s) \\
+  x_s &= f(x_s, u_s) \nonumber \\ 
+  y_s &= h(x_s) \nonumber \\
+  \underline{u} \leq &u_s \leq \overline{u} \nonumber
+\end{align}
+in which the steady-state control input ($u_s$) is the decision variable,  and the model is required to satisfy steady-state constraints. The stage cost for the two reaction system is chosen as:
+\begin{equation*}
+  \ell(y_s, u_s) = p_1C_{Afs} - p_2C_{Bs}
+\end{equation*}
+in which $p_1$ and $p_2$ are the parameters chosen to represent the raw material (species $A$) cost and the product (species $B$) selling price respectively.
+
+First, we plot the economic objective ($\ell(y_s, u_s)$) as a function of the 
+steady-state control input to examine the cost structure obtained with the different models. This plot is shown in Figure \ref{fig:tworeac_sscost}.
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[page=1, height=0.4\textheight, 
+                   width=0.6\textwidth]{tworeac_ssopt.pdf}
+  \caption{Steady-state cost curve $\ell(y_s, u_s)$ as a function of the control input for the plant, grey-box, black-box NN, and Koopman operator models.}
+  \label{fig:tworeac_sscost}
+\end{figure}
+
+We observe that the black-box NN model correctly captures the true plant cost curve after training, and the grey-box and Koopman operator models do not.
+This behavior by the latter two models is due to the linear structure in the dynamic models.
+Upon solving the optimization problem \eqref{eq:ss_opt}, the optimum steady-state control input ($C_{Afs}^0$) values are $1.63$, $2.5$, $1.60$, and $2.5$ in mol/m$^3$ for the plant, grey-box, black-box NN, and Koopman operator models respectively.
+The use of grey-box and Koopman operator models subsequently in real-time optimization of chemical processes can therefore result in suboptimal economic performance.
+
+\subsection{Dynamic economic optimization}
+The performance of the models is now studied with a dynamic economic optimization problem.
+In contrast to the steady-state optimization problem \eqref{eq:ss_opt}, operational strategies based on solving a dynamic optimization problem aim to maximize plant profit during the transients as well as opposed to just steady states.
+The economic optimization problem is:
+
+\begin{align} \label{eq:dyn_opt}
+  \underset{\useq}{\textnormal{min}} & \quad \sum_{k=0}^{N-1} \ell(y(i), u(i)) \\
+  x^+ &= f(x, u) \nonumber \\ 
+  y &= h(x) \nonumber \\
+  x(0) &= x_0 \nonumber \\ 
+  \underline{u} \leq &u \leq \overline{u} \nonumber
+\end{align}
+
+The above optimization problem is similar to the optimization solved in MPC, and the difference is the choice of the economic stage cost $\ell(\cdot)$, which is a setpoint tracking stage cost in a traditional MPC controller.
+This optimization problem can also be solved directly at the regulatory control layers and is often referred to as the economic MPC controller.
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[page=1, height=0.5\textheight, 
+                   width=0.7\textwidth]{tworeac_openloop.pdf}
+  \caption{Optimal open-loop control input and state profiles after solving the dynamic economic MPC problem \eqref{eq:dyn_opt} with the plant, grey-box, black-box NN, and Koopman operator models.}
+  \label{fig:tworeac_openloop}
+\end{figure}
+
+Before analyzing a full closed-loop simulation with the economic MPC controllers based on the different dynamic models, we first examine the open-loop solutions of the optimization problem \eqref{eq:dyn_opt} given one initial state.
+Figure \ref{fig:tworeac_openloop} shows the open-loop measurement and control input profiles for the plant, grey-box, black-box NN, and Koopman operator model.
+We notice that the black-box NN has an optimal control input profile qualitatively similar to that of the plant model.
+Whereas, the optimal profiles for the grey-box and Koopman operator model
+are different than the plant model. This simple analysis shows that the use of the grey-box and Koopman operator models in MPC can result in suboptimal performance.
+
+\section{Large, CSTR and Flash system example}
+
+\subsection{Model validation}
+
+\subsection{Dynamic economic optimization}
 

### 2021-03-14 sysID.tex
write out the model details.

@@
 \usepackage{float}
+\usepackage{units}
 \usepackage[round]{natbib}
@@
 \section{Process modeling}
-We consider the problem of system identification of the following nonlinear plant:
+We consider the problem of system identification of the following nonlinear
+plant:
 \begin{align*}
@@
 \end{align*}
-in which $x_p \in \bbR^{n_p}$ is the plant state, $u \in \bbU \subset \bbR^m$ is the
-manipulated control input, $y \in \bbR^p$ is the measurement, $w$ is the process
-noise, and $v$ is the measurement noise.
-The objective of developing the model is to use it subsequently in model predictive control (MPC).
-Therefore, we consider several types of grey-box, hybrid, and black-box models and analyze their predictive capabilities as well as their usefulness for online optimization in MPC.
+in which $x_p \in \bbR^{n_p}$ is the plant state, $u \in \bbU \subset \bbR^m$ is
+the manipulated control input, $y \in \bbR^p$ is the measurement, $w$ is the
+process noise, and $v$ is the measurement noise. The objective of developing the
+model is to use it subsequently in model predictive control (MPC). Therefore, we
+consider several types of grey-box, hybrid, and black-box models and analyze
+their predictive capabilities as well as their usefulness for online
+optimization in MPC.
 
 \section{Model architectures for system identification}
-We describe several process modeling frameworks which determine parameters in their proposed model architecture by minimizing the multi-step ahead prediction error of the model on training data.
-A nonlinear optimization problem is solved for this identification step.
-The estimated parameters define the model, which is then used subsequently to perform model validation and for use in MPC.
+We describe several process modeling frameworks which determine parameters in
+their proposed model architecture by minimizing the multi-step ahead prediction
+error of the model on training data. A nonlinear optimization problem is solved
+for this identification step. The estimated parameters define the model, which
+is then used subsequently to perform model validation and for use in MPC.
 
 \subsection{Grey-box model}
-We first consider the grey-box modeling strategy, which is the current industrial state-of-the-art system identification technique.
-A parameterized model is defined based on first-principle knowledge about the plant, and parameters in the model are estimated from training data by solving the prediction-error minimization problem.
-The grey-box model is typically defined in continuous time and is of the following form:
+We first consider the grey-box modeling strategy, which is the current
+industrial state-of-the-art system identification technique. A parameterized
+model is defined based on first-principle knowledge about the plant, and
+parameters in the model are estimated from training data by solving the
+prediction-error minimization problem. The grey-box model is typically defined
+in continuous time and is of the following form:
 
@@
 
-The following nonlinear program (NLP) is solved to identify the parameters $\theta_g$:
+The following nonlinear program (NLP) is solved to identify the parameters
+$\theta_g$:
 
@@
 
-in which $\theta_g$ and $x_g^i(0)$ are the decision variables in the optimization problem, which are parameters in the grey-box model and initial state used for forecasting the predicted measurements respectively.
-We assume that multiple trajectories can be used in the training data, and the superscript $i$ is used to denote the trajectory number.
-We use $N_{tr}$ to denote the total number of trajectories in training data and $N_t$ to denote the number of samples in each trajectory.
-The function $f_{g}(\cdot)$ is discretized at the measurement sample time to obtain the function $f_{gd}(\cdot)$ that is used to make forecasts in the above optimization.
+in which $\theta_g$ and $x_g^i(0)$ are the decision variables in the
+optimization problem, which are parameters in the grey-box model and initial
+state used for forecasting the predicted measurements respectively. We assume
+that multiple trajectories can be used in the training data, and the superscript
+$i$ is used to denote the trajectory number. We use $N_{tr}$ to denote the total
+number of trajectories in training data and $N_t$ to denote the number of
+samples in each trajectory. The function $f_{g}(\cdot)$ is discretized at the
+measurement sample time to obtain the function $f_{gd}(\cdot)$ that is used to
+make forecasts in the above optimization.
 
@@
 
-Often in industrial applications, first-principle knowledge about the plant is incomplete.
-Hence, grey-box system identification and subsequent use of the grey-box model in real-time optimization can lead to suboptimal economic performance.
-We propose to augment the grey-box model using a neural network (NN) such that the overall hybrid model can capture any unknown process dynamics in the plant, therefore leading to improved economic performance when used subsequently in real-time optimization.
-We define this hybrid model as follows:
+Often in industrial applications, first-principle knowledge about the plant is
+incomplete. Hence, grey-box system identification and subsequent use of the
+grey-box model in real-time optimization can lead to suboptimal economic
+performance. We propose to augment the grey-box model using a neural network
+(NN) such that the overall hybrid model can capture any unknown process dynamics
+in the plant, therefore leading to improved economic performance when used
+subsequently in real-time optimization. We define this hybrid model as follows:
 
@@
 \end{align*}
-in which $f_N(\cdot)$ is a NN and $\theta_N$ denotes the parameters in the network.
-We use $z$ to denote the vector of past measurements and control inputs.
-This vector is considered as an additional state in the overall dynamic model and is defined as follows:
+in which $f_N(\cdot)$ is a NN and $\theta_N$ denotes the parameters in the
+network. We use $z$ to denote the vector of past measurements and control
+inputs. This vector is considered as an additional state in the overall dynamic
+model and is defined as follows:
 \begin{align} \label{eq:pastyu}
@@
 \end{align}
-in which $N_p$ is the number of past measurements and control inputs used. The function $f_z(\cdot)$ describes the dynamics of the state $z$ and its structure is defined as follows: 
+in which $N_p$ is the number of past measurements and control inputs used. The
+function $f_z(\cdot)$ describes the dynamics of the state $z$ and its structure
+is defined as follows: 
 \begin{align*}
@@
 \end{align*}
-We assume that the parameters in the grey-box model are first estimated by solving the NLP \eqref{eq:gb_training}, and the following similar NLP is solved to determine parameters ($\theta_N$) in the NN.
+We assume that the parameters in the grey-box model are first estimated by
+solving the NLP \eqref{eq:gb_training}, and the following similar NLP is solved
+to determine parameters ($\theta_N$) in the NN.
 
@@
 \end{align}
-The stochastic gradient descent algorithm Adam is used to solve the above optimization using the software tensorflow.
-Note that we do not estimate the initial grey-box state for each trajectory in the optimization problem, and fix those states heuristically to make forecasts during the optimization process.
-The control inputs and measurements are scaled as $u := (u - u_{\textnormal{MEAN}})/u_{\textnormal{STD}}$ and
-$y := (y - y_{\textnormal{MEAN}})/y_{\textnormal{STD}}$, in which $u_{\textnormal{MEAN}}$ and $y_{\textnormal{MEAN}}$ are the respective means computed on training data.
-Similarly, $u_{\textnormal{STD}}$ and $y_{\textnormal{STD}}$ are the respective standard deviations.
-Due to this scaling, we do not consider the penalty $R_v^{-1}$ in the training objective.
+The stochastic gradient descent algorithm Adam is used to solve the above
+optimization using the software tensorflow. Note that we do not estimate the
+initial grey-box state for each trajectory in the optimization problem, and fix
+those states heuristically to make forecasts during the optimization process.
+The control inputs and measurements are scaled as $u := (u -
+u_{\textnormal{MEAN}})/u_{\textnormal{STD}}$ and $y := (y -
+y_{\textnormal{MEAN}})/y_{\textnormal{STD}}$, in which $u_{\textnormal{MEAN}}$
+and $y_{\textnormal{MEAN}}$ are the respective means computed on training data.
+Similarly, $u_{\textnormal{STD}}$ and $y_{\textnormal{STD}}$ are the respective
+standard deviations. Due to this scaling, we do not consider the penalty
+$R_v^{-1}$ in the training objective.
 
 \subsection{Black-box models}
-Next, we consider black-box system identification strategies which do not impose any a-prior knowledge about the plant and develop the dynamic model purely from data.
-We consider two types of black-box models in this section: NNs and dynamic models based on the Koopman operator theory.
-The latter type of model is considered based on the motivation that it imposes a linear structure on the dynamics in a high-dimensional space, therefore leading to a linear or quadratic program to be solved online in MPC.
+Next, we consider black-box system identification strategies which do not impose
+any a-prior knowledge about the plant and develop the dynamic model purely from
+data. We consider two types of black-box models in this section: NNs and dynamic
+models based on the Koopman operator theory. The latter type of model is
+considered based on the motivation that it imposes a linear structure on the
+dynamics in a high-dimensional space, therefore leading to a linear or quadratic
+program to be solved online in MPC.
 
 \subsubsection{Neural network}
-The black-box NN uses the vector of past measurements and control inputs $z$ \eqref{eq:pastyu} as the state in the dynamic model and is structured as follows
+The black-box NN uses the vector of past measurements and control inputs $z$
+\eqref{eq:pastyu} as the state in the dynamic model and is structured as follows
 \begin{align} \label{eq:bbnn_model}
@@
 \end{align}
-in which $h_N(\cdot)$ is a NN and $\theta_N$ are the parameters in the network. The function $f_z(\cdot)$ denotes the dynamics of the state $z$ and is defined as
+in which $h_N(\cdot)$ is a NN and $\theta_N$ are the parameters in the network.
+The function $f_z(\cdot)$ denotes the dynamics of the state $z$ and is defined
+as
 
@@
 
-The Koopman model represents the dynamics as a linear system in a high-dimensional state-space that is a nonlinear transformation of the original state-space.
-This transformation is chosen such that the input-output behavior is linear.
-The overall dynamic model is represented as:
+The Koopman model represents the dynamics as a linear system in a
+high-dimensional state-space that is a nonlinear transformation of the original
+state-space. This transformation is chosen such that the input-output behavior
+is linear. The overall dynamic model is represented as:
 
@@
 \end{align}
-in which $x_{kp} \in \bbR^N$ is the state in a high-dimensional state-space, $f_N(\cdot)$ is a NN used for the nonlinear transformation, and $\theta_N$ denotes the parameters in the network.
+in which $x_{kp} \in \bbR^N$ is the state in a high-dimensional state-space,
+$f_N(\cdot)$ is a NN used for the nonlinear transformation, and $\theta_N$
+denotes the parameters in the network.
 
 \subsubsection{Training}
-The training optimization problem for both the black-box NN and the Koopman operator model is to minimize the mean-squared-error (MSE) similar to the nonlinear program \eqref{eq:hybrid_training}.
-The model equality constraints are replaced with the equations \eqref{eq:bbnn_model} and \eqref{eq:koop_model} for the black-box and Koopman operator model respectively.
-The measurements and control inputs are scaled using the mean and standard deviation computed from the training data.
-The initial state to make model predictions in the optimization process is set based on a window of past measurements and control inputs in the training data.
+The training optimization problem for both the black-box NN and the Koopman
+operator model is to minimize the mean-squared-error (MSE) similar to the
+nonlinear program \eqref{eq:hybrid_training}. The model equality constraints are
+replaced with the equations \eqref{eq:bbnn_model} and \eqref{eq:koop_model} for
+the black-box and Koopman operator model respectively. The measurements and
+control inputs are scaled using the mean and standard deviation computed from
+the training data. The initial state to make model predictions in the
+optimization process is set based on a window of past measurements and control
+inputs in the training data.
 
@@
 
-We next study the performance of the modeling frameworks on a simple two reaction system example in three steps.
-First, we examine the predictive capability of the developed models on open-loop validation data.
-Second, we analyze the steady-state optimum with an economic stage cost.
-Finally, we analyze the optimal control input profiles on a dynamic economic optimization problem.
-These three steps shed light on the predictive abilities of the models and their usefulness for real-time optimization in the operation of chemical processes.
+We now study the performance of the modeling frameworks on a simple two
+reaction system example in three steps. First, we examine the predictive
+capability of the developed models on open-loop validation data. Second, we
+analyze the steady-state optimum with an economic stage cost. Finally, we
+analyze the optimal control input profiles on a dynamic economic optimization
+problem. These three steps shed light on the predictive abilities of the models
+and their usefulness for real-time optimization in the operation of chemical
+processes.
 
-The following plant model with two reactions $A \rightarrow B$ and $3B \rightleftharpoons C$ is considered:
+The following plant model with two reactions $A \rightarrow B$ and $3B
+\rightleftharpoons C$ is considered:
 \begin{align*}
@@
 
-in which, $C_A$, $C_B$, and $C_C$ are the concentrations of the three species $A$, $B$, and $C$. The parameters $k_1$, $k_2$, and $k_3$ are the rate constants and the $\tau$ is the residence time.
+in which, $C_A$, $C_B$, and $C_C$ are the concentrations of the three species
+$A$, $B$, and $C$. The parameters $k_1$, $k_2$, and $k_3$ are the rate constants
+and the $\tau$ is the residence time.
 
-The grey-box model is built as follows based on an assumption that the knowledge of the second side reaction ($3B \rightleftharpoons C $) is not available:
+The grey-box model is built as follows based on an assumption that the knowledge
+of the second side reaction ($3B \rightleftharpoons C $) is not available:
 \begin{align*}
@@
 
-The rate constants and residence time in both the models are set to 
-$k_1 = 1 \ \textnormal{m}^3/\textnormal{min}$, $k_2 = 0.01 \ \textnormal{m}^3/\textnormal{min}$, $k_3 = 0.05 \ \textnormal{m}^3/\textnormal{min}$, and $\tau = 5 \ \textnormal{min}$.
-The sample time for the measurements is $1 \ \textnormal{min}$, and we assume that only the concentrations of the species $A$ and $B$ are measured in real-time.
+The rate constants and residence time in both the models are set to $k_1 = 1 \
+\textnormal{m}^3/\textnormal{min}$, $k_2 = 0.01 \
+\textnormal{m}^3/\textnormal{min}$, $k_3 = 0.05 \
+\textnormal{m}^3/\textnormal{min}$, and $\tau = 5 \ \textnormal{min}$. The
+sample time for the measurements is $1 \ \textnormal{min}$, and we assume that
+only the concentrations of the species $A$ and $B$ are measured in real-time.
 
 \subsection{Model validation}
-In the simulations for this example, we fix the grey-box model parameters heuristically and only train the two black-box models: NN and Koopman operator model.
-We generate 1 day (1440 samples) of data by simulating the plant model with a Pseudo-random-binary-signal (PRBS) control input profile.
-The generated data is split into 12 hours of training data, 6 hours of buffer data, and 6 hours of validation data.
-The MSE metric is monitored on the buffer data during the training process, and model weights are updated only when MSE decreases on the buffer dataset.
-After the training step, the black-box NN and Koopman operator models are used to make predictions on the validation data shown in Figure \ref{fig:tworeac_validation}.
-We observe good quality measurement predictions with both types of dynamic black-box models.
+In the simulations for this example, we fix the grey-box model parameters
+heuristically and only train the two black-box models: NN and Koopman operator
+model. We generate 1 day (1440 samples) of data by simulating the plant model
+with a Pseudo-random-binary-signal (PRBS) control input profile. The generated
+data is split into 12 hours of training data, 6 hours of buffer data, and 6
+hours of validation data. The MSE metric is monitored on the buffer data during
+the training process, and model weights are updated only when MSE decreases on
+the buffer dataset. After the training step, the black-box NN and Koopman
+operator models are used to make predictions on the validation data shown in
+Figure \ref{fig:tworeac_validation}. We observe good quality measurement
+predictions with both types of dynamic black-box models.
 
@@
 \subsection{Steady state optimization}
-The two trained black-box NN and Koopman operator model are used in a steady-state optimization problem to determine the economically optimal
-steady state of the system.
-In a typical chemical plant, this type of optimization is solved periodically every few hours based on changes in raw material costs and product selling prices.
-The obtained steady state is sent to lower-level control layers such as model predictive control (MPC) or the proportional-integral-derivative (PID) controller.
-The optimization problem solved is:
+The two trained black-box NN and Koopman operator model are used in a
+steady-state optimization problem to determine the economically optimal steady
+state of the system. In a typical chemical plant, this type of optimization is
+solved periodically every few hours based on changes in raw material costs and
+product selling prices. The obtained steady state is sent to lower-level control
+layers such as model predictive control (MPC) or the
+proportional-integral-derivative (PID) controller. The optimization problem
+solved is:
 \begin{align} \label{eq:ss_opt}
@@
 \end{align}
-in which the steady-state control input ($u_s$) is the decision variable,  and the model is required to satisfy steady-state constraints. The stage cost for the two reaction system is chosen as:
+in which the steady-state control input ($u_s$) is the decision variable,  and
+the model is required to satisfy steady-state constraints. The stage cost for
+the two reaction system is chosen as:
 \begin{equation*}
@@
 \end{equation*}
-in which $p_1$ and $p_2$ are the parameters chosen to represent the raw material (species $A$) cost and the product (species $B$) selling price respectively.
+in which $p_1$ and $p_2$ are the parameters chosen to represent the raw material
+(species $A$) cost and the product (species $B$) selling price respectively.
 
 First, we plot the economic objective ($\ell(y_s, u_s)$) as a function of the
-steady-state control input to examine the cost structure obtained with the different models. This plot is shown in Figure \ref{fig:tworeac_sscost}.
+steady-state control input to examine the cost structure obtained with the
+different models. This plot is shown in Figure \ref{fig:tworeac_sscost}.
 
@@
 
-We observe that the black-box NN model correctly captures the true plant cost curve after training, and the grey-box and Koopman operator models do not.
-This behavior by the latter two models is due to the linear structure in the dynamic models.
-Upon solving the optimization problem \eqref{eq:ss_opt}, the optimum steady-state control input ($C_{Afs}^0$) values are $1.63$, $2.5$, $1.60$, and $2.5$ in mol/m$^3$ for the plant, grey-box, black-box NN, and Koopman operator models respectively.
-The use of grey-box and Koopman operator models subsequently in real-time optimization of chemical processes can therefore result in suboptimal economic performance.
+We observe that the black-box NN model correctly captures the true plant cost
+curve after training, and the grey-box and Koopman operator models do not. This
+behavior by the latter two models is due to the linear structure in the dynamic
+models. Upon solving the optimization problem \eqref{eq:ss_opt}, the optimum
+steady-state control input ($C_{Afs}^0$) values are $1.63$, $2.5$, $1.60$, and
+$2.5$ in mol/m$^3$ for the plant, grey-box, black-box NN, and Koopman operator
+models respectively. The use of grey-box and Koopman operator models
+subsequently in real-time optimization of chemical processes can therefore
+result in suboptimal economic performance.
 
 \subsection{Dynamic economic optimization}
-The performance of the models is now studied with a dynamic economic optimization problem.
-In contrast to the steady-state optimization problem \eqref{eq:ss_opt}, operational strategies based on solving a dynamic optimization problem aim to maximize plant profit during the transients as well as opposed to just steady states.
-The economic optimization problem is:
+The performance of the models is now studied with a dynamic economic
+optimization problem. In contrast to the steady-state optimization problem
+\eqref{eq:ss_opt}, operational strategies based on solving a dynamic
+optimization problem aim to maximize plant profit during the transients as well
+as opposed to just steady states. The economic optimization problem is:
 
@@
 
-The above optimization problem is similar to the optimization solved in MPC, and the difference is the choice of the economic stage cost $\ell(\cdot)$, which is a setpoint tracking stage cost in a traditional MPC controller.
-This optimization problem can also be solved directly at the regulatory control layers and is often referred to as the economic MPC controller.
+The above optimization problem is similar to the optimization solved in MPC, and
+the difference is the choice of the economic stage cost $\ell(\cdot)$, which is
+a setpoint tracking stage cost in a traditional MPC controller. This
+optimization problem can also be solved directly at the regulatory control
+layers and is often referred to as the economic MPC controller.
 
@@
   \includegraphics[page=1, height=0.5\textheight,
-                   width=0.7\textwidth]{tworeac_openloop.pdf}
-  \caption{Optimal open-loop control input and state profiles after solving the dynamic economic MPC problem \eqref{eq:dyn_opt} with the plant, grey-box, black-box NN, and Koopman operator models.}
+                   width=0.7\textwidth]{tworeac_openloop.pdf} \caption{Optimal
+                   open-loop control input and state profiles after solving the
+                   dynamic economic MPC problem \eqref{eq:dyn_opt} with the
+                   plant, grey-box, black-box NN, and Koopman operator models.}
   \label{fig:tworeac_openloop}
@@
 
-Before analyzing a full closed-loop simulation with the economic MPC controllers based on the different dynamic models, we first examine the open-loop solutions of the optimization problem \eqref{eq:dyn_opt} given one initial state.
-Figure \ref{fig:tworeac_openloop} shows the open-loop measurement and control input profiles for the plant, grey-box, black-box NN, and Koopman operator model.
-We notice that the black-box NN has an optimal control input profile qualitatively similar to that of the plant model.
-Whereas, the optimal profiles for the grey-box and Koopman operator model
-are different than the plant model. This simple analysis shows that the use of the grey-box and Koopman operator models in MPC can result in suboptimal performance.
+Before analyzing a full closed-loop simulation with the economic MPC controllers
+based on the different dynamic models, we first examine the open-loop solutions
+of the optimization problem \eqref{eq:dyn_opt} given one initial state. Figure
+\ref{fig:tworeac_openloop} shows the open-loop state and control input profiles
+for the plant, grey-box, black-box NN, and Koopman operator models. We notice
+that the black-box NN has an optimal control input profile qualitatively similar
+to that of the plant. Whereas, the optimal profiles for the grey-box and Koopman
+operator model are different than the plant model. This simple analysis shows
+that the use of the grey-box and Koopman operator models in MPC can result in
+suboptimal economic performance.
 
-\section{Large, CSTR and Flash system example}
+\section{Large, CSTR and flash system example}
+We next study the performance of the black-box NN modeling framework on a large chemical process that consists of a chemical reactor and flash separator in series.
+The process is depicted in Figure \ref{fig:cstr_flash}, and the ordinary differential equations (ODEs) with parameter values describing the plant model are shown in Appendix \ref{app:cstr_flash}.
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[page=1, height=0.3\textheight,
+                   width=0.6\textwidth]{cstr_flash.pdf}
+  \caption{Schematic of the CSTR and flash chemical process.}
+  \label{fig:cstr_flash}
+\end{figure}
 
@@
 
+\subsection{Steady state optimization}
+
 \subsection{Dynamic economic optimization}
 
-\section{Future work}
+\section{Discussion}
+
+\section*{Appendix} 
+\renewcommand{\thesubsection}{\Alph{subsection}}
+
+\subsection{CSTR and flash system model} \label{app:cstr_flash}
+
+\begin{align*}
+  \dfrac{dH_r}{dt} &= \dfrac{F + D -F_r}{A_r}\\
+  \dfrac{dC_{Ar}}{dt} &= \dfrac{F(C_{Af} -C_{Ar}) +
+                         D(C_{Ad} -C_{Ar})}{A_rH_r} - r_1 \\
+  \dfrac{dC_{Br}}{dt} &= \dfrac{-FC_{Br} + 
+                          D(C_{Bd} -C_{Br})}{A_rH_r} + r_1 -3r_2\\
+  \dfrac{dC_{Cr}}{dt} &= \dfrac{-FC_{Cr} + 
+  D(C_{Cd} -C_{Cr})}{A_rH_r} + r_2\\
+  \dfrac{dT_r}{dt} &= \dfrac{F(T_f - T_r) + D(T_d -T_r)}{A_rH_r} + 
+                      \dfrac{r_1\Delta H_1 + r_2\Delta H_2}{\rho C_p} - 
+                      \dfrac{Q_r}{\rho A_r C_p H_r}\\
+\end{align*}
+
+\begin{align*}
+  \dfrac{dH_b}{dt} &= \dfrac{F_r - F_b - D}{A_b} \\
+  \dfrac{dC_{Ab}}{dt} &= \dfrac{F_r(C_{Ar} -C_{Ab}) + 
+                          D(C_{Ab} -C_{Ad})}{A_bH_b} \\
+  \dfrac{dC_{Bb}}{dt} &= \dfrac{F_r(C_{Br} -C_{Bb}) + 
+                          D(C_{Bb} -C_{Bd})}{A_bH_b} \\
+  \dfrac{dC_{Cb}}{dt} &= \dfrac{F_r(C_{Cr} -C_{Cb}) + 
+                          D(C_{Cb} -C_{Cd})}{A_bH_b} \\
+  \dfrac{dT_b}{dt} &= \dfrac{F_r(T_r - T_b)}{A_bH_b} +
+                      \dfrac{Q_b}{\rho A_b C_p H_b}\\
+\end{align*}
+
+Here, the rate of reactions $r_1$ and $r_2$ are defined as:
+\begin{align*}
+  r_1 = k_1^*e^{-E_1/RT_r}C_{Ar}, \quad r_2 = k_2^*e^{-E_2/RT_r}C_{Br}^3
+\end{align*}
+
+The outlet flowrates from the CSTR and flash are computed from the height of the reaction mixture in each unit as follows:
+\begin{align*}
+  F_r = k_r\sqrt{H_r}, \quad F_b = k_b\sqrt{H_b}
+\end{align*}
+
+The concentrations of the species in the recycle stream are computed using the following equations:
+\begin{align*}
+  C_{Ad} = \dfrac{\alpha_A C_{Ab}}{\sum_{i \in \{ A, B, C\}} \alpha_iC_{ib}} , \quad, C_{Bd} = \dfrac{\alpha_B C_{Bb}}{\sum_{i \in \{ A, B, C\}} \alpha_iC_{ib}}, \quad C_{Cd} = \dfrac{\alpha_C C_{Cb}}{\sum_{i \in \{ A, B, C\}} \alpha_iC_{ib}}
+\end{align*}
+
+The parameters in this model and the control input constraints used for the optimization problems are summarized in Table \ref{table:cstr_flash_pars}.
+
+\begin{table}[!h]
+  \centering
+  \begin{tabular}{|c|c|c|c|c|c|}
+      \hline
+      \multicolumn{6}{ |c| }{Parameters used to simulate the ODEs} \\
+      \hline
+      Parameter & Value & Unit & Parameter & Value & Unit\\
+      \hline
+      $\alpha_A$ & 8 &  & $k_1^*$ & 0.3 & \unit{min$^{-1}$}\\ 
+      $\alpha_B$ & 1 &  & $k_2^*$ & 0.5 & \unit{min$^{-1}$} \\ 
+      $\alpha_C$ & 1 &  & $\Delta H_1$ & 100 & \unitfrac{kJ}{mol} \\ 
+      $\rho$ & 6 & \unitfrac{kg}{m$^3$} & $\Delta H_2$ & 120 &
+      \unitfrac{kJ}{mol} \\ 
+      $C_p$ & 3 & \unitfrac{kJ}{kg-K} & $E_1/R$ & 200 & \unit{K} \\ 
+      $A_r$ & 3 & \unit{m$^2$} & $E_2/R$ & 300 & \unit{K} \\ 
+      $A_b$ & 3 & \unit{m$^2$} & $T_d$ & 310 &
+      \unit{K} \\ 
+      $k_r$ & 4 & \unitfrac{m$^{5/2}$}{min} & $C_{Af}$ & 6 & \unitfrac{mol}{m$^3$} \\ 
+      $k_b$ & 2 & \unitfrac{m$^{5/2}$}{min} &  $Q_r$ &  200 & \unitfrac{kJ}{min} \\ 
+      $T_f$ & 320 & \unit{K} &  & &  \\ 
+      $Q_b$ & 200 & \unitfrac{kJ}{min} & & & \\ 
+      \hline       
+      \multicolumn{6}{ |c| }{Actuator constraints ($\underline{u},
+      \overline{u}$)} \\
+      \hline
+      $F$ & (5, 15) & \unitfrac{m$^3$}{min} & $D$ & (2, 8) &
+      \unitfrac{m$^3$}{min}  \\ 
+      \hline
+  \end{tabular} \caption{Parameters used in the ODEs to simulate the plant in
+  the CSTR in series with a flash separator example, and the actuator constraints
+  for the optimization problems.}
+\label{table:cstr_flash_pars}
+\end{table}
 

### 2021-03-25 sysID.tex
OK, will do one more hybrid model.

@@
 The process is depicted in Figure \ref{fig:cstr_flash}, and the ordinary differential equations (ODEs) with parameter values describing the plant model are shown in Appendix \ref{app:cstr_flash}.
+The model consists of 12 states ($H_r$, $C_{Ar}$, $C_{Br}$, $C_{Cr}$, $T_r$, $H_b$, $C_{Ab}$, $C_{Bb}$, $C_{Cb}$, $T_b$), 6 measurements ($H_r$, $C_{Br}$, $T_r$, $H_b$, $C_{Bb}$, $T_b$), and 2 control inputs ($F$, $D$).
+The sample time for the measurements is 1 min.
 
@@
   \centering
-  \includegraphics[page=1, height=0.3\textheight,
-                   width=0.6\textwidth]{cstr_flash.pdf}
+  \includegraphics[page=1, height=0.5\textheight,
+                   width=0.7\textwidth]{cstr_flash.pdf}
   \caption{Schematic of the CSTR and flash chemical process.}
@@
 \subsection{Model validation}
+
+% \begin{figure}[!h]
+%   \centering
+%   \includegraphics[page=1, height=0.5\textheight,
+%                    width=0.7\textwidth]{cstr_flash_plots.pdf}
+%   \caption{Schematic of the CSTR and flash chemical process.}
+%   \label{fig:cstr_flash_uvalidation}
+% \end{figure}
+
+% \begin{figure}[!h]
+%   \centering
+%   \includegraphics[page=2, height=0.5\textheight,
+%                    width=0.7\textwidth]{cstr_flash_plots.pdf}
+%   \caption{Schematic of the CSTR and flash chemical process.}
+%   \label{fig:cstr_flash_yvalidation}
+% \end{figure}
 
@@
                       \dfrac{Q_r}{\rho A_r C_p H_r}\\
-\end{align*}
-
-\begin{align*}
   \dfrac{dH_b}{dt} &= \dfrac{F_r - F_b - D}{A_b} \\
@@
 
-The parameters in this model and the control input constraints used for the optimization problems are summarized in Table \ref{table:cstr_flash_pars}.
+The parameters used to simulate the model and the control input constraints used for the optimization problems are summarized in Table \ref{table:cstr_flash_pars}.
+
+\newpage
 

### 2021-03-25 cstr_flash_bbtrain.py

@@ -1,6 +1,5 @@
-# [depends] %LIB%/hybridid.py %LIB%/HybridModelLayers.py
-# [depends] %LIB%/../tworeac_nonlin_funcs.py
-# [depends] tworeac_parameters_nonlin.pickle
+# [depends] %LIB%/hybridid.py %LIB%/training_funcs.py
+# [depends] cstr_flash_parameters.pickle
 # [makes] pickle
 """ Script to train the hybrid model for the 
     three reaction system. 
@@ -67,7 +66,7 @@ def main():
                              outputs=train_data['outputs'])
 
         # Train.
-        train_bbmodel(model=model, epochs=30000, batch_size=12, 
+        train_bbmodel(model=model, epochs=30000, batch_size=18, 
                       train_data=train_samples, trainval_data=trainval_data, 
                       stdout_filename=stdout_filename, ckpt_path=ckpt_path)
 

### 2021-03-25 cstr_flash_chk_trained_funcs.py

@@ -1,8 +1,6 @@
-# [depends] %LIB%/hybridid.py %LIB%/linNonlinMPC.py
-# [depends] %LIB%/../tworeac_nonlin.py
-# [depends] tworeac_parameters.pickle
-# [depends] tworeac_train.pickle
-# [makes] pickle
+# [depends] %LIB%/hybridid.py %LIB%/economicopt.py
+# [depends] cstr_flash_parameters.pickle
+# [depends] cstr_flash_bbtrain.pickle
 """ Script to perform closed-loop simulations
     with the trained models.
     Pratyush Kumar, pratyushkumar@ucsb.edu """

### 2021-03-25 cstr_flash_openloop.py

@@ -1,8 +1,8 @@
 # [depends] %LIB%/hybridid.py %LIB%/linNonlinMPC.py
-# [depends] %LIB%/../tworeac_nonlin.py
-# [depends] tworeac_parameters.pickle
-# [depends] tworeac_train.pickle
-# [makes] pickle
+# [depends] %LIB%/cstr_flash_funcs.py %LIB%/economicopt.py
+# [depends] %LIB%/plotting_funcs.py
+# [depends] cstr_flash_parameters.pickle
+# [depends] cstr_flash_bbtrain.pickle
 """ Script to perform closed-loop simulations
     with the trained models.
     Pratyush Kumar, pratyushkumar@ucsb.edu """
@@ -36,7 +36,7 @@ def get_openloop_sol(fxu, hx, model_pars, xuguess):
     Nmpc = 60
 
     # Initial parameters. 
-    init_empc_pars = np.tile(np.array([[1, 2000, 12000]]), (Nmpc, 1))
+    init_empc_pars = np.tile(np.array([[1, 2000, 5000]]), (Nmpc, 1))
 
     # Get upper and lower bounds.
     ulb = model_pars['ulb']
@@ -131,8 +131,10 @@ def main():
     figures = CstrFlashPlots.plot_data(t=t, ulist=ulist, 
                                 ylist=None, xlist=xlist, 
                                 figure_size=PAPER_FIGSIZE, 
-                                ylabel_xcoordinate=-0.1, 
-                                data_type='closed-loop',
+                                u_ylabel_xcoordinate=-0.1, 
+                                y_ylabel_xcoordinate=-0.1, 
+                                x_ylabel_xcoordinate=-0.2, 
+                                plot_ulabel=True,
                                 legend_names=legend_names, 
                                 legend_colors=legend_colors, 
                                 title_loc=(0.25, 0.9), 

### 2021-03-25 cstr_flash_parameters.py

@@ -1,4 +1,4 @@
-# [depends] %LIB%/hybridid.py %LIB%/linNonlinMPC.py
+# [depends] %LIB%/hybridid.py %LIB%/cstr_flash_funcs.py
 # [makes] pickle
 """ Script to generate the necessary
     parameters and training data for the 
@@ -11,10 +11,9 @@ import mpctools as mpc
 import numpy as np
 import scipy.linalg
 from hybridid import (PickleTool, sample_prbs_like, SimData)
-from linNonlinMPC import NonlinearPlantSimulator, NonlinearMHEEstimator
 from cstr_flash_funcs import plant_ode, greybox_ode
 from cstr_flash_funcs import get_plant_pars, get_gb_pars
-from hybridid import get_rectified_xs, get_model, measurement, resample_fast
+from hybridid import get_rectified_xs, get_model
 
 def gen_train_val_data(*, parameters, num_traj,
                           Nsim_train, Nsim_trainval, 
@@ -210,7 +209,7 @@ def main():
     gb_pars = get_gb_pars(plant_pars=plant_pars)
 
     # Generate training data.
-    training_data = gen_train_val_data(parameters=plant_pars, num_traj=14,
+    training_data = gen_train_val_data(parameters=plant_pars, num_traj=20,
                                         Nsim_train=4*60, Nsim_trainval=12*60,
                                         Nsim_val=12*60, seed=2)
 

### 2021-03-25 cstr_flash_plots.py

@@ -1,6 +1,5 @@
-# [depends] cstr_flash_parameters.pickle cstr_flash_train.pickle
-# [depends] cstr_flash_empc.pickle
-# [depends] %LIB%/hybridid.py
+# [depends] %LIB%/hybridid.py %LIB%/plotting_funcs.py
+# [depends] cstr_flash_parameters.pickle cstr_flash_bbtrain.pickle
 """ Script to plot the training data
     and grey-box + NN model predictions on validation data.
     Pratyush Kumar, pratyushkumar@ucsb.edu """
@@ -77,13 +76,16 @@ def main():
     ulist += ulist_train
     ylist += ylist_train
     xlist += xlist_train
-    legend_names = ['Plant']#, 'Black-box']#'Black-box', 'Koopman', 'Koopman-ENC-DEC']
+    legend_names = ['Plant', 'Black-box']#'Black-box', 'Koopman', 'Koopman-ENC-DEC']
     legend_colors = ['b', 'dimgrey']#'dimgrey', 'm', 'tomato']
     figures = []
     figures += CstrFlashPlots.plot_data(t=t, ulist=ulist, 
                                 ylist=ylist, xlist=xlist, 
                                 figure_size=PAPER_FIGSIZE, 
-                                ylabel_xcoordinate=-0.1, data_type='open_loop',
+                                u_ylabel_xcoordinate=-0.1, 
+                                y_ylabel_xcoordinate=-0.1, 
+                                x_ylabel_xcoordinate=-0.2, 
+                                plot_ulabel=False,
                                 legend_names=legend_names, 
                                 legend_colors=legend_colors, 
                                 title_loc=(0.25, 0.9))

### 2021-03-25 lib/cstr_flash_funcs.py

@@ -157,7 +157,7 @@ def get_plant_pars():
     parameters['Qb'] = 200 # kJ/min
     parameters['Qr'] = 200 # kJ/min
 
-    # Store the dimensions.
+    # Store the dimensions. 
     Nx, Nu, Np, Ny = 10, 2, 2, 6
     parameters['Nx'] = Nx
     parameters['Nu'] = Nu

### 2021-03-25 lib/plotting_funcs.py

@@ -49,8 +49,10 @@ class TwoReacPlots:
     @staticmethod
     def plot_xudata(*, t, xlist, ulist, legend_names, 
                        legend_colors, figure_size,
-                       ylabel_xcoordinate, title_loc):
+                       ylabel_xcoordinate, title_loc, 
+                       font_size):
         """ Plot the performance loss economic MPC parameters."""
+        plt.rcParams.update({'font.size': font_size})
         nrow = len(TwoReacPlots.labels)
         (figure, axes) = plt.subplots(nrows=nrow, ncols=1,
                                       sharex=True, figsize=figure_size,
@@ -79,13 +81,15 @@ class TwoReacPlots:
     @staticmethod
     def plot_sscosts(*, us, sscosts, legend_colors, 
                         legend_names, figure_size, 
-                        ylabel_xcoordinate, left_label_frac):
+                        ylabel_xcoordinate, left_label_frac, 
+                        font_size):
         """ Plot the profit curves. """
+        plt.rcParams.update({'font.size': font_size})
         (figure, axes) = plt.subplots(nrows=1, ncols=1, 
                                       sharex=True, 
                                       figsize=figure_size, 
                                       gridspec_kw=dict(left=left_label_frac))
-        xlabel = r'$C_{Af} \ (\textnormal{mol/m}^3)$'
+        xlabel = r'$C_{Afs} \ (\textnormal{mol/m}^3)$'
         ylabel = r'Cost ($\$ $)'
         for (cost, color) in zip(sscosts, legend_colors):
             # Plot the corresponding data.
@@ -121,7 +125,7 @@ class CstrFlashPlots:
                r'$D \ (\textnormal{m}^3/\textnormal{min})$']
 
     def plot_inputs(t, ulist, figure_size, ylabel_xcoordinate, 
-                    data_type, legend_names, legend_colors, 
+                    plot_ulabel, legend_names, legend_colors, 
                     title_loc):
         """ Plot the training input data. """
         nrow = len(CstrFlashPlots.ulabels)
@@ -137,7 +141,7 @@ class CstrFlashPlots:
             axes[row].set_xlabel('Time (hr)')
             axes[row].set_xlim([np.min(t), np.max(t)])
             legend_handles += handle
-        if data_type == 'closed_loop':
+        if plot_ulabel:
             figure.legend(handles = legend_handles,
                         labels = legend_names,
                         loc = title_loc, ncol=len(legend_names))
@@ -190,19 +194,21 @@ class CstrFlashPlots:
         return [figure]
 
[…]

### 2021-03-25 lib/training_funcs.py

@@ -24,7 +24,7 @@ def create_koopmodel(*, Np, Ny, Nu, fN_dims):
     model = KoopmanModel(Np, Ny, Nu, fN_dims)
     # Compile the nn model.
     model.compile(optimizer='adam', loss='mean_squared_error', 
-                  loss_weights=[1., 0.])
+                  loss_weights=[0., 1.])
     # Return the compiled model.
     return model
 

### 2021-03-25 tworeac_kooptrain.py

@@ -66,7 +66,7 @@ def main():
 
         # Train.
         train_koopmodel(model=model, epochs=10000, batch_size=2, 
-                      train_data=train_samples, trainval_data=trainval_data, 
+                      train_data=train_samples, trainval_data=trainval_data,
                       stdout_filename=stdout_filename, ckpt_path=ckpt_path)
 
         # Validate.

### 2021-03-25 tworeac_openloop.py

@@ -156,7 +156,7 @@ def main():
         xlist += [xseq]
 
     # Get figure.
-    t = t*Delta
+    t = t*Delta/60
     legend_names = ['Plant', 'Grey-box', 'Black-box', 'Koopman']
     legend_colors = ['b', 'g', 'dimgrey', 'm']
     figures = TwoReacPlots.plot_xudata(t=t, xlist=xlist, ulist=ulist,
@@ -164,7 +164,8 @@ def main():
                                         legend_colors=legend_colors, 
                                         figure_size=PAPER_FIGSIZE, 
                                         ylabel_xcoordinate=-0.1, 
-                                        title_loc=(0.15, 0.9))
+                                        title_loc=(0.05, 0.9), 
+                                        font_size=12)
 
     # Finally plot.
     with PdfPages('tworeac_openloop.pdf', 'w') as pdf_file:

### 2021-03-25 tworeac_plots.py

@@ -104,11 +104,11 @@ def main():
     #valdata_list += encdeckoopval_predictions
     t, ulist, ylist, xlist = get_plotting_array_list(simdata_list=
                                                      valdata_list[:2],
-                                                     plot_range=(10, 24*60+10))
+                                                     plot_range=(10, 6*60+10))
     (t, ulist_train, 
      ylist_train, xlist_train) = get_plotting_array_list(simdata_list=
                                                      valdata_list[2:],
-                                                     plot_range=(0, 24*60))
+                                                     plot_range=(0, 6*60))
     ulist += ulist_train
     ylist += ylist_train
     xlist += xlist_train
@@ -117,7 +117,8 @@ def main():
                                         legend_colors=legend_colors, 
                                         figure_size=PAPER_FIGSIZE, 
                                         ylabel_xcoordinate=-0.1, 
-                                        title_loc=(0.25, 0.9))
+                                        title_loc=(0.05, 0.9),
+                                        font_size=12)
 
     # Plot validation metrics to show data requirements.
     #num_samples = tworeac_train['num_samples']

### 2021-03-25 tworeac_ssopt.py

@@ -116,7 +116,8 @@ def main():
                                         legend_names=legend_names, 
                                         figure_size=PAPER_FIGSIZE, 
                                         ylabel_xcoordinate=-0.1, 
-                                        left_label_frac=0.15)
+                                        left_label_frac=0.15, 
+                                        font_size=12)
 
     # Finally plot.
     with PdfPages('tworeac_ssopt.pdf', 'w') as pdf_file:

### 2021-04-20 sysID.tex
almost ready for group talk.:

@@
 
-\section{Large, CSTR and flash system example}
-We next study the performance of the black-box NN modeling framework on a large chemical process that consists of a chemical reactor and flash separator in series.
-The process is depicted in Figure \ref{fig:cstr_flash}, and the ordinary differential equations (ODEs) with parameter values describing the plant model are shown in Appendix \ref{app:cstr_flash}.
-The model consists of 12 states ($H_r$, $C_{Ar}$, $C_{Br}$, $C_{Cr}$, $T_r$, $H_b$, $C_{Ab}$, $C_{Bb}$, $C_{Cb}$, $T_b$), 6 measurements ($H_r$, $C_{Br}$, $T_r$, $H_b$, $C_{Bb}$, $T_b$), and 2 control inputs ($F$, $D$).
-The sample time for the measurements is 1 min.
+%\section{Large, CSTR and flash system example}
+%We next study the performance of the black-box NN modeling framework on a 
+%large chemical process that consists of a chemical reactor and flash separator 
+%in series. The process is depicted in Figure \ref{fig:cstr_flash}, and the 
+%ordinary differential equations (ODEs) with parameter values describing the 
+%plant model are shown in Appendix \ref{app:cstr_flash}. The model consists of 
+%12 states ($H_r$, $C_{Ar}$, $C_{Br}$, $C_{Cr}$, $T_r$, $H_b$, $C_{Ab}$, 
+%$C_{Bb}$, $C_{Cb}$, $T_b$), 6 measurements ($H_r$, $C_{Br}$, $T_r$, $H_b$, 
+%$C_{Bb}$, $T_b$), and 2 control inputs ($F$, $D$). The sample time for the 
+%measurements is 1 min.
 
-\begin{figure}[!h]
-  \centering
-  \includegraphics[page=1, height=0.5\textheight,
-                   width=0.7\textwidth]{cstr_flash.pdf}
-  \caption{Schematic of the CSTR and flash chemical process.}
-  \label{fig:cstr_flash}
-\end{figure}
+%\begin{figure}[!h]
+%  \centering
+%  \includegraphics[page=1, height=0.5\textheight,
+%                   width=0.7\textwidth]{cstr_flash.pdf}
+%  \caption{Schematic of the CSTR and flash chemical process.}
+%  \label{fig:cstr_flash}
+%\end{figure}
 
-\subsection{Model validation}
+%\subsection{Model validation}
 
@@
 
-\subsection{Steady state optimization}
+%\subsection{Steady state optimization}
 
-\subsection{Dynamic economic optimization}
+%\subsection{Dynamic economic optimization}
 
-\section{Discussion}
+\section{Hybrid model}
+
+\begin{equation*}
+\begin{bmatrix}
+\dfrac{dC_A}{dt} \\
+\dfrac{dC_B}{dt} \\
+\dfrac{dC_C}{dt}
+\end{bmatrix} = \begin{bmatrix}
+\dfrac{C_{Af} - C_A}{\tau} \\
+- \dfrac{C_B}{\tau} \\
+- \dfrac{C_C}{\tau}
+\end{bmatrix} + f_N(C_A, C_B, C_C)
+\end{equation*}
+
+\section{Learn cost functions with input convex NNs}
+Different proposal. NNs can be made convex from their inputs to outputs. 
+Consider the following architecture (Amos, Xu, Kolter 2017) -
+
+\begin{equation*}
+	z_{i+1} = g_i(W_i^{z}z_{i} + W_i^{y}x + b_i), i=0, 1, ...k-1
+\end{equation*}
+
+If $g_i(\cdot)$ is convex and non-decreasing, $W_0^{z} = 0$, and $W_{1:k-1}^{z}$
+are all non-negative, then the NN architecture is convex, i.e, the function 
+$z_k(x)$ is convex.
+
+\textbf{RTO algorithm}
+\begin{itemize}
+	\item Obtain steady-state data ($u_s$, $y_s$), and samples of the cost 
+	curve ($\ell(u_s, y_s, p)$) based on current economic cost parameters.
+	\item Fit the input convex NN to the cost curve and optimize.
+\end{itemize}
+
+Extensions - Can allow the NN to be non-convex in some inputs and convex in 
+some inputs. So the NN cost function can be non-convex in economic 
+parameters/disturbance estimates, but we will force it to be convex in $u_s$.
 

### 2021-04-20 lib/BlackBoxFuncs.py

@@ -89,7 +89,8 @@ class BlackBoxModel(tf.keras.Model):
         fNLayers = []
         for dim in fNDims[1:-1]:
             fNLayers += [tf.keras.layers.Dense(dim, activation=scaledtanh)]
-        fNLayers += [tf.keras.layers.Dense(fNDims[-1])]
+        fNLayers += [tf.keras.layers.Dense(fNDims[-1], 
+                                           kernel_initializer='zeros')]
 
         # Build model.
         bbCell = BlackBoxCell(Np, Ny, Nu, fNLayers)

### 2021-04-20 lib/KoopmanModelFuncs.py

@@ -1,0 +1,336 @@
+"""
+Custom neural network layers for the 
+data-based completion of grey-box models 
+using neural networks.
+Pratyush Kumar, pratyushkumar@ucsb.edu
+"""
+import sys
+import numpy as np
+import tensorflow as tf
+from BlackBoxFuncs import fnnTF, fnn
+from hybridid import SimData
+
+class KoopmanCell(tf.keras.layers.AbstractRNNCell):
+    """
+    RNN Cell.
+    z = [y_{k-N_p:k-1}', u_{k-N_p:k-1}']
+    x_{kp} = [y;z;NN(y;z)]
+    x_{kp}^+  = Ax_{kp} + Bu
+    H = [I, 0]
+    yz = H*x_{kp}
+    Cell output: yz
+    """
+    def __init__(self, Nxkp, Np, Ny, Nu, A, B, **kwargs):
+        super(KoopmanCell, self).__init__(**kwargs)
+        self.Nxkp, self.Np, self.Ny, self.Nu = Nxkp, Np, Ny, Nu
+        self.Nz = Np*(Ny + Nu)
+        self.A, self.B = A, B
+
+    @property
+    def state_size(self):
+        return self.Nxkp
+    
+    @property
+    def output_size(self):
+        return self.Ny + self.Nz
+
+    def call(self, inputs, states):
+        """ Call function of the hybrid RNN cell.
+            Dimension of states: (None, Nxkp)
+            Dimension of input: (None, Nu)
+            Dimension of output: (None, Ny + Nz)
+        """
+        # Extract important variables.
+        [xkp] = states
+        u = inputs
+
+        # Get the state prediction at the next time step.
+        xkplus = self.A(xkp) + self.B(u)
+        
+        # Get the output at the current timestep.
+        [yz, _] = tf.split(xkp, 
+                           [self.output_size, self.Nxkp-self.output_size], 
+                           axis=-1)
+
+        # Return output and states at the next time-step.
+        return (yz, xkplus)
+
+class KoopmanModel(tf.keras.Model):
+    """ Custom model for the Deep Koopman operator model. """
+    def __init__(self, Np, Ny, Nu, fNDims):
+        
+        # Get a few sizes.
+        Nz = Np*(Ny+Nu)
+        Nxkp = Ny + Nz + fNDims[-1]
+
+        # Create inputs to the layers.
+        useq = tf.keras.Input(name='u', shape=(None, Nu))
+        yz0 = tf.keras.Input(name='yz0', shape=(Ny + Nz, ))
+        
[…]

### 2021-04-20 lib/training_funcs.py

@@ -10,62 +10,3 @@ import numpy as np
 import tensorflow as tf
 from HybridModelLayers import BlackBoxModel, KoopmanModel
 from hybridid import SimData
-
-def create_koopmodel(*, Np, Ny, Nu, fN_dims):
-    """ Create/compile the two reaction model for training. """
-    model = KoopmanModel(Np, Ny, Nu, fN_dims)
-    # Compile the nn model.
-    model.compile(optimizer='adam', loss='mean_squared_error', 
-                  loss_weights=[0., 1.])
-    # Return the compiled model.
-    return model
-
-def train_koopmodel(*, model, epochs, batch_size, train_data, trainval_data, 
-                       stdout_filename, ckpt_path):
-    """ Function to train the NN controller. """
-    # Std out.
-    sys.stdout = open(stdout_filename, 'w')
-    # Create the checkpoint callback.
-    checkpoint_callback = tf.keras.callbacks.ModelCheckpoint(filepath=ckpt_path,
-                                                    monitor='val_loss',
-                                                    save_best_only=True,
-                                                    save_weights_only=True,
-                                                    verbose=1)
-    # Call the fit method to train.
-    model.fit(x=[train_data['inputs'], train_data['yz0']], 
-              y=[train_data['yz'], train_data['outputs']], 
-              epochs=epochs, batch_size=batch_size,
-        validation_data = ([trainval_data['inputs'], trainval_data['yz0']], 
-                           [trainval_data['yz'], trainval_data['outputs']]),
-            callbacks = [checkpoint_callback])
-
-def get_koopval_predictions(*, model, val_data, xuyscales, 
-                               xinsert_indices, ckpt_path):
-    """ Get the validation predictions. """
-
-    # Load best weights.
-    model.load_weights(ckpt_path)
-
-    # Predict.
-    model_predictions = model.predict(x=[val_data['inputs'], val_data['yz0']])
-
-    # Scale.
-    ymean, ystd = xuyscales['yscale']
-    umean, ustd = xuyscales['uscale']
-    ypredictions = model_predictions[1].squeeze()*ystd + ymean
-    uval = val_data['inputs'].squeeze(axis=0)*ustd + umean
-
-    # Get xpredictions.
-    xpredictions = np.insert(ypredictions, xinsert_indices, np.nan, axis=1)
-
-    # Collect data in a Simdata format.
-    Nt = uval.shape[0]
[…]

### 2021-04-20 tworeac_bbNNtrain.py

@@ -36,7 +36,7 @@ def main():
     tthrow = 10
     Np = 0
     tanhScale = 1
-    fNDims = [Ny + Np*(Ny+Nu), 32, Ny]
+    fNDims = [Ny + Np*(Ny+Nu), 64, Ny]
 
     # Create lists to store data.
     trained_weights = []
@@ -66,7 +66,7 @@ def main():
                              outputs=train_data['outputs'])
 
         # Train.
-        train_bbmodel(model=model, epochs=10000, batch_size=6, 
+        train_bbmodel(model=model, epochs=10000, batch_size=2, 
                       train_data=train_samples, trainval_data=trainval_data, 
                       stdout_filename=stdout_filename, ckpt_path=ckpt_path)
 

### 2021-04-20 tworeac_chk_trained_funcs.py

@@ -11,9 +11,10 @@ import sys
 sys.path.append('lib/')
 import numpy as np
 from hybridid import PickleTool, quick_sim
-from BlackBoxFuncs import get_bbNN_pars, bbNN_fxu, bb_hx
+from BlackBoxFuncs import get_bbNN_pars, bbNN_fxu, bb_hx, fnn
 from TwoReacHybridFuncs import (get_tworeacHybrid_pars,
                                 tworeacHybrid_fxu, tworeacHybrid_hx)
+from KoopmanModelFuncs import get_KoopmanModel_pars, koop_fxu, koop_hx
 
 def main():
     """ Main function to be executed. """
@@ -24,6 +25,8 @@ def main():
     tworeac_bbNNtrain = PickleTool.load(filename='tworeac_bbNNtrain.pickle',
                                         type='read')
     tworeac_hybtrain = PickleTool.load(filename='tworeac_hybtrain.pickle',
+                                        type='read')
+    tworeac_kooptrain = PickleTool.load(filename='tworeac_kooptrain.pickle',
                                         type='read')
 
     def check_bbNN(tworeac_bbNNtrain, tworeac_parameters):
@@ -86,52 +89,58 @@ def main():
         # Return 
         return 
 
+    def check_koopman(tworeac_kooptrain, tworeac_parameters):
+        """ Check Black-box functions. """
+        plant_pars = tworeac_parameters['plant_pars']
+
+        # Get some sizes/parameters.
+        tthrow = 10
+        Np = tworeac_kooptrain['Np']
+        Ny, Nu = plant_pars['Ny'], plant_pars['Nu']
+
+        # Get initial state for forecasting.
+        training_data = tworeac_parameters['training_data'][-1]
+        uval = training_data.u[tthrow:, :]
+        y0 = training_data.y[tthrow, :]
+        yp0seq = training_data.y[tthrow-Np:tthrow, :].reshape(Np*Ny, )
+        up0seq = training_data.u[tthrow-Np:tthrow, :].reshape(Np*Ny, )
+        yz0 = np.concatenate((y0, yp0seq, up0seq))
+
+        # Scale initial state and get the lifted state.
+        fNWeights = tworeac_kooptrain['trained_weights'][-1][:-2]
+        xuyscales = tworeac_kooptrain['xuyscales']
+        ymean, ystd = xuyscales['yscale']
[…]

### 2021-04-20 tworeac_hybtrain.py

@@ -36,7 +36,7 @@ def main():
     xinsert_indices = []
     Np = 0
     tthrow = 10
-    fNDims = [Nx, 8, 3]
+    fNDims = [Nx, 16, 3]
 
     # Create lists to store data.
     trained_weights = []
@@ -67,7 +67,7 @@ def main():
                              outputs=train_data['outputs'])
 
         # Train.
-        train_hybrid_model(model=model, epochs=10000, batch_size=2, 
+        train_hybrid_model(model=model, epochs=5000, batch_size=2, 
                       train_data=train_samples, trainval_data=trainval_data,
                       stdout_filename=stdout_filename, ckpt_path=ckpt_path)
 

### 2021-04-20 tworeac_kooptrain.py

@@ -11,8 +11,8 @@ import tensorflow as tf
 import time
 import numpy as np
 from hybridid import PickleTool, get_scaling, get_train_val_data
-from training_funcs import create_koopmodel, train_koopmodel
-from training_funcs import get_koopval_predictions
+from KoopmanModelFuncs import (create_koopmodel, train_koopmodel, 
+                              get_koopval_predictions)
 
 # Set the tensorflow global and graph-level seed.
 tf.random.set_seed(123)
@@ -24,18 +24,18 @@ def main():
                                          type='read')
 
     # Get sizes/raw training data.
-    parameters = tworeac_parameters['parameters']
-    Ny, Nu = parameters['Ny'], parameters['Nu']
+    plant_pars = tworeac_parameters['plant_pars']
+    Ny, Nu = plant_pars['Ny'], plant_pars['Nu']
     training_data = tworeac_parameters['training_data']
     
     # Number of samples.
     num_samples = [hours*60 for hours in [6]]
 
     # Create some parameters.
-    xinsert_indices = [2]
+    xinsert_indices = []
     tthrow = 10
-    Np = 2
-    fN_dims = [Np*(Ny+Nu), 16, 32]
+    Np = 0
+    fNDims = [Ny + Np*(Ny+Nu), 16, 32]
 
     # Create lists to store data.
     trained_weights = []
@@ -56,7 +56,7 @@ def main():
     for num_sample in num_samples:
         
         # Create model.
-        model = create_koopmodel(Np=Np, Ny=Ny, Nu=Nu, fN_dims=fN_dims)
+        model = create_koopmodel(Np=Np, Ny=Ny, Nu=Nu, fNDims=fNDims)
         
         # Use num samples to adjust here the num training samples.
         train_samples = dict(yz0=train_data['yz0'],
@@ -65,7 +65,7 @@ def main():
                              outputs=train_data['outputs'])
 
         # Train.
-        train_koopmodel(model=model, epochs=10000, batch_size=2, 
+        train_koopmodel(model=model, epochs=3000, batch_size=2, 
                       train_data=train_samples, trainval_data=trainval_data,
                       stdout_filename=stdout_filename, ckpt_path=ckpt_path)
 
@@ -87,16 +87,15 @@ def main():
     num_samples = np.asarray(num_samples) + trainval_data['inputs'].shape[1]
 
     # Save the weights.
-    tworeac_training_data = dict(Np=Np,
-                                 fN_dims=fN_dims,
[…]

### 2021-04-20 tworeac_parameters.py

@@ -34,23 +34,23 @@ def gen_train_val_data(*, parameters, num_traj,
         if traj == num_traj-1:
             "Get input for train val simulation."
             Nsim = Nsim_val
-            u = sample_prbs_like(num_change=4, num_steps=Nsim_val, 
+            u = sample_prbs_like(num_change=6, num_steps=Nsim_val, 
                                  lb=ulb, ub=uub,
-                                 mean_change=90, sigma_change=10, 
+                                 mean_change=60, sigma_change=10, 
                                  seed=seed+1)
         elif traj == num_traj-2:
             "Get input for validation simulation."
             Nsim = Nsim_trainval
-            u = sample_prbs_like(num_change=4, num_steps=Nsim_trainval, 
+            u = sample_prbs_like(num_change=12, num_steps=Nsim_trainval, 
                                  lb=ulb, ub=uub,
-                                 mean_change=90, sigma_change=10, 
+                                 mean_change=30, sigma_change=10, 
                                  seed=seed+2)
         else:
             "Get input for training simulation."
             Nsim = Nsim_train
-            u = sample_prbs_like(num_change=4, num_steps=Nsim_train, 
+            u = sample_prbs_like(num_change=12, num_steps=Nsim_train, 
                                  lb=ulb, ub=uub,
-                                 mean_change=90, sigma_change=10, 
+                                 mean_change=30, sigma_change=10, 
                                  seed=seed+3)
 
         seed += 1
@@ -95,7 +95,7 @@ def main():
     
     # Generate training data.
     training_data = gen_train_val_data(parameters=plant_pars,
-                                        num_traj=8, Nsim_train=360,
+                                        num_traj=4, Nsim_train=360,
                                         Nsim_trainval=360, Nsim_val=360,
                                         seed=103)
     

### 2021-04-20 tworeac_plots.py

@@ -78,10 +78,9 @@ def main():
     hyb_predictions = tworeac_hybtrain['val_predictions']
 
     # Load Koopman data after NN training.
-    # tworeac_kooptrain = PickleTool.load(filename="tworeac_kooptrain.pickle",
-    #                                   type='read')
-    # koop_predictions = tworeac_kooptrain['val_predictions']
-
+    tworeac_kooptrain = PickleTool.load(filename="tworeac_kooptrain.pickle",
+                                      type='read')
+    koop_predictions = tworeac_kooptrain['val_predictions']
 
     # Load data after Koopman training.
     #tworeac_kooptrain = PickleTool.load(filename=
@@ -95,16 +94,15 @@ def main():
     #                                type='read')
     #encdeckoopval_predictions = tworeac_encdeckooptrain['val_predictions']
 
-
     # Create a figures list.
     figures = []
 
     # Plot validation data.
-    legend_names = ['Plant', 'Black-box-NN', 'Hybrid']
+    legend_names = ['Plant', 'Black-box-NN', 'Koopman']
     legend_colors = ['b', 'dimgrey', 'm']
     valdata_list = [training_data[-1]]
     valdata_list += bbNN_predictions
-    valdata_list += hyb_predictions
+    valdata_list += koop_predictions
     t, ulist, ylist, xlist = get_plotting_array_list(simdata_list=
                                                      valdata_list[:1],
                                                      plot_range=(10, 6*60+10))
@@ -120,7 +118,7 @@ def main():
                                         legend_colors=legend_colors, 
                                         figure_size=PAPER_FIGSIZE, 
                                         ylabel_xcoordinate=-0.1, 
-                                        title_loc=(0.05, 0.9),
+                                        title_loc=(0.15, 0.9),
                                         font_size=12)
 
     # Plot validation metrics to show data requirements.

### 2021-04-20 tworeac_ssopt.py

@@ -21,6 +21,8 @@ from TwoReacHybridFuncs import (get_tworeacHybrid_pars,
 from economicopt import (get_ss_optimum, get_xuguess, c2dNonlin, 
                          get_sscost)
 from tworeac_funcs import cost_yup, plant_ode
+from KoopmanModelFuncs import (get_KoopmanModel_pars, koop_fxu, koop_hx, 
+                              get_koopman_xguess)
 
 def main():
     """ Main function to be executed. """
@@ -33,12 +35,15 @@ def main():
     tworeac_bbNNtrain = PickleTool.load(filename=
                                     'tworeac_bbNNtrain.pickle',
                                       type='read')
-    tworeac_hybtrain = PickleTool.load(filename=
-                                      'tworeac_hybtrain.pickle',
+    tworeac_kooptrain = PickleTool.load(filename=
+                                    'tworeac_kooptrain.pickle',
                                       type='read')
+    # tworeac_hybtrain = PickleTool.load(filename=
+    #                                   'tworeac_hybtrain.pickle',
+    #                                   type='read')
 
     # Get cost function handle.
-    p = [100, 220]
+    p = [100, 200]
     lyu = lambda y, u: cost_yup(y, u, p)
 
     # Get the black-box model parameters and function handles.
@@ -47,11 +52,17 @@ def main():
     bbNN_Fxu = lambda x, u: bbNN_fxu(x, u, bbNN_pars)
     bbNN_hx = lambda x: bb_hx(x, bbNN_pars)
 
+    # Get the koopman model parameters and function handles.
+    koop_pars = get_KoopmanModel_pars(train=tworeac_kooptrain, 
+                                      plant_pars=plant_pars)
+    koop_Fxu = lambda x, u: koop_fxu(x, u, koop_pars)
+    koop_Hx = lambda x: koop_hx(x, koop_pars)
+
     # Get the black-box model parameters and function handles.
-    hyb_pars = get_tworeacHybrid_pars(train=tworeac_hybtrain, 
-                                      greybox_pars=greybox_pars)
-    hyb_fxu = lambda x, u: tworeacHybrid_fxu(x, u, hyb_pars)
-    hyb_hx = lambda x: tworeacHybrid_hx(x)
[…]

### 2021-08-03 group_talk_apr20_2021.tex
local repo cleanup.

\documentclass[xcolor=dvipsnames, 8pt]{beamer} %
%\setbeamertemplate{navigation symbols}{}

\usetheme{SantaBarbara}

\definecolor{black}{HTML}{0A0A0A}
\definecolor{green}{HTML}{0420eb} 
\definecolor{darkgreen}{HTML}{008000}
\definecolor{gold}{HTML}{FFD000}
\setbeamercolor{normal text}{fg=black,bg=white}
\setbeamercolor{alerted text}{fg=red}
\setbeamercolor{example text}{fg=black}
\setbeamercolor{palette primary}{fg=black, bg=gray!20}
\setbeamercolor{palette secondary}{fg=black, bg=gray!20}
\setbeamercolor{palette tertiary}{fg=white, bg=green!80}
\setbeamercolor{block title}{fg=black,bg=gold!40}
\setbeamercolor{frametitle}{fg=white, bg=green!80}
\setbeamercolor{title}{fg=white, bg=green!80}

\newcommand{\vecc}{\operatorname{vec}}
%\newcommand{\svec}{\operatorname{svec}}
\newcommand{\norm}[1]{\left| #1 \right|}
\newcommand{\xhat}{\hat{x}}
\newcommand{\uhat}{\hat{u}}
\newcommand{\yhat}{\hat{y}}
\newcommand{\wt}{\widehat{\theta}}

\usepackage[utf8]{inputenc}

\usepackage{tikz, tikzsettings}
\usepackage{verbatim}
\usepackage{amssymb}
\usepackage{amsmath}
%\usepackage{amsfonts}
\usepackage{algorithmic}
\graphicspath{{./}{./figures/}{./figures/presentation/}}
%\usepackage[version=4]{mhchem}
\usepackage{subcaption}
\usepackage[authoryear,round]{natbib}
%\usepackage{fancyvrb}
\usepackage{color}
\usepackage{colortbl}
\usepackage{xcolor}
\usepackage{physics}
\usepackage{pgfplots}
\usepackage{ragged2e}
%\pgfplotsset{compat=newest} 
%\pgfplotsset{plot coordinates/math parser=false}
%\usepackage{environ}
%\usetikzlibrary{decorations.markings}
%\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{shapes,calc,spy, calc, backgrounds,arrows, fit, decorations.pathmorphing, decorations.pathreplacing, matrix}
%\usepackage[absolute,overlay]{textpos}
\usepackage{caption}
\usepackage{graphicx}
%\usepackage[controls=true,poster=first]{animate}

\newcommand{\useq}{\mathbf{u}}
\newcommand{\xseq}{\mathbf{x}}
\newcommand{\bbR}{\mathbb{R}}
\newcommand{\bbW}{\mathbb{W}}
\newcommand{\bbU}{\mathbb{U}}
\newcommand{\bbI}{\mathbb{I}}
\newcommand{\bbX}{\mathbb{X}}
\newcommand{\bbP}{\mathbb{P}}

\newcommand{\code}[1]{{\fontfamily{qcr}\selectfont#1}}

\AtBeginSection[]
{\frame<beamer>{\frametitle{Outline}   
	\tableofcontents[currentsection, currentsection]}
	\addtocounter{framenumber}{-1}}
	%
%{
%	\frame<beamer>{ 
%		\frametitle{Outline}   
%		\tableofcontents[currentsection,currentsubsection]}

%}

\makeatother
\setbeamertemplate{footline}
{
	\leavevmode%
	\hbox{%
		\begin{beamercolorbox}[wd=.3\paperwidth,ht=2.25ex,dp=1ex,center]{author in head/foot}%
			\usebeamerfont{author in head/foot}\insertshortauthor
		\end{beamercolorbox}%
		\begin{beamercolorbox}[wd=.6\paperwidth,ht=2.25ex,dp=1ex,center]{title in head/foot}%
			\usebeamerfont{title in head/foot}\insertshorttitle
		\end{beamercolorbox}%
		\begin{beamercolorbox}[wd=.1\paperwidth,ht=2.25ex,dp=1ex,center]{date in head/foot}%
			\insertframenumber{} / \inserttotalframenumber\hspace*{1ex}
	\end{beamercolorbox}}%
	\vskip0pt%
}

\title{Hybrid system identification using neural networks}
\date{Jan 19, 2021}
\author[Pratyush Kumar]{\large Pratyush Kumar}
\institute[UCSB]{
	\begin{minipage}{4in}
		\vspace{-10pt}
		\centering
		\raisebox{-0.1\height}{\includegraphics[width=0.25\textwidth]{UCSB_seal}}
		%\hspace*{.2in}
		%\raisebox{-0.5\height}{\includegraphics[width=0.25\textwidth]{jci_logo}}
	\end{minipage}
	\vspace{10pt}
	\newline
	{\large Department of chemical engineering}
	\vspace{10pt}
	\newline
	{\large Group talk}}

\begin{document}

%\frame{\titlepage}

\begin{frame}{Outline}
\tableofcontents
\end{frame}

\section{System identification}

\subsection{Koopman dynamic model}

\begin{frame}{Plant}
\begin{align*}
\dot x_p &= f_p(x_p, u, d, w) \\
y &= h_p(x_p) + v
\end{align*}

\begin{itemize}
	\setlength{\itemsep}{10pt}%
	\item $x_p \in \bbR^{n_p}$ (state), $y \in \bbR^{n_y}$ (measurement), 
	and $u \in \bbU$ (actuator), $d \in \bbR^{n_d}$ (disturbance). 
	\item $w \in \bbR^{n_w}$ (process noise), $v \in \bbR^{n_y}$ (measurement 
	noise).
	\item Goal: Obtain a dynamic model from data for subsequent use in 
	steady-state optimization or dynamic economic MPC.
\end{itemize}
\end{frame}

\begin{frame}{Black-box NN model}
\begin{align*}
\mathbf{y}_{k-N_p:k-1} &= [y(k-N_p)', y(k-N_p+1)', ..., y(k-1)']' \\
\mathbf{u}_{k-N_p:k-1} &= [u(k-N_p)', u(k-N_p+1)', ..., u(k-1)']' \\
z(k) &= [\mathbf{y}'_{k-N_p:k-1}, \mathbf{u}'_{k-N_p:k-1}]'
\end{align*}
in which, $N_p$ is the number of past inputs and outputs used.

\begin{align*}
x &= [y', z']' \\
x^+ &= \begin{bmatrix}
f_N(x, u, \theta_N) \\
y(k-N_p+1) \\
\vdots \\
y(k) \\
u(k-N_p+1) \\ 
\vdots \\
u(k)
\end{bmatrix} = f(x, u)\\
y &= [I, \ 0,\  0, \cdots]x = h(x)
\end{align*}

Note: Keeping the full state dynamics structure is helpful to make multi-step 
ahead forecasts.
\end{frame}

\begin{frame}{Koopman dynamic model}

Approach: Transform the state onto a manifold such that the system dynamics are 
approximately linear in that manifold \citep{korda:mezic:2018}.

\begin{align*} 
x_{kp} &= [y', z', f_N(y, z, \theta_N)']' \\
x_{kp}^+ &= Ax_{kp} + Bu = f(x, u)\\
y &= [I, \ 0] x_{kp} = h(x)
\end{align*}

\begin{itemize}
	\item $x_{kp} \in \bbR^N$, $f_N(\cdot)$ is a NN used for the nonlinear 
	transformation.
	\item Find $A$, $B$, and $\theta_N$ from data using the training 
	optimization problem.
\end{itemize}
\end{frame}

\subsection{Reactor example}

\begin{frame}{Nonlinear reactor model}

	\textbf{Plant}
	\begin{align*}
	\dfrac{dC_A}{dt} &= \dfrac{C_{Af} - C_A}{\tau} - k_1C_A\\
	\dfrac{dC_B}{dt} &= k_1C_A - 3k_2C^3_B + 3k_3C_C- \dfrac{C_B}{\tau}\\
	\dfrac{dC_C}{dt} &= k_2C^3_B - k_3C_C - \dfrac{C_C}{\tau}
	\end{align*}
	$u = C_{Af}$, $x = [C_A, C_B, C_C]'$.

\begin{itemize}
\item Assume full state feedback.
\item Use 18 hours (1080 samples) of training data, 6 hours (360 samples) of 
validation data.
\item Black-box NN architecture: $[3, \ 64, \ 3]$
\item Architecture of NN used in Koopman model lifting: $[3, \ 32, \ 32]$
\end{itemize}
\end{frame}

\begin{frame}{Model validation -- open loop data}
	\vspace{-0.1in}
	\begin{figure}
		\centering
		\includegraphics[page=1, height=0.8\textheight, 
		width=0.7\textwidth]{tworeac_plots.pdf}
	\end{figure}
\end{frame}

\begin{frame}{Steady-state optimization problem}
	\begin{align*}
	\underset{u_s}{\textnormal{min}} & \quad \ell(y_s, u_s) \\
	x_s &= f(x_s, u_s) \nonumber \\ 
	y_s &= h(x_s) \nonumber \\
	\underline{u} \leq &u_s \leq \overline{u} \nonumber
	\end{align*}

	For the reactor model, choose the stage cost as: $\ell(y_s, u_s) = 
	p_1C_{Afs} - p_2C_{Bs}$

\end{frame}

\begin{frame}{Steady-state cost curves}
	\begin{figure}
		\centering
		\includegraphics[page=1, height=0.7\textheight, 
		width=0.6\textwidth]{tworeac_ssopt.pdf}
	\end{figure}

\begin{itemize}
	\item The Koopman dynamic model has a linear structure, therefore, the cost 
	curve is linear as well. 
	\item The benefit of Koopman models, as discussed in 
	\cite{korda:mezic:2018} for tracking MPC, is that they would better 
	prediction accuracy than standard linear models.
\end{itemize}

\end{frame}

\section{Paresto}

\subsection{NLP setup with casadi}

\begin{frame}{Parameter estimation problem}

\begin{align*}
\underset{x(0), \theta}{\textnormal{min}} \sum_{k=0}^{N_t} 
\dfrac{1}{N_t} &\norm{\tilde{y}(k) - y(k)}^2_{R_v^{-1}}  \\
\textnormal{subject to} \quad \quad x^{+} &= f(x, u, \theta) \\
0 &= g(x, z, \theta) \\
\tilde{y} &= h(x, z, \theta)
\end{align*}

\begin{itemize}
	\setlength{\itemsep}{10pt}%
	\item Multi-step-ahead prediction-error minimization. 
	\item Multiple-shooting method: All the states ($x$ and $z$) 
	at each measurement instant are kept as decision variables in the 
	optimization. 
	\item The states are constrained by model equations, and states at each 
	time point are used to make forecasts at the next time point. 
	\item Confidence intervals are computed using Hessian of the objective. 
\end{itemize}
\end{frame}

\begin{frame}{Nonlinear programming in Casadi}

General NLP in Casadi
\begin{align*}
\underset{x}{\textnormal{min}} \quad &f(x, p) \\
\textnormal{subject to} \quad \underline{x} &\leq x \leq \overline{x} \\
\underline{g} &\leq g(x, p) \leq \overline{g}\\
\end{align*}

Casadi usage

\begin{center}

\code{nlpInfo = struct('x', x, 'f', fxp, 'g', gxp, 'p', p)} \\
\medskip
\code{testNlp = casadi.nlpsol('testNlp', 'solverName', nlpInfo, optsStruct)} \\
\medskip
\code{testNlp('x0', xguess, 'lbx', lbx, 'ubx', ubx, 'lbg', lbg, 'ubg', ubg, 
'p', pval)}
\end{center}

\begin{itemize}
	\item Now setup the previous parameter estimation problem in this format.
	\item The decision variable in the NLP as grey-box model parameters 
	($\theta$) and states ($x$ and $z$) at each time point, and parameters in 
	the NLP as measurements.
\end{itemize}
\end{frame}

\begin{frame}{Problem setup}

Create casadi symbolics (decision variables, objective function, constraints, 
parameters). \\
\medskip

	\code{theta = casadi.MX.sym('theta', Ntheta);} \\
	\code{y = casadi.MX.sym('y', Ny*(Nt+1));}\\
	\code{xk = casadi.MX.sym('x0', Nx);}\\
	\code{zk = casadi.MX.sym('z0', Nz);}\\
	\code{w = \{ theta, x0, z0 \} \% decision variable list} \\
	\code{lsq = \{ y(1:Ny) - hxz('x', xk, 'z', zk, 'theta', theta) \}} \\
	\code{g = \{alg('x', xk, 'z', zk, 'theta', theta) \}} \\

	\medskip
	\code{for k=1:Nt} \\

	\medskip
	\code{\% Simulate via a casadi function.} \\
	\code{xkplus = fxz('x', xk, 'z', zk, 'theta', theta)}

	\medskip 
	\code{\% Create new decision variables.} \\
	\code{xk = casadi.MX.sym('xk', Nx);}\\
	\code{zk = casadi.MX.sym('zk', Nz);}\\
	\code{w\{end+1\} = xk;} \\
	\code{w\{end+1\} = zk;} \\

	\medskip 
	\code{\% Get objective function contribution.} \\
	\code{lsq\{end+1\} = y(k*Ny+1:(k+1)*Ny) - hxz('x', xk, 'z', zk, 'theta', 
	theta);} \\

	\medskip 
	\code{\% Get equality constraints.} \\
	\code{g\{end+1\} = xplus - xk;} \\
	\code{g\{end+1\} = alg('x', xk, 'z', zk, 'theta', theta);} \\

	\medskip
	\code{end\%for}

\end{frame}

\begin{frame}{Create NLP solver and solve}

	\medskip

	\code{w = vertcat(w\{:\});} \\
	\code{J = sumsqr(lsq);} \\
	\code{g = vertcat(g\{:\});} \\
	\code{nlpInfo = struct('x', w, 'f', J, 'g', g, 'p', p)} \\
	\medskip
	\code{nlp = casadi.nlpsol('testNlp', 'ipopt', nlpInfo, optsStruct)} \\
	\medskip
	\code{nlp('x0', xguess, 'lbx', lbx, 'ubx', ubx, 'lbg', lbg, 'ubg', ubg, 
	'p', pval)} \\
	\medskip

An update needed in paresto: Handle data generated by time-varying control 
input profiles.
\end{frame}

\subsection{KKT conditions and DAE bug}
\begin{frame}{KKT conditions}

Formulate Lagrangian \citep{andersson:rawlings:2018} -

\begin{align*}
L(x, p, \lambda_x, \lambda_g) &= f(x, p) + \lambda_x' x + \lambda_g'g(x, p) \\
\lambda_x = \overline{\lambda}_x - \underline{\lambda}_x \\
\lambda_g = \overline{\lambda}_g - \underline{\lambda}_g \\
\end{align*}

KKT conditions - 
\begin{align*}
&\grad_x L(x, p, \lambda_x, \lambda_g) =0 \\
\underline{x} &\leq x \leq \overline{x} \\
\underline{g} &\leq g(x, p) \leq \overline{g} \\
&I^0_x \lambda_x + \overline{I}_x\overline{x} + \underline{I}_x\underline{x} - 
I_xx = 0 \\
&I^0_g \lambda_g + \overline{I}_g\overline{g} + \underline{I}_g\underline{g} - 
I_gg(x, p) = 0
\end{align*}

\begin{itemize}
	\item $\overline{I}_x$ and $\underline{I}_x$ are diagonal matrices with 
	zeros and ones denoting active and inactive constraints respectively.
	\item $I_x = \overline{I}_x + \underline{I}_x$ and $I^0_x = I - I_x$ denote 
	either and neither lower/upper bound active respectively.
	\item Casadi uses these active constraint informations to compute the 
	Hessian.
\end{itemize}

\end{frame}

\begin{frame}{$\lambda = 0$ doesn't mean constraint is inactive}

Consider the following optimization problem -

\begin{align*}
\underset{x, y}{\textnormal{min}} & \ (x-2)^2 \\
\textnormal{subject to} & x-y = 0
\end{align*}

Lagrangian - 
\begin{equation*}
	L(x, y, \lambda) = (x-2)^2 + \lambda (x-y)
\end{equation*}

KKT conditions - 
\begin{align*}
	2(x-2) + \lambda &= 0 \\
	-\lambda &= 0 \\
	x - y &= 0 \\
	\lambda &\geq 0 \\
	\lambda(x-y) &= 0
\end{align*}

\begin{itemize}
	\item Optimal solution is $x=y=2$ and $\lambda = 0$, but the equality 
	constraint is active. 
	\item Casadi was interpreting zero Lagrange multipliers as inactive 
	constraint, and therefore, was computing incorrect Hessian/confidence 
	intervals.
\end{itemize}

\end{frame}

\begin{frame}{References}
\bibliographystyle{abbrvnat}
\bibliography{articles,proceedings,books,unpub}
\end{frame}

\end{document}

### 2021-08-03 cstr_flash_encdeckooptrain.py

@@ -1,0 +1,157 @@
+# [depends] %LIB%/hybridid.py %LIB%/HybridModelLayers.py
+# [depends] cstr_flash_parameters.pickle
+# [makes] pickle
+""" Script to train the hybrid model for the 
+    three reaction system.
+    Pratyush Kumar, pratyushkumar@ucsb.edu """
+
+import sys
+sys.path.append('lib/')
+import tensorflow as tf
+import time
+import numpy as np
+from hybridid import (PickleTool, SimData)
+from HybridModelLayers import KoopmanEncDecModel
+from hybridid import get_scaling, get_train_val_data
+
+# Set the tensorflow global and graph-level seed.
+tf.random.set_seed(2)
+
+def create_model(*, Np, enc_dims, dec_dims, cstr_flash_parameters):
+    """ Create/compile the two reaction model for training. """
+
+    Ny, Nu = cstr_flash_parameters['Ny'], cstr_flash_parameters['Nu']
+    model = KoopmanEncDecModel(Np=Np, Ny=Ny, Nu=Nu,
+                               enc_dims=enc_dims, dec_dims=dec_dims)
+    model.compile(optimizer='adam', loss='mean_squared_error',
+                  loss_weights = [1., 0.])
+    
+    # Return the compiled model.
+    return model
+
+def train_model(model, train_data, trainval_data, stdout_filename, ckpt_path):
+    """ Function to train the NN controller. """
+    # Std out.
+    sys.stdout = open(stdout_filename, 'w')
+    # Create the checkpoint callback.
+    checkpoint_callback = tf.keras.callbacks.ModelCheckpoint(filepath=ckpt_path,
+                                                    monitor='val_loss',
+                                                    save_best_only=True,
+                                                    save_weights_only=True,
+                                                    verbose=1)
+    # Call the fit method to train.
+    model.fit(x = [train_data['inputs'], train_data['yz0']],
+              y = [train_data['yz'], train_data['outputs']],
+              epochs=10000, batch_size=24,
+        validation_data = ([trainval_data['inputs'], trainval_data['yz0']], 
+                           [trainval_data['yz'], trainval_data['outputs']]),
+        callbacks = [checkpoint_callback])
+
+def get_val_predictions(model, val_data, xuyscales, ckpt_path):
+    """ Get the validation predictions. """
+    
+    # Load best weights during training.
+    model.load_weights(ckpt_path)
+    
+    # Predict.
+    model_predictions = model.predict(x=[val_data['inputs'], 
[…]

### 2021-08-03 tworeac_distmodels_nonlin.py

@@ -1,0 +1,135 @@
+# [depends] %LIB%/hybridid.py tworeac_parameters_nonlin.pickle
+# [depends] tworeac_train_nonlin.pickle
+# [makes] pickle
+""" Script to use the grey-box models to evaluate 
+    different disturbance models.
+    Pratyush Kumar, pratyushkumar@ucsb.edu """
+
+import sys
+sys.path.append('lib/')
+sys.path.append('../')
+import numpy as np
+import casadi
+import matplotlib.pyplot as plt
+import mpctools as mpc
+import itertools
+from matplotlib.backends.backend_pdf import PdfPages
+from hybridid import (PickleTool, PAPER_FIGSIZE, c2d,
+                      plot_profit_curve, KalmanFilter,
+                      get_augmented_matrices_for_filter)
+from tworeac_parameters_nonlin import _get_tworeac_model
+
+def _get_linear_greybox_model(*, parameters):
+    """ Check the observability of the original linear system
+        and compute the matrix required to predict the correct 
+        grey-box state evolution. """
+    # Measurement matrix for the plant.
+    sample_time = parameters['sample_time']
+    tau = parameters['ps'].squeeze()
+    Ng = parameters['Ng']
+    C = np.eye(Ng)
+    # Get the continuous time A/B matrices.
+    k1 = parameters['k1']
+    A = np.array([[-k1-(1/tau), 0.], 
+                  [k1, -(1/tau)]])
+    B = np.array([[1/tau], [0.]]) 
+    (Ad, Bd) = c2d(A, B, sample_time)
+    return (Ad, Bd, C)
+
+def _get_yps_ygs_dists_cost_estimates(*, us, dist_model, parameters):
+    """ Do a closed-loop simulation and get disturbance estimates. """
+    # Construct and return the plant.
+    (Bd, Cd) = dist_model
+    ps = parameters['ps']
+    (Nx, Ng, Nu, Ny) = (parameters['Nx'], parameters['Ng'], 
+                        parameters['Nu'], parameters['Ny'])
+    (Nsim, Nd) = (120, Ny)
+
+    # Get the grey-box/dist model and Kalman filter.
+    (Bd, Cd) = dist_model
+    (A, B, C) = _get_linear_greybox_model(parameters=parameters)
+    (Qwx, Qwd, Rv) = (1e-6*np.eye(Ng), 1e-3*np.eye(Nd), 1e-8*np.eye(Ny))
+    (Aaug, Baug, 
[…]

### 2021-08-03 tworeac_encdeckooptrain_nonlin.py

@@ -1,0 +1,157 @@
+# [depends] %LIB%/hybridid.py %LIB%/HybridModelLayers.py
+# [depends] %LIB%/../tworeac_nonlin_funcs.py
+# [depends] tworeac_parameters_nonlin.pickle
+# [makes] pickle
+""" Script to train the deep Koopman model for the
+    three reaction system.
+    Pratyush Kumar, pratyushkumar@ucsb.edu """
+
+import sys
+sys.path.append('lib/')
+import tensorflow as tf
+import time
+import numpy as np
+from hybridid import PickleTool, SimData, get_scaling
+from hybridid import get_train_val_data, get_scaling
+from HybridModelLayers import KoopmanEncDecModel
+
+# Set the tensorflow global and graph-level seed.
+tf.random.set_seed(123)
+
+def create_model(*, Np, enc_dims, dec_dims, tworeac_parameters):
+    """ Create/compile the two reaction model for training. """
+
+    Ny, Nu = tworeac_parameters['Ny'], tworeac_parameters['Nu']
+    model = KoopmanEncDecModel(Np=Np, Ny=Ny, Nu=Nu,
+                               enc_dims=enc_dims, dec_dims=dec_dims)
+    model.compile(optimizer='adam', loss='mean_squared_error',
+                  loss_weights = [1., 0.])
+
+    # Return the compiled model.
+    return model
+
+def train_model(model, train_data, trainval_data, 
+                stdout_filename, ckpt_path):
+    """ Function to train the Koopman model."""
+    # Std out.
+    sys.stdout = open(stdout_filename, 'w')
+
+    # Create the checkpoint callback.
+    checkpoint_callback = tf.keras.callbacks.ModelCheckpoint(filepath=ckpt_path,
+                                                    monitor='val_loss',
+                                                    save_best_only=True,
+                                                    save_weights_only=True,
+                                                    verbose=1)
+
+    # Call the fit method to train.
+    model.fit(x=[train_data['inputs'], train_data['yz0']], 
+              y=[train_data['yz'], train_data['outputs']], 
+              epochs=3000, batch_size=2,
+        validation_data = ([trainval_data['inputs'], trainval_data['yz0']], 
+                           [trainval_data['yz'], trainval_data['outputs']]),
+            callbacks = [checkpoint_callback])
+
+    # Return.
+    return 
+
+def get_val_predictions(model, val_data, xuyscales, ckpt_path):
+    """ Get validation predictions and metric. """
+
+    model.load_weights(ckpt_path)
[…]

### 2021-08-03 archiveScripts/blackboxID.tex
move some more files.

\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{authblk}
\usepackage{amssymb}
%\usepackage{lucidbry}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{subcaption}
\usepackage{physics}
\usepackage[dvipsnames]{xcolor}
\usepackage{tikz,tikzsettings}
\usepackage{verbatim}
\usetikzlibrary{arrows,shapes}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{verbatim}
\usepackage{float}
\usepackage[round]{natbib}
\usepackage[english]{babel}
\bibliographystyle{plainnat}
%\setcitestyle{numeric,open={(},close={)}} \usepackage{color}
%\usepackage{colortbl}
\usetikzlibrary{decorations.markings}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{shapes,calc,backgrounds,arrows,fit,decorations.pathmorphing,matrix}
\usepackage[margin=0.79in]{geometry}

\graphicspath{{./}{./figures/}{./figures/paper/}}

\newcommand{\Is}{\bm{\mathrm{I}}} % Resources.
\newcommand{\Js}{\bm{\mathrm{J}}} % Units.
\newcommand{\Ks}{\bm{\mathrm{K}}} % Resources.
\newcommand{\Ms}{\bm{\mathrm{M}}} % Transition catalog.
\newcommand{\Ns}{\bm{\mathrm{N}}} % Transition counter.

\newcommand{\useq}{\mathbf{u}} \newcommand{\xseq}{\mathbf{x}}
\newcommand{\bbR}{\mathbb{R}} \newcommand{\bbW}{\mathbb{W}}
\newcommand{\bbU}{\mathbb{U}} \newcommand{\bbI}{\mathbb{I}}
\newcommand{\bbX}{\mathbb{X}} \newcommand{\bbP}{\mathbb{P}}

\title{Black-box system identification using neural networks and Koopman operators}
\author{Pratyush Kumar}
\date{\today}

\begin{document}

\maketitle

\section{Process modeling}
Consider a nonlinear plant evolving in continuous time as
\begin{align*}
  \dot{x}_p &= f_p(x_p, u, w) \\
  y &= h_p(x_p) + v
\end{align*}
in which $x_p \in \bbR^{n_p}$ is the plant state, $u \in \bbU$ is the
manipulated control input, $y \in \bbR^p$ is the measurement, $w$ is the process
noise, and $v$ is the measurement noise.

\subsection{Neural network model}
First, we model the input-output relationship of the plant using a black-box
neural network that uses past control inputs and measurements to predict the
future measurements of the plant. We denote this past history as follows:
\begin{align*}
  \mathbf{y}_{k-N_p:k-1} &= [y(k-N_p)', ... \ , y(k-1)']' \\
  \mathbf{u}_{k-N_p:k-1} &= [u(k-N_p)', ... \ , u(k-1)']' \\
  z(k) &= [\mathbf{y}_{k-N_p:k-1}', \mathbf{u}_{k-N_p:k-1}']'
\end{align*}

in which $N_p$ is the number of past measurements and control inputs used. The
state of the dynamic model represented by the neural network is 
\begin{align*}
  x(k) &= [y(k)', z(k)']'
\end{align*}
and the overall model is
\begin{align*}
  x^+ = \begin{bmatrix}
    y^+ \\
    z^+
  \end{bmatrix} &= \begin{bmatrix}
    f_N(x, u) \\
    f_z(x)
  \end{bmatrix}, \quad y = \begin{bmatrix}
    I & 0
  \end{bmatrix}x
\end{align*}

\subsubsection{Training}
The training problem is to find the parameters in the neural network
$f_N$ by minimizing the following mean squared error:

\begin{align*}
  \underset{W_i, b_i}{\textnormal{min}} \sum_{k=0}^{N_t-1} 
  \dfrac{1}{N_t}(\hat{y}(k) - y(k))^2
\end{align*}

\subsection{Deep Koopman operator model}

Next, we study black-box linear system identification using neural networks as Koopman operators to lift the state-space in a higher dimensional space. The
state of the dynamic model is represented as follows
\begin{align*}
  x(k) &= [y(k)', z(k)']'
\end{align*}
and the overall model is
\begin{align*}
  x_{kp}^+ = Ax_{kp} + Bu, \quad x_{kp} = [x', f_N(x)']', \quad x = \begin{bmatrix}
    I & 0
  \end{bmatrix}x_{kp}, \quad y = \begin{bmatrix}
    I & 0
  \end{bmatrix}x
\end{align*}

\subsubsection{Training}
The training problem is to find the parameters in the neural network
$f_N$ and the linear model matrices $A$ and $B$ by minimizing the following mean squared error:

\begin{align*}
  \underset{A, B, W_i, b_i}{\textnormal{min}} \sum_{k=0}^{N_t-1} 
  \dfrac{1}{N_t}(\hat{x}(k) - x(k))^2
\end{align*}

\subsection{Encoder-Decoder Koopman operator model}

\begin{align*}
  x(k) &= [y(k)', z(k)']'
\end{align*}
The overall model is
\begin{align*}
  x_{kp}^+ = Ax_{kp} + Bu, \quad x_{kp} = f_{EN}(x), \quad x = f_{DN}(x_{kp}), \quad y = \begin{bmatrix}
    I & 0
  \end{bmatrix}x
\end{align*}

\subsubsection{Training}
The training problem is to find the parameters in the neural networks
$f_{EN}$, $f_{DN}$ and the linear model matrices $A$ and $B$ by minimizing the following mean squared error:

\begin{align*}
  \underset{A, B, W_i, b_i}{\textnormal{min}} \sum_{k=0}^{N_t-1} 
  \dfrac{1}{N_t}(\hat{x}(k) - x(k))^2
\end{align*}

\end{document}

### 2021-10-06 paper.tex
add tex files.
add a plain paper.

--- 2021-03-14 sysID.tex
@@
-\documentclass[10pt]{article}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{authblk}
-\usepackage{amssymb}
-%\usepackage{lucidbry}
+\documentclass[preprint,5p, twocolumn, authoryear]{elsarticle}
+
+%% Use the option review to obtain double line spacing %
+%\documentclass[preprint,review,12pt]{elsarticle}
+
+%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn % for a
+%journal layout: % \documentclass[final,1p,times]{elsarticle} %
+%\documentclass[final,1p,times,twocolumn]{elsarticle} %
+%\documentclass[final,3p,times]{elsarticle} %
+%\documentclass[final,3p,times,twocolumn]{elsarticle} %
+%\documentclass[final,5p,times]{elsarticle} %
+%\documentclass[final,5p,times,twocolumn]{elsarticle}
+
+%% The graphicx package provides the includegraphics command.
+\usepackage[dvipsnames]{xcolor}
+%\usepackage{color} \usepackage{colortbl} \usepackage{xcolor}
 \usepackage{graphicx}
-\usepackage{latexsym}
-\usepackage{subcaption}
-\usepackage{physics}
-\usepackage[dvipsnames]{xcolor}
-\usepackage{tikz,tikzsettings}
-\usepackage{verbatim}
-\usetikzlibrary{arrows,shapes}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{verbatim}
-\usepackage{float}
-\usepackage{units}
-\usepackage[round]{natbib}
-\usepackage[english]{babel}
-\bibliographystyle{plainnat}
-%\setcitestyle{numeric,open={(},close={)}} \usepackage{color}
-%\usepackage{colortbl}
-\usetikzlibrary{decorations.markings}
-\usetikzlibrary{decorations.pathreplacing}
-\usetikzlibrary{shapes,calc,backgrounds,arrows,fit,decorations.pathmorphing,matrix}
-\usepackage[margin=0.79in]{geometry}
+\usepackage{makecell}
+%% The amssymb package provides various useful mathematical symbols
+\usepackage{amsmath}   % Extra math commands and environments from the AMS
+\usepackage{amssymb}   % Special symbols from the AMS
+\usepackage{amsthm}    % Enhanced theorem and proof environments from the AMS
+%\usepackage{latexsym}  % A few extra LaTeX symbols \usepackage{array}
+%\usepackage{makecell} \usepackage{units} \usepackage{xcolor} % The lineno
+%packages adds line numbers. Start line numbering with % \begin{linenumbers},
+%end it with \end{linenumbers}. Or switch it on % for the whole article with
+%\linenumbers after \end{frontmatter}. \usepackage{lineno}
+%\usepackage[singlelinecheck=false]{caption}
+
+\RequirePackage{booktabs} \RequirePackage{mpcsymbols} \RequirePackage{units_jbr}
+\usepackage{tikz, tikzsettings}
+\usetikzlibrary{shapes,calc,spy, calc, backgrounds,arrows, fit,
+  decorations.pathmorphing, decorations.pathreplacing, matrix, positioning}
+
+%\biboptions{longnamesfirst, semicolon} % natbib.sty is loaded by default.
+%However, natbib options can be % provided with \biboptions{...} command.
+%Following options are % valid:
+
+%%   round  -  round parentheses are used (default) %   square -  square
+%brackets are used   [option] %   curly  -  curly braces are used      {option}
+%%   angle  -  angle brackets are used    <option> %   semicolon  -  multiple
+%citations separated by semi-colon %   colon  - same as semicolon, an earlier
+%confusion %   comma  -  separated by comma %   numbers-  selects numerical
+%citations %   super  -  numerical citations as superscripts %   sort   -  sorts
+%multiple citations according to order in ref. list %   sort&compress   -  like
+%sort, but also compresses numerical citations %   compress - compresses without
+%sorting
+%%
+%% \biboptions{comma,round}
+
+% \biboptions{}
+
+\newcommand{\underl}[1]{\underline{#1}} \newcommand{\overl}[1]{\overline{#1}}
+\newcommand{\satlb}[1]{\textnormal{sat}(#1, \underline{#1}, \overline{#1})}
+\newcommand{\tif}{\textnormal{if}}
+
+\newcommand{\degC}{\ensuremath{^\circ}{\rm C}} \newcommand{\oT}{\overline{T}}
+\newcommand{\uT}{\underline{T}}
 
@@
 
-\newcommand{\Is}{\bm{\mathrm{I}}} % Resources.
-\newcommand{\Js}{\bm{\mathrm{J}}} % Units.
-\newcommand{\Ks}{\bm{\mathrm{K}}} % Resources.
-\newcommand{\Ms}{\bm{\mathrm{M}}} % Transition catalog.
-\newcommand{\Ns}{\bm{\mathrm{N}}} % Transition counter.
+%\newdefinition{assumption}{Assumption} \newdefinition{prop}{Proposition}
+%\newdefinition{definition}{Definition} \newdefinition{rmk}{Remark}
+%\newtheorem{thm}{Theorem}
 
-\newcommand{\useq}{\mathbf{u}} \newcommand{\xseq}{\mathbf{x}}
-\newcommand{\bbR}{\mathbb{R}} \newcommand{\bbW}{\mathbb{W}}
-\newcommand{\bbU}{\mathbb{U}} \newcommand{\bbI}{\mathbb{I}}
-\newcommand{\bbX}{\mathbb{X}} \newcommand{\bbP}{\mathbb{P}}
-
-\title{Nonlinear system identification using machine learning tools}
-\author{Pratyush Kumar}
-\date{\today}
+\journal{Computers and Chemical Engineering}
 
@@
 
-\maketitle
+\begin{frontmatter}
 
-\section{Process modeling}
-We consider the problem of system identification of the following nonlinear
-plant:
-\begin{align*}
-  x_p^+ &= f_p(x_p, u, w) \\
-  y &= h_p(x_p) + v
-\end{align*}
-in which $x_p \in \bbR^{n_p}$ is the plant state, $u \in \bbU \subset \bbR^m$ is
-the manipulated control input, $y \in \bbR^p$ is the measurement, $w$ is the
-process noise, and $v$ is the measurement noise. The objective of developing the
-model is to use it subsequently in model predictive control (MPC). Therefore, we
-consider several types of grey-box, hybrid, and black-box models and analyze
-their predictive capabilities as well as their usefulness for online
-optimization in MPC.
+%% Title, authors and addresses
 
-\section{Model architectures for system identification}
-We describe several process modeling frameworks which determine parameters in
-their proposed model architecture by minimizing the multi-step ahead prediction
-error of the model on training data. A nonlinear optimization problem is solved
-for this identification step. The estimated parameters define the model, which
-is then used subsequently to perform model validation and for use in MPC.
+\title{Hybrid process modeling strategies for real-time optimization and control}
 
-\subsection{Grey-box model}
-We first consider the grey-box modeling strategy, which is the current
-industrial state-of-the-art system identification technique. A parameterized
-model is defined based on first-principle knowledge about the plant, and
-parameters in the model are estimated from training data by solving the
-prediction-error minimization problem. The grey-box model is typically defined
-in continuous time and is of the following form:
+%% use the tnoteref command within \title for footnotes; % use the tnotetext
+%command for the associated footnote; % use the fnref command within \author or
+%\address for footnotes; % use the fntext command for the associated footnote; %
+%use the corref command within \author for corresponding author footnotes; % use
+%the cortext command for the associated footnote; % use the ead command for the
+%email address, % and the form \ead[url] for the home page:
+%%
+%% \title{Title\tnoteref{label1}} % \tnotetext[label1]{}
+\author[label1]{Pratyush Kumar\corref{cor1}}
+\ead{pratyushkumar@ucsb.edu}
+%% \ead[url]{home page} % \fntext[label2]{}
+\cortext[cor1]{Corresponding Author}
+%% \fntext[label3]{}
 
-\begin{align*}
-  \dot{x}_g &= f_g(x_g, u, \theta_g) \\
-  y &= h_g(x_g, \theta_g)
-\end{align*}
+\author[label1]{James B. Rawlings}
+\ead{jbraw@ucsb.edu}
 
-The following nonlinear program (NLP) is solved to identify the parameters
-$\theta_g$:
+\address[label1]{Department of Chemical Engineering, University of California,
+        Santa Barbara, CA 93106, United States}
 
-\begin{align} \label{eq:gb_training}
-  \underset{x_g^i(0), \ \theta_g}{\textnormal{min}} \sum_{i=1}^{N_{tr}} \sum_{k=1}^{N_t} 
-  \dfrac{1}{N_tN_{tr}} &\norm{\tilde{y}^i(k) - y^i(k)}^2_{R_v^{-1}}  \\
-  \textnormal{subject to} \quad \quad x_g^{i+} &= f_{gd}(x_g^i, u^i, \theta_g) \nonumber \\
-   \tilde{y}^i &= h_g(x_g^i, \theta_g) \nonumber
-\end{align}
+\begin{abstract}
 
-in which $\theta_g$ and $x_g^i(0)$ are the decision variables in the
-optimization problem, which are parameters in the grey-box model and initial
-state used for forecasting the predicted measurements respectively. We assume
-that multiple trajectories can be used in the training data, and the superscript
-$i$ is used to denote the trajectory number. We use $N_{tr}$ to denote the total
-number of trajectories in training data and $N_t$ to denote the number of
-samples in each trajectory. The function $f_{g}(\cdot)$ is discretized at the
-measurement sample time to obtain the function $f_{gd}(\cdot)$ that is used to
-make forecasts in the above optimization.
+In industrial applications, model predictive control (MPC) is installed in a
+hierarchical operational framework above regulatory layer proportional-integral
+(PI) controllers. Traditional MPC implementations do not include closed-loop
+dynamics of feedback loops regulated by the PI controllers in MPC optimization
+problems. This paper presents case studies to demonstrate the advantages of
+modeling the closed-loop dynamics of PI feedback loops in MPC optimization
+formulations. We show that this modeling step has two key advantages, (i)
+systematic treatment of unmeasured disturbances in the PI feedback loops, (ii)
+improving closed-loop transient dynamics of cascade control architectures for
+slow PI feedback loops. Case studies on a continuous stirred tank reactor (CSTR)
+example and an heating, ventilation, and air-conditioning (HVAC) application are
+presented to highlight both these advantages of modeling PI controllers in MPC
+formulations.
 
-\subsection{Hybrid model}
+\end{abstract}
 
-Often in industrial applications, first-principle knowledge about the plant is
-incomplete. Hence, grey-box system identification and subsequent use of the
-grey-box model in real-time optimization can lead to suboptimal economic
-performance. We propose to augment the grey-box model using a neural network
-(NN) such that the overall hybrid model can capture any unknown process dynamics
-in the plant, therefore leading to improved economic performance when used
-subsequently in real-time optimization. We define this hybrid model as follows:
+\begin{keyword}
+Cascade control \sep model predictive control \sep proportional-integral control
+\sep economic MPC
+\end{keyword}
 
-\begin{align*}
-  x_g^+ &= f_{gd}(x_g, u, \theta_g) + f_N(z, u, \theta_N) \\
-  z^+ &= f_z(x_g, z, u) \\
-  y &= h_g(x_g, \theta_g)
-\end{align*}
-in which $f_N(\cdot)$ is a NN and $\theta_N$ denotes the parameters in the
-network. We use $z$ to denote the vector of past measurements and control
-inputs. This vector is considered as an additional state in the overall dynamic
-model and is defined as follows:
-\begin{align} \label{eq:pastyu}
-  \mathbf{y}_{k-N_p:k-1} &= [y(k-N_p)', ... \ , y(k-1)']' \nonumber \\
-  \mathbf{u}_{k-N_p:k-1} &= [u(k-N_p)', ... \ , u(k-1)']' \nonumber \\
-  z(k) &= [\mathbf{y}_{k-N_p:k-1}', \mathbf{u}_{k-N_p:k-1}']'
-\end{align}
-in which $N_p$ is the number of past measurements and control inputs used. The
-function $f_z(\cdot)$ describes the dynamics of the state $z$ and its structure
-is defined as follows: 
-\begin{align*}
-  f_z(x_g, z, u) = \begin{bmatrix}
-    y(k-N_p+1) \\
-    y(k-N_p+2) \\
-    \vdots \\
-    h_g(x_g, \theta_g) \\
-    u(k-N_p+1) \\ 
-    u(k-N_p+2) \\
-    \vdots \\
-    u(k)
-  \end{bmatrix}
-\end{align*}
-We assume that the parameters in the grey-box model are first estimated by
-solving the NLP \eqref{eq:gb_training}, and the following similar NLP is solved
-to determine parameters ($\theta_N$) in the NN.
+\end{frontmatter}
 
-\begin{align} \label{eq:hybrid_training}
-  \underset{\theta_N}{\textnormal{min}} \sum_{i=1}^{N_{tr}} \sum_{k=1}^{N_t} 
-  \dfrac{1}{N_tN_{tr}} &\norm{\tilde{y}^i(k) - y^i(k)}^2  \\
-  \textnormal{subject to} \quad \quad x_g^{i+} &= f_{gd}(x_g^i, u^i, \theta_g) + f_z(z^i, u^i, \theta_N) \nonumber \\
-  z^{i+} &= f_z(x_g^i, z^i, u^i) \nonumber \\
-   \tilde{y}^i &= h_g(x_g^i, \theta_g) \nonumber
-\end{align}
-The stochastic gradient descent algorithm Adam is used to solve the above
-optimization using the software tensorflow. Note that we do not estimate the
-initial grey-box state for each trajectory in the optimization problem, and fix
-those states heuristically to make forecasts during the optimization process.
-The control inputs and measurements are scaled as $u := (u -
-u_{\textnormal{MEAN}})/u_{\textnormal{STD}}$ and $y := (y -
-y_{\textnormal{MEAN}})/y_{\textnormal{STD}}$, in which $u_{\textnormal{MEAN}}$
-and $y_{\textnormal{MEAN}}$ are the respective means computed on training data.
-Similarly, $u_{\textnormal{STD}}$ and $y_{\textnormal{STD}}$ are the respective
-standard deviations. Due to this scaling, we do not consider the penalty
-$R_v^{-1}$ in the training objective.
+%% main text
+\section{Introduction} \label{sec:introduction}
 
-\subsection{Black-box models}
-Next, we consider black-box system identification strategies which do not impose
-any a-prior knowledge about the plant and develop the dynamic model purely from
-data. We consider two types of black-box models in this section: NNs and dynamic
-models based on the Koopman operator theory. The latter type of model is
-considered based on the motivation that it imposes a linear structure on the
-dynamics in a high-dimensional space, therefore leading to a linear or quadratic
-program to be solved online in MPC.
+\section*{Acknowledgment}
+The funding for this work was provided by the industrial members of the
+Texas-Wisconsin-California Control Consortium (TWCCC).
 
-\subsubsection{Neural network}
-The black-box NN uses the vector of past measurements and control inputs $z$
-\eqref{eq:pastyu} as the state in the dynamic model and is structured as follows
-\begin{align} \label{eq:bbnn_model}
-  z^+ &= f_z(z, u) \\ 
-  y &= h_N(z, \theta_N) \nonumber
-\end{align}
-in which $h_N(\cdot)$ is a NN and $\theta_N$ are the parameters in the network.
-The function $f_z(\cdot)$ denotes the dynamics of the state $z$ and is defined
-as
-
-\begin{align*}
-  f_z(z, u) = \begin{bmatrix}
-    y(k-N_p+1) \\
-    y(k-N_p+2) \\
-    \vdots \\
-    h_N(z, \theta_N) \\
-    u(k-N_p+1) \\ 
-    u(k-N_p+2) \\
-    \vdots \\
-    u(k)
-  \end{bmatrix}
-\end{align*}
-
-\subsubsection{Koopman operator model}
-
-The Koopman model represents the dynamics as a linear system in a
-high-dimensional state-space that is a nonlinear transformation of the original
-state-space. This transformation is chosen such that the input-output behavior
-is linear. The overall dynamic model is represented as:
-
-\begin{align} \label{eq:koop_model}
-  x_{kp} &= [y', z', f_N(y, z, \theta_N)']' \\
-  x_{kp}^+ &= Ax_{kp} + Bu \nonumber \\
-  y &= [I, \ 0] x_{kp} \nonumber
-\end{align}
-in which $x_{kp} \in \bbR^N$ is the state in a high-dimensional state-space,
-$f_N(\cdot)$ is a NN used for the nonlinear transformation, and $\theta_N$
-denotes the parameters in the network.
-
-\subsubsection{Training}
-The training optimization problem for both the black-box NN and the Koopman
-operator model is to minimize the mean-squared-error (MSE) similar to the
-nonlinear program \eqref{eq:hybrid_training}. The model equality constraints are
-replaced with the equations \eqref{eq:bbnn_model} and \eqref{eq:koop_model} for
-the black-box and Koopman operator model respectively. The measurements and
-control inputs are scaled using the mean and standard deviation computed from
-the training data. The initial state to make model predictions in the
-optimization process is set based on a window of past measurements and control
-inputs in the training data.
-
-\section{Two reaction system example}
-
-We now study the performance of the modeling frameworks on a simple two
-reaction system example in three steps. First, we examine the predictive
-capability of the developed models on open-loop validation data. Second, we
-analyze the steady-state optimum with an economic stage cost. Finally, we
-analyze the optimal control input profiles on a dynamic economic optimization
-problem. These three steps shed light on the predictive abilities of the models
-and their usefulness for real-time optimization in the operation of chemical
-processes.
-
-The following plant model with two reactions $A \rightarrow B$ and $3B
-\rightleftharpoons C$ is considered:
-\begin{align*}
-  \dfrac{dC_A}{dt} &= \dfrac{C_{Af} - C_A}{\tau} - k_1C_A\\
-  \dfrac{dC_B}{dt} &= k_1C_A - 3k_2C^3_B + 3k_3C_C- \dfrac{C_B}{\tau}\\
-  \dfrac{dC_C}{dt} &= k_2C^3_B - k_3C_C - \dfrac{C_C}{\tau}
-\end{align*}
-
-in which, $C_A$, $C_B$, and $C_C$ are the concentrations of the three species
-$A$, $B$, and $C$. The parameters $k_1$, $k_2$, and $k_3$ are the rate constants
-and the $\tau$ is the residence time.
-
-The grey-box model is built as follows based on an assumption that the knowledge
-of the second side reaction ($3B \rightleftharpoons C $) is not available:
-\begin{align*}
-  \dfrac{dC_A}{dt} &= \dfrac{C_{Af} - C_A}{\tau} - k_1C_A\\
-  \dfrac{dC_B}{dt} &= k_1C_A - \dfrac{C_B}{\tau}
-\end{align*}
-
-The rate constants and residence time in both the models are set to $k_1 = 1 \
-\textnormal{m}^3/\textnormal{min}$, $k_2 = 0.01 \
-\textnormal{m}^3/\textnormal{min}$, $k_3 = 0.05 \
-\textnormal{m}^3/\textnormal{min}$, and $\tau = 5 \ \textnormal{min}$. The
-sample time for the measurements is $1 \ \textnormal{min}$, and we assume that
-only the concentrations of the species $A$ and $B$ are measured in real-time.
-
-\subsection{Model validation}
-In the simulations for this example, we fix the grey-box model parameters
-heuristically and only train the two black-box models: NN and Koopman operator
-model. We generate 1 day (1440 samples) of data by simulating the plant model
-with a Pseudo-random-binary-signal (PRBS) control input profile. The generated
-data is split into 12 hours of training data, 6 hours of buffer data, and 6
-hours of validation data. The MSE metric is monitored on the buffer data during
-the training process, and model weights are updated only when MSE decreases on
-the buffer dataset. After the training step, the black-box NN and Koopman
-operator models are used to make predictions on the validation data shown in
-Figure \ref{fig:tworeac_validation}. We observe good quality measurement
-predictions with both types of dynamic black-box models.
-
-\begin{figure}[!h]
-  \centering
-  \includegraphics[page=1, height=0.5\textheight,
-                   width=0.7\textwidth]{tworeac_plots.pdf}
-  \caption{Model validation on open-loop data with the black-box NN and Koopman operator model.}
-  \label{fig:tworeac_validation}
-\end{figure}
-
-\subsection{Steady state optimization}
-The two trained black-box NN and Koopman operator model are used in a
-steady-state optimization problem to determine the economically optimal steady
-state of the system. In a typical chemical plant, this type of optimization is
-solved periodically every few hours based on changes in raw material costs and
-product selling prices. The obtained steady state is sent to lower-level control
-layers such as model predictive control (MPC) or the
-proportional-integral-derivative (PID) controller. The optimization problem
-solved is:
-\begin{align} \label{eq:ss_opt}
-  \underset{u_s}{\textnormal{min}} & \quad \ell(y_s, u_s) \\
-  x_s &= f(x_s, u_s) \nonumber \\ 
-  y_s &= h(x_s) \nonumber \\
-  \underline{u} \leq &u_s \leq \overline{u} \nonumber
-\end{align}
-in which the steady-state control input ($u_s$) is the decision variable,  and
-the model is required to satisfy steady-state constraints. The stage cost for
-the two reaction system is chosen as:
-\begin{equation*}
-  \ell(y_s, u_s) = p_1C_{Afs} - p_2C_{Bs}
-\end{equation*}
-in which $p_1$ and $p_2$ are the parameters chosen to represent the raw material
-(species $A$) cost and the product (species $B$) selling price respectively.
-
-First, we plot the economic objective ($\ell(y_s, u_s)$) as a function of the
-steady-state control input to examine the cost structure obtained with the
-different models. This plot is shown in Figure \ref{fig:tworeac_sscost}.
-
-\begin{figure}[!h]
-  \centering
-  \includegraphics[page=1, height=0.4\textheight,
-                   width=0.6\textwidth]{tworeac_ssopt.pdf}
-  \caption{Steady-state cost curve $\ell(y_s, u_s)$ as a function of the control input for the plant, grey-box, black-box NN, and Koopman operator models.}
-  \label{fig:tworeac_sscost}
-\end{figure}
-
-We observe that the black-box NN model correctly captures the true plant cost
-curve after training, and the grey-box and Koopman operator models do not. This
-behavior by the latter two models is due to the linear structure in the dynamic
-models. Upon solving the optimization problem \eqref{eq:ss_opt}, the optimum
-steady-state control input ($C_{Afs}^0$) values are $1.63$, $2.5$, $1.60$, and
-$2.5$ in mol/m$^3$ for the plant, grey-box, black-box NN, and Koopman operator
-models respectively. The use of grey-box and Koopman operator models
-subsequently in real-time optimization of chemical processes can therefore
-result in suboptimal economic performance.
-
-\subsection{Dynamic economic optimization}
-The performance of the models is now studied with a dynamic economic
-optimization problem. In contrast to the steady-state optimization problem
-\eqref{eq:ss_opt}, operational strategies based on solving a dynamic
-optimization problem aim to maximize plant profit during the transients as well
-as opposed to just steady states. The economic optimization problem is:
-
-\begin{align} \label{eq:dyn_opt}
-  \underset{\useq}{\textnormal{min}} & \quad \sum_{k=0}^{N-1} \ell(y(i), u(i)) \\
-  x^+ &= f(x, u) \nonumber \\ 
-  y &= h(x) \nonumber \\
-  x(0) &= x_0 \nonumber \\ 
-  \underline{u} \leq &u \leq \overline{u} \nonumber
-\end{align}
-
-The above optimization problem is similar to the optimization solved in MPC, and
-the difference is the choice of the economic stage cost $\ell(\cdot)$, which is
-a setpoint tracking stage cost in a traditional MPC controller. This
-optimization problem can also be solved directly at the regulatory control
-layers and is often referred to as the economic MPC controller.
-
-\begin{figure}[!h]
-  \centering
-  \includegraphics[page=1, height=0.5\textheight,
-                   width=0.7\textwidth]{tworeac_openloop.pdf} \caption{Optimal
-                   open-loop control input and state profiles after solving the
-                   dynamic economic MPC problem \eqref{eq:dyn_opt} with the
-                   plant, grey-box, black-box NN, and Koopman operator models.}
-  \label{fig:tworeac_openloop}
-\end{figure}
-
-Before analyzing a full closed-loop simulation with the economic MPC controllers
-based on the different dynamic models, we first examine the open-loop solutions
-of the optimization problem \eqref{eq:dyn_opt} given one initial state. Figure
-\ref{fig:tworeac_openloop} shows the open-loop state and control input profiles
-for the plant, grey-box, black-box NN, and Koopman operator models. We notice
-that the black-box NN has an optimal control input profile qualitatively similar
-to that of the plant. Whereas, the optimal profiles for the grey-box and Koopman
-operator model are different than the plant model. This simple analysis shows
-that the use of the grey-box and Koopman operator models in MPC can result in
-suboptimal economic performance.
-
-\section{Large, CSTR and flash system example}
-We next study the performance of the black-box NN modeling framework on a large chemical process that consists of a chemical reactor and flash separator in series.
-The process is depicted in Figure \ref{fig:cstr_flash}, and the ordinary differential equations (ODEs) with parameter values describing the plant model are shown in Appendix \ref{app:cstr_flash}.
-
-\begin{figure}[!h]
-  \centering
-  \includegraphics[page=1, height=0.3\textheight,
-                   width=0.6\textwidth]{cstr_flash.pdf}
-  \caption{Schematic of the CSTR and flash chemical process.}
-  \label{fig:cstr_flash}
-\end{figure}
-
-\subsection{Model validation}
-
-\subsection{Steady state optimization}
-
-\subsection{Dynamic economic optimization}
-
-\section{Discussion}
-
-\section*{Appendix} 
-\renewcommand{\thesubsection}{\Alph{subsection}}
-
-\subsection{CSTR and flash system model} \label{app:cstr_flash}
-
-\begin{align*}
-  \dfrac{dH_r}{dt} &= \dfrac{F + D -F_r}{A_r}\\
-  \dfrac{dC_{Ar}}{dt} &= \dfrac{F(C_{Af} -C_{Ar}) +
-                         D(C_{Ad} -C_{Ar})}{A_rH_r} - r_1 \\
-  \dfrac{dC_{Br}}{dt} &= \dfrac{-FC_{Br} + 
-                          D(C_{Bd} -C_{Br})}{A_rH_r} + r_1 -3r_2\\
-  \dfrac{dC_{Cr}}{dt} &= \dfrac{-FC_{Cr} + 
-  D(C_{Cd} -C_{Cr})}{A_rH_r} + r_2\\
-  \dfrac{dT_r}{dt} &= \dfrac{F(T_f - T_r) + D(T_d -T_r)}{A_rH_r} + 
-                      \dfrac{r_1\Delta H_1 + r_2\Delta H_2}{\rho C_p} - 
-                      \dfrac{Q_r}{\rho A_r C_p H_r}\\
-\end{align*}
-
-\begin{align*}
-  \dfrac{dH_b}{dt} &= \dfrac{F_r - F_b - D}{A_b} \\
-  \dfrac{dC_{Ab}}{dt} &= \dfrac{F_r(C_{Ar} -C_{Ab}) + 
-                          D(C_{Ab} -C_{Ad})}{A_bH_b} \\
-  \dfrac{dC_{Bb}}{dt} &= \dfrac{F_r(C_{Br} -C_{Bb}) + 
-                          D(C_{Bb} -C_{Bd})}{A_bH_b} \\
-  \dfrac{dC_{Cb}}{dt} &= \dfrac{F_r(C_{Cr} -C_{Cb}) + 
-                          D(C_{Cb} -C_{Cd})}{A_bH_b} \\
-  \dfrac{dT_b}{dt} &= \dfrac{F_r(T_r - T_b)}{A_bH_b} +
-                      \dfrac{Q_b}{\rho A_b C_p H_b}\\
-\end{align*}
-
-Here, the rate of reactions $r_1$ and $r_2$ are defined as:
-\begin{align*}
-  r_1 = k_1^*e^{-E_1/RT_r}C_{Ar}, \quad r_2 = k_2^*e^{-E_2/RT_r}C_{Br}^3
-\end{align*}
-
-The outlet flowrates from the CSTR and flash are computed from the height of the reaction mixture in each unit as follows:
-\begin{align*}
-  F_r = k_r\sqrt{H_r}, \quad F_b = k_b\sqrt{H_b}
-\end{align*}
-
-The concentrations of the species in the recycle stream are computed using the following equations:
-\begin{align*}
-  C_{Ad} = \dfrac{\alpha_A C_{Ab}}{\sum_{i \in \{ A, B, C\}} \alpha_iC_{ib}} , \quad, C_{Bd} = \dfrac{\alpha_B C_{Bb}}{\sum_{i \in \{ A, B, C\}} \alpha_iC_{ib}}, \quad C_{Cd} = \dfrac{\alpha_C C_{Cb}}{\sum_{i \in \{ A, B, C\}} \alpha_iC_{ib}}
-\end{align*}
-
-The parameters in this model and the control input constraints used for the optimization problems are summarized in Table \ref{table:cstr_flash_pars}.
-
-\begin{table}[!h]
-  \centering
-  \begin{tabular}{|c|c|c|c|c|c|}
-      \hline
-      \multicolumn{6}{ |c| }{Parameters used to simulate the ODEs} \\
-      \hline
-      Parameter & Value & Unit & Parameter & Value & Unit\\
-      \hline
-      $\alpha_A$ & 8 &  & $k_1^*$ & 0.3 & \unit{min$^{-1}$}\\ 
-      $\alpha_B$ & 1 &  & $k_2^*$ & 0.5 & \unit{min$^{-1}$} \\ 
-      $\alpha_C$ & 1 &  & $\Delta H_1$ & 100 & \unitfrac{kJ}{mol} \\ 
-      $\rho$ & 6 & \unitfrac{kg}{m$^3$} & $\Delta H_2$ & 120 &
-      \unitfrac{kJ}{mol} \\ 
-      $C_p$ & 3 & \unitfrac{kJ}{kg-K} & $E_1/R$ & 200 & \unit{K} \\ 
-      $A_r$ & 3 & \unit{m$^2$} & $E_2/R$ & 300 & \unit{K} \\ 
-      $A_b$ & 3 & \unit{m$^2$} & $T_d$ & 310 &
-      \unit{K} \\ 
-      $k_r$ & 4 & \unitfrac{m$^{5/2}$}{min} & $C_{Af}$ & 6 & \unitfrac{mol}{m$^3$} \\ 
-      $k_b$ & 2 & \unitfrac{m$^{5/2}$}{min} &  $Q_r$ &  200 & \unitfrac{kJ}{min} \\ 
-      $T_f$ & 320 & \unit{K} &  & &  \\ 
-      $Q_b$ & 200 & \unitfrac{kJ}{min} & & & \\ 
-      \hline       
-      \multicolumn{6}{ |c| }{Actuator constraints ($\underline{u},
-      \overline{u}$)} \\
-      \hline
-      $F$ & (5, 15) & \unitfrac{m$^3$}{min} & $D$ & (2, 8) &
-      \unitfrac{m$^3$}{min}  \\ 
-      \hline
-  \end{tabular} \caption{Parameters used in the ODEs to simulate the plant in
-  the CSTR in series with a flash separator example, and the actuator constraints
-  for the optimization problems.}
-\label{table:cstr_flash_pars}
-\end{table}
+%% New version of the num-names style
+\bibliographystyle{elsarticle-num-names}
+\bibliography{abbreviations,articles,books,unpub,proceedings}
 

### 2021-10-08 minireport.tex
add minireport.

--- 2021-03-14 sysID.tex
@@
 
-\title{Nonlinear system identification using machine learning tools}
+\title{Hybrid process modeling strategies for real time optimization and control}
 \author{Pratyush Kumar}
@@
 
-Often in industrial applications, first-principle knowledge about the plant is
+Often in industrial applicationns, first-principle knowledge about the plant is
 incomplete. Hence, grey-box system identification and subsequent use of the
@@
 
-\section{Two reaction system example}
-
-We now study the performance of the modeling frameworks on a simple two
-reaction system example in three steps. First, we examine the predictive
-capability of the developed models on open-loop validation data. Second, we
-analyze the steady-state optimum with an economic stage cost. Finally, we
-analyze the optimal control input profiles on a dynamic economic optimization
-problem. These three steps shed light on the predictive abilities of the models
-and their usefulness for real-time optimization in the operation of chemical
-processes.
-
-The following plant model with two reactions $A \rightarrow B$ and $3B
-\rightleftharpoons C$ is considered:
-\begin{align*}
-  \dfrac{dC_A}{dt} &= \dfrac{C_{Af} - C_A}{\tau} - k_1C_A\\
-  \dfrac{dC_B}{dt} &= k_1C_A - 3k_2C^3_B + 3k_3C_C- \dfrac{C_B}{\tau}\\
-  \dfrac{dC_C}{dt} &= k_2C^3_B - k_3C_C - \dfrac{C_C}{\tau}
-\end{align*}
-
-in which, $C_A$, $C_B$, and $C_C$ are the concentrations of the three species
-$A$, $B$, and $C$. The parameters $k_1$, $k_2$, and $k_3$ are the rate constants
-and the $\tau$ is the residence time.
-
-The grey-box model is built as follows based on an assumption that the knowledge
-of the second side reaction ($3B \rightleftharpoons C $) is not available:
-\begin{align*}
-  \dfrac{dC_A}{dt} &= \dfrac{C_{Af} - C_A}{\tau} - k_1C_A\\
-  \dfrac{dC_B}{dt} &= k_1C_A - \dfrac{C_B}{\tau}
-\end{align*}
-
-The rate constants and residence time in both the models are set to $k_1 = 1 \
-\textnormal{m}^3/\textnormal{min}$, $k_2 = 0.01 \
-\textnormal{m}^3/\textnormal{min}$, $k_3 = 0.05 \
-\textnormal{m}^3/\textnormal{min}$, and $\tau = 5 \ \textnormal{min}$. The
-sample time for the measurements is $1 \ \textnormal{min}$, and we assume that
-only the concentrations of the species $A$ and $B$ are measured in real-time.
-
-\subsection{Model validation}
-In the simulations for this example, we fix the grey-box model parameters
-heuristically and only train the two black-box models: NN and Koopman operator
-model. We generate 1 day (1440 samples) of data by simulating the plant model
-with a Pseudo-random-binary-signal (PRBS) control input profile. The generated
-data is split into 12 hours of training data, 6 hours of buffer data, and 6
-hours of validation data. The MSE metric is monitored on the buffer data during
-the training process, and model weights are updated only when MSE decreases on
-the buffer dataset. After the training step, the black-box NN and Koopman
-operator models are used to make predictions on the validation data shown in
-Figure \ref{fig:tworeac_validation}. We observe good quality measurement
-predictions with both types of dynamic black-box models.
-
-\begin{figure}[!h]
-  \centering
-  \includegraphics[page=1, height=0.5\textheight,
-                   width=0.7\textwidth]{tworeac_plots.pdf}
-  \caption{Model validation on open-loop data with the black-box NN and Koopman operator model.}
-  \label{fig:tworeac_validation}
-\end{figure}
-
-\subsection{Steady state optimization}
-The two trained black-box NN and Koopman operator model are used in a
-steady-state optimization problem to determine the economically optimal steady
-state of the system. In a typical chemical plant, this type of optimization is
-solved periodically every few hours based on changes in raw material costs and
-product selling prices. The obtained steady state is sent to lower-level control
-layers such as model predictive control (MPC) or the
-proportional-integral-derivative (PID) controller. The optimization problem
-solved is:
-\begin{align} \label{eq:ss_opt}
-  \underset{u_s}{\textnormal{min}} & \quad \ell(y_s, u_s) \\
-  x_s &= f(x_s, u_s) \nonumber \\ 
-  y_s &= h(x_s) \nonumber \\
-  \underline{u} \leq &u_s \leq \overline{u} \nonumber
-\end{align}
-in which the steady-state control input ($u_s$) is the decision variable,  and
-the model is required to satisfy steady-state constraints. The stage cost for
-the two reaction system is chosen as:
-\begin{equation*}
-  \ell(y_s, u_s) = p_1C_{Afs} - p_2C_{Bs}
-\end{equation*}
-in which $p_1$ and $p_2$ are the parameters chosen to represent the raw material
-(species $A$) cost and the product (species $B$) selling price respectively.
-
-First, we plot the economic objective ($\ell(y_s, u_s)$) as a function of the
-steady-state control input to examine the cost structure obtained with the
-different models. This plot is shown in Figure \ref{fig:tworeac_sscost}.
-
-\begin{figure}[!h]
-  \centering
-  \includegraphics[page=1, height=0.4\textheight,
-                   width=0.6\textwidth]{tworeac_ssopt.pdf}
-  \caption{Steady-state cost curve $\ell(y_s, u_s)$ as a function of the control input for the plant, grey-box, black-box NN, and Koopman operator models.}
-  \label{fig:tworeac_sscost}
-\end{figure}
-
-We observe that the black-box NN model correctly captures the true plant cost
-curve after training, and the grey-box and Koopman operator models do not. This
-behavior by the latter two models is due to the linear structure in the dynamic
-models. Upon solving the optimization problem \eqref{eq:ss_opt}, the optimum
-steady-state control input ($C_{Afs}^0$) values are $1.63$, $2.5$, $1.60$, and
-$2.5$ in mol/m$^3$ for the plant, grey-box, black-box NN, and Koopman operator
-models respectively. The use of grey-box and Koopman operator models
-subsequently in real-time optimization of chemical processes can therefore
-result in suboptimal economic performance.
-
-\subsection{Dynamic economic optimization}
-The performance of the models is now studied with a dynamic economic
-optimization problem. In contrast to the steady-state optimization problem
-\eqref{eq:ss_opt}, operational strategies based on solving a dynamic
-optimization problem aim to maximize plant profit during the transients as well
-as opposed to just steady states. The economic optimization problem is:
-
-\begin{align} \label{eq:dyn_opt}
-  \underset{\useq}{\textnormal{min}} & \quad \sum_{k=0}^{N-1} \ell(y(i), u(i)) \\
-  x^+ &= f(x, u) \nonumber \\ 
-  y &= h(x) \nonumber \\
-  x(0) &= x_0 \nonumber \\ 
-  \underline{u} \leq &u \leq \overline{u} \nonumber
-\end{align}
-
-The above optimization problem is similar to the optimization solved in MPC, and
-the difference is the choice of the economic stage cost $\ell(\cdot)$, which is
-a setpoint tracking stage cost in a traditional MPC controller. This
-optimization problem can also be solved directly at the regulatory control
-layers and is often referred to as the economic MPC controller.
-
-\begin{figure}[!h]
-  \centering
-  \includegraphics[page=1, height=0.5\textheight,
-                   width=0.7\textwidth]{tworeac_openloop.pdf} \caption{Optimal
-                   open-loop control input and state profiles after solving the
-                   dynamic economic MPC problem \eqref{eq:dyn_opt} with the
-                   plant, grey-box, black-box NN, and Koopman operator models.}
-  \label{fig:tworeac_openloop}
-\end{figure}
-
-Before analyzing a full closed-loop simulation with the economic MPC controllers
-based on the different dynamic models, we first examine the open-loop solutions
-of the optimization problem \eqref{eq:dyn_opt} given one initial state. Figure
-\ref{fig:tworeac_openloop} shows the open-loop state and control input profiles
-for the plant, grey-box, black-box NN, and Koopman operator models. We notice
-that the black-box NN has an optimal control input profile qualitatively similar
-to that of the plant. Whereas, the optimal profiles for the grey-box and Koopman
-operator model are different than the plant model. This simple analysis shows
-that the use of the grey-box and Koopman operator models in MPC can result in
-suboptimal economic performance.
-
-\section{Large, CSTR and flash system example}
-We next study the performance of the black-box NN modeling framework on a large chemical process that consists of a chemical reactor and flash separator in series.
-The process is depicted in Figure \ref{fig:cstr_flash}, and the ordinary differential equations (ODEs) with parameter values describing the plant model are shown in Appendix \ref{app:cstr_flash}.
-
-\begin{figure}[!h]
-  \centering
-  \includegraphics[page=1, height=0.3\textheight,
-                   width=0.6\textwidth]{cstr_flash.pdf}
-  \caption{Schematic of the CSTR and flash chemical process.}
-  \label{fig:cstr_flash}
-\end{figure}
-
-\subsection{Model validation}
-
-\subsection{Steady state optimization}
-
-\subsection{Dynamic economic optimization}
-
-\section{Discussion}
-
-\section*{Appendix} 
-\renewcommand{\thesubsection}{\Alph{subsection}}
-
-\subsection{CSTR and flash system model} \label{app:cstr_flash}
-
-\begin{align*}
-  \dfrac{dH_r}{dt} &= \dfrac{F + D -F_r}{A_r}\\
-  \dfrac{dC_{Ar}}{dt} &= \dfrac{F(C_{Af} -C_{Ar}) +
-                         D(C_{Ad} -C_{Ar})}{A_rH_r} - r_1 \\
-  \dfrac{dC_{Br}}{dt} &= \dfrac{-FC_{Br} + 
-                          D(C_{Bd} -C_{Br})}{A_rH_r} + r_1 -3r_2\\
-  \dfrac{dC_{Cr}}{dt} &= \dfrac{-FC_{Cr} + 
-  D(C_{Cd} -C_{Cr})}{A_rH_r} + r_2\\
-  \dfrac{dT_r}{dt} &= \dfrac{F(T_f - T_r) + D(T_d -T_r)}{A_rH_r} + 
-                      \dfrac{r_1\Delta H_1 + r_2\Delta H_2}{\rho C_p} - 
-                      \dfrac{Q_r}{\rho A_r C_p H_r}\\
-\end{align*}
-
-\begin{align*}
-  \dfrac{dH_b}{dt} &= \dfrac{F_r - F_b - D}{A_b} \\
-  \dfrac{dC_{Ab}}{dt} &= \dfrac{F_r(C_{Ar} -C_{Ab}) + 
-                          D(C_{Ab} -C_{Ad})}{A_bH_b} \\
-  \dfrac{dC_{Bb}}{dt} &= \dfrac{F_r(C_{Br} -C_{Bb}) + 
-                          D(C_{Bb} -C_{Bd})}{A_bH_b} \\
-  \dfrac{dC_{Cb}}{dt} &= \dfrac{F_r(C_{Cr} -C_{Cb}) + 
-                          D(C_{Cb} -C_{Cd})}{A_bH_b} \\
-  \dfrac{dT_b}{dt} &= \dfrac{F_r(T_r - T_b)}{A_bH_b} +
-                      \dfrac{Q_b}{\rho A_b C_p H_b}\\
-\end{align*}
-
-Here, the rate of reactions $r_1$ and $r_2$ are defined as:
-\begin{align*}
-  r_1 = k_1^*e^{-E_1/RT_r}C_{Ar}, \quad r_2 = k_2^*e^{-E_2/RT_r}C_{Br}^3
-\end{align*}
-
-The outlet flowrates from the CSTR and flash are computed from the height of the reaction mixture in each unit as follows:
-\begin{align*}
-  F_r = k_r\sqrt{H_r}, \quad F_b = k_b\sqrt{H_b}
-\end{align*}
-
-The concentrations of the species in the recycle stream are computed using the following equations:
-\begin{align*}
-  C_{Ad} = \dfrac{\alpha_A C_{Ab}}{\sum_{i \in \{ A, B, C\}} \alpha_iC_{ib}} , \quad, C_{Bd} = \dfrac{\alpha_B C_{Bb}}{\sum_{i \in \{ A, B, C\}} \alpha_iC_{ib}}, \quad C_{Cd} = \dfrac{\alpha_C C_{Cb}}{\sum_{i \in \{ A, B, C\}} \alpha_iC_{ib}}
-\end{align*}
-
-The parameters in this model and the control input constraints used for the optimization problems are summarized in Table \ref{table:cstr_flash_pars}.
-
-\begin{table}[!h]
-  \centering
-  \begin{tabular}{|c|c|c|c|c|c|}
-      \hline
-      \multicolumn{6}{ |c| }{Parameters used to simulate the ODEs} \\
-      \hline
-      Parameter & Value & Unit & Parameter & Value & Unit\\
-      \hline
-      $\alpha_A$ & 8 &  & $k_1^*$ & 0.3 & \unit{min$^{-1}$}\\ 
-      $\alpha_B$ & 1 &  & $k_2^*$ & 0.5 & \unit{min$^{-1}$} \\ 
-      $\alpha_C$ & 1 &  & $\Delta H_1$ & 100 & \unitfrac{kJ}{mol} \\ 
-      $\rho$ & 6 & \unitfrac{kg}{m$^3$} & $\Delta H_2$ & 120 &
-      \unitfrac{kJ}{mol} \\ 
-      $C_p$ & 3 & \unitfrac{kJ}{kg-K} & $E_1/R$ & 200 & \unit{K} \\ 
-      $A_r$ & 3 & \unit{m$^2$} & $E_2/R$ & 300 & \unit{K} \\ 
-      $A_b$ & 3 & \unit{m$^2$} & $T_d$ & 310 &
-      \unit{K} \\ 
-      $k_r$ & 4 & \unitfrac{m$^{5/2}$}{min} & $C_{Af}$ & 6 & \unitfrac{mol}{m$^3$} \\ 
-      $k_b$ & 2 & \unitfrac{m$^{5/2}$}{min} &  $Q_r$ &  200 & \unitfrac{kJ}{min} \\ 
-      $T_f$ & 320 & \unit{K} &  & &  \\ 
-      $Q_b$ & 200 & \unitfrac{kJ}{min} & & & \\ 
-      \hline       
-      \multicolumn{6}{ |c| }{Actuator constraints ($\underline{u},
-      \overline{u}$)} \\
-      \hline
-      $F$ & (5, 15) & \unitfrac{m$^3$}{min} & $D$ & (2, 8) &
-      \unitfrac{m$^3$}{min}  \\ 
-      \hline
-  \end{tabular} \caption{Parameters used in the ODEs to simulate the plant in
-  the CSTR in series with a flash separator example, and the actuator constraints
-  for the optimization problems.}
-\label{table:cstr_flash_pars}
-\end{table}
-
 \end{document}

### 2021-10-09 SimStudiesAug26-Sept8/BlackBoxFuncs.py
add some files.

@@ -1,0 +1,286 @@
+# [depends] hybridId.py
+import sys
+import numpy as np
+import tensorflow as tf
+from hybridId import SimData
+
+def tanh(x, tF=True, a=1):
+    """ Custom tanh function. The input to tanh 
+        can be scaled if required. """
+    if tF:
+        num = tf.math.exp(a*x) - tf.math.exp(-a*x)
+        den = tf.math.exp(a*x) + tf.math.exp(-a*x)
+    else:
+        num = np.exp(a*x) - np.exp(-a*x)
+        den = np.exp(a*x) + np.exp(-a*x)
+    # Return.
+    return num/den
+
+def fnnTf(nnInput, nnLayers):
+    """ Compute the output of a feedforward network, 
+        with tensorflow inputs. """
+    nnOutput = nnInput
+    for layer in nnLayers:
+        nnOutput = layer(nnOutput)
+    # Return the final output.
+    return nnOutput
+
+def createDenseLayers(nnDims):
+    """ Create dense layers based on the feed-forward NN layer dimensions.
+        nnDims: List that contains the dimensions of the feed forward NN.
+        nnLayers: Output of the feedforward NN.
+    """
+    nnLayers = []
+    for dim in nnDims[1:-1]:
+        nnLayers += [tf.keras.layers.Dense(dim, activation=tanh)]
+    nnLayers += [tf.keras.layers.Dense(nnDims[-1])]
+    # Return.
+    return nnLayers
+
+def fnn(nnInput, nnWeights):
+    """ Compute the output of a feedforward network, 
+        with inputs and weights as numpy arrays. """
+
+    # Check that the input has only one dimension, and add one extra.
+    assert nnInput.ndim == 1 
+    nnOutput = nnInput[:, np.newaxis]
+
+    # Loop over layers.
+    for i in range(0, len(nnWeights)-2, 2):
+        W, b = nnWeights[i:i+2]
+        nnOutput = W.T @ nnOutput + b[:, np.newaxis]
+        nnOutput = tanh(nnOutput, tF=False)
+    Wf, bf = nnWeights[-2:]
+    
+    # Return output in the same number of dimensions as input.
+    nnOutput = Wf.T @ nnOutput + bf[:, np.newaxis]
+    nnOutput = nnOutput[:, 0]
+
+    # Return.
+    return nnOutput
+
[…]

### 2021-10-09 SimStudiesAug26-Sept8/ReacHybridFullGbFuncs.py

@@ -1,0 +1,350 @@
+# [depends] BlackBoxFuncs.py hybridId.py
+import sys
+import numpy as np
+import mpctools as mpc
+import tensorflow as tf
+from BlackBoxFuncs import fnnTf, fnn, createDenseLayers
+from hybridId import SimData
+
+class ReacFullGbCell(tf.keras.layers.AbstractRNNCell):
+    """
+    RNN Cell
+    x_g^+ = f_g(x_g, u; NN(x_g))
+    """
+    def __init__(self, r1Layers, r2Layers, r3Layers,
+                       xuyscales, hyb_fullgb_pars, **kwargs):
+        super(ReacFullGbCell, self).__init__(**kwargs)
+        
+        # Attributes.
+        self.r1Layers = r1Layers
+        self.r2Layers = r2Layers
+        self.r3Layers = r3Layers
+        self.xuyscales = xuyscales
+        self.hyb_fullgb_pars = hyb_fullgb_pars
+
+    @property
+    def state_size(self):
+        """ Number of states in the model. """
+        # Return.
+        return self.hyb_fullgb_pars['Nx']
+    
+    @property
+    def output_size(self):
+        """ Number of outputs of the model. """
+        # Return.
+        return self.hyb_fullgb_pars['Nx']
+
+    def _fxu(self, x, u):
+        """ Function to compute the 
+            derivative (RHS of the ODE)
+            for the two reaction model.
+            
+            dCa/dt = F*(Caf - Ca)/V - r1
+            dCb/dt = -F*Cb/V + r1 - 3*r2 + r3
+            dCc/dt = -F*Cc/V + r2 - r3
+        """
+        
+        # Extract the parameters (nominal value of unmeasured disturbance).
+        F = self.hyb_fullgb_pars['ps'].squeeze()
+        V = self.hyb_fullgb_pars['V']
+
+        # Get the states before scaling.
+        Ca, Cb, Cc = x[..., 0:1], x[..., 1:2], x[..., 2:3]
+        Caf = u[..., 0:1]
+
+        # Get scaling factors.
+        ymean, ystd = self.xuyscales['yscale']
+        Camean, Cbmean = ymean[0:1], ymean[1:2]
+        Castd, Cbstd = ystd[0:1], ystd[1:2]
+        Cafmean, Cafstd = self.xuyscales['uscale']
+
+        # Compute NN reaction rates.
[…]

### 2021-10-09 SimStudiesAug26-Sept8/ReacHybridFullGbFuncsWithExtraLoss.py

@@ -1,0 +1,651 @@
+# [depends] BlackBoxFuncs.py hybridId.py
+import sys
+import numpy as np
+import mpctools as mpc
+import tensorflow as tf
+from BlackBoxFuncs import fnnTf, fnn, createDenseLayers
+from hybridId import SimData
+
+# class InterpolationLayer(tf.keras.layers.Layer):
+#     """
+#     The layer to perform interpolation for RK4 predictions.
+#     Nvar: Number of variables.
+#     Np + 1: Number of variables.
+#     """
+#     def __init__(self, Nvar, Np, trainable=False, name=None):
+#         super(InterpolationLayer, self).__init__(trainable, name)
+#         self.Nvar = Nvar
+#         self.Np = Np
+
+#     def call(self, yseq):
+#         """ The main call function of the interpolation layer.
+#             yseq is of dimension: (None, (Np+1)*Nvar)
+#             Return y of dimension: (None, Np*Nvar)
+#         """
+#         yseq_interp = []
+#         for t in range(self.Np):
+#             yseq_interp += [0.5*(yseq[..., t*self.Nvar:(t+1)*self.Nvar] + 
+#                                  yseq[..., (t+1)*self.Nvar:(t+2)*self.Nvar])]
+#         # Return.
+#         return tf.concat(yseq_interp, axis=-1)
+
+class FullGbLoss(tf.keras.losses.Loss):
+
+    """ 
+        Loss function for the full grey-box model. 
+
+        lam: scaling factor for the extra cost term.
+
+        All the indices are lists.
+        yi: indices relevant to predict the measurements.
+
+        unmeasGbPredi: indices relevant to predict the unmeasured grey-box
+        states by the predictor model.
+
+        unmeasGbEsti: indices relevant to predict the unmeasured grey-box states
+        by the estimator model.
+
+    """
+
+    def __init__(self, lam, yis, unmeasGbPredis, unmeasGbEstis):
+
+        # Lambda, yi, indices relevant to the unmeasured grey-box states.
+        self.lam = lam
+        self.yis = yis
+        self.unmeasGbPredis = unmeasGbPredis
+        self.unmeasGbEstis = unmeasGbEstis
+
+        # Initialize.
+        super(FullGbLoss, self).__init__()
+
[…]

### 2021-10-09 SimStudiesAug26-Sept8/ReacHybridPartialGbFuncs.py

@@ -1,0 +1,409 @@
+# [depends] BlackBoxFuncs.py hybridId.py
+import sys
+import numpy as np
+import mpctools as mpc
+import tensorflow as tf
+from BlackBoxFuncs import fnnTf, fnn, createDenseLayers
+from hybridId import SimData
+
+class InterpolationLayer(tf.keras.layers.Layer):
+    """
+    The layer to perform interpolation for RK4 predictions.
+    Nvar: Number of variables.
+    Np + 1: Number of variables.
+    """
+    def __init__(self, Nvar, Np, trainable=False, name=None):
+        super(InterpolationLayer, self).__init__(trainable, name)
+        self.Nvar = Nvar
+        self.Np = Np
+
+    def call(self, yseq):
+        """ The main call function of the interpolation layer.
+            yseq is of dimension: (None, (Np+1)*Nvar)
+            Return y of dimension: (None, Np*Nvar)
+        """
+        yseq_interp = []
+        for t in range(self.Np):
+            yseq_interp += [0.5*(yseq[..., t*self.Nvar:(t+1)*self.Nvar] + 
+                                 yseq[..., (t+1)*self.Nvar:(t+2)*self.Nvar])]
+        # Return.
+        return tf.concat(yseq_interp, axis=-1)
+
+def getInterpolatedVals(yseq, Nvar, Np):
+    """ y is of dimension: (None, (Np+1)*Nvar)
+        Return y of dimension: (None, Np*Nvar). """
+    yseq_interp = []
+    for t in range(Np):
+        yseq_interp += [0.5*(yseq[t*Nvar:(t+1)*Nvar] + 
+                             yseq[(t+1)*Nvar:(t+2)*Nvar])]
+    # Return.
+    return np.concatenate(yseq_interp)
+
+class ReacPartialGbCell(tf.keras.layers.AbstractRNNCell):
+    """
+    RNN Cell:
+    dx_g/dt  = f_g(x_g, u) + (chosen functions).
+    y = x_g
+    r1 = NN1(Ca)
+    r2 = NN2(Cb)
+    r3 = NN3(z)
+    """
+    def __init__(self, r1Layers, r2Layers, r3Layers, Np, interpLayer,
+                       xuyscales, hyb_partialgb_pars, **kwargs):
+        super(ReacPartialGbCell, self).__init__(**kwargs)
+
+        # r1, r2, and r3 layers.
+        self.r1Layers = r1Layers
+        self.r2Layers = r2Layers
+        self.r3Layers = r3Layers
[…]

### 2021-10-09 SimStudiesAug26-Sept8/hybridId.py

@@ -1,0 +1,327 @@
+# [depends] linNonlinMPC.py
+import sys
+import random
+import numpy as np
+import mpctools as mpc
+import scipy.linalg
+import matplotlib.pyplot as plt
+import casadi
+import collections
+import pickle
+import plottools
+import time
+from linNonlinMPC import getXsYsSscost, c2dNonlin
+
+# Custom class to store datasets.
+SimData = collections.namedtuple('SimData',
+                                ['t', 'x', 'u', 'y', 'p'])
+
+class PickleTool:
+    """ Class which contains a few static methods for saving and
+        loading pkl data files conveniently. """
+
+    @staticmethod
+    def load(filename, type='write'):
+        """Wrapper to load data."""
+        if type == 'read':
+            with open(filename, "rb") as stream:
+                return pickle.load(stream)
+        if type == 'write':
+            with open(filename, "wb") as stream:
+                return pickle.load(stream)
+    
+    @staticmethod
+    def save(data_object, filename):
+        """Wrapper to pickle a data object."""
+        with open(filename, "wb") as stream:
+            pickle.dump(data_object, stream)
+
+def _sample_repeats(num_change, num_simulation_steps,
+                    mean_change, sigma_change):
+    """ Sample the number of times a repeat in each
+        of the sampled vector is required."""
+    repeat = sigma_change*np.random.randn(num_change-1) + mean_change
+    repeat = np.floor(repeat)
+    repeat = np.where(repeat<=0., 0., repeat)
+    repeat = np.append(repeat, num_simulation_steps-np.int(np.sum(repeat)))
+    return repeat.astype(int)
+
+def sample_prbs_like(*, num_change, num_steps, 
+                        lb, ub, mean_change, sigma_change, 
+                        num_constraint=0, seed=1):
+    """ Sample a PRBS like sequence.
+    num_change: Number of changes in the signal.
+    num_steps: Number of steps in the signal.
+    mean_change: mean_value after which a 
+                 change in the signal is desired.
+    sigma_change: standard deviation of changes in the signal.
+    """
[…]

### 2021-10-09 SimStudiesAug26-Sept8/reacFuncs.py

@@ -1,0 +1,138 @@
+import numpy as np
+import mpctools as mpc
+
+def plant_ode(x, u, p, parameters):
+    """ Simple ODE describing a 3D system. """
+
+    # Extract the parameters.
+    k1 = parameters['k1']
+    k2 = parameters['k2']
+    k3 = parameters['k3']
+    V = parameters['V']
+
+    # Extract the plant states into meaningful names.
+    Ca, Cb, Cc = x[0], x[1], x[2]
+    Caf = u[0]
+    F = p[0]
+
+    # Rate laws.
+    r1 = k1*Ca
+    r2 = k2*(Cb**3)
+    r3 = k3*Cc
+
+    # Write the ODEs.
+    dCabydt = F*(Caf - Ca)/V - r1
+    dCbbydt = -F*Cb/V + r1 - 3*r2 + r3
+    dCcbydt = -F*Cc/V + r2 - r3
+
+    # Return.
+    return mpc.vcat([dCabydt, dCbbydt, dCcbydt])
+
+def get_plant_pars():
+    """ Plant model parameters. """
+    
+    # Parameters.
+    parameters = {}
+    parameters['k1'] = 2e-1 # m^3/min
+    parameters['k2'] = 5e-1 # m^3/min
+    parameters['k3'] = 1e-1 # m^3/min
+    parameters['V'] = 15 # m^3
+
+    # Store the dimensions.
+    parameters['Nx'] = 3
+    parameters['Nu'] = 1
+    parameters['Ny'] = 2
+    parameters['Np'] = 1
+
+    # Sample time.
+    parameters['Delta'] = 1. # min.
+
+    # Get the steady states.
+    parameters['xs'] = np.array([1., 0.5, 0.5]) # to be rectified.
+    parameters['us'] = np.array([2.0]) # mol/m^3
+    parameters['ps'] = np.array([0.8]) # m^3/min
+
+    # Get the constraints.
+    ulb = np.array([1.0])
+    uub = np.array([3.0])
+    parameters['ulb'] = ulb
+    parameters['uub'] = uub
+
+    # Measurement indices and noise.
+    parameters['yindices'] = [0, 1]
+    parameters['Rv'] = 0*np.diag([1e-3, 1e-4])
+
+    # Return the parameters dict.
+    return parameters
+
+def get_hyb_pars(*, plant_pars, Nx):
+    """ GreyBox parameters for the partial hybrid model. """
+    
+    # Parameters.
[…]

### 2021-10-26 reac_plots.py
update the rate analysis calculation.

@@ -57,7 +57,7 @@ def plot_xudata(*, t, ylist, xlist, ulist, legend_names,
                 handle = axes[row].plot(t, u[:, 0], color)
 
             # Axes labels.
-            axes[row].set_ylabel(ylabels[row])
+            axes[row].set_ylabel(ylabels[row], rotation=False)
             axes[row].get_yaxis().set_label_coords(ylabel_xcoordinate, 0.5)
         
         # Store handles.
@@ -110,7 +110,7 @@ def plot_xsvus(*, us, xs_list, legend_names,
             handle = axes[row].plot(us, xs[:, row], color)
             
             # Y-axis label.
-            axes[row].set_ylabel(ylabels[row])
+            axes[row].set_ylabel(ylabels[row], rotation=False)
             axes[row].get_yaxis().set_label_coords(ylabel_xcoordinate, 0.5)
         
         # Store handle.
@@ -196,7 +196,7 @@ def plot_fullGbR2Errors(*, r2XGrid, r2YGrid, r2Errors,
                            ylabel_xcoordinate, left_frac, 
                            wspace, right_frac, title_y):
     """ Plot errors in reaction rate 2. """
-        
+    
     # Create figure/axes.
     figure, axes = plt.subplots(nrows=1, ncols=1, 
                                 sharex=True, figsize=figure_size,
@@ -209,8 +209,8 @@ def plot_fullGbR2Errors(*, r2XGrid, r2YGrid, r2Errors,
     figure.colorbar(mesh, ax=axes)  
 
     # Labels.
-    axes.set_ylabel(ylabel)
     axes.set_xlabel(xlabel)
+    axes.set_ylabel(ylabel, rotation=False)
     axes.set_title(title, loc='center', y=title_y)
 
     # Limits.
@@ -327,7 +327,7 @@ def main():
                             legend_colors=legend_colors, 
                             figure_size=PRESENTATION_FIGSIZE, 
                             ylabel_xcoordinate=-0.1, 
-                            title_loc=(0.25, 0.9))
+                            title_loc=(0.22, 0.9))
 
     # Steady state Concentrations.
     us = reac_ssopt[0]['us']
@@ -341,7 +341,7 @@ def main():
                           legend_names=legend_names, 
                           figure_size=PRESENTATION_FIGSIZE, 
                           ylabel_xcoordinate=-0.1, 
[…]

### 2021-10-26 reac_rateanalysis.py

@@ -134,6 +134,39 @@ def getFullGbRateErrorsInStateSpace(*, CaRange, CbRange, CcRange,
     # Return.
     return r1Errors, r2Errors
 
+def getPartialGbRateErrorsInStateSpace(*, CaRange, r1Weights, r2Weights,
+                                          Np, Ny, Nu, xuyscales, k1, k2f, k2b):
+    """ Get errors in the reaction rate laws of the 
+        Partial Grey-Box model in a chosen region of A concentration. """
+
+    # Get an empty zero array.
+    zeroArray = np.zeros((1, ))
+    zeroZArray = np.zeros((Np*(Ny + Nu), ))
+
+    # Create a list to store errors in r1.
+    r1Errors = []
+
+    # Get errors in r1.
+    for i, Ca in enumerate(CaRange):
+
+        # True rate.
+        r1, _ = getTrueR1R2(Ca=Ca, Cb=zeroArray, Cc=zeroArray,
+                            k1=k1, k2f=k2f, k2b=k2b)
+
+        # NN rate.
+        r1NN, _ = getPartialGbR1R2(Ca=Ca, Cb=zeroArray, z=zeroZArray, Np=Np,
+                                   r1Weights=r1Weights, r2Weights=r2Weights,
+                                   xuyscales=xuyscales)
+
+        # Get the error.
+        r1Errors += [np.abs(r1 - r1NN)/np.abs(r1)]
+
+    # Get the errors in an array form.
+    r1Errors = np.array(r1Errors).squeeze()
+
+    # Return.
+    return r1Errors
+
 def getRateErrorsOnGeneratedData(*, training_data, Ntstart, Np, Ny, Nu, 
                                     fGbWeights, pGbWeights, yindices, 
                                     xuyscales, k1, k2f, k2b):
@@ -228,7 +261,8 @@ def main():
     Np = reac_hybpartialgbtrain[1]['Np']
     xuyscales = reac_hybfullgbtrain[1]['xuyscales']
 
-    # Get errors in reactions 1 and 2 for the full GB model. 
+    # Get errors in Reactions 1 and 2 over the state-space for 
+    # the full GB model. 
     CaRange = list(np.arange(0.1, 1.0, 0.01)[:, np.newaxis])
     CbRange = list(np.arange(0.1, 0.7, 0.01)[:, np.newaxis])
     CcRange = list(np.arange(0.1, 0.7, 0.01)[:, np.newaxis])
[…]

### 2021-10-26 reac_plots.py
working on the python script.

@@ -420,7 +420,7 @@ def main():
     r2Errors = fGbErrorsInStateSpace['r2Errors']
     xlabel = r'$c_B$'
     ylabel = r'$c_C$'
-    title = 'Modified error, Reaction-2, Hybrid-FullGb'
+    title = 'Modified error, Reaction-2, Hybrid - FullGb'
     figures += plot_fullGbR2Errors(r2XGrid=r2XGrid,
                                    r2YGrid=r2YGrid, r2Errors=r2Errors, 
                                    figure_size=PRESENTATION_FIGSIZE, 

### 2021-10-26 reac_rateanalysis.py

@@ -263,9 +263,9 @@ def main():
 
     # Get errors in Reactions 1 and 2 over the state-space for 
     # the full GB model. 
-    CaRange = list(np.arange(0.1, 1.2, 0.01)[:, np.newaxis])
-    CbRange = list(np.arange(0.1, 0.8, 0.01)[:, np.newaxis])
-    CcRange = list(np.arange(0.1, 0.8, 0.01)[:, np.newaxis])
+    CaRange = list(np.arange(0.05, 1.2, 0.01)[:, np.newaxis])
+    CbRange = list(np.arange(0.05, 0.8, 0.01)[:, np.newaxis])
+    CcRange = list(np.arange(0.05, 0.8, 0.01)[:, np.newaxis])
     r2XGrid, r2YGrid = np.meshgrid(CbRange, CcRange)
     r1Errors, r2Errors = getFullGbRateErrorsInStateSpace(CaRange=CaRange, 
                             CbRange=CbRange, CcRange=CcRange, 

### 2021-10-26 reac_rateanalysis.py
update the talk.

@@ -105,7 +105,7 @@ def getFullGbRateErrorsInStateSpace(*, CaRange, CbRange, CcRange,
     r1Errors = np.array(r1Errors).squeeze()
 
     # Absolute and relative tolerance. 
-    r2Reltol = 1e-4
+    r2Reltol = 1e-6
     r2Abstol = 1e-4
 
     # Create empty arrays to store errors in r2.

### 2021-11-23 lib/trainingFuncs.py
working on the reactor example.

@@ -1,93 +1,0 @@
-# [depends] linNonlinMPC.py
-import sys
-import random
-import numpy as np
-import scipy.linalg
-import matplotlib.pyplot as plt
-import collections
-import pickle
-import plottools
-import time
-
-# Custom class to store datasets.
-SimData = collections.namedtuple('SimData',
-                                ['t', 'x', 'u', 'y', 'p'])
-
-def get_scaling(*, data):
-    """ Scale the input/output. """
-    
-    # Xmean.
-    xmean = np.mean(data.x, axis=0)
-    xstd = np.std(data.x, axis=0)
-    
-    # Umean.
-    umean = np.mean(data.u, axis=0)
-    ustd = np.std(data.u, axis=0)
-    
-    # Ymean.
-    ymean = np.mean(data.y, axis=0)
-    ystd = np.std(data.y, axis=0)
-    
-    # Return.
-    return dict(xscale = (xmean, xstd), 
-                uscale = (umean, ustd), 
-                yscale = (ymean, ystd))
-
-def get_train_val_data(*, Ntstart, Np, xuyscales, data_list):
-    """ Get the data for training and validation in 
-        appropriate format for training.
-        All data are already scaled.
-    """
-
-    # Get scaling parameters.
-    xmean, xstd = xuyscales['xscale']
-    umean, ustd = xuyscales['uscale']
-    ymean, ystd = xuyscales['yscale']
-    Nx, Nu, Ny = len(xmean), len(umean), len(ymean)
-
-    # Lists to store data.
-    xseq, useq, yseq = [], [], []
-    y0, z0, yz0 = [], [], []
-
-    # Loop through the data list.
-    for data in data_list:
-        
-        # Scale data.
-        x = (data.x - xmean)/xstd
-        u = (data.u - umean)/ustd
-        y = (data.y - ymean)/ystd
-
-        # Get the input and output trajectory.
-        x_traj = x[Ntstart:, :][np.newaxis, ...]
-        u_traj = u[Ntstart:, :][np.newaxis, ...]
-        y_traj = y[Ntstart:, :][np.newaxis, ...]
-
-        # Get initial states.
-        yp0seq = y[Ntstart-Np:Ntstart, :].reshape(Np*Ny, )[np.newaxis, :]
-        up0seq = u[Ntstart-Np:Ntstart, :].reshape(Np*Nu, )[np.newaxis, :]
[…]

### 2021-11-23 reac_rateanalysis.py

@@ -1,0 +1,304 @@
+# [depends] %LIB%/hybridId.py %LIB%/BlackBoxFuncs.py
+# [depends] reac_parameters.pickle reac_hybfullgbtrain.pickle
+# [depends] reac_hybpartialgbtrain.pickle
+# [makes] pickle
+import sys
+sys.path.append('lib/')
+import itertools
+import numpy as np
+from hybridId import PickleTool
+from BlackBoxFuncs import fnn
+
+def getTrueR1R2(*, Ca, Cb, Cc, k1, k2f, k2b):
+    """ Function to get the true reaction rates. """
+
+    # Get r1.
+    r1 = k1*Ca
+
+    # Get r2. 
+    r2 = k2f*(Cb**3) - k2b*Cc
+
+    # Return.
+    return r1, r2
+
+def getFullGbR1R2(*, Ca, Cb, Cc, r1Weights, r2Weights, xuyscales):
+    """ Function to get the rate laws for the full GB hybrid model. 
+        Rates for this hybrid model are parametrized such that
+        r1 = NN_r1(Ca), r2 = NN_r2(Cb, Cc)
+    """
+
+    # Get scaling.
+    ymean, ystd = xuyscales['yscale']
+    Camean, Cbmean = ymean[0:1], ymean[1:2]
+    Castd, Cbstd = ystd[0:1], ystd[1:2]
+
+    # Get r1.
+    Ca = (Ca - Camean)/Castd
+    r1 = fnn(Ca, r1Weights)*Castd 
+
+    # Get r2.
+    Cb = (Cb - Cbmean)/Cbstd
+    Cc = (Cc - Cbmean)/Cbstd
+    CbCc = np.concatenate((Cb, Cc))
+    r2 = fnn(CbCc, r2Weights)*Cbstd 
+
+    # Return.
+    return r1, r2
+
+def getPartialGbR1R2(*, Ca, Cb, z, Np, r1Weights, r2Weights, xuyscales):
+    """ Function to get the rate laws for the partial GB hybrid model. 
+        Rates for this hybrid model are parametrized such that
+        r1 = NN_r1(Ca), r2 = NN_r2(Cb, z)
+    """
+
+    # Get scaling.
+    ymean, ystd = xuyscales['yscale']
+    umean, ustd = xuyscales['uscale']
+    Camean, Cbmean = ymean[0:1], ymean[1:2]
+    Castd, Cbstd = ystd[0:1], ystd[1:2]
[…]

### 2021-11-24 lib/styrenePolyHybridFuncs.py
update the paper.

@@ -1,0 +1,324 @@
+# [depends] BlackBoxFuncs.py hybridId.py
+import sys
+import numpy as np
+import mpctools as mpc
+import tensorflow as tf
+from BlackBoxFuncs import fnnTf, fnn, createDenseLayers
+from hybridId import SimData
+
+class ReacFullGbCell(tf.keras.layers.AbstractRNNCell):
+    """
+    RNN Cell
+    x_g^+ = f_g(x_g, u; NN(x_g))
+    """
+    def __init__(self, r1Layers, r2Layers,
+                       xuyscales, hyb_fullgb_pars, **kwargs):
+        super(ReacFullGbCell, self).__init__(**kwargs)
+        
+        # Attributes.
+        self.r1Layers = r1Layers
+        self.r2Layers = r2Layers
+        self.xuyscales = xuyscales
+        self.hyb_fullgb_pars = hyb_fullgb_pars
+
+        # State size.
+        self.Nx = hyb_fullgb_pars['Ng']
+
+    @property
+    def state_size(self):
+        """ Number of states in the model. """
+        # Return.
+        return self.Nx
+    
+    @property
+    def output_size(self):
+        """ Number of outputs of the model. """
+        # Return.
+        return self.Nx
+
+    def _ode(self, x, u):
+        """ Function to compute the 
+            derivative (RHS of the ODE)
+            for the two reaction model.
+            
+            dCa/dt = F*(Caf - Ca)/V - r1
+            dCb/dt = -F*Cb/V + r1 - 3*r2
+            dCc/dt = -F*Cc/V + r2
+        """
+        
+        # Extract hybrid model parameters.
+        F = self.hyb_fullgb_pars['ps'].squeeze()
+        V = self.hyb_fullgb_pars['V']
+
+        # Get the states before scaling.
+        Ca, Cb, Cc = x[..., 0:1], x[..., 1:2], x[..., 2:3]
+        Caf = u[..., 0:1]
+
+        # Get scaling factors.
+        ymean, ystd = self.xuyscales['yscale']
+        Camean, Cbmean = ymean[0:1], ymean[1:2]
+        Castd, Cbstd = ystd[0:1], ystd[1:2]
+        Cafmean, Cafstd = self.xuyscales['uscale']
+
+        # Compute NN reaction rates.
+        r1 = fnnTf(Ca, self.r1Layers)*Castd
+        r2Input = tf.concat((Cb, Cc), axis=-1)
[…]
